### 2019-03-20 Rapport/TG_ITA/TG_ITA/ApeA/apendiceA.tex
Tranfering files from PC to Github
First time uploading files from PC to Github.

\section{Uma Primeira Seção para o Apêndice}

A matriz de Dilema Linear $M$ e o vetor de torques inerciais $b$,
utilizados na simulação são calculados segundo a formulação 
abaixo:
\begin{equation}
M=\left[ \begin{array}{ccc}
M_{11} & M_{12} & M_{13} \\
M_{21} & M_{22} & M_{23} \\
M_{31} & M_{32} & M_{33}
\end{array} \right]
\end{equation}

\begin{figure}[h]
\centering
\includegraphics[height=5cm, width=5cm]{ApeA/pragas_ciclo_cupim}
\caption{Uma figura que está no apêndice}\label{FD}
\end{figure}

### 2019-03-20 Rapport/TG_ITA/TG_ITA/Cap0/agradecimentos.tex
Tranfering files from PC to Github
First time uploading files from PC to Github.

Primeiramente, gostaria de agradecer ao Dr. Donald E. Knuth, por ter desenvolvido o \TeX.

Ao Dr. Leslie Lamport, por ter criado o \LaTeX, facilitando muito a utilização do \TeX, e assim, eu não ter que usar o Word.

Ao Prof. Dr. Meu Orientador, pela orientação e confiança depositada na realização deste trabalho.

Ao Dr. Nelson D'Ávilla, por emprestar seu nome a essa importante via de trânsito na cidade de São José dos Campos.

Ah, já estava esquecendo... agradeço também, mais uma vez ao \TeX, por ele não possuir vírus de macro :-)

### 2019-03-20 Rapport/TG_ITA/TG_ITA/Cap0/resumo.tex
Tranfering files from PC to Github
First time uploading files from PC to Github.

Aqui começa o resumo do referido trabalho. Não tenho a menor idéia do que colocar aqui. Sendo assim, vou inventar. Lá vai: Este trabalho apresenta uma metodologia de controle de posição das juntas passivas de um manipulador subatuado de uma maneira subótima. O termo subatuado se refere ao fato de que nem todas as juntas ou graus de liberdade do sistema são equipados com atuadores, o que ocorre na prática devido a falhas ou como resultado de projeto. As juntas passivas de manipuladores desse tipo são indiretamente controladas pelo movimento das juntas ativas usando as características de acoplamento da dinâmica de manipuladores. A utilização de redundância de atuação das juntas ativas permite a minimização de alguns critérios, como consumo de energia, por exemplo.
Apesar da estrutura cinemática de manipuladores subatuados ser idêntica a do totalmente atuado, em geral suas caraterísticas dinâmicas diferem devido a presença de juntas passivas. Assim, apresentamos a modelagem dinâmica de um manipulador subatuado e o conceito de índice de acoplamento. Este índice é utilizado na sequência de controle ótimo do \mbox{manipulador}.
A hipótese de que o número de juntas ativas seja maior que o número de
passivas  $(n_{a} > n_{p})$  permite o controle ótimo das juntas passivas, uma vez que na etapa de controle destas há mais entradas (torques nos atuadores das juntas ativas), que elementos a controlar (posição das juntas passivas).

### 2019-03-20 Rapport/TG_ITA/TG_ITA/Cap1/cap1.tex
Tranfering files from PC to Github
First time uploading files from PC to Github.

\section{Objetivo}
O objetivo deste projeto de mestrado é desenvolver técnicas de controle subótimo das juntas passivas (não atuadas) de um robô subatuado, incluindo o estudo teórico do tema, proposição de um método de controle e sua verificação
experimental em um manipulador de três graus de liberdade \cite{Nascimento1970}.

O teste \cite{Patagonios2001} e validação das técnicas de controle propostas foram realizados em um ambiente de simulação e no manipulador
experimental, adquirido através do projeto FAPESP $N^{\circ}$ 98/00649-5, que se encontra em funcionamento no Laboratório de Sistemas Inteligentes (LASI) do Departamento de Engenharia Elétrica da USP em São Carlos. Pode-se citar \cite{Furmento1995}:
\begin{itemize}
\item Isso;
\item Aquilo; e
\item Aquele outro.
\end{itemize}

\section{Motivação}
Manipuladores mecânicos \cite{Sbornian2002} vêm sendo utilizados há várias décadas para a automação de tarefas
repetitivas em ambientes industriais, ambientes estes de fácil acesso tanto em termos físicos quanto em termos de baixo
risco à saúde humana. Nos últimos anos, verifica-se uma utilização cada vez maior de manipuladores em
ambientes de difícil acesso ou inóspitos, como no interior de usinas nucleares, no fundo dos oceanos e no
espaço. A localização dos manipuladores nesta nova gama de aplicações faz com que sua manutenção,
Dpós uma falha mecânica ou elétrica, seja custosa e demorada, portanto estes mecanismos requerem sofisticadas
metodologias de controle tolerante a falhas \cite{ITALUS2004}.

Após a ocorrência de uma falha em um de seus atuadores, o manipulador torna-se um sistema subatuado. Um sistema também pode se tornar subatuado quando é projetado  dessa maneira, ou quando o operador deliberadamente mantém um ou mais atuadores disponíveis inoperantes durante uma tarefa. Reduzindo o número de atuadores sem reduzir o número de graus de
liberdade e ajustando-se o sistema de controle adequado, pode-se obter um mecanismo cujo consumo de energia é menor, mas cujas propriedades são mantidas \cite{Arystides1994}.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{Cap1/cupim}
\caption{Proibido estacionar cupins. Legenda grande, com o objetivo de demonstrar a indentação na lista de figuras.}
\label{cupim}
\end{figure}

Controle do manipulador após uma falha é fundamental do ponto de vista de operação, principalmente nos casos descritos acima, em que a localização do manipulador impede sua manutenção de forma fácil. Recentemente tem havido a combinação
de algorítmos de detecção e isolação de falhas com os de controle pós-falha em um método unificado. Uma extensão desse trabalho, que vê o problema de controle tolerante a falhas através de uma perspectiva integrada, foi proposta por
{marcel4}. Os autores apresentam um ambiente híbrido consistindo de três unidades básicas que garantem a compleição de tarefas na presença de qualquer número de juntas falhas (Figura \ref{cupim}). A primeira unidade é um esquema de detecção
e isolação de falhas que continuamente monitora o manipulador para detectar e identificar possíveis falhas nas juntas. A segunda unidade é responsável pela reconfiguração do controle. A terceira unidade é composta de algorítmos de
controle apropriados para cada tipo de configuração do robô, baseado na informação da unidade de reconfiguração \cite{COFFEE2000}.

No presente trabalho nos concentramos na unidade de algorítmo de controle, e mais especificamente no problema de controle da posição  angular de uma junta falha para qualquer posição desejada de uma maneira subótima, quando dispomos
de redundância de atuação para a realização dessa tarefa. O termo subótimo se deve ao fato de que não há garantias de otimalidade em vista das não-linearidades inerentes ao sistema e de outros fatores que serão abordados nos capítulos posteriores. Ao longo do texto, para simplificação, usaremos tanto o termo subótimo como ótimo para nos referirmos à metodologia utilizada.

Segundo, o critério de otimização utilizado será o acoplamento entre as juntas do
manipulador e neste caso, temos um sistema redundante quando ocorre falha de uma das juntas do manipulador de três juntas, e seu posicionamento é controlado pelas duas restantes. Nossa solução para o problema é baseada na formulação
de redundância local, extensivamente estudada no contexto de cinemática inversa ({nakamura}). A principal contribuição deste trabalho é a extensão deste método usando as equações dinâmicas de manipuladores subatuados e a utilização do índice de acoplamento como um critério para a minimização do torque e da energia gasta pelo sistema durante o controle das juntas falhas.

\begin{figure}[ht!]
\centering
\includegraphics[width=1\textwidth]{Cap1/cupimconcreto}
\caption{Exemplo real de cupim frente ao seu dilema.}
\label{FDII}
\end{figure}

\section{Organização do trabalho}
\subsection{Sub-organização}
O capítulo 1 contém a introdução do trabalho, onde são expostos o objetivo, a motivação do mesmo, a descrição do sistema e a formulação do problema com a nomenclatura utilizada; além de uma revisão bibliográfica da literatura relacionada ao tema do trabalho.

\subsubsection{SubSub-organização}

No capítulo 2 apresentamos a modelagem dinâmica de um manipulador subatuado e o conceito de índice de acoplamento para medir o acoplamento dinâmico entre as juntas ativas e passivas. Este índice é utilizado para a análise e projeto de uma metodologia de controle subótimo do manipulador.

\subsubsection{Outra subsub-organizacao}

O capítulo 3 apresenta o controle subótimo de manipuladores através de redundância de atuação. Descreve-se a técnica de controle ponto a ponto de manipuladores subatuados. A seguir mostramos  a linearização destes por realimentação, cujo efeito é linearizar e desacoplar o sistema não linear. Finalmente é proposta uma sequência de controle subótimo local das juntas passivas visando a minimização de certos critérios como torque, velocidade e em particular a energia consumida pelo sistema. Este é de fato o tema principal deste mestrado.

É também apresentado no capítulo 4 um resumo do projeto de controladores  $H_{2}$ e $H_{\infty}$, cuja principal vantagem é a robustez na presença de incertezas paramétricas e distúrbios externos.

O capítulo 5 mostra as características e a operação do robô e do ambiente de simulação utilizados nos testes e experimentação da metodologia apresentada.

Os procedimentos da metodologia e os resultados obtidos para algumas configurações e diferentes controladores encontram-se no capítulo 6.

No capítulo 7 são apresentadas as conclusões do trabalho.

Quatro apêndices fazem parte do trabalho. O apêndice A apresenta alguns tópicos de álgebra linear que são a base do método proposto. No apêndice B são mostradas as equações da matriz de inércia e do vetor de torques não-inerciais
utilizados na modelagem dinâmica do manipulador. No apêndice C temos as expressões literais dessas equações feitas no software MAPLE e no apêndice D alguns programas feitos no software MATLAB utilizados no projeto \cite{Furmento1995}\cite{Morgado2003}.

### 2019-03-20 Rapport/TG_ITA/TG_ITA/Cap2/cap2.tex
Tranfering files from PC to Github
First time uploading files from PC to Github.

\section{Modelagem no espaço das juntas}
Manipuladores subatuados diferem dos totalmente atuados pois são equipados com um número de atuadores que é sempre menor que o número de graus de liberdade (GDL). Portanto, nem todos os GDL podem ser controlados ativamente ao mesmo tempo \cite{Sbornian2004}. Por exemplo, com um manipulador planar de 3 juntas equipado com dois atuadores, ou seja, duas juntas ativas e
uma passiva, pode-se controlar ao mesmo tempo duas das juntas a qualquer instante, mas não todas. Para controlar todas as juntas de um manipulador subatuado, deve-se usar um controle sequencial. Este princípio foi provado pela primeira vez por {arai} usando  argumentos dinâmicos linearizados \cite{Joea2003}, e é a base para a modelagem no espaço das juntas e no espaço Cartesiano. A Tabela \ref{minhatab} apresenta os resultados \cite{Assenmacher1993,Silberschatz1991,Caromel1998}.

\begin{table}
\caption{Exemplo de uma Tabela}
\label{minhatab}

\center
\begin{tabular}{cccc}
  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
  \hline
	Parâmetro & Unidade & Valor da simulação & Valor experimental   \\
	\hline
  Comprimento, $\alpha$ & $m$ &  $8,23$  & $8,54$ \\
  Altura, $\beta$ & $m$     &  $29,1$ & $28,3$\\
	Velocidade, $v$ & $m/s$  &  $60,2$ & $67,3$\\
	\hline
\end{tabular}
\end{table}

Devido ao fato de que no máximo $n_{a}$ coordenadas generalizadas (ângulos das juntas ou variáveis cartesianas) podem ser controladas num dado instante, o vetor de coordenadas generalizadas é dividido em duas partes, representando as coordenadas generalizadas ativas e as coordenadas generalizadas passivas \cite{Callaghan1995}.

\begin{figure}[ht]
\centering
\includegraphics[width=0.75\textwidth]{Cap2/spiderrobot}
\caption{Cupim cibernético.}\label{FDIII}
\end{figure}

Considerando um robô manipulador rígido, malha aberta, e de $n$-juntas em série. Seja $q$ a representação de seu vetor de posição angular das juntas  e $\tau$ a representação de seu vetor de torque. A equação dinâmica pelo método de
Lagrange é dada por:
\begin{equation} \label{eq:lagr1}
\frac{d}{dt}(\frac{\partial L}{\partial \dot{q}})-\frac{\partial L}{\partial q}=\tau^{T}.
\end{equation}
O Lagrangiano $L$ é definido como a diferença entre as energias cinética e potencial do sistema:
\begin{equation} \label{L}
L=T-P
\end{equation}

A energia cinética total dos ligamentos é representada:
\begin{equation} \label{energT}
T=\frac{1}{2}\dot{q}^{T}M(q)\dot{q}
\end{equation}

### 2019-03-20 Rapport/TG_ITA/TG_ITA/Cap3/cap3.tex
Tranfering files from PC to Github
First time uploading files from PC to Github.

\section{Controle combinado}
Conforme vimos na seção \ref{ectq3} podemos controlar um sistema nao linear como  através da técnica do torque computado, usando um controlador PD dado por:
\begin{equation} \label{ectq3}
\tau'=\ddot{q}_d+K_v(\dot{q}_d-\dot{q})+K_p(q_d-q) \; ,
\end{equation}
sendo $q_{d}$, $\dot{q}_{d}$ e $\ddot{q}_{d}$ a posição desejada, a velocidade desejada e a aceleração desejada; $K_p$
e $K_v$ são matrizes diagonais $n \times n$, sendo que cada elemento da diagonal é um ganho positivo e escalar.

Aqui $M_{est}$ e $b_{est}$ são modelos estimados da matriz de inércia, $M$, e do vetor de torques não inerciais, $b$, do robô real,  respectivamente. A equação de malha fechada do sistema é:
\begin{equation} \label{ectq4}
\ddot{e}+K_v\dot{e}+K_pe=M_{est}^{-1}[(M-M_{est})\ddot{q}+(b-b_{est})] \; .
\end{equation}

Em um manipulador real, podem existir distúrbios externos tais como atrito, variação de torque dos atuadores, e perturbações em virtude  das cargas no robô. Se a soma destes distúrbios for definida como $d_{ext}$ e adicionada à (\ref{ectq4}), teremos
\begin{equation} \label{ectq5}
\ddot{e}+K_v\dot{e}+K_pe=M_{est}^{-1}[(M-M_{est})\ddot{q}+(b-b_{est})+d_{ext}] \; .
\end{equation}

### 2019-03-20 Rapport/TG_ITA/TG_ITA/Cap4/cap4.tex
Tranfering files from PC to Github
First time uploading files from PC to Github.

%\section{Conclusão}

Neste trabalho realizou-se o projeto de uma metodologia de controle subótimo redundante da junta passiva de um manipulador com três graus de liberdade instantaneamente. Para este propósito usou-se nas formulações o vetor gradiente de uma função escalar que estima o acoplamento entre a junta passiva e as ativas desse manipulador. Aqui a redundância
foi usada da melhor maneira possível sem focalizar o efeito global. Portanto, este método deve ser denominado de \emph{controle ótimo local por redundância}. A principal vantagem dessa formulação é a computação em tempo real, que é
necessária para o controle do manipulador experimental. Além disso esse método pode ser usado com diferentes tipos de controladores, uma vez que as alterações são feitas nas equações dinâmicas do manipulador.

A consequência direta observada nessa formulação é a redução dos torques na fase de controle da junta passiva, e consequente redução da energia elétrica gasta. Isso ocorre devido ao fato de que ao longo da trajetória do manipulador
o índice de acoplamento de torque tende a ser maximizado, e portanto, menor é o torque necessário nos atuadores para se conseguir o posicionamento da junta passiva do manipulador.

Outros resultados indiretos obtidos são: um movimento mais uniforme e suave do manipulador e um tempo de acomodação menor tanto no posicionamento da junta passiva quanto das ativas, conforme podemos obervar nos gráficos de desempenho dos resultados apresentados. Isso ocorre porque a maximização do acoplamento entre as juntas facilita o controle. Assim
ocorrem menos picos de torque, e como as juntas ativas tem ``menos trabalho'' para posicionar a passiva estas se movem menos na direçao contrária ao movimento daquelas, diminuindo assim as velocidades alcançadas e os tempos de posicionamento.

Uma extensão deste trabalho pode ser a implementação de um \emph{controle ótimo global por redundância} da junta passiva do manipulador. Para isto pode-se fazer o planejamento \emph{off-line} da trajetória das juntas de modo a minimizar a energia consumida. Alguns estudos foram feitos nesse sentido, usando o Princípio Mínimo de Pontryagin, mas sem resultados satisfatórios até o momento.

### 2019-03-20 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
Week 1 of Coursera
A few notes added.

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}
\section{Coursera}
\subsubsection{Week 1}
Machine Learning algorithms :
	Supervised learning
	Unsupervised learning
	Reinforcement Learning
	Recommender systems

Supervised Learning
	based on right answers
	Regression problem: predict continuous valued output (price)
	Classification problem: output value of 0 or 1
	%% COURSERA TEXT
	To describe the supervised learning problem slightly more formally, our goal is, given a training set, to learn a function h : X → Y so that h(x) is a “good” predictor for the corresponding value of y. For historical reasons, this function h is called a hypothesis. Seen pictorially, the process is therefore like this:
%%%%

Unsupervised Learning
%% COURSERA TEXT
	Unsupervised learning allows us to approach problems with little or no idea what our results should look like. We can derive structure from data where we don't necessarily know the effect of the variables.

We can derive this structure by clustering the data based on relationships among the variables in the data.

With unsupervised learning there is no feedback based on the prediction results.
%%%%
	Not told what to do -> find some structure in data
	Clustering Algorithm
	not give the algorithm the right answers
	Cocktail party problem: take a data that seems dated together and separate them (two audio sources),

Hypothesis Function

Cost Function
	Cost Function measures the accuracy of our hypothesis function.

\end{document}

### 2019-03-21 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
Assignment Week 2 Coursera
Linear Regression. Single and multivariables.

@@
 \documentclass[12pt,a4paper]{report}
+\newcommand{\dd}[1]{\mathrm{d}#1}
 \usepackage[utf8]{inputenc}
@@
 	%% COURSERA TEXT
-	To describe the supervised learning problem slightly more formally, our goal is, given a training set, to learn a function h : X → Y so that h(x) is a “good” predictor for the corresponding value of y.
+	To describe the supervised learning problem slightly more formally, our goal is, given a training set, to learn a function h : X -> Y so that h(x) is a good predictor for the corresponding value of y.
 For historical reasons, this function h is called a hypothesis.
@@
 
+Gradient descent Algorithm
+	Optimize parameters in hypothesis function to minimize the cost function
+	Convex functions has only a global optima, no local optima (good for working with)
+	"Batch" gradient descent: each step of gradient descent uses all the training examples
+
+Learning rate
+	Controls how big the step is in the optimization process
+
+\subsubsection{Week 2}
+Fix a notation thorughout the work
+	m -> the number of training examples i
+	n -> the number of features j\\
+	$ x^{(i)}_{j} $\\
+	$ i = 1,...,m $\\
+	$ j = 0,...,n $
+	i are the lines and j the columns.
+Gradient descent for multiple variables
+
+Feature scale
+	mean normalize data to converge faster (do not apply to inserted $ x_{0} $ that equals 1 for example). Adjust input values as shown in this formula.
+	\begin{equation}
+	x_{i} := \frac{x_{i}-\mu_{i}}{s_{i}}
+	\end{equation}
+
+	where $ \mu_{i} $ is the average of all values for feature (i) and $ s_{i} $ is the range of values ($ max(x_{i})-min(x_{i}) $) or the standard deviation of a given feature $ i $.
+	get all features between -1 and 1 range or 0 and 1
+
+Gradient descent $\mathcal{O}(n^{2})$
+	\begin{equation}
+	\theta_{j} := \theta_{j} - \alpha \frac{\partial}{\partial \theta_{j}}J(\theta)
+	\end{equation}
+	works well when n is large
+
+Learning rate
+	how to choose: look graph min J vs No. of iterations. It should decrease at every iteration.
+
+Polynomial regression
+	Create new features to fit a polynomial regression to a linear regression.
+	E.g.
+
+	\begin{equation}
+	x_{0} = (feature)
+	\end{equation}
+	\begin{equation}
+	x_{1} = (feature)^{2}
+	\end{equation}
+
+Analytic solution (faster convergence)
+	The value of $ \theta = (X^{T}X)^{-1}X^{T}y $ (normal equation) gives the optimal value of $\theta$ that minimizes $ J(\theta) $
+	Don't need to iterate nor choose $\alpha$.
+	Compute the inverse costs a lot $\mathcal{O}(n^{3})$. So it is preferably to use if n is small over the iterative method.
+	Noninvertibility: pinv -> pseudo inverse: calculates the value of the inverse even if it is non inversible.
+
 \end{document}

### 2019-03-21 Coursera/coursera_week2.py

@@ -1,0 +1,27 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Mar 21 13:33:43 2019
+
+@author: Eduardo D
+"""
+# Import libraries
+import numpy as np
+import random as rd
+import matplotlib.pyplot as plt
+
+a = np.linspace(1,100,100)
+y = np.sin(a)
+plt.plot(a,y)
+plt.figure()
+plt.plot(y,a)
+
+a.dot(y) # equivalent to dot product
+
+b = np.array([1, 2, 3, 4])
+a3 = np.array(b, b ,b)
+print(a3)
+
+
+for i in range(10):
+    print(i*rd.randint(0,1))
+    

### 2019-03-21 Coursera/machine-learning-ex1/machine-learning-ex1/ex1/computeCostMulti.m

@@ -1,0 +1,27 @@
+function J = computeCostMulti(X, y, theta)
+%COMPUTECOSTMULTI Compute cost for linear regression with multiple variables
+%   J = COMPUTECOSTMULTI(X, y, theta) computes the cost of using theta as the
+%   parameter for linear regression to fit the data points in X and y
+
+% Initialize some useful values
+m = length(y); % number of training examples
+
+% You need to return the following variables correctly 
+J = 0;
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Compute the cost of a particular choice of theta
+%               You should set J to the cost.
+
+
+J = 1/(2*m)*(X*theta-y)'*(X*theta-y);
+
+% equivalent in noob code
+% for i = 1:m
+%     J = J + 1/(2*m) * ((theta'*X(i,:)'-y(i)).^2);
+% end
+
+
+% =========================================================================
+
+end

### 2019-03-21 Coursera/machine-learning-ex1/machine-learning-ex1/ex1/ex1.m

@@ -1,0 +1,140 @@
+%% Machine Learning Online Class - Exercise 1: Linear Regression
+
+%  Instructions
+%  ------------
+%
+%  This file contains code that helps you get started on the
+%  linear exercise. You will need to complete the following functions
+%  in this exericse:
+%
+%     warmUpExercise.m
+%     plotData.m
+%     gradientDescent.m
+%     computeCost.m
+%     gradientDescentMulti.m
+%     computeCostMulti.m
+%     featureNormalize.m
+%     normalEqn.m
+%
+%  For this exercise, you will not need to change any code in this file,
+%  or any other files other than those mentioned above.
+%
+% x refers to the population size in 10,000s
+% y refers to the profit in $10,000s
+%
+
+%% Initialization
+clear ; close all; clc
+
+%% ==================== Part 1: Basic Function ====================
+% Complete warmUpExercise.m
+fprintf('Running warmUpExercise ... \n');
+fprintf('5x5 Identity Matrix: \n');
+warmUpExercise()
+
+fprintf('Program paused. Press enter to continue.\n');
+pause;
+
+
+%% ======================= Part 2: Plotting =======================
+fprintf('Plotting Data ...\n')
+data = load('ex1data1.txt');
+X = data(:, 1); y = data(:, 2);
+m = length(y); % number of training examples
+
+% Plot Data
+% Note: You have to complete the code in plotData.m
+plotData(X, y);
+
+fprintf('Program paused. Press enter to continue.\n');
+pause;
+
+%% =================== Part 3: Cost and Gradient descent ===================
+
+X = [ones(m, 1), data(:,1)]; % Add a column of ones to x
+theta = zeros(2, 1); % initialize fitting parameters
+
+% Some gradient descent settings
+iterations = 1500;
+alpha = 0.01;
+
+fprintf('\nTesting the cost function ...\n')
+% compute and display initial cost
+J = computeCost(X, y, theta);
+fprintf('With theta = [0 ; 0]\nCost computed = %f\n', J);
+fprintf('Expected cost value (approx) 32.07\n');
+
+% further testing of the cost function
+J = computeCost(X, y, [-1 ; 2]);
+fprintf('\nWith theta = [-1 ; 2]\nCost computed = %f\n', J);
+fprintf('Expected cost value (approx) 54.24\n');
+
[…]

### 2019-03-21 Coursera/machine-learning-ex1/machine-learning-ex1/ex1/ex1_multi.m

@@ -1,0 +1,172 @@
+%% Machine Learning Online Class
+%  Exercise 1: Linear regression with multiple variables
+%
+%  Instructions
+%  ------------
+% 
+%  This file contains code that helps you get started on the
+%  linear regression exercise. 
+%
+%  You will need to complete the following functions in this 
+%  exericse:
+%
+%     warmUpExercise.m
+%     plotData.m
+%     gradientDescent.m
+%     computeCost.m
+%     gradientDescentMulti.m
+%     computeCostMulti.m
+%     featureNormalize.m
+%     normalEqn.m
+%
+%  For this part of the exercise, you will need to change some
+%  parts of the code below for various experiments (e.g., changing
+%  learning rates).
+%
+
+%% Initialization
+
+%% ================ Part 1: Feature Normalization ================
+
+%% Clear and Close Figures
+clear ; close all; clc
+
+fprintf('Loading data ...\n');
+
+%% Load Data
+data = load('ex1data2.txt');
+X = data(:, 1:2);
+y = data(:, 3);
+m = length(y);
+
+% Print out some data points
+fprintf('First 10 examples from the dataset: \n');
+fprintf(' x = [%.0f %.0f], y = %.0f \n', [X(1:10,:) y(1:10,:)]');
+
+fprintf('Program paused. Press enter to continue.\n');
+pause;
+
+% Scale features and set them to zero mean
+fprintf('Normalizing Features ...\n');
+
+[X mu sigma] = featureNormalize(X);
+
+% Add intercept term to X
+X = [ones(m, 1) X];
+
+
+%% ================ Part 2: Gradient Descent ================
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: We have provided you with the following starter
+%               code that runs gradient descent with a particular
+%               learning rate (alpha). 
+%
+%               Your task is to first make sure that your functions - 
+%               computeCost and gradientDescent already work with 
+%               this starter code and support multiple variables.
+%
+%               After that, try running gradient descent with 
+%               different values of alpha and see which one gives
+%               you the best result.
+%
+%               Finally, you should complete the code at the end
+%               to predict the price of a 1650 sq-ft, 3 br house.
+%
+% Hint: By using the 'hold on' command, you can plot multiple
+%       graphs on the same figure.
+%
[…]

### 2019-03-21 Coursera/machine-learning-ex1/machine-learning-ex1/ex1/featureNormalize.m

@@ -1,0 +1,41 @@
+function [X_norm, mu, sigma] = featureNormalize(X)
+%FEATURENORMALIZE Normalizes the features in X 
+%   FEATURENORMALIZE(X) returns a normalized version of X where
+%   the mean value of each feature is 0 and the standard deviation
+%   is 1. This is often a good preprocessing step to do when
+%   working with learning algorithms.
+
+% You need to set these values correctly
+X_norm = X;
+mu = zeros(1, size(X, 2));
+sigma = zeros(1, size(X, 2));
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: First, for each feature dimension, compute the mean
+%               of the feature and subtract it from the dataset,
+%               storing the mean value in mu. Next, compute the 
+%               standard deviation of each feature and divide
+%               each feature by it's standard deviation, storing
+%               the standard deviation in sigma. 
+%
+%               Note that X is a matrix where each column is a 
+%               feature and each row is an example. You need 
+%               to perform the normalization separately for 
+%               each feature. 
+%
+% Hint: You might find the 'mean' and 'std' functions useful.
+%       
+
+
+mu = mean(X);
+sigma = std(X);
+
+X_norm = (X-mu)./sigma;
+
+
+
+
+
+% ============================================================
+
+end

### 2019-03-21 Coursera/machine-learning-ex1/machine-learning-ex1/ex1/gradientDescent.m

@@ -1,0 +1,39 @@
+function [theta, J_history] = gradientDescent(X, y, theta, alpha, num_iters)
+%GRADIENTDESCENT Performs gradient descent to learn theta
+%   theta = GRADIENTDESCENT(X, y, theta, alpha, num_iters) updates theta by 
+%   taking num_iters gradient steps with learning rate alpha
+
+% Initialize some useful values
+m = length(y); % number of training examples
+J_history = zeros(num_iters, 1);
+
+for iter = 1:num_iters
+
+    % ====================== YOUR CODE HERE ======================
+    % Instructions: Perform a single gradient step on the parameter vector
+    %               theta. 
+    %
+    % Hint: While debugging, it can be useful to print out the values
+    %       of the cost function (computeCost) and gradient here.
+    %
+    thetaTemp = zeros(length(theta),1);
+
+    for j = 1:length(theta)
+        for i = 1:m
+            thetaTemp(j) = thetaTemp(j) - alpha*1/m*(theta'*X(i,:)'-y(i))*X(i,j);
+        end
+    end
+    for j = 1:length(theta)
+       theta(j) = theta(j)+thetaTemp(j); 
+    end
+
+
+
+    % ============================================================
+
+    % Save the cost J in every iteration    
+    J_history(iter) = computeCost(X, y, theta);
+
+end
+
+end

### 2019-03-21 Coursera/machine-learning-ex1/machine-learning-ex1/ex1/gradientDescentMulti.m

@@ -1,0 +1,47 @@
+function [theta, J_history] = gradientDescentMulti(X, y, theta, alpha, num_iters)
+%GRADIENTDESCENTMULTI Performs gradient descent to learn theta
+%   theta = GRADIENTDESCENTMULTI(x, y, theta, alpha, num_iters) updates theta by
+%   taking num_iters gradient steps with learning rate alpha
+
+% Initialize some useful values
+m = length(y); % number of training examples
+J_history = zeros(num_iters, 1);
+
+for iter = 1:num_iters
+
+    % ====================== YOUR CODE HERE ======================
+    % Instructions: Perform a single gradient step on the parameter vector
+    %               theta. 
+    %
+    % Hint: While debugging, it can be useful to print out the values
+    %       of the cost function (computeCostMulti) and gradient here.
+    %
+
+
+   thetaTemp = zeros(length(theta),1);
+
+    for j = 1:length(theta)
+        for i = 1:m
+            thetaTemp(j) = thetaTemp(j) - alpha*1/m*(theta'*X(i,:)'-y(i))*X(i,j);
+        end
+    end
+    for j = 1:length(theta)
+       theta(j) = theta(j)+thetaTemp(j); 
+    end
+
+
+
+
+
+
+
+
+
+    % ============================================================
+
+    % Save the cost J in every iteration    
+    J_history(iter) = computeCostMulti(X, y, theta);
+
+end
+
+end

### 2019-03-21 Coursera/machine-learning-ex1/machine-learning-ex1/ex1/plotData.m

@@ -1,0 +1,28 @@
+function plotData(x, y)
+%PLOTDATA Plots the data points x and y into a new figure 
+%   PLOTDATA(x,y) plots the data points and gives the figure axes labels of
+%   population and profit.
+
+figure; % open a new figure window
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Plot the training data into a figure using the 
+%               "figure" and "plot" commands. Set the axes labels using
+%               the "xlabel" and "ylabel" commands. Assume the 
+%               population and revenue data have been passed in
+%               as the x and y arguments of this function.
+%
+% Hint: You can use the 'rx' option with plot to have the markers
+%       appear as red crosses. Furthermore, you can make the
+%       markers larger by using plot(..., 'rx', 'MarkerSize', 10);
+
+plot(x, y, 'rx', 'MarkerSize', 10);
+ylabel('Profit in $10,000s'); % Set the y?axis label
+xlabel('Population of City in 10,000s');
+
+
+
+
+% ============================================================
+
+end

### 2019-03-22 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
First coding of communication system
And notes chapter 1 book intro to comm systems

@@
 Analytic solution (faster convergence)
-	The value of $ \theta = (X^{T}X)^{-1}X^{T}y $ (normal equation) gives the optimal value of $\theta$ that minimizes $ J(\theta) $
+	The value of $ \theta = (X^{T}X)^{-1}X^{T}y $ (normal equation) gives the optimal value of $\theta$ that minimizes $ J(\theta) $ and doesn't need feature scaling.
 	Don't need to iterate nor choose $\alpha$.

### 2019-03-22 Rapport/TG_ITA/Bibliographic_analysis/General_research.tex
First coding of communication system
And notes chapter 1 book intro to comm systems

\documentclass[12pt,a4paper]{report}
\newcommand{\dd}[1]{\mathrm{d}#1}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{tikz}
\begin{document}
\section{General Research}
\subsubsection{Communication Chain}

Communication is the act of conveying intended meanings from one entity or group to another through the use of mutually understood signs and semiotic rules

\subsection{Introduction to digital communication}

Claude Channon - creation of Informational Technology

Step by step of a communication system: 
	1. Communication sources  : speech waveforms,
image waveforms, text files, as being representable by binary sequences
	2. convert the source output into a binary sequence
	3. convert that binary sequence into a form suitable for transmission over particular physical media
such as cable, twisted wire pair, optical fiber, or electromagnetic radiation through space.

	\textbf{Digital communication systems}, by definition, are communication systems that use such a digital1
sequence as an interface between the source and the channel input (and similarly between the
channel output and final destination)

Placing a binary interface between source and channel. The source encoder converts the source output to a binary sequence and the channel encoder (often
called a modulator) processes the binary sequence for transmission over the channel.
The channel decoder (demodulator) recreates the incoming binary sequence (hopefully
reliably), and the source decoder recreates the source output

Benefits of a digital interface between source and channel:
\begin{itemize}
\item Digital hardware has become so cheap, reliable, and miniaturized, that digital interfaces are
eminently practical.
\item A standardized binary interface between source and channel simplifies implementation and
understanding, since source coding/decoding can be done independently of the channel,
and, similarly, channel coding/decoding can be done independently of the source
\item A standardized binary interface between source and channel simplifies networking, which
now reduces to sending binary sequences through the network
\item One of the most important of Shannon’s information theoretic results is that if a source
can be transmitted over a channel in any way at all, it can be transmitted using a binary
interface between source and channel. This is known as the source/channel separation
theorem
\end{itemize}

Definition of a digital sequence: A digital sequence is a sequence made up of elements from a finite alphabet (e.g., the binary digits {0, 1},
the decimal digits {0, 1, . . . , 9} )

Fundamental architectural principles:
	\textit{Standarized interfaces} and \textit{layering}
	A standardized interface allows the user or equipment on one side of the interface to ignore all
details about the other side of the interface except for certain specified interface characteristics
	The idea of layering in communication systems is to break up communication functions into a
string of separate layers as illustrated in Figure 1.2.
Each layer consists of an input module at the input end of a communcation system and a peer
output module at the other end. The input module at layer i processes the information received
from layer i+1 and sends the processed information on to layer i-1. The peer output module at
layer i works in the opposite direction, processing the received information from layer i-1 and
sending it on to layer i.

Encoding and decoding - basics: layer modulator as converting a
binary sequence to a waveform, with the peer demodulator converting the waveform back to the binary sequence

\subsubsection{Source coding}
	Discrete source:
	function of converting the input from its original
form into a sequence of bits

	The simplest source coding techniques apply to discrete sources and simply involve representing
each succesive source symbol by a sequence of binary digits
	upper-case letters, lower-case
letters, and a great many special symbols may be converted into 8-bit blocks (“bytes”) using
the standard ASCII code

\subsubsection{Communication channels}

	Channel encoding and decoding when the channel is the physical medium (either with or without amplifiers, antennas, lasers, etc.) is usually called (digital) modulation and demodulation
respectively.

	binary sequence is first converted to a baseband
waveform and the resulting baseband waveform is converted to bandpass by the same type of
procedure used for analog modulation. As will be seen, the challenging part of this problem is the conversion of binary data to baseband waveforms.

	As in the study of any type of system, a channel is usually viewed in terms of its possible inputs,
its possible outputs, and a description of how the input affects the output. This description is
usually probabilistic. If a channel were simply a linear time-invariant system (e.g., a filter), then
it could be completely characterized by its impulse response or frequency response. However,
the channels here (and channels in practice) always have an extra ingredient— noise

	it was Shannon, in 1948, who realized that noise provides
the fundamental limitation to performance in communication systems

	Channel model: $ y(t) = u(t) + w(t) $
	where $u(t)$ is a waveform input and $w(t)$ an added noise waveform. Both waveform are viewed as random processes.
	Types of channels: Additive channel model, Linear Gaussian channel (the waveform u is filtered), random variably channels, etc.
	The channel with encoder and decoder is called \textit{physical layer}.

	\subsubsection*{Modulations}

	2-PAM modulation: binary pulse amplitude modulation. Bits -1/1 instead of 0/1.

	\subsubsection{Error correction}

	channel encoder separated in two layers: error-correcting code and a simple modulator.

	Shannon showed was the very unintuitive fact that more sophisticated coding schemes can
achieve arbitrarily low error probability at any data rate above a value known as the \textit{channel capacity.}

	The channel capacity is a function of the probabilistic description of the output conditional on each possible input. Conversely, it is not possible to achieve low error probability at rates above the channel capacity.

	\subsubsection{Definitions}
	\begin{itemize}
	\item $E_{b}$, energy per bit.
	\item $E_{s}=nE_{b}$, energy per symbol with $n$ bits.
	\item $T_{b}$, bit duration.
	\item $\frac{1}{2}N_{0}$, noise power spectral density (W/Hz)
	\item $P_{b}$, probability of bit-error.
	\item $P_{s}$, probability of symbol-error.
	\end{itemize}

\end{document}

### 2019-03-24 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
Assigment week 2

@@
 
+Vectorized implementation of Gradient descent $\mathcal{O}(n^{2})$
+	\begin{equation}
+	\theta := \theta - \frac{\alpha}{m} X^{T}(g(X\theta)-Y)	
+		\end{equation}
 Learning rate
@@
 
+	\subsubsection{Week 3}
+
+Classification problem: take a bunch of data and outputs discrete  values
+	multi class: several discrete outcomes
+	binary class: 2 outputs (0,1)
+
+Logistic regression (classification algorithm): output is always between 0 and 1.
+sigmoid function or logistic function: $g(z) = \frac{1}{1+e^{-z}}$.
+Hypothesis: $h(x)=g(\theta^{T}x)$ maps any real number to the (0,l) interval.
+$h(x)$ will give the probability that our output is 1.
+
+Decision boundary: line where the hypothesis equals to 0.5. Separates 2 regions, greater than and less than 0.5 to the hypothesis function. The decision boundaries can be linear or not.
+	The decision boundary is a property of the hypothesis function. The greater the order of the polynomial, the more complex is the shape.
+
+Cost function: find a function that is convex to arrive in a global minimum.
+
+Optimization algorithms:
+	\begin{itemize}
+	\item Gradient descent
+	\item Conjugate gradient
+	\item BFGS
+	\item L-BFGS
+	\end{itemize}
+
+No need to manually pick $\alpha$ in the 3 last algorithms. They are fasten than gradient descent. However, they are more complex
+
+Multiclass classification
+	One vs all (one vs rest)
+
+Avoid overfitting: doesn't add too much features, otherwise the training data will not predict well values.	The cost function is almost zero to the training example.
+	To fight this problem: reduce the number of features (select which features to keep), regularization (keep all features but reduce the magnitude of parameters theta), Regularization works well when we have a lot of slightly useful features.
+	Add a regularization term in the end of the cost function. A regularization parameter is set to make a trade off between how well the data is fitted and weakening the parameters theta.
+
+In the end we have a regularized linear regression that works much better than the regular one.
+It even avoid that the normal equation is non-reversible.
+The calculus is the same, but we add an extra term to the cost function.	To calculate the gradient descent, one must actualize the value of theta 0 and of others thetas, since the added sum takes from indexes 1 to n.
+The value of theta 0 is known as bias.
+
+\begin{equation}
+J(\theta) = \frac{1}{2m} \sum_{n=1}^{m} (h(x^{(i)})-y^{(i)})^{2} + \lambda\sum_{n=1}^{n} \theta_{j}^{2}
+\end{equation}
+
+Cost Function in logistic regression
+
+\begin{equation}
+J(\theta) = \frac{1}{m} \sum_{n=1}^{m} [-y^{(i)}}\log(h(x^{(i)}))-(1-y^{(i)})\log(1-h(x^{(i)})]
+\end{equation}
+
 \end{document}

### 2019-03-24 Rapport/TG_ITA/Bibliographic_analysis/General_research.tex
Assigment week 2

@@
 
-\subsection{Introduction to digital communication}
+\subsection{Introduction to digital communication book}
 
@@
 
+	 digital transmission, the number of bit errors is the number of received bits of a data stream over a communication channel that have been altered due to noise, interference, distortion or bit synchronization errors.
+	 bit error rate (BER) is the number of bit errors per unit time.
+	 The bit error ratio (also BER) is the number of bit errors divided by the total number of transferred bits during a studied time interval. 
+	  bit error probability pe is the expectation value of the bit error ratio.
+The bit error ratio can be considered as an approximate estimate of the bit error probability.
+This estimate is accurate for a long time interval and a high number of bit errors.
+
+	\subsubsection{Chapter 2 - Coding for discrete sources}
+
 \end{document}

### 2019-03-24 Coursera/machine-learning-ex1/machine-learning-ex1/ex1/gradientDescentMulti.m

@@ -17,23 +17,20 @@ for iter = 1:num_iters
     %       of the cost function (computeCostMulti) and gradient here.
     %
 
+%   Noob code
+%    thetaTemp = zeros(length(theta),1);
+% 
+%     for j = 1:length(theta)
+%         for i = 1:m
+%             thetaTemp(j) = thetaTemp(j) - alpha*1/m*(theta'*X(i,:)'-y(i))*X(i,j);
+%         end
+%     end
+%     for j = 1:length(theta)
+%        theta(j) = theta(j)+thetaTemp(j); 
+%     end
 
-   thetaTemp = zeros(length(theta),1);
-
-    for j = 1:length(theta)
-        for i = 1:m
-            thetaTemp(j) = thetaTemp(j) - alpha*1/m*(theta'*X(i,:)'-y(i))*X(i,j);
-        end
-    end
-    for j = 1:length(theta)
-       theta(j) = theta(j)+thetaTemp(j); 
-    end
-
-
-
-
-
-
+    % Vectorization:
+    theta = theta - alpha/m*(((theta'*X')'-y)'*X)';
 
 
 

### 2019-03-24 Coursera/machine-learning-ex2/machine-learning-ex2/ex2/costFunction.m

@@ -1,0 +1,35 @@
+function [J, grad] = costFunction(theta, X, y)
+%COSTFUNCTION Compute cost and gradient for logistic regression
+%   J = COSTFUNCTION(theta, X, y) computes the cost of using theta as the
+%   parameter for logistic regression and the gradient of the cost
+%   w.r.t. to the parameters.
+
+% Initialize some useful values
+m = length(y); % number of training examples
+
+% You need to return the following variables correctly 
+J = 0;
+grad = zeros(size(theta));
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Compute the cost of a particular choice of theta.
+%               You should set J to the cost.
+%               Compute the partial derivatives and set grad to the partial
+%               derivatives of the cost w.r.t. each parameter in theta
+%
+% Note: grad should have the same dimensions as theta
+%
+
+h = sigmoid(theta'*X')';
+
+J = 1/m*sum(-(y'*log(h))-((1-y)'*log(1-h)));
+
+grad = (1/m*((h-y)'*X))';
+
+
+
+
+
+% =============================================================
+
+end

### 2019-03-24 Coursera/machine-learning-ex2/machine-learning-ex2/ex2/costFunctionReg.m

@@ -1,0 +1,31 @@
+function [J, grad] = costFunctionReg(theta, X, y, lambda)
+%COSTFUNCTIONREG Compute cost and gradient for logistic regression with regularization
+%   J = COSTFUNCTIONREG(theta, X, y, lambda) computes the cost of using
+%   theta as the parameter for regularized logistic regression and the
+%   gradient of the cost w.r.t. to the parameters. 
+
+% Initialize some useful values
+m = length(y); % number of training examples
+
+% You need to return the following variables correctly 
+J = 0;
+grad = zeros(size(theta));
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Compute the cost of a particular choice of theta.
+%               You should set J to the cost.
+%               Compute the partial derivatives and set grad to the partial
+%               derivatives of the cost w.r.t. each parameter in theta
+
+
+h = sigmoid(theta'*X')';
+
+J = 1/m*sum(-(y'*log(h))-((1-y)'*log(1-h))) + lambda/(2*m)*[0;theta(2:end,1)]'*[0;theta(2:end,1)];
+
+grad = (1/m*((h-y)'*X))' + lambda/m*[0;theta(2:end,1)];
+
+
+
+% =============================================================
+
+end

### 2019-03-24 Coursera/machine-learning-ex2/machine-learning-ex2/ex2/ex2_reg.m

@@ -1,0 +1,136 @@
+%% Machine Learning Online Class - Exercise 2: Logistic Regression
+%
+%  Instructions
+%  ------------
+%
+%  This file contains code that helps you get started on the second part
+%  of the exercise which covers regularization with logistic regression.
+%
+%  You will need to complete the following functions in this exericse:
+%
+%     sigmoid.m
+%     costFunction.m
+%     predict.m
+%     costFunctionReg.m
+%
+%  For this exercise, you will not need to change any code in this file,
+%  or any other files other than those mentioned above.
+%
+
+%% Initialization
+clear ; close all; clc
+
+%% Load Data
+%  The first two columns contains the X values and the third column
+%  contains the label (y).
+
+data = load('ex2data2.txt');
+X = data(:, [1, 2]); y = data(:, 3);
+
+plotData(X, y);
+
+% Put some labels
+hold on;
+
+% Labels and Legend
+xlabel('Microchip Test 1')
+ylabel('Microchip Test 2')
+
+% Specified in plot order
+legend('y = 1', 'y = 0')
+hold off;
+
+
+%% =========== Part 1: Regularized Logistic Regression ============
+%  In this part, you are given a dataset with data points that are not
+%  linearly separable. However, you would still like to use logistic
+%  regression to classify the data points.
+%
+%  To do so, you introduce more features to use -- in particular, you add
+%  polynomial features to our data matrix (similar to polynomial
+%  regression).
+%
+
+% Add Polynomial Features
+
+% Note that mapFeature also adds a column of ones for us, so the intercept
+% term is handled
+X = mapFeature(X(:,1), X(:,2));
+
+% Initialize fitting parameters
+initial_theta = zeros(size(X, 2), 1);
+
+% Set regularization parameter lambda to 1
+lambda = 1;
+
+% Compute and display initial cost and gradient for regularized logistic
+% regression
+[cost, grad] = costFunctionReg(initial_theta, X, y, lambda);
+
+fprintf('Cost at initial theta (zeros): %f\n', cost);
+fprintf('Expected cost (approx): 0.693\n');
+fprintf('Gradient at initial theta (zeros) - first five values only:\n');
+fprintf(' %f \n', grad(1:5));
+fprintf('Expected gradients (approx) - first five values only:\n');
[…]

### 2019-03-24 Coursera/machine-learning-ex2/machine-learning-ex2/ex2/plotData.m

@@ -1,0 +1,30 @@
+function plotData(X, y)
+%PLOTDATA Plots the data points X and y into a new figure 
+%   PLOTDATA(x,y) plots the data points with + for the positive examples
+%   and o for the negative examples. X is assumed to be a Mx2 matrix.
+
+% Create New Figure
+figure; hold on;
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Plot the positive and negative examples on a
+%               2D plot, using the option 'k+' for the positive
+%               examples and 'ko' for the negative examples.
+%
+
+% Find Indices of Positive and Negative Examples
+pos = find(y==1); neg = find(y == 0);
+% Plot Examples
+plot(X(pos, 1), X(pos, 2), 'k+','LineWidth', 2, ...
+'MarkerSize', 7); % plot all positive results (admitted students)
+plot(X(neg, 1), X(neg, 2), 'ko', 'MarkerFaceColor', 'y', ...
+'MarkerSize', 7); % plot all negative results (denied students)
+
+
+% =========================================================================
+
+
+
+hold off;
+
+end

### 2019-03-24 Coursera/machine-learning-ex2/machine-learning-ex2/ex2/plotDecisionBoundary.m

@@ -1,0 +1,48 @@
+function plotDecisionBoundary(theta, X, y)
+%PLOTDECISIONBOUNDARY Plots the data points X and y into a new figure with
+%the decision boundary defined by theta
+%   PLOTDECISIONBOUNDARY(theta, X,y) plots the data points with + for the 
+%   positive examples and o for the negative examples. X is assumed to be 
+%   a either 
+%   1) Mx3 matrix, where the first column is an all-ones column for the 
+%      intercept.
+%   2) MxN, N>3 matrix, where the first column is all-ones
+
+% Plot Data
+plotData(X(:,2:3), y);
+hold on
+
+if size(X, 2) <= 3
+    % Only need 2 points to define a line, so choose two endpoints
+    plot_x = [min(X(:,2))-20,  max(X(:,2))+20];
+
+    % Calculate the decision boundary line
+    plot_y = (-1./theta(3)).*(theta(2).*plot_x + theta(1));
+
+    % Plot, and adjust axes for better viewing
+    plot(plot_x, plot_y)
+    
+    % Legend, specific for the exercise
+    legend('Admitted', 'Not admitted', 'Decision Boundary')
+    axis([30, 100, 30, 100])
+else
+    % Here is the grid range
+    u = linspace(-1, 1.5, 50);
+    v = linspace(-1, 1.5, 50);
+
+    z = zeros(length(u), length(v));
+    % Evaluate z = theta*x over the grid
+    for i = 1:length(u)
+        for j = 1:length(v)
+            z(i,j) = mapFeature(u(i), v(j))*theta;
+        end
+    end
+    z = z'; % important to transpose z before calling contour
+
+    % Plot z = 0
+    % Notice you need to specify the range [0, 0]
+    contour(u, v, z, [0, 0], 'LineWidth', 2)
+end
+hold off
+
+end

### 2019-03-24 Coursera/machine-learning-ex2/machine-learning-ex2/ex2/predict.m

@@ -1,0 +1,24 @@
+function p = predict(theta, X)
+%PREDICT Predict whether the label is 0 or 1 using learned logistic 
+%regression parameters theta
+%   p = PREDICT(theta, X) computes the predictions for X using a 
+%   threshold at 0.5 (i.e., if sigmoid(theta'*x) >= 0.5, predict 1)
+
+m = size(X, 1); % Number of training examples
+
+% You need to return the following variables correctly
+p = zeros(m, 1);
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Complete the following code to make predictions using
+%               your learned logistic regression parameters. 
+%               You should set p to a vector of 0's and 1's
+%
+
+p = 1.0*(sigmoid(theta'*X') >= 0.5)';
+
+
+% =========================================================================
+
+
+end

### 2019-03-24 Coursera/machine-learning-ex2/machine-learning-ex2/ex2/sigmoid.m

@@ -1,0 +1,17 @@
+function g = sigmoid(z)
+%SIGMOID Compute sigmoid function
+%   g = SIGMOID(z) computes the sigmoid of z.
+
+% You need to return the following variables correctly 
+g = zeros(size(z));
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Compute the sigmoid of each value of z (z can be a matrix,
+%               vector or scalar).
+
+g = 1.0./(1+exp(-z));
+
+
+% =============================================================
+
+end

### 2019-03-26 Rapport/TG_ITA/Bibliographic_analysis/Coursera-v2.tex
Week 4 assignment

--- 2019-03-24 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
@@
 	\end{equation}
+
 	\begin{equation}
@@
 
-\begin{equation}
-J(\theta) = \frac{1}{m} \sum_{n=1}^{m} [-y^{(i)}}\log(h(x^{(i)}))-(1-y^{(i)})\log(1-h(x^{(i)})]
-\end{equation}
+...
+
+\subsubsection{Week 4 - Neural Networks}
+
+	Neural network: more efficient way to learn complex hypothesis.
+
+	Model 1 representation: 
+	\begin{itemize}
+	\item Dendrite: input wires
+	\item Axon: output wire
+	\item Cell body: do computations
+	\item axon send info to another neuron
+	\end{itemize}
+
+	Hypothesis function = activation function
+
+	$\theta$ = parameters = weights
+
+	Layers: first layer: input layer, last layer: output layer, hidden layer: values you don't observe between the input and output.
+
+	Notation: $\Theta$ - matrix of weights controlling function mapping from layer $j$ to layer $j+1$.
+The dimension is equal to $s_{j+1} \times (s_{j}+1)$; $a_{i}^{(j)}$ - activation of unit $i$ in layer $j$;
 

### 2019-03-26 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
Week 4 assignment

@@
 
+Regularized cost function:
 \begin{equation}
@@
 
-\begin{equation}
-J(\theta) = \frac{1}{m} \sum_{n=1}^{m} [-y^{(i)}}\log(h(x^{(i)}))-(1-y^{(i)})\log(1-h(x^{(i)})]
-\end{equation}
+...
 
+\subsubsection{Week 4 - Neural Networks}
+
+	Neural network: more efficient way to learn complex hypothesis.
+
+	\textbf{Model 1 representation: }
+	\begin{itemize}
+	\item Dendrite: input wires
+	\item Axon: output wire
+	\item Cell body: do computations
+	\item axon send info to another neuron
+	\end{itemize}
+
+	Hypothesis function = activation function
+
+	$\theta$ = parameters = weights
+
+	Layers: first layer: input layer, last layer: output layer, hidden layer: values you don't observe between the input and output.
+
+	Notation: $\Theta^{j}$ - matrix of weights controlling function mapping from layer $j$ to layer $j+1$.
+The dimension is equal to $s_{j+1} \times (s_{j}+1)$.
+The index +1 takes into account the bias added to the inputs and to the computed layer.; $a_{i}^{(j)}$ - activation of unit $i$ in layer $j$;
+
+	\textbf{Model Representation II}
+	Vectorized: forward propagation:
+
+	\begin{align}\label{eq:my_eq}
+	z^{(j)} = & \Theta^{(j-1)}a^{j-1} \\
+	a^{(j)} = & g(z^{(j)})
+	\end{align}
+
+Last layer: 	
+\begin{align}\label{eq:my_eq}
+	z^{(j+1)} = & \Theta^{(j)}a^{j} \\
+	a^{(j+1)} = & g(z^{(j+1)}) = h_{\Theta}(x)
+	\end{align}
+
+	The away a NN is connected is called an architecture.	
+	The first layers computes task of a certain complexity that increases when we go deeper into the network. The last layer usually computes the most demanding task.
+
+	Multiclass Classification
+	To classify data into multiple classes, we let our hypothesis function return a vector of values.
+
+	Task 3: Multi-class Classification and Neural Networks (matlab)
+
+	\subsubsection{Week 5}
 \end{document}

### 2019-03-26 Coursera/machine-learning-ex3/machine-learning-ex3/ex3/ex3.m

@@ -1,0 +1,91 @@
+%% Machine Learning Online Class - Exercise 3 | Part 1: One-vs-all
+
+%  Instructions
+%  ------------
+%
+%  This file contains code that helps you get started on the
+%  linear exercise. You will need to complete the following functions
+%  in this exericse:
+%
+%     lrCostFunction.m (logistic regression cost function)
+%     oneVsAll.m
+%     predictOneVsAll.m
+%     predict.m
+%
+%  For this exercise, you will not need to change any code in this file,
+%  or any other files other than those mentioned above.
+%
+
+%% Initialization
+clear ; close all; clc
+
+%% Setup the parameters you will use for this part of the exercise
+input_layer_size  = 400;  % 20x20 Input Images of Digits
+num_labels = 10;          % 10 labels, from 1 to 10
+                          % (note that we have mapped "0" to label 10)
+
+%% =========== Part 1: Loading and Visualizing Data =============
+%  We start the exercise by first loading and visualizing the dataset.
+%  You will be working with a dataset that contains handwritten digits.
+%
+
+% Load Training Data
+fprintf('Loading and Visualizing Data ...\n')
+
+load('ex3data1.mat'); % training data stored in arrays X, y
+m = size(X, 1);
+
+% Randomly select 100 data points to display
+rand_indices = randperm(m);
+sel = X(rand_indices(1:100), :);
+
+displayData(sel);
+
+fprintf('Program paused. Press enter to continue.\n');
+pause;
+
+%% ============ Part 2a: Vectorize Logistic Regression ============
+%  In this part of the exercise, you will reuse your logistic regression
+%  code from the last exercise. You task here is to make sure that your
+%  regularized logistic regression implementation is vectorized. After
+%  that, you will implement one-vs-all classification for the handwritten
+%  digit dataset.
+%
+
+% Test case for lrCostFunction
+fprintf('\nTesting lrCostFunction() with regularization');
+
+theta_t = [-2; -1; 1; 2];
+X_t = [ones(5,1) reshape(1:15,5,3)/10];
+y_t = ([1;0;1;0;1] >= 0.5);
+lambda_t = 3;
+[J grad] = lrCostFunction(theta_t, X_t, y_t, lambda_t);
+
+fprintf('\nCost: %f\n', J);
+fprintf('Expected cost: 2.534819\n');
+fprintf('Gradients:\n');
[…]

### 2019-03-26 Coursera/machine-learning-ex3/machine-learning-ex3/ex3/ex3_nn.m

@@ -1,0 +1,91 @@
+%% Machine Learning Online Class - Exercise 3 | Part 2: Neural Networks
+
+%  Instructions
+%  ------------
+% 
+%  This file contains code that helps you get started on the
+%  linear exercise. You will need to complete the following functions 
+%  in this exericse:
+%
+%     lrCostFunction.m (logistic regression cost function)
+%     oneVsAll.m
+%     predictOneVsAll.m
+%     predict.m
+%
+%  For this exercise, you will not need to change any code in this file,
+%  or any other files other than those mentioned above.
+%
+
+%% Initialization
+clear ; close all; clc
+
+%% Setup the parameters you will use for this exercise
+input_layer_size  = 400;  % 20x20 Input Images of Digits
+hidden_layer_size = 25;   % 25 hidden units
+num_labels = 10;          % 10 labels, from 1 to 10   
+                          % (note that we have mapped "0" to label 10)
+
+%% =========== Part 1: Loading and Visualizing Data =============
+%  We start the exercise by first loading and visualizing the dataset. 
+%  You will be working with a dataset that contains handwritten digits.
+%
+
+% Load Training Data
+fprintf('Loading and Visualizing Data ...\n')
+
+load('ex3data1.mat');
+m = size(X, 1);
+
+% Randomly select 100 data points to display
+sel = randperm(size(X, 1));
+sel = sel(1:100);
+
+displayData(X(sel, :));
+
+fprintf('Program paused. Press enter to continue.\n');
+pause;
+
+%% ================ Part 2: Loading Pameters ================
+% In this part of the exercise, we load some pre-initialized 
+% neural network parameters.
+
+fprintf('\nLoading Saved Neural Network Parameters ...\n')
+
+% Load the weights into variables Theta1 and Theta2
+load('ex3weights.mat');
+
+
+%% ================= Part 3: Implement Predict =================
+%  After training the neural network, we would like to use it to predict
+%  the labels. You will now implement the "predict" function to use the
+%  neural network to predict the labels of the training set. This lets
+%  you compute the training set accuracy.
+
+pred = predict(Theta1, Theta2, X);
+
+fprintf('\nTraining Set Accuracy: %f\n', mean(double(pred == y)) * 100);
+
+fprintf('Program paused. Press enter to continue.\n');
+pause;
+
+%  To give you an idea of the network's output, you can also run
[…]

### 2019-03-26 Coursera/machine-learning-ex3/machine-learning-ex3/ex3/lrCostFunction.m

@@ -1,0 +1,49 @@
+function [J, grad] = lrCostFunction(theta, X, y, lambda)
+%LRCOSTFUNCTION Compute cost and gradient for logistic regression with 
+%regularization
+%   J = LRCOSTFUNCTION(theta, X, y, lambda) computes the cost of using
+%   theta as the parameter for regularized logistic regression and the
+%   gradient of the cost w.r.t. to the parameters. 
+
+% Initialize some useful values
+m = length(y); % number of training examples
+
+% You need to return the following variables correctly 
+J = 0;
+grad = zeros(size(theta));
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Compute the cost of a particular choice of theta.
+%               You should set J to the cost.
+%               Compute the partial derivatives and set grad to the partial
+%               derivatives of the cost w.r.t. each parameter in theta
+%
+% Hint: The computation of the cost function and gradients can be
+%       efficiently vectorized. For example, consider the computation
+%
+%           sigmoid(X * theta)
+%
+%       Each row of the resulting matrix will contain the value of the
+%       prediction for that example. You can make use of this to vectorize
+%       the cost function and gradient computations. 
+%
+% Hint: When computing the gradient of the regularized cost function, 
+%       there're many possible vectorized solutions, but one solution
+%       looks like:
+%           grad = (unregularized gradient for logistic regression)
+%           temp = theta; 
+%           temp(1) = 0;   % because we don't add anything for j = 0  
+%           grad = grad + YOUR_CODE_HERE (using the temp variable)
+%
+h = sigmoid(X*theta);
+
+J = 1/m*sum(-(y'*log(h))-((1-y)'*log(1-h))) + lambda/(2*m)*[0;theta(2:end,1)]'*[0;theta(2:end,1)];
+
+grad = (1/m*(X'*(h-y))) + lambda/m*[0;theta(2:end,1)];
+
+
+% =============================================================
+
+grad = grad(:);
+
+end

### 2019-03-26 Coursera/machine-learning-ex3/machine-learning-ex3/ex3/predict.m

@@ -1,0 +1,36 @@
+function p = predict(Theta1, Theta2, X)
+%PREDICT Predict the label of an input given a trained neural network
+%   p = PREDICT(Theta1, Theta2, X) outputs the predicted label of X given the
+%   trained weights of a neural network (Theta1, Theta2)
+
+% Useful values
+m = size(X, 1);
+num_labels = size(Theta2, 1);
+
+% You need to return the following variables correctly 
+p = zeros(size(X, 1), 1);
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Complete the following code to make predictions using
+%               your learned neural network. You should set p to a 
+%               vector containing labels between 1 to num_labels.
+%
+% Hint: The max function might come in useful. In particular, the max
+%       function can also return the index of the max element, for more
+%       information see 'help max'. If your examples are in rows, then, you
+%       can use max(A, [], 2) to obtain the max for each row.
+%
+% Add ones to the X data matrix
+X = [ones(m, 1) X];
+
+z_2 = sigmoid(X*Theta1');
+h = sigmoid([ones(m,1) z_2]*Theta2');
+[M,I] = max(h, [], 2);
+
+p = I
+
+
+% =========================================================================
+
+
+end

### 2019-03-26 Coursera/machine-learning-ex3/machine-learning-ex3/ex3/predictOneVsAll.m

@@ -1,0 +1,41 @@
+function p = predictOneVsAll(all_theta, X)
+%PREDICT Predict the label for a trained one-vs-all classifier. The labels 
+%are in the range 1..K, where K = size(all_theta, 1). 
+%  p = PREDICTONEVSALL(all_theta, X) will return a vector of predictions
+%  for each example in the matrix X. Note that X contains the examples in
+%  rows. all_theta is a matrix where the i-th row is a trained logistic
+%  regression theta vector for the i-th class. You should set p to a vector
+%  of values from 1..K (e.g., p = [1; 3; 1; 2] predicts classes 1, 3, 1, 2
+%  for 4 examples) 
+
+m = size(X, 1);
+num_labels = size(all_theta, 1);
+
+% You need to return the following variables correctly 
+p = zeros(size(X, 1), 1);
+
+% Add ones to the X data matrix
+X = [ones(m, 1) X];
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Complete the following code to make predictions using
+%               your learned logistic regression parameters (one-vs-all).
+%               You should set p to a vector of predictions (from 1 to
+%               num_labels).
+%
+% Hint: This code can be done all vectorized using the max function.
+%       In particular, the max function can also return the index of the 
+%       max element, for more information see 'help max'. If your examples 
+%       are in rows, then, you can use max(A, [], 2) to obtain the max 
+%       for each row.
+%       
+[M,I] = max(sigmoid(X*all_theta'), [], 2);
+
+p = I;
+
+
+
+% =========================================================================
+
+
+end

### 2019-03-31 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
Git commands
Backpropagation lab implemented

@@
 \usepackage{amssymb}
+\usepackage{tabto}
+\usepackage{xcolor}
 \begin{document}
@@
 
+	The training examples of a training set must be independent!
 	\subsubsection{Week 5}
+
+	Backpropagation. "Backpropagation" is neural-network terminology for minimizing our cost function
+
+	Generalized linear regression for neural networks
+
+	 \begin{multline}
+	 J(\Theta) = -\frac{1}{m} \left[ \sum_{i = 1}^{m}\sum_{k = 1}^{K} y_{k}^{(i)} log(h_{\Theta}(x^{(i)}))_{k} + (1-y_{k}^{(i)})log(1-h_{\Theta}(x^{(i)}))_{k}\right]\\
+	  + \frac{\lambda}{2m} \sum_{l = 1}^{L-1} \sum_{i = 1}^{s_{l}} \sum_{j = 1}^{s_{l+1}} (\Theta_{ij}^{(j)})^{2}
+	 \end{multline}
+	where L = total number of layers in the network, sl = number of units (not counting bias unit) in layer l, K = number of output units/classes.
+
+	We denote $h_{\Theta}(x)_{k}$ as being a hypothesis that results in the $k^{th}$ output.
+
+	We have added a few nested summations to account for our multiple output nodes.
+
+	The number of columns in our current theta matrix is equal to the number of nodes in our current layer (including the bias unit).
+The number of rows in our current theta matrix is equal to the number of nodes in the next layer (excluding the bias unit).
+
+	\begin{itemize}
+	\item the double sum simply adds up the logistic regression costs calculated for each cell in the output layer
+\item the triple sum simply adds up the squares of all the individual $\Theta$s in the entire network.
+\item the i in the triple sum does not refer to training example i
+	\end{itemize}
+
+	The back-propagation algorithm is used to minimize J. So we need to compute J and the partial derivatives of J.
+	We compute the delta error for each layer (but the input layer because we set $a^{(1)} = x^{(i)}$), beginning in the last layer, where: $\delta_{j}^{L} = a_{L}^{l}-y_{j}$, $\delta_{j}^{l-1} = (\Theta^{l-1}) ^{T}\delta^{(l+1)}.*g'(z^{(l-1)})$, $g'(z^{(l-1)}) = a^{(l-1)}.*(1-a^{(l-1)}) $.
+Step by step
+	\begin{enumerate}
+	\item Training set $\left\lbrace  (x^{1},y^{(1)}), ..., (x^{m},y^{(m)})\right\rbrace $.
+
+	\item $\Delta_{ij}^{l} = 0$ for all $l,i,j$. (Useful to compute $\frac{\partial}{\partial \Theta^{(l)}_{ij}} J(\Theta)$)
+
+	\item $\delta_{j}^{l-1} = (\Theta^{l-1}) ^{T}\delta^{(l)}.*g'(z^{(l-1)})$
+
+	\item $g'(z^{(l-1)}) = a^{(l-1)}.*(1-a^{(l-1)}) $ 
+	\end{enumerate}
+	\begin{enumerate}
+	\item For $i = 1$ to $m$\hspace{10mm}\\
+	\textcolor{white}{.} \hspace{10mm}	 Set $a^{(1)} = x^{(i)}$\\
+	\textcolor{white}{.} \hspace{10mm} Perform forward propagation to compute $a^{(l)}$ for $l = 2, 3, ..., L$\\
+	\textcolor{white}{.} \hspace{10mm} Using $y^{(i)}$, compute  $\delta^{(L)} = a^{(L)}-y^{(i)}$\\
+	\textcolor{white}{.} \hspace{10mm} Compute $\delta^{(L-1)},..., \delta^{(2)}$ where $\delta_{j}^{l-1} = (\Theta^{(l-1)}) ^{T}\delta^{(l)}.*g'(z^{(l-1)})$.\\
+	\textcolor{white}{.} \hspace{10mm} Compute $\Delta^{(l)}:=\Delta^{(l)}+\delta^{(l+1)}(a^{(l)})^{T}$.
+Contribution of each training set to the gradient that is calculated with the general Theta for the NN.
+
+	\item Compute $D^{(l)}:=\frac{1}{m}(\Delta^{(l)}+\lambda\Theta^{(l)})$ if $j\neq 0$.
+
+	\item Compute $D^{(l)}:=\frac{1}{m}\Delta^{(l)}$ if $j = 0$ (bias term).
+
+	\item $\frac{\partial}{\partial \Theta^{(l)}_{ij}} J(\Theta) = D^{(l)}_{ij}$
+	\end{enumerate}
+
+	So in general you do a forward propagation with a training example, than a back propagation with the same training example, than changes the training example and repeats.
+
+	Implement gradient checking for make sure that forward and backpropagation are working correctly. After you know they are working, comment the code.
+
+	The neural network used to check back-propagation is usually small, that's why it is important to code general functions.
+
 \end{document}

### 2019-03-31 Coursera/machine-learning-ex4/machine-learning-ex4/machine-learning-ex4/ex4/ex4.m

@@ -1,0 +1,235 @@
+%% Machine Learning Online Class - Exercise 4 Neural Network Learning
+
+%  Instructions
+%  ------------
+% 
+%  This file contains code that helps you get started on the
+%  linear exercise. You will need to complete the following functions 
+%  in this exericse:
+%
+%     sigmoidGradient.m
+%     randInitializeWeights.m
+%     nnCostFunction.m
+%
+%  For this exercise, you will not need to change any code in this file,
+%  or any other files other than those mentioned above.
+%
+
+%% Initialization
+clear ; close all; clc
+
+%% Setup the parameters you will use for this exercise
+input_layer_size  = 400;  % 20x20 Input Images of Digits
+hidden_layer_size = 25;   % 25 hidden units
+num_labels = 10;          % 10 labels, from 1 to 10   
+                          % (note that we have mapped "0" to label 10)
+
+%% =========== Part 1: Loading and Visualizing Data =============
+%  We start the exercise by first loading and visualizing the dataset. 
+%  You will be working with a dataset that contains handwritten digits.
+%
+
+% Load Training Data
+fprintf('Loading and Visualizing Data ...\n')
+
+load('ex4data1.mat');
+m = size(X, 1);
+
+% Randomly select 100 data points to display
+sel = randperm(size(X, 1));
+sel = sel(1:100);
+
+displayData(X(sel, :));
+
+fprintf('Program paused. Press enter to continue.\n');
+pause;
+
+
+%% ================ Part 2: Loading Parameters ================
+% In this part of the exercise, we load some pre-initialized 
+% neural network parameters.
+
+fprintf('\nLoading Saved Neural Network Parameters ...\n')
+
+% Load the weights into variables Theta1 and Theta2
+load('ex4weights.mat');
+
+% Unroll parameters 
+nn_params = [Theta1(:) ; Theta2(:)];
+
+%% ================ Part 3: Compute Cost (Feedforward) ================
+%  To the neural network, you should first start by implementing the
+%  feedforward part of the neural network that returns the cost only. You
+%  should complete the code in nnCostFunction.m to return cost. After
+%  implementing the feedforward to compute the cost, you can verify that
+%  your implementation is correct by verifying that you get the same cost
+%  as us for the fixed debugging parameters.
+%
+%  We suggest implementing the feedforward cost *without* regularization
+%  first so that it will be easier for you to debug. Later, in part 4, you
[…]

### 2019-03-31 Coursera/machine-learning-ex4/machine-learning-ex4/machine-learning-ex4/ex4/randInitializeWeights.m

@@ -1,0 +1,29 @@
+function W = randInitializeWeights(L_in, L_out)
+%RANDINITIALIZEWEIGHTS Randomly initialize the weights of a layer with L_in
+%incoming connections and L_out outgoing connections
+%   W = RANDINITIALIZEWEIGHTS(L_in, L_out) randomly initializes the weights 
+%   of a layer with L_in incoming connections and L_out outgoing 
+%   connections. 
+%
+%   Note that W should be set to a matrix of size(L_out, 1 + L_in) as
+%   the first column of W handles the "bias" terms
+%
+
+% You need to return the following variables correctly 
+W = zeros(L_out, 1 + L_in);
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Initialize W randomly so that we break the symmetry while
+%               training the neural network.
+%
+% Note: The first column of W corresponds to the parameters for the bias unit
+%
+
+epsilon_init = 0.12; % choosing eps: eps = sqrt(6)/(sqrt(Lin + Lout))
+W = rand(L_out, 1 + L_in) * 2 * epsilon_init - epsilon_init;
+
+
+
+% =========================================================================
+
+end

### 2019-04-15 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
Coursera Ex 5

@@
 
+	\subsubsection{Week 6}
+Improving the Learning Algorithm:
+
+	\begin{itemize}
+	\item Get more training examples
+	\item Try smaller setof features
+	\item Try getting additional features
+	\item Try adding polynomial features
+	\item Try decreasing $\lambda$
+	\item Try increasing $\lambda$
+	\end{itemize}
+
+	Diagnostic: show what is working well and what is not in the algorithm. Show what is more fruitful to try to improve next.
+
+	Evaluating a hypothesis: split the dataset in trining set (80\%), cross validation set (20\%) and test set (20\%).
+Each set must be randomly distributed.
+You can try several hypothesis models and generate several $\theta$ and test on the cross validation set to find the best degree d of the polynomial used to estimate $\theta$.
+Then, estimate the generalization error using the test set with $J_{test}(\Theta^{(d)})$.
+
+	Missclassification error: 
+\begin{equation}
+	err(h_\theta(x),y) = 
+	\begin{cases}
+      1 & $ if $ h_\theta(x) \geq 0.5$ and $y=0$ or $h_\theta(x) < 0.5$ and $y = 1 \\
+      0 & $otherwise$ 
+   \end{cases}
+
+\end{equation}
+
+Classification Test error = $1/m_{test}\sum_{i=1}^ {m_{test}}err(h_\theta(x_{test}^{(i)}, y_{test}^{(i)})$
+
+Cross validation error: $J_{cv}(\theta) = \frac{1}{2m_{cv}}\sum_{i=1}^ {m_{cv}}(h_\theta(x_{cv}^{(i)}) - y_{cv})^{2}$
+
+Usually a learning algorithm may have a variance problem or a bias problem.
+
+If Jtest is high, you are facing a bias error (underfit) Jcv approx Jtest. If Jcv is high, you are facing a variance error (overfitting) Jcv $>>$ Jtrain.
+
+	We can observe this phenomena with the help of the graph Cost ($J(\theta)$ vs $d$ (polynomial degree))
+
+	Choose a good value for $\lambda$ to solve these problems. 
+\begin{enumerate}
+\item Try several values.
+\item minimize $J(\theta)$
+\item calculates Jcv
+\item Take the smalles Jcv
+\item Evaluate Jtest of the chosen set
+\item plot the errors in function of $\lambda$
+\end{enumerate}
+
+	Overfittin the data  = low $\lambda$
+	\begin{enumerate}
+	\item Create a list of lambdas (i.e. λ∈{0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24});
+\item Create a set of models with different degrees or any other variants.
+\item Iterate through the $\lambda$ and for each $\lambda$ go through all the models to learn some $\Theta$.
+\item Compute the cross validation error using the learned $\Theta$ (computed with $\lambda$) on the $J_{cv}(\Theta)$ without regularization or $\lambda = 0$.
+\item Select the best combo that produces the lowest error on the cross validation set.
+\item Using the best combo $\Theta$ and $\lambda$, apply it on $J_{test}(\Theta)$ to see if it has a good generalization of the problem.
+	\end{enumerate}
+
+	\textbf{Learning curves}
+
+	Error vs training size.
+We can find out if the algorithm suffers from variance or bias.
+If the learning algorithm suffers from variance, adding training examples may help.
+Adding more layers doesnt help with high variance error.
+
+	Resume:
+
+	\begin{enumerate}
+	\item Getting more training examples: Fixes high variance
+\item Trying smaller sets of features: Fixes high variance
+\item Adding features: Fixes high bias
+\item Adding polynomial features: Fixes high bias
+\item Decreasing λ: Fixes high bias
+\item Increasing λ: Fixes high variance.
+	\end{enumerate}
+
+	A neural network with fewer parameters is prone to underfitting. It is also computationally cheaper.
+
+A large neural network with more parameters is prone to overfitting. It is also computationally expensive. In this case you can use regularization (increase λ) to address the overfitting.
+
+Lower-order polynomials (low model complexity) have high bias and low variance. In this case, the model fits poorly consistently.
+
+Higher-order polynomials (high model complexity) fit the training data extremely well and the test data extremely poorly. These have low bias on the training data, but very high variance.
+
+In reality, we would want to choose a model somewhere in between, that can generalize well but also fits the data reasonably well.
+
 \end{document}

### 2019-04-15 Coursera/machine-learning-ex5/machine-learning-ex5/ex5/learningCurve.m

@@ -1,0 +1,69 @@
+function [error_train, error_val] = ...
+    learningCurve(X, y, Xval, yval, lambda)
+%LEARNINGCURVE Generates the train and cross validation set errors needed
+%to plot a learning curve
+%   [error_train, error_val] = ...
+%       LEARNINGCURVE(X, y, Xval, yval, lambda) returns the train and
+%       cross validation set errors for a learning curve. In particular,
+%       it returns two vectors of the same length - error_train and
+%       error_val. Then, error_train(i) contains the training error for
+%       i examples (and similarly for error_val(i)).
+%
+%   In this function, you will compute the train and test errors for
+%   dataset sizes from 1 up to m. In practice, when working with larger
+%   datasets, you might want to do this in larger intervals.
+%
+
+% Number of training examples
+m = size(X, 1);
+
+% You need to return these values correctly
+error_train = zeros(m, 1);
+error_val   = zeros(m, 1);
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Fill in this function to return training errors in
+%               error_train and the cross validation errors in error_val.
+%               i.e., error_train(i) and
+%               error_val(i) should give you the errors
+%               obtained after training on i examples.
+%
+% Note: You should evaluate the training error on the first i training
+%       examples (i.e., X(1:i, :) and y(1:i)).
+%
+%       For the cross-validation error, you should instead evaluate on
+%       the _entire_ cross validation set (Xval and yval).
+%
+% Note: If you are using your cost function (linearRegCostFunction)
+%       to compute the training and cross validation error, you should
+%       call the function with the lambda argument set to 0.
+%       Do note that you will still need to use lambda when running
+%       the training to obtain the theta parameters.
+%
+% Hint: You can loop over the examples with the following:
+%
+%       for i = 1:m
+%           % Compute train/cross validation errors using training examples
+%           % X(1:i, :) and y(1:i), storing the result in
+%           % error_train(i) and error_val(i)
+%           ....
+%
+%       end
+%
+
+% ---------------------- Sample Solution ----------------------
+n = size(Xval,1);
+for i = 1:m
+    [theta] = trainLinearReg([ones(i, 1) X(1:i, :)], y(1:i), lambda);
+    [J, grad] = linearRegCostFunction([ones(i, 1) X(1:i, :)], y(1:i), theta, 0);
+    error_train(i) = J;
+    [J, grad] = linearRegCostFunction([ones(n, 1) Xval], yval, theta, 0);
+    error_val(i) = J;
[…]

### 2019-04-15 Coursera/machine-learning-ex5/machine-learning-ex5/ex5/validationCurve.m

@@ -1,0 +1,58 @@
+function [lambda_vec, error_train, error_val] = ...
+    validationCurve(X, y, Xval, yval)
+%VALIDATIONCURVE Generate the train and validation errors needed to
+%plot a validation curve that we can use to select lambda
+%   [lambda_vec, error_train, error_val] = ...
+%       VALIDATIONCURVE(X, y, Xval, yval) returns the train
+%       and validation errors (in error_train, error_val)
+%       for different values of lambda. You are given the training set (X,
+%       y) and validation set (Xval, yval).
+%
+
+% Selected values of lambda (you should not change this)
+lambda_vec = [0 0.001 0.003 0.01 0.03 0.1 0.3 1 3 10]';
+
+% You need to return these variables correctly.
+error_train = zeros(length(lambda_vec), 1);
+error_val = zeros(length(lambda_vec), 1);
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Fill in this function to return training errors in 
+%               error_train and the validation errors in error_val. The 
+%               vector lambda_vec contains the different lambda parameters 
+%               to use for each calculation of the errors, i.e, 
+%               error_train(i), and error_val(i) should give 
+%               you the errors obtained after training with 
+%               lambda = lambda_vec(i)
+%
+% Note: You can loop over lambda_vec with the following:
+%
+%       for i = 1:length(lambda_vec)
+%           lambda = lambda_vec(i);
+%           % Compute train / val errors when training linear 
+%           % regression with regularization parameter lambda
+%           % You should store the result in error_train(i)
+%           % and error_val(i)
+%           ....
+%           
+%       end
+%
+%
+
+for i = 1:length(lambda_vec)
+   lambda = lambda_vec(i);
+   [theta] = trainLinearReg(X, y, lambda);
+   error_train(i) = linearRegCostFunction(X, y, theta, 0);
+   error_val(i) = linearRegCostFunction(Xval, yval, theta, 0);
+end
+
+
+
+
+
+
+
+
+% =========================================================================
+
+end

### 2019-04-15 MyCode/my_functions.py

@@ -7,10 +7,9 @@ Created on Fri Apr  5 17:28:03 2019
 
 from numpy import *
 import matplotlib.pyplot as plt
-
 from scipy.special import erfc # complementary error function
 
-def plotit(t, u, title="Title", X="X", Y="Y"):
+def plotit(t, u, title="Title", X="time (s)", Y="Amplitude"):
     plt.figure()
     plt.plot(t,u,'b')
     
@@ -21,4 +20,76 @@ def plotit(t, u, title="Title", X="X", Y="Y"):
     plt.ylabel(Y)
     plt.show()
 
+def scatterit(t, u, title="Title", X="time (s)", Y="Amplitude"):
+    plt.figure()
+    plt.plot(t,u, 'o')
+    
+    plt.grid(True)
+    
+    plt.title(title)
+    plt.xlabel(X)
+    plt.ylabel(Y)
+    plt.show()
+
+def sineWave(t, mean, amplitude, f):
+    omega = 2 * pi * f # fs > 2f
+    return amplitude * sin(omega * t + 0) + mean
+
+def analog2binary(t, bCodification, inputSignal):
+    # Example: we divide the amplitude in a scale of 8 bits and each sequence of 8 bits represents a point
+    # in the end we should have t/T * 8 bits
+    u = inputSignal
+    M = len(u)
+    x = 2**bCodification * (
+    u - mean(u) - min(u - mean(u))) / max(u - mean(u) - min(u - mean(u)))
+    
+    x = x.astype(int)
+    
+    scatterit(t, x, "Quantized signal")
+    b = ''  # strings of all bits togheter
+    for i in range(M):
+        str = "{0:b}".format(x[i])  # variable length source code word
+        str = str.zfill(bCodification)  
+        b = b + str  # transform b in an array of symbols 0's and 1's
+    
+    return b
+    
+def PSK2 (b, V):
+    # 2 - PSK: 2 different phase representations. Maps binary symbols unsing the map 0 -> -1 and 1-> 1. Constellation of size 2
+    um = empty(len(b))  # u - modulated (symbols modulated)
+    for i in range(len(b)): # step function
+        if (b[i] == '0'):
+            um[i] = -1*V
+        else:
+            um[i] = 1*V
+            
+    return um
+
+def pulsesignal(A, d): 
+    # d is the discretization of the pulse
[…]

### 2019-04-16 Rapport/TG_ITA/Bibliographic_analysis/Coursera.tex
Coursera Ex 6

@@
 	Missclassification error: 
-\begin{equation}
-	err(h_\theta(x),y) = 
-	\begin{cases}
-      1 & $ if $ h_\theta(x) \geq 0.5$ and $y=0$ or $h_\theta(x) < 0.5$ and $y = 1 \\
-      0 & $otherwise$ 
-   \end{cases}
+%\begin{equation}
+%	err(h_\theta(x),y) = 
+%	\begin{cases}
+%      1 & \  if \  h_\theta(x) \geq 0.5 \  and \  y=0 \  or\ h_\theta(x) < 0.5 \  and  \  y = 1 \\
+%      0 & \ otherwise
+%   \end{cases}
 
-\end{equation}
+%\end{equation}
 
@@
 	\begin{enumerate}
-	\item Create a list of lambdas (i.e. λ∈{0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24});
+	\item Create a list of lambdas (i.e. $\lambda\in {0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24}$);
 \item Create a set of models with different degrees or any other variants.
@@
 \item Adding polynomial features: Fixes high bias
-\item Decreasing λ: Fixes high bias
-\item Increasing λ: Fixes high variance.
+\item Decreasing $\lambda$: Fixes high bias
+\item Increasing $\lambda$: Fixes high variance.
 	\end{enumerate}
@@
 
-A large neural network with more parameters is prone to overfitting. It is also computationally expensive. In this case you can use regularization (increase λ) to address the overfitting.
+A large neural network with more parameters is prone to overfitting. It is also computationally expensive. In this case you can use regularization (increase $\lambda$) to address the overfitting.
 
@@
 
+	Skewed classes: events that are rare in a given set. Example: 0.5 \% of people have cancer.
+
+	The metrics to measure skewed classes are called precision and recall.
+
+	Precision is defined as: 
+	\begin{equation}
+	precision = \frac{true \ positives}{true \ positives + false \ positives}
+	\end{equation}
+
+	Recall is defined as:
+	\begin{equation}
+	precision = \frac{true \ positives}{true \ positives + false \ negatives}
+	\end{equation}
+
+	We can change the threshold value for h for classifying y = 1 or y = 0. if our threshold value is > 0.5, we are valuing confidence, so high \textbf{precision}, lower recall.
+
+	However, the opposite means the false negatives are important and should NOT happen. So, when in doubt, y = 1. You end up with higher recall and lower precision.
+
+	You want precision and recall to be both elevated.
+
+	How to choose the threshold? You can compute the F1 score $2\frac{PR}{P+R}$. There are others too.
+
+	\subsubsection{Week 7}
+
+	Large margin calssifiers
+
+	We can rewrite the cost function to optimize the calculations, choose functions that are very close to the optimal but are easier to calculate.
+
+	For example, the logistic regression cost with Support Vector  Machine (SVM) is known as: SVM hypothesis and is written as follows for the logistic regression problem: 
+	\begin{equation}
+	min_{\theta} C \sum_{i=1}^{m} [y^{(i)}cost_1(\theta^Tx^{(i)}) + (1- y^{(i)})cost_0(\theta^Tx^{(i)})] + \frac{1}{2}\sum_{j=1}^{n}\theta_j^2
+	\end{equation}
+	Cost1 and Cost0 are  the linearization of the cost curve.
+
+	The parameter C plays with bias and variance. Large C means lower bias and high variance. While small C means higher bias and low variance.
+
+	So in the end you want to predict the value of y juist basing over the value of $\theta^Tx$.
+
+	\textbf{Kernels: }
+
+	Suppose you have a non-linear decision boundary.
+Given x, compute new features based on the distance with given landmarks.
+The functions that calculate is called a similarity function (kernel function) and may be, for example, a gaussian kernel, where:
+	\begin{equation}
+	f_1=exp(-\frac{||x-l^{(k)}||^2}{2\sigma^2})
+	\end{equation}
+
+	Landmark: $l^{(k)}$	. Kernel Function: andmark: $f^{(k)}(x)$.
+
+	With them, we can define new features and make complex boundary curves. Instead of basing ourselves in the product of $\theta^Tx$, we compute $\theta^Tf(x)$ to make our decisions.
+
+	The variance of the kernel plays over bias and variance.
+
+	Large $\sigma^2$: features $f_i$ vary more smoothly, high bias, lower variance.
+
+	Small $\sigma^2$: features $f_i$ vary less smoothly, lower bias, higher variance.
+
+	If SVM is overfittinf your data, it is good to DECREASE C and INCREASE $\sigma^2$.
+
+	There are libraries to solve SVM optimization problems. DO NOT  implement code to optimize stuff, like minimizing for theta.
+
+	To use the algorithms, you should:
+	\begin{enumerate}
+	\item Choose C.
+	\item Choose Kernel (similarity function). If you have lots of features but small training set, than don't use kernels and just use logistic regression).
+	\item Choose $\sigma^2$.
+	\item Implement the kernel function (DO PERFORM feature scaling before using the Kernel, otherwise one feature will take over).
+	\item The choice of a Kernel shoul satisfy "Mercer's Theorem".
+	\item Fix the parameter in the ones that gives the best cross validation error.
+	\end{enumerate}
+
+	Available Kernels: Polynomial, Gaussian, String, chi-squared, etc. SVM is a convex optimization problem.
+
+	If the number of features is small and the number of training examples is large, create features and then use logistic regression, or a linear Kernel, because PCs struggle to calculate a Gaussian Kernel for example.
+
+	Neural Network should work in all conditions, but should be slower to train.
+
 \end{document}

### 2019-04-16 Coursera/machine-learning-ex6/machine-learning-ex6/ex6/dataset3Params.m

@@ -1,0 +1,45 @@
+function [C, sigma] = dataset3Params(X, y, Xval, yval)
+%DATASET3PARAMS returns your choice of C and sigma for Part 3 of the exercise
+%where you select the optimal (C, sigma) learning parameters to use for SVM
+%with RBF kernel
+%   [C, sigma] = DATASET3PARAMS(X, y, Xval, yval) returns your choice of C and 
+%   sigma. You should complete this function to return the optimal C and 
+%   sigma based on a cross-validation set.
+%
+
+% You need to return the following variables correctly.
+C = 1;
+sigma = 0.3;
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Fill in this function to return the optimal C and sigma
+%               learning parameters found using the cross validation set.
+%               You can use svmPredict to predict the labels on the cross
+%               validation set. For example, 
+%                   predictions = svmPredict(model, Xval);
+%               will return the predictions on the cross validation set.
+%
+%  Note: You can compute the prediction error using 
+%        mean(double(predictions ~= yval))
+%
+
+C1 = [0.01; 0.03; 0.1; 0.3; 1; 3; 10; 30];
+sigma1 = C1;
+error = zeros(length(C1), length(sigma1));
+for i = 1:length(C1)
+    for j = 1:length(sigma1)
+        model = svmTrain(X, y, C1(i), @(X, Y) gaussianKernel(X, Y, sigma1(j)));
+        pred = svmPredict(model, Xval);
+        error(i,j) = mean(double(pred ~= yval));
+    end
+end
+
+
+[M,c] = min(min(error,[],1));
+[M,l] = min(min(error,[],2));
+C = C1(l);
+sigma = sigma1(c);
+
+% =========================================================================
+
+end

### 2019-04-16 Coursera/machine-learning-ex6/machine-learning-ex6/ex6/emailFeatures.m

@@ -1,0 +1,59 @@
+function x = emailFeatures(word_indices)
+%EMAILFEATURES takes in a word_indices vector and produces a feature vector
+%from the word indices
+%   x = EMAILFEATURES(word_indices) takes in a word_indices vector and 
+%   produces a feature vector from the word indices. 
+
+% Total number of words in the dictionary
+n = 1899;
+
+% You need to return the following variables correctly.
+x = zeros(n, 1);
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Fill in this function to return a feature vector for the
+%               given email (word_indices). To help make it easier to 
+%               process the emails, we have have already pre-processed each
+%               email and converted each word in the email into an index in
+%               a fixed dictionary (of 1899 words). The variable
+%               word_indices contains the list of indices of the words
+%               which occur in one email.
+% 
+%               Concretely, if an email has the text:
+%
+%                  The quick brown fox jumped over the lazy dog.
+%
+%               Then, the word_indices vector for this text might look 
+%               like:
+%               
+%                   60  100   33   44   10     53  60  58   5
+%
+%               where, we have mapped each word onto a number, for example:
+%
+%                   the   -- 60
+%                   quick -- 100
+%                   ...
+%
+%              (note: the above numbers are just an example and are not the
+%               actual mappings).
+%
+%              Your task is take one such word_indices vector and construct
+%              a binary feature vector that indicates whether a particular
+%              word occurs in the email. That is, x(i) = 1 when word i
+%              is present in the email. Concretely, if the word 'the' (say,
+%              index 60) appears in the email, then x(60) = 1. The feature
+%              vector should look like:
+%
+%              x = [ 0 0 0 0 1 0 0 0 ... 0 0 0 0 1 ... 0 0 0 1 0 ..];
+%
+%
+
+
+for i = 1:length(word_indices)
+   x(word_indices(i)) = 1;
+end
+
+% =========================================================================
+    
+
+end

### 2019-04-16 Coursera/machine-learning-ex6/machine-learning-ex6/ex6/gaussianKernel.m

@@ -1,0 +1,26 @@
+function sim = gaussianKernel(x1, x2, sigma)
+%RBFKERNEL returns a radial basis function kernel between x1 and x2
+%   sim = gaussianKernel(x1, x2) returns a gaussian kernel between x1 and x2
+%   and returns the value in sim
+
+% Ensure that x1 and x2 are column vectors
+x1 = x1(:); x2 = x2(:);
+
+% You need to return the following variables correctly.
+sim = 0;
+
+% ====================== YOUR CODE HERE ======================
+% Instructions: Fill in this function to return the similarity between x1
+%               and x2 computed using a Gaussian kernel with bandwidth
+%               sigma
+%
+%
+
+sim = exp(-(x1-x2)'*(x1-x2)/(2*sigma^2));
+
+
+
+
+% =============================================================
+    
+end

### 2019-04-16 Coursera/machine-learning-ex6/machine-learning-ex6/ex6/processEmail.m

@@ -1,0 +1,126 @@
+function word_indices = processEmail(email_contents)
+%PROCESSEMAIL preprocesses a the body of an email and
+%returns a list of word_indices 
+%   word_indices = PROCESSEMAIL(email_contents) preprocesses 
+%   the body of an email and returns a list of indices of the 
+%   words contained in the email. 
+%
+
+% Load Vocabulary
+vocabList = getVocabList();
+
+% Init return value
+word_indices = [];
+
+% ========================== Preprocess Email ===========================
+
+% Find the Headers ( \n\n and remove )
+% Uncomment the following lines if you are working with raw emails with the
+% full headers
+
+% hdrstart = strfind(email_contents, ([char(10) char(10)]));
+% email_contents = email_contents(hdrstart(1):end);
+
+% Lower case
+email_contents = lower(email_contents);
+
+% Strip all HTML
+% Looks for any expression that starts with < and ends with > and replace
+% and does not have any < or > in the tag it with a space
+email_contents = regexprep(email_contents, '<[^<>]+>', ' ');
+
+% Handle Numbers
+% Look for one or more characters between 0-9
+email_contents = regexprep(email_contents, '[0-9]+', 'number');
+
+% Handle URLS
+% Look for strings starting with http:// or https://
+email_contents = regexprep(email_contents, ...
+                           '(http|https)://[^\s]*', 'httpaddr');
+
+% Handle Email Addresses
+% Look for strings with @ in the middle
+email_contents = regexprep(email_contents, '[^\s]+@[^\s]+', 'emailaddr');
+
+% Handle $ sign
+email_contents = regexprep(email_contents, '[$]+', 'dollar');
+
+
+% ========================== Tokenize Email ===========================
+
+% Output the email to screen as well
+fprintf('\n==== Processed Email ====\n\n');
+
+% Process file
+l = 0;
+
+while ~isempty(email_contents)
+
+    % Tokenize and also get rid of any punctuation
+    [str, email_contents] = ...
+       strtok(email_contents, ...
+              [' @$/#.-:&*+=[]?!(){},''">_<;%' char(10) char(13)]);
+   
+    % Remove any non alphanumeric characters
+    str = regexprep(str, '[^a-zA-Z0-9]', '');
+
+    % Stem the word 
+    % (the porterStemmer sometimes has issues, so we use a try catch block)
+    try str = porterStemmer(strtrim(str)); 
+    catch str = ''; continue;
+    end;
+
[…]

### 2019-04-24 Bibliography/Machine Learning/neuralnetworksanddeeplearning.com.pdf-master/neuralnetworksanddeeplearning.com.pdf-master/ch2_backprop.tex
Ameliorate bibliography

\documentclass[a4paper,twoside,10pt]{book}
\include{header}
\begin{document}
\chapter{How the backpropagation algorithm works}
In the last chapter we saw how neural networks can learn their weights and biases using the gradient descent algorithm. There was, however, a gap in our explanation: we didn't discuss how to compute the gradient of the cost function. That's quite a gap! In this chapter I'll explain a fast algorithm for computing such gradients, an algorithm known as\textit{ backpropagation}.

The backpropagation algorithm was originally introduced in the 1970s, but its importance wasn't fully appreciated until a famous 1986 \href{http://www.nature.com/nature/journal/v323/n6088/pdf/323533a0.pdf}{paper} by David Rumelhart, Geoffrey Hinton, and Ronald Williams. That paper describes several neural networks where backpropagation works far faster than earlier approaches to learning, making it possible to use neural nets to solve problems which had previously been insoluble. Today, the backpropagation algorithm is the workhorse of learning in neural networks.

This chapter is more mathematically involved than the rest of the book. If you're not crazy about mathematics you may be tempted to skip the chapter, and to treat backpropagation as a black box whose details you're willing to ignore. Why take the time to study those details?

The reason, of course, is understanding. At the heart of backpropagation is an expression for the partial derivative $\partial{}C/\partial{}w$ of the cost function $C$ with respect to any weight $w$ (or bias $b$) in the network. The expression tells us how quickly the cost changes when we change the weights and biases. And while the expression is somewhat complex, it also has a beauty to it, with each element having a natural, intuitive interpretation. And so backpropagation isn't just a fast algorithm for learning. It actually gives us detailed insights into how changing the weights and biases changes the overall behaviour of the network. That's well worth studying in detail.

With that said, if you want to skim the chapter, or jump straight to the next chapter, that's fine. I've written the rest of the book to be accessible even if you treat backpropagation as a black box. There are, of course, points later in the book where I refer back to results from this chapter. But at those points you should still be able to understand the main conclusions, even if you don't follow all the reasoning.

\section{Warm up: a fast matrix-based approach to computing the output from a neural network}
Before discussing backpropagation, let's warm up with a fast matrix-based algorithm to compute the output from a neural network. We actually already briefly saw this algorithm near the end of the last chapter (section \ref{sec:1.7}), but I described it quickly, so it's worth revisiting in detail. In particular, this is a good way of getting comfortable with the notation used in backpropagation, in a familiar context.

Let's begin with a notation which lets us refer to weights in the network in an unambiguous way. We'll use $w^l_{jk}$ to denote the weight for the connection from the $k$-th neuron in the $(l-1)$-th layer to the $j$-th neuron in the $l$-th layer. So, for example, the diagram below shows the weight on a connection from the fourth neuron in the second layer to the second neuron in the third layer of a network:

\begin{center}
	\includegraphics[width=0.7\linewidth]{./figures/ch2/tikz16}
\end{center}
This notation is cumbersome at first, and it does take some work to master. But with a little effort you'll find the notation becomes easy and natural. One quirk of the notation is the ordering of the j and k indices. You might think that it makes more sense to use j to refer to the input neuron, and k to the output neuron, not vice versa, as is actually done. I'll explain the reason for this quirk below.
We use a similar notation for the network's biases and activations. Explicitly, we use $b^l_j$ for the bias of the $j$-th neuron in the $l$-th layer. And we use $a^l_j$ for the activation of the $j$-th neuron in the $l$-th layer. The following diagram shows examples of these notations in use:
\begin{center}
\includegraphics[width=0.45\linewidth]{./figures/ch2/tikz17}
\end{center}
With these notations, the activation $a^l_j$ of the $j$-th neuron in the $l$-th layer is related to the activations in the $(l-1)$-th layer by the equation (compare Equation~(\ref{eq:4}) and surrounding discussion in the last chapter)

\begin{equation}
	a^{l}_j = \sigma\left( \sum_k w^{l}_{jk} a^{l-1}_k + b^l_j \right),\tag{23}\label{eq:23}
\end{equation}
where the sum is over all neurons $k$ in the $(l-1)$-th layer. To rewrite this expression in a matrix form we define a \textit{weight matrix} $w^l$ for each layer, $l$. The entries of the weight matrix $w^l$ are just the weights connecting to the $l$-th layer of neurons, that is, the entry in the $j$-th row and $k$-th column is $w^l_{jk}$. Similarly, for each layer $l$ we define a \textit{bias vector}, $b^l$. You can probably guess how this works -- the components of the bias vector are just the values $b^l_j$, one component for each neuron in the $l$-th layer. And finally, we define an activation vector al whose components are the activations $a^l_j$.
The last ingredient we need to rewrite (\ref{eq:23}) in a matrix form is the idea of vectorizing a function such as $\sigma$. We met vectorization briefly in the last chapter, but to recap, the idea is that we want to apply a function such as $\sigma$ to every element in a vector $v$. We use the obvious notation $\sigma(v)$ to denote this kind of elementwise application of a function. That is, the components of $\sigma(v)$ are just $\sigma(v)_j = \sigma(v_j)$. As an example, if we have the function $f(x)=x^2$ then the vectorized form of $f$ has the effect
\begin{equation}
f\left(\left[ \begin{array}{c} 2 \\ 3 \end{array} \right] \right)
= \left[ \begin{array}{c} f(2) \\ f(3) \end{array} \right]
= \left[ \begin{array}{c} 4 \\ 9 \end{array} \right],
\tag{24}\label{eq:24}
\end{equation}
that is, the vectorized $f$ just squares every element of the vector.

With these notations in mind, Equation~(\ref{eq:23}) can be rewritten in the beautiful and compact vectorized form
\begin{equation}
a^{l} = \sigma(w^l a^{l-1}+b^l).
\tag{25}\label{eq:25}
\end{equation}
This expression gives us a much more global way of thinking about how the activations in one layer relate to activations in the previous layer: we just apply the weight matrix to the activations, then add the bias vector, and finally apply the $\sigma$ function\footnote{By the way, it's this expression that motivates the quirk in the $w^l_{jk}$ notation mentioned earlier. If we used $j$ to index the input neuron, and $k$ to index the output neuron, then we'd need to replace the weight matrix in Equation~(\ref{eq:25}) by the transpose of the weight matrix. That's a small change, but annoying, and we'd lose the easy simplicity of saying (and thinking) ``apply the weight matrix to the activations''.}. That global view is often easier and more succinct (and involves fewer indices!) than the neuron-by-neuron view we've taken to now. Think of it as a way of escaping index hell, while remaining precise about what's going on. The expression is also useful in practice, because most matrix libraries provide fast ways of implementing matrix multiplication, vector addition, and vectorization. Indeed, the code (see \ref{sec:1.7}) in the last chapter made implicit use of this expression to compute the behaviour of the network.

When using Equation~(\ref{eq:25}) to compute $a^l$, we compute the intermediate quantity $z^l \equiv w^l a^{l-1}+b^l$ along the way. This quantity turns out to be useful enough to be worth naming: we call $z^l$ the weighted input to the neurons in layer $l$. We'll make considerable use of the weighted input $z^l$ later in the chapter. Equation~(\ref{eq:25}) is sometimes written in terms of the weighted input, as $a^l=\sigma(z^l)$. It's also worth noting that $z^l$ has components $z^l_j= \sum_k w^l_{jk} a^{l-1}_k+b^l_j$, that is, $z^l_j$ is just the weighted input to the activation function for neuron $j$ in layer $l$.

\section{The two assumptions we need about the cost function}
The goal of backpropagation is to compute the partial derivatives $\partial{}C/\partial{}w$ and $\partial{}C/\partial{}b$ of the cost function C with respect to any weight w or bias b in the network. For backpropagation to work we need to make two main assumptions about the form of the cost function. Before stating those assumptions, though, it's useful to have an example cost function in mind. We'll use the quadratic cost function from last chapter (c.f. Equation~(\ref{eq:6})). In the notation of the last section, the quadratic cost has the form
\begin{equation}
C = \frac{1}{2n} \sum_x \left\|y(x)-a^L(x)\right\|^2,
\tag{26}\label{eq:26}
\end{equation}
where: $n$ is the total number of training examples; the sum is over individual training examples, $x$; $y=y(x)$ is the corresponding desired output; $L$ denotes the number of layers in the network; and $a^L=a^L(x)$ is the vector of activations output from the network when $x$ is input.

Okay, so what assumptions do we need to make about our cost function, $C$, in order that backpropagation can be applied? The first assumption we need is that the cost function can be written as an average $C=\frac1n\sum_xC_x$ over cost functions $C_x$ for individual training examples, $x$. This is the case for the quadratic cost function, where the cost for a single training example is $C_x=\frac12\|y-a^L\|^2$. This assumption will also hold true for all the other cost functions we'll meet in this book.

The reason we need this assumption is because what backpropagation actually lets us do is compute the partial derivatives $\partial{}C_x/\partial{}w$ and $\partial{}C_x/\partial{}b$ for a single training example. We then recover $\partial{}C/\partial{}w$ and $\partial{}C/\partial{}b$ by averaging over training examples. In fact, with this assumption in mind, we'll suppose the training example $x$ has been fixed, and drop the $x$ subscript, writing the cost $C_x$ as $C$. We'll eventually put the $x$ back in, but for now it's a notational nuisance that is better left implicit.

The second assumption we make about the cost is that it can be written as a function of the outputs from the neural network:

\begin{center}
\includegraphics[width=0.7\linewidth]{./figures/ch2/tikz18}
\end{center}
For example, the quadratic cost function satisfies this requirement, since the quadratic cost for a single training example $x$ may be written as
\begin{equation}
C = \frac{1}{2} \left\|y-a^L\right\|^2 = \frac{1}{2} \sum_j \left(y_j-a^L_j\right)^2,
\tag{27}\label{eq:27}
\end{equation}
and thus is a function of the output activations. Of course, this cost function also depends on the desired output $y$, and you may wonder why we're not regarding the cost also as a function of $y$. Remember, though, that the input training example $x$ is fixed, and so the output $y$ is also a fixed parameter. In particular, it's not something we can modify by changing the weights and biases in any way, i.e., it's not something which the neural network learns. And so it makes sense to regard $C$ as a function of the output activations $a^L$ alone, with $y$ merely a parameter that helps define that function.
\section{The Hadamard product, $s\odot{}t$}
The backpropagation algorithm is based on common linear algebraic operations -- things like vector addition, multiplying a vector by a matrix, and so on. But one of the operations is a little less commonly used. In particular, suppose $s$ and $t$ are two vectors of the same dimension. Then we use $s\odot{}t$ to denote the elementwise product of the two vectors. Thus the components of $s\odot{}t$ are just $(s \odot t)_j = s_j t_j$. As an example,
\begin{equation}
	\left[\begin{array}{c} 1 \\ 2 \end{array}\right] 
	\odot \left[\begin{array}{c} 3 \\ 4\end{array} \right]
	= \left[ \begin{array}{c} 1 * 3 \\ 2 * 4 \end{array} \right]
	= \left[ \begin{array}{c} 3 \\ 8 \end{array} \right].
	\tag{28}\label{eq:28}
\end{equation}
This kind of elementwise multiplication is sometimes called the \textit{Hadamard product} or \textit{Schur product}. We'll refer to it as the Hadamard product. Good matrix libraries usually provide fast implementations of the Hadamard product, and that comes in handy when implementing backpropagation.

\section{The four fundamental equations behind backpropagation}
\label{sec:2.4}
Backpropagation is about understanding how changing the weights and biases in a network changes the cost function. Ultimately, this means computing the partial derivatives $\partial{}C/\partial{}w^l_{jk}$ and $\partial{}C/\partial{}b^l_j$. But to compute those, we first introduce an intermediate quantity, $\delta^l_j$, which we call the \textit{error} in the $j$-th neuron in the $l$-th layer. Backpropagation will give us a procedure to compute the error $\delta^l_j$, and then will relate $\delta^l_j$ to $\partial{}C/\partial{}w^l_{jk}$ and $\partial{}C/\partial{}b^l_j$.

To understand how the error is defined, imagine there is a demon in our neural network:

\begin{center}
	\includegraphics[width=0.6\linewidth]{./figures/ch2/tikz19}
\end{center}
The demon sits at the $j$-th neuron in layer $l$. As the input to the neuron comes in, the demon messes with the neuron's operation. It adds a little change $\Delta{}z^l_j$ to the neuron's weighted input, so that instead of outputting $\sigma(z^l_j)$, the neuron instead outputs $\sigma(z^l_j+\Delta{}z^l_j)$. This change propagates through later layers in the network, finally causing the overall cost to change by an amount $\frac{\partial{}C}{\partial{}z^l_j}\Delta{}z^l_j$.

Now, this demon is a good demon, and is trying to help you improve the cost, i.e., they're trying to find a $\Delta{}z^l_j$ which makes the cost smaller. Suppose $\partial{}C/\partial{}z^l_j$ has a large value (either positive or negative). Then the demon can lower the cost quite a bit by choosing $\Delta{}z^l_j$ to have the opposite sign to $\partial{}C/\partial{}z^l_j$. By contrast, if $\partial{}C/\partial{}z^l_j$ is close to zero, then the demon can't improve the cost much at all by perturbing the weighted input $z^l_j$. So far as the demon can tell, the neuron is already pretty near optimal\footnote{This is only the case for small changes $\Delta{}z^l_j$, of course. We'll assume that the demon is constrained to make such small changes.}. And so there's a heuristic sense in which $\partial{}C/\partial{}z^l_j$ is a measure of the error in the neuron.

Motivated by this story, we define the error $\delta^l_j$ of neuron $j$ in layer $l$ by
\begin{equation}
	\delta^l_j\equiv\frac{\partial{}C}{\partial{}z^l_j}.\tag{29}\label{eq:29}
\end{equation}
As per our usual conventions, we use $\delta{}^l$ to denote the vector of errors associated with layer $l$. Backpropagation will give us a way of computing $\delta{}^l$ for every layer, and then relating those errors to the quantities of real interest, $\partial{}C/\partial{}w^l_{jk}$ and $\partial{}C/\partial{}b^l_j$.

You might wonder why the demon is changing the weighted input $z^l_j$. Surely it'd be more natural to imagine the demon changing the output activation $a^l_j$, with the result that we'd be using $\frac{\partial{}C}{\partial{}a^l_j}$ as our measure of error. In fact, if you do this things work out quite similarly to the discussion below. But it turns out to make the presentation of backpropagation a little more algebraically complicated. So we'll stick with $\delta^l_j=\frac{\partial{}C}{\partial{}z^l_j}$ as our measure of error\footnote{In classification problems like MNIST the term ``error'' is sometimes used to mean the classification failure rate. E.g., if the neural net correctly classifies 96.0 percent of the digits, then the error is 4.0 percent. Obviously, this has quite a different meaning from our $\delta$ vectors. In practice, you shouldn't have trouble telling which meaning is intended in any given usage.}.

\textbf{Plan of attack}: Backpropagation is based around four fundamental equations. Together, those equations give us a way of computing both the error $\delta{}^l$ and the gradient of the cost function. I state the four equations below. Be warned, though: you shouldn't expect to instantaneously assimilate the equations. Such an expectation will lead to disappointment. In fact, the backpropagation equations are so rich that understanding them well requires considerable time and patience as you gradually delve deeper into the equations. The good news is that such patience is repaid many times over. And so the discussion in this section is merely a beginning, helping you on the way to a thorough understanding of the equations.

Here's a preview of the ways we'll delve more deeply into the equations later in the chapter: I'll give a \hyperref[sec:2.5]{short proof of the equations}, which helps explain why they are true; we'll restate the equations in algorithmic form as \hyperref[sec:2.6]{pseudocode}, and see how the pseudocode can be implemented as real, \hyperref[sec:2.7]{running Python code}; and, in the final section of the chapter, we'll develop an intuitive picture of what the backpropagation equations mean, and how someone \hyperref[sec:2.9]{might discover them from scratch}. Along the way we'll return repeatedly to the four fundamental equations, and as you deepen your understanding those equations will come to seem comfortable and, perhaps, even beautiful and natural.

\textbf{An equation for the error in the output layer,} $\delta^L$: The components of $\delta^L$ are given by
\begin{equation}
	\delta^L_j = \frac{\partial C}{\partial a^L_j} \sigma'(z^L_j).
	\tag{BP1}\label{eq:BP1}
\end{equation}
This is a very natural expression. The first term on the right, $\partial{}C/\partial{}a^L_j$, just measures how fast the cost is changing as a function of the $j$-th output activation. If, for example, $C$ doesn't depend much on a particular output neuron, $j$, then $\delta^L_j$ will be small, which is what we'd expect. The second term on the right, $\sigma'(z^L_j)$, measures how fast the activation function $\sigma$ is changing at $z^L_j$.

Notice that everything in Eq.~(\ref{eq:BP1}) is easily computed. In particular, we compute $z^L_j$ while computing the behaviour of the network, and it's only a small additional overhead to compute $\sigma'(z^L_j)$. The exact form of $\partial{}C/\partial{}a^L_j$ will, of course, depend on the form of the cost function. However, provided the cost function is known there should be little trouble computing $\partial{}C/\partial{}a^L_j$. For example, if we're using the quadratic cost function then $C=\frac12\sum_j(y_j-a^L_j)^2$, and so $\partial{}C/\partial{}a^L_j=(a^L_j-y_j)$, which obviously is easily computable.

Equation~(\ref{eq:BP1}) is a componentwise expression for $\delta^L$. It's a perfectly good expression, but not the matrix-based form we want for backpropagation. However, it's easy to rewrite the equation in a matrix-based form, as
%$\delta^l$=∇aC⊙$\sigma$′(zL).(BP1a)
\begin{equation}
\delta^L = \nabla_a C \odot \sigma'(z^L).
\tag{BP1a}\label{eq:BP1a}
\end{equation}
Here, $\nabla_a C$ is defined to be a vector whose components are the partial derivatives $\partial{}C/\partial{}a^L_j$. You can think of $\nabla_a C$ as expressing the rate of change of $C$ with respect to the output activations. It's easy to see that Equations (\ref{eq:BP1a}) and (\ref{eq:BP1}) are equivalent, and for that reason from now on we'll use (\ref{eq:BP1}) interchangeably to refer to both equations. As an example, in the case of the quadratic cost we have $\nabla_a C =(a^L-y)$, and so the fully matrix-based form of (\ref{eq:BP1}) becomes
\begin{equation}
\delta^L = (a^L-y) \odot \sigma'(z^L).
\tag{30}\label{eq:30}
\end{equation}
As you can see, everything in this expression has a nice vector form, and is easily computed using a library such as Numpy.

\textbf{An equation for the error $\delta^l$ in terms of the error in the next layer}, $\delta^{l+1}$: In particular
\begin{equation}
	\delta^l = \left((w^{l+1})^T \delta^{l+1}\right) \odot \sigma'(z^l),
	\tag{BP2}\label{eq:BP2}
\end{equation}
%$\delta^l$=(($w^{l+1}$)T$\delta^{l+1}$)⊙$\sigma$′(zl),(\ref{eq:BP2})
where $(w^{l+1})^T$ is the transpose of the weight matrix $w^{l+1}$ for the ($l$+1)-th layer. This equation appears complicated, but each element has a nice interpretation. Suppose we know the error $\delta^{l+1}$ at the ($l$+1)-th layer. When we apply the transpose weight matrix, $(w^{l+1})^T$, we can think intuitively of this as moving the error \textit{backward} through the network, giving us some sort of measure of the error at the output of the $l$-th layer. We then take the Hadamard product $\odot \sigma'(z^l)$. This moves the error backward through the activation function in layer $l$, giving us the error $\delta^l$ in the weighted input to layer $l$.

By combining (\ref{eq:BP2}) with (\ref{eq:BP1}) we can compute the error $\delta^l$ for any layer in the network. We start by using (\ref{eq:BP1}) to compute $\delta^L$, then apply Equation~(\ref{eq:BP2}) to compute $\delta^{L-1}$, then Equation~(\ref{eq:BP2}) again to compute $\delta^{L-2}$, and so on, all the way back through the network.

\textbf{An equation for the rate of change of the cost with respect to any bias in the network:} In particular:
%\partial{}C\partial{}b^l_j=\delta^l_j.(\ref{eq:BP3})
\begin{equation}
\frac{\partial C}{\partial b^l_j} = \delta^l_j.
\tag{BP3}\label{eq:BP3}
\end{equation}
That is, the error $\delta^l_j$ is \textit{exactly equal} to the rate of change $\partial{}C/\partial{}b^l_j$. This is great news, since (\ref{eq:BP1}) and (\ref{eq:BP2}) have already told us how to compute $\delta^l_j$. We can rewrite (\ref{eq:BP3}) in shorthand as
\begin{equation}
\frac{\partial C}{\partial b} = \delta,\tag{31}\label{eq:31}
\end{equation}
where it is understood that $\delta$ is being evaluated at the same neuron as the bias $b$.

\textbf{An equation for the rate of change of the cost with respect to any weight in the network:} In particular:
%\partial{}C\partial{}w^l_{jk}=al-1k\delta^l_j.(\ref{eq:BP4})
\begin{equation}
\frac{\partial C}{\partial w^l_{jk}} = a^{l-1}_k \delta^l_j.
\tag{BP4}\label{eq:BP4}
\end{equation}
This tells us how to compute the partial derivatives $\partial{}C/\partial{}w^l_{jk}$ in terms of the quantities $\delta^l$ and $a^{l-1}$, which we already know how to compute. The equation can be rewritten in a less index-heavy notation as
%\partial{}C\partial{}w=ain\Delta{}out,(32)
\begin{equation}
\frac{\partial C}{\partial w} = a_{\rm in} \delta_{\rm out},
\tag{32}\label{eq:32}
\end{equation}
where it's understood that $a_\mathrm{in}$ is the activation of the neuron input to the weight $w$, and $\delta_\mathrm{out}$ is the error of the neuron output from the weight $w$. Zooming in to look at just the weight $w$, and the two neurons connected by that weight, we can depict this as:

\begin{center}
	\includegraphics[width=0.35\linewidth]{./figures/ch2/tikz20}
\end{center}
A nice consequence of Equation~(\ref{eq:32}) is that when the activation $a_\mathrm{in}$ is small, $a_\mathrm{in}\approx0$, the gradient term $\partial{}C/\partial{}w$ will also tend to be small. In this case, we'll say the weight \textit{learns slowly}, meaning that it's not changing much during gradient descent. In other words, one consequence of (\ref{eq:BP4}) is that weights output from low-activation neurons learn slowly.

There are other insights along these lines which can be obtained from (\ref{eq:BP1})--(\ref{eq:BP4}). Let's start by looking at the output layer. Consider the term $\sigma'(z^L_j)$ in (\ref{eq:BP1}). Recall from the graph of the sigmoid function in the last chapter that the $\sigma$ function becomes very flat when $\sigma(z^L_j)$ is approximately 0 or 1. When this occurs we will have $\sigma'(z^L_j)\approx0$. And so the lesson is that a weight in the final layer will learn slowly if the output neuron is either low activation ($\approx0$) or high activation ($\approx1$). In this case it's common to say the output neuron has \textit{saturated} and, as a result, the weight has stopped learning (or is learning slowly). Similar remarks hold also for the biases of output neuron.

We can obtain similar insights for earlier layers. In particular, note the $\sigma'(z^l)$ term in (\ref{eq:BP2}). This means that $\delta^l_j$ is likely to get small if the neuron is near saturation. And this, in turn, means that any weights input to a saturated neuron will learn slowly\footnote{This reasoning won't hold if $(w^{l+1})^T\delta^{l+1}$ has large enough entries to compensate for the smallness of $\sigma'(z^l_j)$. But I'm speaking of the general tendency.}.

Summing up, we've learnt that a weight will learn slowly if either the input neuron is low-activation, or if the output neuron has saturated, i.e., is either high- or low-activation.

None of these observations is too greatly surprising. Still, they help improve our mental model of what's going on as a neural network learns. Furthermore, we can turn this type of reasoning around. The four fundamental equations turn out to hold for any activation function, not just the standard sigmoid function (that's because, as we'll see in a moment, the proofs don't use any special properties of $\sigma$). And so we can use these equations to design activation functions which have particular desired learning properties. As an example to give you the idea, suppose we were to choose a (non-sigmoid) activation function $\sigma$ so that $\sigma'$ is always positive, and never gets close to zero. That would prevent the slow-down of learning that occurs when ordinary sigmoid neurons saturate. Later in the book we'll see examples where this kind of modification is made to the activation function. Keeping the four equations (\ref{eq:BP1})--(\ref{eq:BP4}) in mind can help explain why such modifications are tried, and what impact they can have.
\begin{center}
	\includegraphics[width=0.65\linewidth]{figures/ch2/tikz21}
\end{center}

\begin{exercize}{Problem}
	\item \textbf{Alternate presentation of the equations of backpropagation:} I've stated the equations of backpropagation (notably (\ref{eq:BP1}) and (\ref{eq:BP2})) using the Hadamard product. This presentation may be disconcerting if you're unused to the Hadamard product. There's an alternative approach, based on conventional matrix multiplication, which some readers may find enlightening.
	\begin{enumerate}
		\item[(1)]  Show that (\ref{eq:BP1}) may be rewritten as
		\begin{equation}
			\delta^L = \Sigma'(z^L) \nabla_a C,
			\tag{33}\label{eq:33}
		\end{equation}
		where $\Sigma'(z^L)$ is a square matrix whose diagonal entries are the values $\sigma'(z^L_j)$, and whose off-diagonal entries are zero. Note that this matrix acts on $\nabla_a C$ by conventional matrix multiplication.
		\item[(2)] Show that (\ref{eq:BP2}) may be rewritten as
		\begin{equation}
			\delta^l = \Sigma'(z^l) (w^{l+1})^T \delta^{l+1}.
			\tag{34}\label{eq:34}
		\end{equation}
		\item[(3)] By combining observations (1) and (2) show that
		\begin{equation}
			\delta^l = \Sigma'(z^l) (w^{l+1})^T \ldots \Sigma'(z^{L-1}) (w^L)^T \Sigma'(z^L) \nabla_a C
			\tag{35}\label{eq:35}
		\end{equation}
		For readers comfortable with matrix multiplication this equation may be easier to understand than (\ref{eq:BP1}) and (\ref{eq:BP2}). The reason I've focused on (\ref{eq:BP1}) and (\ref{eq:BP2}) is because that approach turns out to be faster to implement numerically.
	\end{enumerate}
\end{exercize}
\section{Proof of the four fundamental equations (optional)}
\label{sec:2.5}
We'll now prove the four fundamental equations (\ref{eq:BP1})--(\ref{eq:BP4}). All four are consequences of the chain rule from multivariable calculus. If you're comfortable with the chain rule, then I strongly encourage you to attempt the derivation yourself before reading on.

Let's begin with Equation~(\ref{eq:BP1}), which gives an expression for the output error, $\delta^l$. To prove this equation, recall that by definition
%$\delta^l$j=\partial{}C\partial{}z^L_j.(36)
\begin{equation}
	\delta^L_j = \frac{\partial C}{\partial z^L_j}.
	\tag{36}\label{eq:36}
\end{equation}
Applying the chain rule, we can re-express the partial derivative above in terms of partial derivatives with respect to the output activations,
\begin{equation}
	\delta^L_j = \sum_k \frac{\partial C}{\partial a^L_k} \frac{\partial a^L_k}{\partial z^L_j},
	\tag{37}\label{eq:37}
\end{equation}%$\delta^l$j=∑k\partial{}C\partial{}aLk\partial{}aLk\partial{}z^L_j,(37)
where the sum is over all neurons $k$ in the output layer. Of course, the output activation $a^L_k$ of the $k$-th neuron depends only on the weighted input $z^L_j$ for the $j$-th neuron when $k=j$. And so $\partial{}a^L_k/\partial{}z^L_j$ vanishes when $k\ne{}j$. As a result we can simplify the previous equation to
\begin{equation}
	\delta^L_j = \frac{\partial C}{\partial a^L_j} \frac{\partial a^L_j}{\partial z^L_j}.
	\tag{38}\label{eq:38}
\end{equation}%$\delta^l$j=\partial{}C\partial{}a^L_j\partial{}a^L_j\partial{}z^L_j.(38)
Recalling that $a^L_j=\sigma(z^L_j)$ the second term on the right can be written as $\sigma'(z^L_j)$, and the equation becomes
%$\delta^l$j=\partial{}C\partial{}a^L_j$\sigma$′(z^L_j),(39)
\begin{equation}
	\delta^L_j = \frac{\partial C}{\partial a^L_j} \sigma'(z^L_j),
	\tag{39}\label{eq:39}
\end{equation}
which is just (\ref{eq:BP1}), in component form.
%
Next, we'll prove (\ref{eq:BP2}), which gives an equation for the error $\delta^l$ in terms of the error in the next layer, $\delta^{l+1}$. To do this, we want to rewrite $\delta^l_j=\partial{}C/\partial{}z^l_j$ in terms of $\delta^{l+1}_k=\partial{}C/\partial{}z^{l+1}_k$. We can do this using the chain rule,
\begin{equation}
\delta^l_j =  \frac{\partial C}{\partial z^l_j} =  \sum_k \frac{\partial C}{\partial z^{l+1}_k} \frac{\partial z^{l+1}_k}{\partial z^l_j} = \sum_k \frac{\partial z^{l+1}_k}{\partial z^l_j} \delta^{l+1}_k,
\tag{42}\label{eq:42}
\end{equation}
%\delta^l_j===\partial{}C\partial{}z^l_j∑k\partial{}C\partial{}zl+1k\partial{}zl+1k\partial{}z^l_j∑k\partial{}zl+1k\partial{}z^l_j$\delta^{l+1}$k,(40)(41)(42)
where in the last line we have interchanged the two terms on the right-hand side, and substituted the definition of $\delta^{l+1}_k$. To evaluate the first term on the last line, note that
%zl+1k=∑j$w^{l+1}$kjalj+bl+1k=∑j$w^{l+1}$kj$\sigma$(z^l_j)+bl+1k.(43)
\begin{equation}
	z^{l+1}_k = \sum_j w^{l+1}_{kj} a^l_j +b^{l+1}_k = \sum_j w^{l+1}_{kj} \sigma(z^l_j) +b^{l+1}_k.
	\tag{43}\label{eq:43}
\end{equation}
Differentiating, we obtain
%\partial{}zl+1k\partial{}z^l_j=$w^{l+1}$kj$\sigma$′(z^l_j).(44)
\begin{equation}
	\frac{\partial z^{l+1}_k}{\partial z^l_j} = w^{l+1}_{kj} \sigma'(z^l_j).
	\tag{44}\label{eq:44}
\end{equation}
Substituting back into (\ref{eq:42}) we obtain
\begin{equation}
	\delta^l_j = \sum_k w^{l+1}_{kj}  \delta^{l+1}_k \sigma'(z^l_j).
	\tag{45}\label{eq:45}
\end{equation}%\delta^l_j=∑k$w^{l+1}$kj$\delta^{l+1}$k$\sigma$′(z^l_j).(45)
This is just (\ref{eq:BP2}) written in component form.

The final two equations we want to prove are (\ref{eq:BP3}) and (\ref{eq:BP4}). These also follow from the chain rule, in a manner similar to the proofs of the two equations above. I leave them to you as an exercise.

\begin{exercize}{Exercise}
	\item Prove Equations (\ref{eq:BP3}) and (\ref{eq:BP4}).
\end{exercize}
That completes the proof of the four fundamental equations of backpropagation. The proof may seem complicated. But it's really just the outcome of carefully applying the chain rule. A little less succinctly, we can think of backpropagation as a way of computing the gradient of the cost function by systematically applying the chain rule from multi-variable calculus. That's all there really is to backpropagation -- the rest is details.

\section{The backpropagation algorithm}
\label{sec:2.6}
The backpropagation equations provide us with a way of computing the gradient of the cost function. Let's explicitly write this out in the form of an algorithm:
\begin{enumerate}
\item \textbf{Input} $x$: Set the corresponding activation $a^1$ for the input layer.
\item \textbf{Feedforward:} For each $l=2,3,\ldots{},L$ compute $z^l=w^la^{l-1}+b^l$ and $a^l=\sigma(z_l)$.
\item \textbf{Output error} $\delta^L$: Compute the vector $\delta^{L} = \nabla_a C \odot \sigma'(z^L)$.
\item \textbf{Backpropagate the error}: For each $l = L-1, L-2,\ldots, 2$ compute $\delta^{l} = ((w^{l+1})^T \delta^{l+1}) \odot \sigma'(z^{l})$.
\item \textbf{Output}: The gradient of the cost function is given by $\frac{\partial{}C}{\partial{}w^l_{jk}}=a^{l-1}_k\delta^l_j$ and $\frac{\partial{}C}{\partial{}b^l_j}=\delta^l_j$.
\end{enumerate}
Examining the algorithm you can see why it's called\textit{ back}propagation. We compute the error vectors $\delta^l$ backward, starting from the final layer. It may seem peculiar that we're going through the network backward. But if you think about the proof of backpropagation, the backward movement is a consequence of the fact that the cost is a function of outputs from the network. To understand how the cost varies with earlier weights and biases we need to repeatedly apply the chain rule, working backward through the layers to obtain usable expressions.

\begin{exercize}{Exercises}
	\item \textbf{Backpropagation with a single modified neuron} Suppose we modify a single neuron in a feedforward network so that the output from the neuron is given by $f(\sum_j w_j x_j + b)$, where $f$ is some function other than the sigmoid. How should we modify the backpropagation algorithm in this case?
	\item \textbf{Backpropagation with linear neurons} Suppose we replace the usual non-linear $\sigma$ function with $\sigma(z)=z$ throughout the network. Rewrite the backpropagation algorithm for this case.
\end{exercize}
As I've described it above, the backpropagation algorithm computes the gradient of the cost function for a single training example, $C=C_x$. In practice, it's common to combine backpropagation with a learning algorithm such as stochastic gradient descent, in which we compute the gradient for many training examples. In particular, given a mini-batch of $m$ training examples, the following algorithm applies a gradient descent learning step based on that mini-batch:
\begin{enumerate}
\item Input a set of training examples
\item For each training example $x$: Set the corresponding input activation $a^{x,1}$, and perform the following steps:
	\begin{itemize} 
		\item Feedforward: For each l$=2,3,\ldots,L$ compute $z^{x,l}=w^la^{x,l-1}+b^l$ and $a^{x,l}=\sigma(z^{x,l})$.
		\item Output error $\delta^{x,L}$: Compute the vector $\delta^{x,L} = \nabla_a C_x \odot \sigma'(z^{x,L})$.
		\item Backpropagate the error: For each $l=L-1,L-2,\ldots,2$ compute $\delta^{x,l} = ((w^{l+1})^T \delta^{x,l+1}) \odot \sigma'(z^{x,l})$.
	\end{itemize}
\item Gradient descent: For each $l=L,L-1,\ldots,2$ update the weights according to the rule $w^l \to
w^l-\frac{\eta}{m} \sum_x \delta^{x,l} (a^{x,l-1})^T$, and the biases according to the rule $b^l \to b^l-\frac{\eta}{m}\sum_x \delta^{x,l}$.
\end{enumerate}
Of course, to implement stochastic gradient descent in practice you also need an outer loop generating mini-batches of training examples, and an outer loop stepping through multiple epochs of training. I've omitted those for simplicity.
\section{The code for backpropagation}
\label{sec:2.7}
Having understood backpropagation in the abstract, we can now understand the code used in the last chapter to implement backpropagation. Recall from that chapter that the code was contained in the \inline{update_mini_batch} and \inline{backprop} methods of the \inline{Network} class. The code for these methods is a direct translation of the algorithm described above. In particular, the \inline{update_mini_batch}  method updates the \inline{Network}'s weights and biases by computing the gradient for the current \inline{mini_batch} of training examples:
\begin{lstlisting}
class Network(object):
...
	def update_mini_batch(self, mini_batch, eta):
		"""Update the network's weights and biases by applying
		gradient descent using backpropagation to a single mini batch.
		The "mini_batch" is a list of tuples "(x, y)", and "eta"
		is the learning rate."""
		nabla_b = [np.zeros(b.shape) for b in self.biases]
		nabla_w = [np.zeros(w.shape) for w in self.weights]
		for x, y in mini_batch:
			delta_nabla_b, delta_nabla_w = self.backprop(x, y)
			nabla_b = [nb+dnb for nb, dnb in zip(nabla_b, delta_nabla_b)]
			nabla_w = [nw+dnw for nw, dnw in zip(nabla_w, delta_nabla_w)]
		self.weights = [w-(eta/len(mini_batch))*nw 
						for w, nw in zip(self.weights, nabla_w)]
		self.biases = [b-(eta/len(mini_batch))*nb 
						for b, nb in zip(self.biases, nabla_b)]

\end{lstlisting}
Most of the work is done by the line \inline{delta_nabla_b}, \inline{delta_nabla_w = self.backprop(x, y)} which uses the \inline{backprop} method to figure out the partial derivatives $\partial{}C_x/\partial{}b^l_j$ and $\partial{}C_x/\partial{}w^l_{jk}$. The \inline{backprop} method follows the algorithm in the last section closely. There is one small change -- we use a slightly different approach to indexing the layers. This change is made to take advantage of a feature of Python, namely the use of negative list indices to count backward from the end of a list, so, e.g., \inline{l[-3]} is the third last entry in a list \inline{l}. The code for \inline{backprop} is below, together with a few helper functions, which are used to compute the $\sigma$ function, the derivative $\sigma'$, and the derivative of the cost function. With these inclusions you should be able to understand the code in a self-contained way. If something's tripping you up, you may find it helpful to consult the original description (and complete listing) of the code.

\begin{lstlisting}
class Network(object):
...
	def backprop(self, x, y):
		"""Return a tuple "(nabla_b, nabla_w)" representing the
		gradient for the cost function C_x.  "nabla_b" and
		"nabla_w" are layer-by-layer lists of numpy arrays, similar
		to "self.biases" and "self.weights"."""
		nabla_b = [np.zeros(b.shape) for b in self.biases]
		nabla_w = [np.zeros(w.shape) for w in self.weights]
		# feedforward
		activation = x
		activations = [x] # list to store all the activations, layer by layer
		zs = [] # list to store all the z vectors, layer by layer
		for b, w in zip(self.biases, self.weights):
			z = np.dot(w, activation)+b
			zs.append(z)
			activation = sigmoid(z)
			activations.append(activation)
		# backward pass
		delta = self.cost_derivative(activations[-1], y) * sigmoid_prime(zs[-1])
		nabla_b[-1] = delta
		nabla_w[-1] = np.dot(delta, activations[-2].transpose())
		# Note that the variable l in the loop below is used a little
		# differently to the notation in Chapter 2 of the book.  Here,
		# l = 1 means the last layer of neurons, l = 2 is the
		# second-last layer, and so on.  It's a renumbering of the
		# scheme in the book, used here to take advantage of the fact
		# that Python can use negative indices in lists.
		for l in xrange(2, self.num_layers):
			z = zs[-l]
			sp = sigmoid_prime(z)
			delta = np.dot(self.weights[-l+1].transpose(), delta) * sp
			nabla_b[-l] = delta
			nabla_w[-l] = np.dot(delta, activations[-l-1].transpose())
		return (nabla_b, nabla_w)
...
	def cost_derivative(self, output_activations, y):
		"""Return the vector of partial derivatives \partial{} C_x /
		\partial{} a for the output activations."""
		return (output_activations-y) 
	def sigmoid(z):
		"""The sigmoid function."""
		return 1.0/(1.0+np.exp(-z))
	def sigmoid_prime(z):
		"""Derivative of the sigmoid function."""
		return sigmoid(z)*(1-sigmoid(z))
\end{lstlisting}

\begin{exercize}{Problem}
\label{backprop_over_minibatch}
\item \textbf{Fully matrix-based approach to backpropagation over a mini-batch} Our implementation of stochastic gradient descent loops over training examples in a mini-batch. It's possible to modify the backpropagation algorithm so that it computes the gradients for all training examples in a mini-batch simultaneously. The idea is that instead of beginning with a single input vector, $x$, we can begin with a matrix $X=[x_1x_2\ldots{}x_m]$ whose columns are the vectors in the mini-batch. We forward-propagate by multiplying by the weight matrices, adding a suitable matrix for the bias terms, and applying the sigmoid function everywhere. We backpropagate along similar lines. Explicitly write out pseudocode for this approach to the backpropagation algorithm. Modify \inline{network.py} so that it uses this fully matrix-based approach. The advantage of this approach is that it takes full advantage of modern libraries for linear algebra. As a result it can be quite a bit faster than looping over the mini-batch. (On my laptop, for example, the speedup is about a factor of two when run on MNIST classification problems like those we considered in the last chapter.) In practice, all serious libraries for backpropagation use this fully matrix-based approach or some variant.
\end{exercize}
\section{In what sense is backpropagation a fast algorithm?}
\label{sec:2.8}
In what sense is backpropagation a fast algorithm? To answer this question, let's consider another approach to computing the gradient. Imagine it's the early days of neural networks research. Maybe it's the 1950s or 1960s, and you're the first person in the world to think of using gradient descent to learn! But to make the idea work you need a way of computing the gradient of the cost function. You think back to your knowledge of calculus, and decide to see if you can use the chain rule to compute the gradient. But after playing around a bit, the algebra looks complicated, and you get discouraged. So you try to find another approach. You decide to regard the cost as a function of the weights $C=C(w)$ alone (we'll get back to the biases in a moment). You number the weights $w_1,w_2,\ldots$, and want to compute $\partial{}C/\partial{}w_j$ for some particular weight $w_j$. An obvious way of doing that is to use the approximation
\begin{equation}
	\frac{\partial C}{\partial w_{j}} \approx \frac{C(w+\epsilon e_j)-C(w)}{\epsilon},
	\tag{46}\label{eq:46}
\end{equation}
%\partial{}C\partial{}wj≈C(w+\epsilon{}ej)-C(w)\epsilon{},(46)
where $\epsilon>0$ is a small positive number, and $e_j$ is the unit vector in the $j$-th direction. In other words, we can estimate $\partial{}C/\partial{}w_j$ by computing the cost $C$ for two slightly different values of $w_j$, and then applying Equation~(\ref{eq:46}). The same idea will let us compute the partial derivatives $\partial{}C/\partial{}b$ with respect to the biases.

This approach looks very promising. It's simple conceptually, and extremely easy to implement, using just a few lines of code. Certainly, it looks much more promising than the idea of using the chain rule to compute the gradient!

Unfortunately, while this approach appears promising, when you implement the code it turns out to be extremely slow. To understand why, imagine we have a million weights in our network. Then for each distinct weight $w_j$ we need to compute $C(w+\epsilon{}e_j)$ in order to compute $\partial{}C/\partial{}w_j$. That means that to compute the gradient we need to compute the cost function a million different times, requiring a million forward passes through the network (per training example). We need to compute $C(w)$ as well, so that's a total of a million and one passes through the network.

What's clever about backpropagation is that it enables us to simultaneously compute all the partial derivatives $\partial{}C/\partial{}w_j$ using just one forward pass through the network, followed by one backward pass through the network. Roughly speaking, the computational cost of the backward pass is about the same as the forward pass\footnote{This should be plausible, but it requires some analysis to make a careful statement. It's plausible because the dominant computational cost in the forward pass is multiplying by the weight matrices, while in the backward pass it's multiplying by the transposes of the weight matrices. These operations obviously have similar computational cost.}. And so the total cost of backpropagation is roughly the same as making just two forward passes through the network. Compare that to the million and one forward passes we needed for the approach based on (\ref{eq:46})! And so even though backpropagation appears superficially more complex than the approach based on (\ref{eq:46}), it's actually much, much faster.

This speedup was first fully appreciated in 1986, and it greatly expanded the range of problems that neural networks could solve. That, in turn, caused a rush of people using neural networks. Of course, backpropagation is not a panacea. Even in the late 1980s people ran up against limits, especially when attempting to use backpropagation to train deep neural networks, i.e., networks with many hidden layers. Later in the book we'll see how modern computers and some clever new ideas now make it possible to use backpropagation to train such deep neural networks.
%
\section{Backpropagation: the big picture}
\label{sec:2.9}
As I've explained it, backpropagation presents two mysteries. First, what's the algorithm really doing? We've developed a picture of the error being backpropagated from the output. But can we go any deeper, and build up more intuition about what is going on when we do all these matrix and vector multiplications? The second mystery is how someone could ever have discovered backpropagation in the first place? It's one thing to follow the steps in an algorithm, or even to follow the proof that the algorithm works. But that doesn't mean you understand the problem so well that you could have discovered the algorithm in the first place. Is there a plausible line of reasoning that could have led you to discover the backpropagation algorithm? In this section I'll address both these mysteries.

To improve our intuition about what the algorithm is doing, let's imagine that we've made a small change $\Delta{}w^l_{jk}$ to some weight in the network, $w^l_{jk}$:
\begin{center}
	\includegraphics[scale=0.5]{./figures/ch2/tikz22}
\end{center}
That change in weight will cause a change in the output activation from the corresponding neuron:
\begin{center}
	\includegraphics[scale=0.5]{./figures/ch2/tikz23}
\end{center}
That, in turn, will cause a change in all the activations in the next layer:
\begin{center}
	\includegraphics[scale=0.5]{./figures/ch2/tikz24}
\end{center}
Those changes will in turn cause changes in the next layer, and then the next, and so on all the way through to causing a change in the final layer, and then in the cost function:
\begin{center}
	\includegraphics[scale=0.5]{./figures/ch2/tikz25}
\end{center}
The change $\Delta{}C$ in the cost is related to the change $\Delta{}w^l_{jk}$ in the weight by the equation
\begin{equation}
	\Delta C \approx \frac{\partial C}{\partial w^l_{jk}} \Delta w^l_{jk}.
	\tag{47}\label{eq:47}
\end{equation}
%\Delta{}C≈\partial{}C\partial{}w^l_{jk}\Delta{}w^l_{jk}.(47)
This suggests that a possible approach to computing $\partial{}C/\partial{}w^l_{jk}$ is to carefully track how a small change in $w^l_{jk}$ propagates to cause a small change in $C$. If we can do that, being careful to express everything along the way in terms of easily computable quantities, then we should be able to compute $\partial{}C/\partial{}w^l_{jk}$.

Let's try to carry this out. The change $\Delta{}w^l_{jk}$ causes a small change $\Delta{}a^l_j$ in the activation of the $j$-th neuron in the $l$-th layer. This change is given by
%\Delta{}alj≈\partial{}alj\partial{}w^l_{jk}\Delta{}w^l_{jk}.(48)
\begin{equation}
	\Delta a^l_j \approx \frac{\partial a^l_j}{\partial w^l_{jk}} \Delta w^l_{jk}.
	\tag{48}\label{eq:48}
\end{equation}
The change in activation $\Delta{}a^l_j$ will cause changes in all the activations in the next layer, i.e., the ($l$+1)-th layer. We'll concentrate on the way just a single one of those activations is affected, say $a^{l+1}_q$,
\begin{center}
	\includegraphics[scale=0.5]{./figures/ch2/tikz26}
\end{center}
In fact, it'll cause the following change:
\begin{equation}
	\Delta a^{l+1}_q \approx \frac{\partial a^{l+1}_q}{\partial a^l_j} \Delta a^l_j.
	\tag{49}\label{eq:49}
\end{equation}%\Delta{}al+1q≈\partial{}al+1q\partial{}alj\Delta{}alj.(49)
Substituting in the expression from Equation~(\ref{eq:48}), we get:
%\Delta{}al+1q≈\partial{}al+1q\partial{}alj\partial{}alj\partial{}w^l_{jk}\Delta{}w^l_{jk}.(50)
\begin{equation}
	\Delta a^{l+1}_q \approx \frac{\partial a^{l+1}_q}{\partial a^l_j} \frac{\partial a^l_j}{\partial w^l_{jk}} \Delta w^l_{jk}.
	\tag{50}\label{eq:50}
\end{equation}
Of course, the change $\Delta{}a^{l+1}_q$ will, in turn, cause changes in the activations in the next layer. In fact, we can imagine a path all the way through the network from $w^l_{jk}$ to $C$, with each change in activation causing a change in the next activation, and, finally, a change in the cost at the output. If the path goes through activations $a^l_j,a^{l+1}_q,\cdots,a^{L-1}_n,a^L_m$ then the resulting expression is
\begin{equation}
	\Delta C \approx \frac{\partial C}{\partial a^L_m} 
	\frac{\partial a^L_m}{\partial a^{L-1}_n}
	\frac{\partial a^{L-1}_n}{\partial a^{L-2}_p} \ldots
	\frac{\partial a^{l+1}_q}{\partial a^l_j}
	\frac{\partial a^l_j}{\partial w^l_{jk}} \Delta w^l_{jk},
	\tag{51}\label{eq:51}
\end{equation}
%\Delta{}C≈\partial{}C\partial{}aLm\partial{}aLm\partial{}aL-1n\partial{}aL-1n\partial{}aL-2p…\partial{}al+1q\partial{}alj\partial{}alj\partial{}w^l_{jk}\Delta{}w^l_{jk},(51)
that is, we've picked up a $\partial{}a/\partial{}a$ type term for each additional neuron we've passed through, as well as the $\partial{}C/\partial{}a^L_m$ term at the end. This represents the change in $C$ due to changes in the activations along this particular path through the network. Of course, there's many paths by which a change in $w^l_{jk}$ can propagate to affect the cost, and we've been considering just a single path. To compute the total change in $C$ it is plausible that we should sum over all the possible paths between the weight and the final cost, i.e.,
\begin{equation}
	\Delta C \approx \sum_{mnp\ldots q} \frac{\partial C}{\partial a^L_m} 
	\frac{\partial a^L_m}{\partial a^{L-1}_n}
	\frac{\partial a^{L-1}_n}{\partial a^{L-2}_p} \ldots
	\frac{\partial a^{l+1}_q}{\partial a^l_j} 
	\frac{\partial a^l_j}{\partial w^l_{jk}} \Delta w^l_{jk},
	\tag{52}\label{eq:52}
\end{equation}
%\Delta{}C≈∑mnp…q\partial{}C\partial{}aLm\partial{}aLm\partial{}aL-1n\partial{}aL-1n\partial{}aL-2p…\partial{}al+1q\partial{}alj\partial{}alj\partial{}w^l_{jk}\Delta{}w^l_{jk},(52)
where we've summed over all possible choices for the intermediate neurons along the path. Comparing with (\ref{eq:47}) we see that
%\partial{}C\partial{}w^l_{jk}=∑mnp…q\partial{}C\partial{}aLm\partial{}aLm\partial{}aL-1n\partial{}aL-1n\partial{}aL-2p…\partial{}al+1q\partial{}alj\partial{}alj\partial{}w^l_{jk}.(53)
\begin{equation}
	\frac{\partial C}{\partial w^l_{jk}} = \sum_{mnp\ldots q} \frac{\partial C}{\partial a^L_m} 
	\frac{\partial a^L_m}{\partial a^{L-1}_n}
	\frac{\partial a^{L-1}_n}{\partial a^{L-2}_p} \ldots
	\frac{\partial a^{l+1}_q}{\partial a^l_j} 
	\frac{\partial a^l_j}{\partial w^l_{jk}}.
	\tag{53}\label{eq:53}
\end{equation}
Now, Equation~(\ref{eq:53}) looks complicated. However, it has a nice intuitive interpretation. We're computing the rate of change of $C$ with respect to a weight in the network. What the equation tells us is that every edge between two neurons in the network is associated with a rate factor which is just the partial derivative of one neuron's activation with respect to the other neuron's activation. The edge from the first weight to the first neuron has a rate factor $\partial{}a^l_j/\partial{}w^l_{jk}$. The rate factor for a path is just the product of the rate factors along the path. And the total rate of change $\partial{}C/\partial{}w^l_{jk}$ is just the sum of the rate factors of all paths from the initial weight to the final cost. This procedure is illustrated here, for a single path:
\begin{center}
	\includegraphics[scale=0.5]{./figures/ch2/tikz27}
\end{center}
What I've been providing up to now is a heuristic argument, a way of thinking about what's going on when you perturb a weight in a network. Let me sketch out a line of thinking you could use to further develop this argument. First, you could derive explicit expressions for all the individual partial derivatives in Equation~(\ref{eq:53}). That's easy to do with a bit of calculus. Having done that, you could then try to figure out how to write all the sums over indices as matrix multiplications. This turns out to be tedious, and requires some persistence, but not extraordinary insight. After doing all this, and then simplifying as much as possible, what you discover is that you end up with exactly the backpropagation algorithm! And so you can think of the backpropagation algorithm as providing a way of computing the sum over the rate factor for all these paths. Or, to put it slightly differently, the backpropagation algorithm is a clever way of keeping track of small perturbations to the weights (and biases) as they propagate through the network, reach the output, and then affect the cost.

Now, I'm not going to work through all this here. It's messy and requires considerable care to work through all the details. If you're up for a challenge, you may enjoy attempting it. And even if not, I hope this line of thinking gives you some insight into what backpropagation is accomplishing.

What about the other mystery -- how backpropagation could have been discovered in the first place? In fact, if you follow the approach I just sketched you will discover a proof of backpropagation. Unfortunately, the proof is quite a bit longer and more complicated than the one I described earlier in this chapter. So how was that short (but more mysterious) proof discovered? What you find when you write out all the details of the long proof is that, after the fact, there are several obvious simplifications staring you in the face. You make those simplifications, get a shorter proof, and write that out. And then several more obvious simplifications jump out at you. So you repeat again. The result after a few iterations is the proof we saw earlier\footnote{There is one clever step required. In Equation~(\ref{eq:53}) the intermediate variables are activations like $a^{l+1}_q$. The clever idea is to switch to using weighted inputs, like $z^{l+1}_q$, as the intermediate variables. If you don't have this idea, and instead continue using the activations $a^{l+1}_q$, the proof you obtain turns out to be slightly more complex than the proof given earlier in the chapter.} -- short, but somewhat obscure, because all the signposts to its construction have been removed! I am, of course, asking you to trust me on this, but there really is no great mystery to the origin of the earlier proof. It's just a lot of hard work simplifying the proof I've sketched in this section.
\end{document}

### 2019-04-24 Bibliography/Machine Learning/neuralnetworksanddeeplearning.com.pdf-master/neuralnetworksanddeeplearning.com.pdf-master/ch4_visualproof.tex
Ameliorate bibliography

\documentclass[a4paper,twoside,10pt]{book}
\include{header}
\begin{document}
\chapter{A visual proof that neural nets can compute any function}
One of the most striking facts about neural networks is that they can compute any function at all. That is, suppose someone hands you some complicated, wiggly function, $f(x)$:
\begin{center} 
	\begin{tikzpicture}
	\begin{axis}[title={$f(x)$},axis lines=middle,width=0.75\linewidth,height=0.5\linewidth,ymin=-0.05,ymax=1,scaled y ticks = false,y tick label style={/pgf/number format/fixed},xmin=-0.05,xmax=1.05,ticks=none]
	\addplot[blue,domain=0:1,samples=101]{0.2 + 0.4*x*x + 0.3*x*sin(15*deg(x)) + 0.05*cos(50*deg(x)) };
	\end{axis}
	\end{tikzpicture}
\end{center}
\label{basic_network_precursor}%
No matter what the function, there is guaranteed to be a neural network so that for every possible input, $x$, the value $f(x)$ (or some close approximation) is output from the network, e.g.:
\begin{center}
	\includegraphics[width=0.43\linewidth]{figures/ch4/tikz400}
\end{center}
This result holds even if the function has many inputs, $f=f(x_1,\ldots,x_m)$, and many outputs. For instance, here's a network computing a function with $m=3$ inputs and $n=2$ outputs:
\begin{center}
	\includegraphics[width=0.45\linewidth]{figures/ch4/tikz401}
\end{center}
This result tells us that neural networks have a kind of \textit{universality}. No matter what function we want to compute, we know that there is a neural network which can do the job.

What's more, this universality theorem holds even if we restrict our networks to have just a single layer intermediate between the input and the output neurons -- a so-called single hidden layer. So even very simple network architectures can be extremely powerful.

The universality theorem is well known by people who use neural networks. But why it's true is not so widely understood. Most of the explanations available are quite technical. For instance, one of the original papers proving the result\footnote{\href{http://www.dartmouth.edu/~gvc/Cybenko_MCSS.pdf}{Approximation by superpositions of a sigmoidal function}, by George Cybenko (1989). The result was very much in the air at the time, and several groups proved closely related results. Cybenko's paper contains a useful discussion of much of that work. Another important early paper is \href{http://www.sciencedirect.com/science/article/pii/0893608089900208}{Multilayer feedforward networks are universal approximators}, by Kurt Hornik, Maxwell Stinchcombe, and Halbert White (1989). This paper uses the Stone-Weierstrass theorem to arrive at similar results.} did so using the Hahn-Banach theorem, the Riesz Representation theorem, and some Fourier analysis. If you're a mathematician the argument is not difficult to follow, but it's not so easy for most people. That's a pity, since the underlying reasons for universality are simple and beautiful.

In this chapter I give a simple and mostly visual explanation of the universality theorem. We'll go step by step through the underlying ideas. You'll understand why it's true that neural networks can compute any function. You'll understand some of the limitations of the result. And you'll understand how the result relates to deep neural networks.

To follow the material in the chapter, you do not need to have read earlier chapters in this book. Instead, the chapter is structured to be enjoyable as a self-contained essay. Provided you have just a little basic familiarity with neural networks, you should be able to follow the explanation. I will, however, provide occasional links to earlier material, to help fill in any gaps in your knowledge.

Universality theorems are a commonplace in computer science, so much so that we sometimes forget how astonishing they are. But it's worth reminding ourselves: the ability to compute an arbitrary function is truly remarkable. Almost any process you can imagine can be thought of as function computation. Consider the problem of naming a piece of music based on a short sample of the piece. That can be thought of as computing a function. Or consider the problem of translating a Chinese text into English. Again, that can be thought of as computing a function\footnote{Actually, computing one of many functions, since there are often many acceptable translations of a given piece of text.}. Or consider the problem of taking an mp4 movie file and generating a description of the plot of the movie, and a discussion of the quality of the acting. Again, that can be thought of as a kind of function computation\footnote{Ditto the remark about translation and there being many possible functions..} Universality means that, in principle, neural networks can do all these things and many more.

Of course, just because we know a neural network exists that can (say) translate Chinese text into English, that doesn't mean we have good techniques for constructing or even recognizing such a network. This limitation applies also to traditional universality theorems for models such as Boolean circuits. But, as we've seen earlier in the book, neural networks have powerful algorithms for learning functions. That combination of learning algorithms + universality is an attractive mix. Up to now, the book has focused on the learning algorithms. In this chapter, we focus on universality, and what it means.

\section{Two caveats}
Before explaining why the universality theorem is true, I want to mention two caveats to the informal statement ``a neural network can compute any function''.

First, this doesn't mean that a network can be used to \textit{exactly} compute any function. Rather, we can get an \textit{approximation} that is as good as we want. By increasing the number of hidden neurons we can improve the approximation. For instance, earlier (see \ref{basic_network_precursor}) I illustrated a network computing some function $f(x)$ using three hidden neurons. For most functions only a low-quality approximation will be possible using three hidden neurons. By increasing the number of hidden neurons (say, to five) we can typically get a better approximation:
\begin{center}
	\includegraphics[width=0.43\linewidth]{figures/ch4/tikz402}
\end{center}
And we can do still better by further increasing the number of hidden neurons.

To make this statement more precise, suppose we're given a function $f(x)$ which we'd like to compute to within some desired accuracy $\epsilon>0$. The guarantee is that by using enough hidden neurons we can always find a neural network whose output $g(x)$ satisfies $|g(x)-f(x)|<\epsilon$, for all inputs $x$. In other words, the approximation will be good to within the desired accuracy for every possible input.

The second caveat is that the class of functions which can be approximated in the way described are the \textit{continuous} functions. If a function is discontinuous, i.e., makes sudden, sharp jumps, then it won't in general be possible to approximate using a neural net. This is not surprising, since our neural networks compute continuous functions of their input. However, even if the function we'd really like to compute is discontinuous, it's often the case that a continuous approximation is good enough. If that's so, then we can use a neural network. In practice, this is not usually an important limitation.

Summing up, a more precise statement of the universality theorem is that neural networks with a single hidden layer can be used to approximate any continuous function to any desired precision. In this chapter we'll actually prove a slightly weaker version of this result, using two hidden layers instead of one. In the problems I'll briefly outline how the explanation can, with a few tweaks, be adapted to give a proof which uses only a single hidden layer.

\section{Universality with one input and one output}
To understand why the universality theorem is true, let's start by understanding how to construct a neural network which approximates a function with just one input and one output:
\begin{center} 
	\begin{tikzpicture}
	\begin{axis}[title={$f(x)$},axis lines=middle,width=0.75\linewidth,height=0.5\linewidth,ymin=-0.05,ymax=1,scaled y ticks = false,y tick label style={/pgf/number format/fixed},xmin=-0.05,xmax=1.05,ticks=none]
	\addplot[blue,domain=0:1,samples=101]{0.2 + 0.4*x*x + 0.3*x*sin(15*deg(x)) + 0.05*cos(50*deg(x)) };
	\end{axis}
	\end{tikzpicture}
\end{center}
It turns out that this is the core of the problem of universality. Once we've understood this special case it's actually pretty easy to extend to functions with many inputs and many outputs.

To build insight into how to construct a network to compute $f$, let's start with a network containing just a single hidden layer, with two hidden neurons, and an output layer containing a single output neuron:
\begin{center}
	\includegraphics[width=0.45\linewidth]{figures/ch4/tikz403x}
\end{center}
To get a feel for how components in the network work, let's focus on the top hidden neuron. In the diagram below, click on the weight, $w$, and drag the mouse a little ways to the right to increase $w$. You can immediately see how the function computed by the top hidden neuron changes:
\begin{center}
\begin{tikzpicture}[declare function = {sigma(\x) = 1/(1+exp(-\x)); f(\x,\w,\b) = sigma(\w*\x+\b);}]
\pgfplotsset{every axis legend/.append style={at={(1.03,1)},anchor=north west}}

\begin{axis}[width=0.75\linewidth,height=0.45\linewidth,title={Output from neuron},axis x line=bottom,axis y line=left,xtick={0,1},ytick={0,1},xmax=1.05,ymax=1.05]
\addplot[blue!10!red,domain=0:1,samples=101]{f(x,7,-2)};
\addplot[blue!25!red,domain=0:1,samples=101]{f(x,7,-3)};
\addplot[blue!50!red,domain=0:1,samples=101]{f(x,7,-4)};
\addplot[blue!75!red,domain=0:1,samples=101]{f(x,7,-5)};
\addplot[blue!90!red,domain=0:1,samples=101]{f(x,7,-6)};
\addlegendentry{$w=7, b = -2$}
\addlegendentry{$w=7, b = -3$}
\addlegendentry{$w=7, b = -4$}
\addlegendentry{$w=7, b = -5$}
\addlegendentry{$w=7, b = -6$}
\end{axis}
\end{tikzpicture}
\end{center}
\begin{center}
	\begin{tikzpicture}[declare function = {sigma(\x) = 1/(1+exp(-\x)); f(\x,\w,\b) = sigma(\w*\x+\b);}]
	\pgfplotsset{every axis legend/.append style={at={(1.03,1)},anchor=north west}}

	\begin{axis}[width=0.75\linewidth,height=0.45\linewidth,title={Output from neuron},axis x line=bottom,axis y line=left,xtick={0,1},ytick={0,1},xmax=1.05,ymax=1.05]
	\addplot[blue!90!red,domain=0:1,samples=101]{f(x,9,-4)};
	\addplot[blue!75!red,domain=0:1,samples=101]{f(x,8,-4)};
	\addplot[blue!50!red,domain=0:1,samples=101]{f(x,7,-4)};
	\addplot[blue!25!red,domain=0:1,samples=101]{f(x,6,-4)};
	\addplot[blue!10!red,domain=0:1,samples=101]{f(x,5,-4)};
	\addlegendentry{$w=9, b = -4$}
	\addlegendentry{$w=8, b = -4$}
	\addlegendentry{$w=7, b = -4$}
	\addlegendentry{$w=6, b = -4$}
	\addlegendentry{$w=5, b = -4$}
	\end{axis}
	\end{tikzpicture}
\end{center}
As we learnt earlier in the book, what's being computed by the hidden neuron is $\sigma(wx+b)$, where $\sigma(z)\equiv1/(1+e^{-z})$ is the sigmoid function. Up to now, we've made frequent use of this algebraic form. But for the proof of universality we will obtain more insight by ignoring the algebra entirely, and instead manipulating and observing the shape shown in the graph. This won't just give us a better feel for what's going on, it will also give us a proof\footnote{Strictly speaking, the visual approach I'm taking isn't what's traditionally thought of as a proof. But I believe the visual approach gives more insight into why the result is true than a traditional proof. And, of course, that kind of insight is the real purpose behind a proof. Occasionally, there will be small gaps in the reasoning I present: places where I make a visual argument that is plausible, but not quite rigorous. If this bothers you, then consider it a challenge to fill in the missing steps. But don't lose sight of the real purpose: to understand why the universality theorem is true.} of universality that applies to activation functions other than the sigmoid function.
%
%To get started on this proof, try clicking on the bias, $b$, in the diagram above, and dragging to the right to increase it. You'll see that as the bias increases the graph moves to the left, but its shape doesn't change.
%
%Next, click and drag to the left in order to decrease the bias. You'll see that as the bias decreases the graph moves to the right, but, again, its shape doesn't change.
%
%Next, decrease the weight to around 2 or 3. You'll see that as you decrease the weight, the curve broadens out. You might need to change the bias as well, in order to keep the curve in-frame.
%
%Finally, increase the weight up past $w=100$. As you do, the curve gets steeper, until eventually it begins to look like a step function. Try to adjust the bias so the step occurs near $x=0.3$. The following short clip shows what your result should look like. Click on the play button to play (or replay) the video:
%\begin{lstlisting}
%ANIMATION
%\end{lstlisting}
We can simplify our analysis quite a bit by increasing the weight so much that the output really is a step function, to a very good approximation. Below I've plotted the output from the top hidden neuron when the weight is $w=999$. %Note that this plot is static, and you can't change parameters such as the weight.
%\begin{center}
%	\includegraphics[width=0.7\linewidth]{figures/ch4/high_weight_function}
%\end{center}
\begin{center}
	\includegraphics[width=0.35\linewidth]{figures/ch4/tikz403y}
	\begin{tikzpicture}[declare function = {sigma(\x,\w,\b) = 1/(1+exp(-\w*\x-\b));}]
		\begin{axis}[width=0.4\linewidth,height=0.4\linewidth,align=center,title={Output from top hidden neuron},axis x line=middle,axis y line=middle,xtick={0,1},ytick={0,1},xmax=1.15,ymax=1.15,ymin=-0.15,xlabel={$x$}]
		\addplot[blue!90!red,domain=0:1,samples=201]{sigma(x,999,-400)};
		\end{axis}
	\end{tikzpicture}
\end{center}
It's actually quite a bit easier to work with step functions than general sigmoid functions. The reason is that in the output layer we add up contributions from all the hidden neurons. It's easy to analyze the sum of a bunch of step functions, but rather more difficult to reason about what happens when you add up a bunch of sigmoid shaped curves. And so it makes things much easier to assume that our hidden neurons are outputting step functions. More concretely, we do this by fixing the weight $w$ to be some very large value, and then setting the position of the step by modifying the bias. Of course, treating the output as a step function is an approximation, but it's a very good approximation, and for now we'll treat it as exact. I'll come back later to discuss the impact of deviations from this approximation.

At what value of $x$ does the step occur? Put another way, how does the position of the step depend upon the weight and bias?

To answer this question, try modifying the weight and bias in the diagram above (you may need to scroll back a bit). Can you figure out how the position of the step depends on $w$ and $b$? With a little work you should be able to convince yourself that the position of the step is \textit{proportional} to $b$, and \textit{inversely proportional} to $w$.

In fact, the step is at position $s=-b/w$, as you can see by modifying the weight and bias in the following diagram:
\begin{center}
	\includegraphics[width=0.35\linewidth]{figures/ch4/tikz404x}
	\begin{tikzpicture}[declare function = {sigma(\x,\w,\b) = 1/(1+exp(-\w*\x-\b));}]
		\begin{axis}[width=0.4\linewidth,height=0.4\linewidth,align=center,title={Output from top hidden neuron},axis x line=middle,axis y line=middle,xtick={0,0.4,1},xticklabels={0,$-b/w=0.4$,1},ytick={0,1},xmax=1.15,ymax=1.15,ymin=-0.15,xlabel={$x$}]
		\addplot[blue!90!red,domain=0:1,samples=201]{sigma(x,100,-40)};
		\end{axis}
	\end{tikzpicture}
\end{center}
It will greatly simplify our lives to describe hidden neurons using just a single parameter, $s$, which is the step position, $s=-b/w$. Try modifying $s$ in the following diagram, in order to get used to the new parameterization:
\begin{center}
	\includegraphics[width=0.35\linewidth]{figures/ch4/tikz405}
	\begin{tikzpicture}[declare function = {step(\x,\s)=0.5*(sign(\x-\s)+1);}]
		\begin{axis}[width=0.4\linewidth,height=0.4\linewidth,align=center,title={Output from top hidden neuron},axis x line=middle,axis y line=middle,xtick={0,0.4,1},xticklabels={0,$s$,1},ytick={0,1},xmax=1.15,ymax=1.15,ymin=-0.15,xlabel={$x$}]
		\addplot[blue!90!red,domain=0:1,samples=201]{step(x,0.4)};
		\end{axis}
	\end{tikzpicture}
\end{center}
%\begin{lstlisting}
%ANIMATIOIN
%\end{lstlisting}
As noted above, we've implicitly set the weight $w$ on the input to be some large value -- big enough that the step function is a very good approximation. We can easily convert a neuron parameterized in this way back into the conventional model, by choosing the bias $b=-ws$.

Up to now we've been focusing on the output from just the top hidden neuron. Let's take a look at the behavior of the entire network. In particular, we'll suppose the hidden neurons are computing step functions parameterized by step points $s_1$ (top neuron) and $s_2$ (bottom neuron). And they'll have respective output weights $w_1$ and $w_2$. Here's the network:
\begin{center}
	\includegraphics[width=0.35\linewidth]{figures/ch4/tikz407}
	\begin{tikzpicture}[declare function = {step(\x,\s,\h)=0.5*\h*(sign(\x-\s)+1);}]
		\begin{axis}[width=0.4\linewidth,height=0.4\linewidth,align=center,title={Weighted output from hidden layer},axis x line=middle,axis y line=middle,xtick={0,0.4,0.6,1},xticklabels={0,$s_1$,$s_2$,1},ytick={0,0.6,1,1.8},yticklabels={0,$w_1$,1,$w_1+w_2$},xmax=1.15,ymax=2.15,ymin=-0.15,xlabel={$x$}]
		\addplot[blue!90!red,domain=0:1,samples=201]{step(x,0.4,0.6)+step(x,0.6,1.2)};
		\end{axis}
	\end{tikzpicture}
\end{center}
%\begin{lstlisting}
%ANIMATIOIN
%\end{lstlisting}
What's being plotted on the right is the weighted output $w_1a_1+w_2a_2$ from the hidden layer. Here, $a_1$ and $a_2$ are the outputs from the top and bottom hidden neurons, respectively\footnote{Note, by the way, that the output from the whole network is $\sigma(w_1a_1+w_2a_2+b)$, where $b$ is the bias on the output neuron. Obviously, this isn't the same as the weighted output from the hidden layer, which is what we're plotting here. We're going to focus on the weighted output from the hidden layer right now, and only later will we think about how that relates to the output from the whole network.}. These outputs are denoted with as because they're often known as the neurons' activations.

Try increasing and decreasing the step point $s_1$ of the top hidden neuron. Get a feel for how this changes the weighted output from the hidden layer. It's particularly worth understanding what happens when $s_1$ goes past $s_2$. You'll see that the graph changes shape when this happens, since we have moved from a situation where the top hidden neuron is the first to be activated to a situation where the bottom hidden neuron is the first to be activated.

Similarly, try manipulating the step point $s_2$ of the bottom hidden neuron, and get a feel for how this changes the combined output from the hidden neurons.

Try increasing and decreasing each of the output weights. Notice how this rescales the contribution from the respective hidden neurons. What happens when one of the weights is zero?

Finally, try setting $w_1$ to be 0.8 and $w_2$ to be -0.8. You get a ``bump'' function, which starts at point $s_1$, ends at point $s_2$, and has height 0.8. For instance, the weighted output might look like this:
\begin{center}
	\includegraphics[width=0.35\linewidth]{figures/ch4/tikz408}
	\begin{tikzpicture}[declare function = {step(\x,\s,\h)=0.5*\h*(sign(\x-\s)+1);}]
		\begin{axis}[width=0.4\linewidth,height=0.4\linewidth,align=center,title={Weighted output from hidden layer},axis x line=middle,axis y line=middle,xtick={0,0.4,0.6,1},xticklabels={0,$s_1$,$s_2$,1},ytick={0,0.8,1},yticklabels={0,$w_1=-w_2$,1},xmax=1.15,ymax=2.15,ymin=-0.15,xlabel={$x$}]
		\addplot[blue!90!red,domain=0:1,samples=201]{step(x,0.4,0.8)+step(x,0.6,-0.8)};
		\end{axis}
	\end{tikzpicture}
\end{center}
Of course, we can rescale the bump to have any height at all. Let's use a single parameter, $h$, to denote the height. To reduce clutter I'll also remove the ``$s_1=\ldots$'' and ``$w_1=\ldots$'' notations.
\begin{center}
	\includegraphics[width=0.35\linewidth]{figures/ch4/tikz406}
	\begin{tikzpicture}[declare function = {bump(\x,\s,\t,\h)=0.5*\h*(sign(\x-\s)-sign(\x-\t));}]
	\begin{axis}[width=0.4\linewidth,height=0.4\linewidth,align=center,title={Weighted output from hidden layer},axis x line=middle,axis y line=middle,xtick={0,0.4,0.6,1},xticklabels={0,$s_1$,$s_2$,1},ytick={-1,0,0.6,1},yticklabels={$-1$,0,$h$,1},xmax=1.25,ymax=1.25,ymin=-1.25,xlabel={$x$}]
	\addplot[blue!90!red,domain=0:1,samples=201]{bump(x,0.4,0.6,0.6};
	\end{axis}
	\end{tikzpicture}
\end{center}
Try changing the value of $h$ up and down, to see how the height of the bump changes. Try changing the height so it's negative, and observe what happens. And try changing the step points to see how that changes the shape of the bump.

You'll notice, by the way, that we're using our neurons in a way that can be thought of not just in graphical terms, but in more conventional programming terms, as a kind of \texttt{if-then-else} statement, e.g.:
\begin{lstlisting}
if input >= step point:
	add 1 to the weighted output
else:
	add 0 to the weighted output
\end{lstlisting}
For the most part I'm going to stick with the graphical point of view. But in what follows you may sometimes find it helpful to switch points of view, and think about things in terms of \texttt{if-then-else}.

We can use our bump-making trick to get two bumps, by gluing two pairs of hidden neurons together into the same network:
\begin{center}
	\includegraphics[width=0.35\linewidth]{figures/ch4/tikz421}
	\begin{tikzpicture}[declare function = {step(\x,\s,\t)=0.5*(sign(\x-\s)-sign(\x-\t));twobumps(\x,\s,\t,\h,\ss,\tt,\hh) = \h*step(\x,\s,\t)+\hh*step(\x,\ss,\tt); }]
	\begin{axis}[width=0.4\linewidth,height=0.4\linewidth,align=center,title={Weighted output from hidden layer},axis x line=bottom,axis y line=left,xtick={0,0.4,0.6,0.7,0.9,1},xticklabels={0,$s_1^1$,$s_2^1$,$s_1^2$,$s_2^2$,1},ytick={-1.2,-1,0,0.3,1},yticklabels={$h_1$,$-1$,0,$h_2$,1},xmax=1.15,ymax=1.25,ymin=-1.25]
	\addplot[blue!90!red,domain=0:1,samples=201]{twobumps(x,0.4,0.6,-1.2,0.7,0.9,0.3};
	\end{axis}
	\end{tikzpicture}
\end{center}
I've suppressed the weights here, simply writing the $h$ values for each pair of hidden neurons. Try increasing and decreasing both $h$ values, and observe how it changes the graph. Move the bumps around by changing the step points.

More generally, we can use this idea to get as many peaks as we want, of any height. In particular, we can divide the interval [0,1] up into a large number, $N$, of subintervals, and use $N$ pairs of hidden neurons to set up peaks of any desired height. Let's see how this works for $N=5$. That's quite a few neurons, so I'm going to pack things in a bit. Apologies for the complexity of the diagram: I could hide the complexity by abstracting away further, but I think it's worth putting up with a little complexity, for the sake of getting a more concrete feel for how these networks work.
\begin{center}
\includegraphics[height=0.65\linewidth]{./figures/ch4/tikz438}
\begin{tikzpicture}[declare function = {step(\x,\s,\t)=0.5*(sign(\x-\s)-sign(\x-\t));}]
\begin{axis}[width=0.42\linewidth,height=0.65\linewidth,align=center,title={Weighted output},axis x line=middle,axis y line=left,xtick={0,1},ytick={-1,0,1},xmax=1.15,ymax=2.2,ymin=-1.5]
	\addplot[blue!90!red,domain=0:1.05,samples=201]{step(x,0,0.2)};
	\addplot[blue!90!red,domain=0:1.05,samples=201]{-0.5*step(x,0.2,0.4)};
	\addplot[blue!90!red,domain=0:1.05,samples=201]{+1.5*step(x,0.4,0.6)};
	\addplot[blue!90!red,domain=0:1.05,samples=201]{+0.6*step(x,0.6,0.8)};
	\addplot[blue!90!red,domain=0:1.05,samples=201]{-0.2*step(x,0.8,1.0)};
	\end{axis}
\end{tikzpicture}	
\end{center}
You can see that there are five pairs of hidden neurons. The step points for the respective pairs of neurons are $0,1/5$, then $1/5,2/5$, and so on, out to $4/5,5/5$. These values are fixed -- they make it so we get five evenly spaced bumps on the graph.

Each pair of neurons has a value of $h$ associated to it. Remember, the connections output from the neurons have weights $h$ and $-h$ (not marked). Click on one of the $h$ values, and drag the mouse to the right or left to change the value. As you do so, watch the function change. By changing the output weights we're actually \textit{designing} the function!

Contrariwise, try clicking on the graph, and dragging up or down to change the height of any of the bump functions. As you change the heights, you can see the corresponding change in $h$ values. And, although it's not shown, there is also a change in the corresponding output weights, which are $+h$ and $-h$.

In other words, we can directly manipulate the function appearing in the graph on the right, and see that reflected in the $h$ values on the left. A fun thing to do is to hold the mouse button down and drag the mouse from one side of the graph to the other. As you do this you draw out a function, and get to watch the parameters in the neural network adapt.

Time for a challenge.

Let's think back to the function I plotted at the beginning of the chapter:
\begin{center} 
	\begin{tikzpicture}
	\begin{axis}[title={$f(x)$},axis lines=middle,width=0.6\linewidth,height=0.5\linewidth,ymin=-0.05,ymax=1,scaled y ticks = false,y tick label style={/pgf/number format/fixed},xmin=-0.05,xmax=1.05,ticks=none]
	\addplot[blue,domain=0:1,samples=101]{0.2 + 0.4*x*x + 0.3*x*sin(15*deg(x)) + 0.05*cos(50*deg(x)) };
	\end{axis}
	\end{tikzpicture}
\end{center}
I didn't say it at the time, but what I plotted is actually the function
\begin{equation}
f(x)=0.2+0.4x^2+0.3x\sin(15x)+0.05\cos(50x),\tag{113}\label{eq:113}
\end{equation}
plotted over $x$ from 0 to 1, and with the $y$ axis taking values from 0 to 1.

That's obviously not a trivial function.

You're going to figure out how to compute it using a neural network.

In our networks above we've been analyzing the weighted combination $\sum_jw_ja_j$ output from the hidden neurons. We now know how to get a lot of control over this quantity. But, as I noted earlier, this quantity is not what's output from the network. What's output from the network is $\sigma\left(\sum_jw_ja_j+b\right)$ where $b$ is the bias on the output neuron. Is there some way we can achieve control over the actual output from the network?

The solution is to design a neural network whose hidden layer has a weighted output given by $\sigma^{-1} \circ f(x)$, where $\sigma^{-1}$ is just the inverse of the $\sigma$ function. That is, we want the weighted output from the hidden layer to be:
\begin{center} 
	\begin{tikzpicture}
	\begin{axis}[axis lines=middle,width=0.6\linewidth,height=0.5\linewidth,ymin=-2.05,ymax=2.25,scaled y ticks = false,y tick label style={/pgf/number format/fixed},xmin=-0.05,xmax=1.05, xtick={0,1}]
	\addplot[blue,domain=0:1,samples=101]{4*(-0.3 + 0.4*x*x + 0.3*x*sin(15*deg(x)) + 0.05*cos(50*deg(x))) };
	\node[anchor=north,align=center] at (axis cs: 0.5,2) {$\sigma^{-1}\circ f(x)$};
	\end{axis}
	\end{tikzpicture}
\end{center}
If we can do this, then the output from the network as a whole will be a good approximation to $f(x)$\footnote{Note that I have set the bias on the output neuron to 0.}.

Your challenge, then, is to design a neural network to approximate the goal function shown just above. To learn as much as possible, I want you to solve the problem twice. The first time, please click on the graph, directly adjusting the heights of the different bump functions. You should find it fairly easy to get a good match to the goal function. How well you're doing is measured by the average deviation between the goal function and the function the network is actually computing. Your challenge is to drive the average deviation as low as possible. You complete the challenge when you drive the average deviation to 0.40 or below\footnote{This paragraph refers to interactive element, available online. The graph shows the final result of manual minimization of average deviation. }

%Once you've done that, click on ``Reset'' to randomly re-initialize the bumps. The second time you solve the problem, resist the urge to click on the graph. Instead, modify the h values on the left-hand side, and again attempt to drive the average deviation to 0.40 or below.
\begin{center} 
\includegraphics[height=0.65\textwidth]{./figures/ch4/tikz439}
\begin{tikzpicture}[declare function = {step(\x,\s,\t)=0.5*(sign(\x-\s)-sign(\x-\t));}]
	\begin{axis}[axis lines=middle,width=0.6\textwidth,height=0.65\textwidth,ymin=-2.05,ymax=2.25,scaled y ticks = false,y tick label style={/pgf/number format/fixed},xmin=-0.05,xmax=1.05, xtick={0,1}]
		\addplot[blue,domain=0:1,samples=101]{4*(-0.3 + 0.4*x*x + 0.3*x*sin(15*deg(x)) + 0.05*cos(50*deg(x))) };
		\addplot[red!50!black,domain=-0.05:0.35,samples=100]{-1.1*step(x,0,0.2)};
		\addplot[red!50!black,domain=0:0.6,samples=100]{-1.4*step(x,0.2,0.4)};
		\addplot[red!50!black,domain=0.2:0.8,samples=100]{-0.3*step(x,0.4,0.6)};
		\addplot[red!50!black,domain=0.4:1,samples=100]{-1*step(x,0.6,0.8)};
		\addplot[red!50!black,domain=0.6:1.05,samples=100]{+1.2*step(x,0.8,1.0)};
		\node[anchor=north,align=center] at (axis cs: 0.5,2) {Weighted output\\$\sigma = 0.38$};
	\end{axis}
\end{tikzpicture}
\end{center}
You've now figured out all the elements necessary for the network to approximately compute the function $f(x)$! It's only a coarse approximation, but we could easily do much better, merely by increasing the number of pairs of hidden neurons, allowing more bumps.

In particular, it's easy to convert all the data we have found back into the standard parametrization used for neural networks. Let me just recap quickly how that works.

The first layer of weights all have some large, constant value, say $w=1000$.

The biases on the hidden neurons are just $b=-ws$. So, for instance, for the second hidden neuron $s=0.2$ becomes $b=-1000\times0.2=-200$.

The final layer of weights are determined by the $h$ values. So, for instance, the value you've chosen above for the first $h$, $h= -0.6$, means that the output weights from the top two hidden neurons are $-0.6$ and 0.6, respectively. And so on, for the entire layer of output weights.

Finally, the bias on the output neuron is 0.

That's everything: we now have a complete description of a neural network which does a pretty good job computing our original goal function. And we understand how to improve the quality of the approximation by improving the number of hidden neurons.

What's more, there was nothing special about our original goal function, $f(x)=0.2+0.4x^2+0.3\sin(15x)+0.05\cos(50x)$. We could have used this procedure for any continuous function from $[0,1]$ to $[0,1]$. In essence, we're using our single-layer neural networks to build a lookup table for the function. And we'll be able to build on this idea to provide a general proof of universality.

\section{Many input variables}
Let's extend our results to the case of many input variables. This sounds complicated, but all the ideas we need can be understood in the case of just two inputs. So let's address the two-input case.

We'll start by considering what happens when we have two inputs to a neuron:
\begin{center}
	\includegraphics[width=0.4\linewidth]{figures/ch4/tikz411}
\end{center}
Here, we have inputs $x$ and $y$, with corresponding weights $w_1$ and $w_2$, and a bias $b$ on the neuron. Let's set the weight $w_2$ to 0, and then play around with the first weight, $w_1$, and the bias, $b$, to see how they affect the output from the neuron:
\begin{center}
	\begin{tikzpicture}[scale=0.8]
		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)}, grid=major,width=5cm,height=6cm,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$w_1=8,b=-8$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-(8*x-8)))};
			\end{axis}
	\end{tikzpicture}
	\begin{tikzpicture}[scale=0.8]
		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=6cm,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$w_1=8,b=-5$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-(8*x-5)))};
		\end{axis}
	\end{tikzpicture}
	\begin{tikzpicture}[scale=0.8]
		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=6cm,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$w_1=8,b=-2$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-(8*x-2)))};
		\end{axis}
	\end{tikzpicture}
\end{center}

\begin{center}
	\begin{tikzpicture}[scale=0.8]
		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=6cm,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$w_1=11,b=-5$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-(11*x-5)))};
			\end{axis}
	\end{tikzpicture}
	\begin{tikzpicture}[scale=0.8]
		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=6cm,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$w_1=8,b=-5$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-(8*x-5)))};
		\end{axis}
	\end{tikzpicture}
	\begin{tikzpicture}[scale=0.8]
		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=6cm,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$w_1=5,b=-5$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-(5*x-5)))};
		\end{axis}
	\end{tikzpicture}
\end{center}
%
%x=1y=1Output
%
As you can see, with $w_2=0$ the input $y$ makes no difference to the output from the neuron. It's as though $x$ is the only input.

Given this, what do you think happens when we increase the weight $w_1$ to $w_1=100$, with $w_2$ remaining 0? If you don't immediately see the answer, ponder the question for a bit, and see if you can figure out what happens. Then try it out and see if you're right. I've shown what happens in the following movie:

Just as in our earlier discussion, as the input weight gets larger the output approaches a step function. The difference is that now the step function is in three dimensions. Also as before, we can move the location of the step point around by modifying the bias. The actual location of the step point is $s_x\equiv-b/w_1$.

Let's redo the above using the position of the step as the parameter:
\begin{center}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/stepx1}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/stepx2}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/stepx3}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$s_x=0.25$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-1000*(x-0.25)))};
%			\end{axis}
%	\end{tikzpicture}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s_x=0.5$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-1000*(x-0.5)))};
%		\end{axis}
%	\end{tikzpicture}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s_x=0.7$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-1000*(x-0.7)))};
%		\end{axis}
%	\end{tikzpicture}
\end{center}
Here, we assume the weight on the $x$ input has some large value -- I've used $w_1=1000$ -- and the weight $w_2=0$. The number on the neuron is the step point, and the little $x$ above the number reminds us that the step is in the $x$ direction. Of course, it's also possible to get a step function in the $y$ direction, by making the weight on the $y$ input very large (say, $w_2=1000$), and the weight on the $x$ equal to 0, i.e., $w_1=0$:
\begin{center}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$s_y=0.25$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-1000*(y-0.25)))};
%			\end{axis}
%	\end{tikzpicture}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s_y=0.5$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-1000*(y-0.5)))};
%		\end{axis}
%	\end{tikzpicture}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s_y=0.7$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-1000*(y-0.7)))};
%		\end{axis}
%	\end{tikzpicture}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/stepy1}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/stepy2}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/stepy3}
\end{center}
The number on the neuron is again the step point, and in this case the little $y$ above the number reminds us that the step is in the $y$ direction. I could have explicitly marked the weights on the $x$ and $y$ inputs, but decided not to, since it would make the diagram rather cluttered. But do keep in mind that the little $y$ marker implicitly tells us that the $y$ weight is large, and the $x$ weight is 0.

We can use the step functions we've just constructed to compute a three-dimensional bump function. To do this, we use two neurons, each computing a step function in the $x$ direction. Then we combine those step functions with weight $h$ and $-h$, respectively, where $h$ is the desired height of the bump. It's all illustrated in the following diagram:
\begin{center}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$s^1_x=0.25,s^2_x=0.75, h=1$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-100*(x-0.25)))-1/(1+exp(-100*(x-0.75)))};
%			\end{axis}
%	\end{tikzpicture}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s^1_x=0.25,s^2_x=0.75, h=0.75$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{0.75/(1+exp(-100*(x-0.25)))-0.75/(1+exp(-100*(x-0.75)))};
%		\end{axis}
%	\end{tikzpicture}
%%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s^1_x=0.5,s^2_x=0.75, h=1$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-100*(x-0.5)))-1/(1+exp(-100*(x-0.75)))};
%		\end{axis}
%	\end{tikzpicture}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/bumpx1}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/bumpx2}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/bumpx3}
\end{center}
%
%x=1y=1Weighted output from hidden layer
%
Try changing the value of the height, h. Observe how it relates to the weights in the network. And see how it changes the height of the bump function on the right.

Also, try changing the step point 0.30 associated to the top hidden neuron. Witness how it changes the shape of the bump. What happens when you move it past the step point 0.70 associated to the bottom hidden neuron?

We've figured out how to make a bump function in the $x$ direction. Of course, we can easily make a bump function in the $y$ direction, by using two step functions in the $y$ direction. Recall that we do this by making the weight large on the $y$ input, and the weight 0 on the $x$ input. Here's the result:
\begin{center}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/bumpy1}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/bumpy2}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/bumpy3}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$s^1_y=0.25,s^2_y=0.75, h=1$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{1/(1+exp(-100*(y-0.25)))-1/(1+exp(-100*(y-0.75)))};
%			\end{axis}
%	\end{tikzpicture}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s^1_y=0.25,s^2_y=0.75, h=0.75$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{0.75/(1+exp(-100*(y-0.25)))-0.75/(1+exp(-100*(y-0.75)))};
%		\end{axis}
%	\end{tikzpicture}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s^1_y=0.25,s^2_y=0.45, h=0.8$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{0.8/(1+exp(-100*(y-0.25)))-0.8/(1+exp(-100*(y-0.45)))};
%		\end{axis}
%	\end{tikzpicture}
\end{center}
%
%x=1y=1Weighted output from hidden layer
%
This looks nearly identical to the earlier network! The only thing explicitly shown as changing is that there's now little $y$ markers on our hidden neurons. That reminds us that they're producing $y$ step functions, not $x$ step functions, and so the weight is very large on the $y$ input, and zero on the $x$ input, not vice versa. As before, I decided not to show this explicitly, in order to avoid clutter.
%
Let's consider what happens when we add up two bump functions, one in the $x$ direction, the other in the $y$ direction, both of height $h$:
\begin{center}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={$s^1_x=0.4,s^2_x=0.6,s^1_y=0.3,s^2_y=0.7, h=1$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{0.5/(1+exp(-100*(y-0.3)))-0.5/(1+exp(-100*(y-0.7)))+0.5/(1+exp(-100*(x-0.4)))-0.5/(1+exp(-100*(x-0.6)))};
%			\end{axis}
%	\end{tikzpicture}
%	\begin{tikzpicture}[scale=0.8]
%		\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},title={$s^1_x=0.4,s^2_x=0.6,s^1_y=0.3,s^2_y=0.7, h=0.5$},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%			\addplot3[surf, domain=0:1,samples=26,shader=faceted]{0.25/(1+exp(-100*(y-0.3)))-0.25/(1+exp(-100*(y-0.7)))+0.25/(1+exp(-100*(x-0.4)))-0.25/(1+exp(-100*(x-0.6)))};
%		\end{axis}
%	\end{tikzpicture}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/bumpxy1}
	\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/bumpxy2}
\end{center}
To simplify the diagram I've dropped the connections with zero weight. For now, I've left in the little $x$ and $y$ markers on the hidden neurons, to remind you in what directions the bump functions are being computed. We'll drop even those markers later, since they're implied by the input variable.
%
Try varying the parameter $h$. As you can see, this causes the output weights to change, and also the heights of both the $x$ and $y$ bump functions.
%
What we've built looks a little like a \textit{tower} function:
\begin{center}
\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/towerfunction}%	\begin{tikzpicture}[scale=0.8]
%			\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={Tower function},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%				\addplot3[surf, domain=0:1,samples=26,shader=faceted]{(1/(1+exp(-100*(y-0.4)))-1/(1+exp(-100*(y-0.6))))*(1/(1+exp(-100*(x-0.4)))-1/(1+exp(-100*(x-0.6))))};
%				\end{axis}
%		\end{tikzpicture}
\end{center}
%
%x=1y=1Tower function
If we could build such tower functions, then we could use them to approximate arbitrary functions, just by adding up many towers of different heights, and in different locations:
\begin{center}
\includegraphics[width=0.32\linewidth]{./figures/ch4/3d/manytowers}
%	\begin{tikzpicture}[scale=0.8]
%			\begin{axis}[colormap={justred}{rgb255=(255,127,127);rgb255=(127,31,31)},grid=major,width=5cm,height=0.45\linewidth,view={-15}{10},xlabel={$x$},ylabel={$y$},zlabel={Output},title={Many towers},zmin=0,zmax=1,xtick={0,1},ytick={0,1},zmajorticks=false]
%				\addplot3[surf, domain=0:1,samples=26,shader=faceted]{0.2*(1/(1+exp(-100*(y-0.1)))-1/(1+exp(-100*(y-0.4))))*(1/(1+exp(-100*(x-0.1)))-1/(1+exp(-100*(x-0.4))))+0.4*(1/(1+exp(-100*(y-0.4)))-1/(1+exp(-100*(y-0.7))))*(1/(1+exp(-100*(x-0.1)))-1/(1+exp(-100*(x-0.4))))+0.6*(1/(1+exp(-100*(y-0.1)))-1/(1+exp(-100*(y-0.4))))*(1/(1+exp(-100*(x-0.4)))-1/(1+exp(-100*(x-0.7))))+0.8*(1/(1+exp(-100*(y-0.4)))-1/(1+exp(-100*(y-0.7))))*(1/(1+exp(-100*(x-0.4)))-1/(1+exp(-100*(x-0.7))))};
%				\end{axis}
%		\end{tikzpicture}
\end{center}
%x=1y=1Many towers
Of course, we haven't yet figured out how to build a tower function. What we have constructed looks like a central tower, of height $2h$, with a surrounding plateau, of height $h$.

But we can make a tower function. Remember that earlier we saw neurons can be used to implement a type of 'inline{if-then-else} statement:
\begin{lstlisting}
if input >= threshold: 
	output 1
else:
	output 0
\end{lstlisting}
That was for a neuron with just a single input. What we want is to apply a similar idea to the combined output from the hidden neurons:
\begin{lstlisting}
if combined output from hidden neurons >= threshold:
	output 1
else:
	output 0
\end{lstlisting}
If we choose the threshold appropriately --- say, a value of $3h/2$, which is sandwiched between the height of the plateau and the height of the central tower -- we could squash the plateau down to zero, and leave just the tower standing.

Can you see how to do this? Try experimenting with the following network to figure it out. Note that we're now plotting the output from the entire network, not just the weighted output from the hidden layer. This means we add a bias term to the weighted output from the hidden layer, and apply the sigma function. Can you find values for $h$ and $b$ which produce a tower? This is a bit tricky, so if you think about this for a while and remain stuck, here's two hints: (1)~To get the output neuron to show the right kind of \texttt{if-then-else} behaviour, we need the input weights (all $h$ or $-h$) to be large; and (2)~the value of $b$ determines the scale of the \texttt{if-then-else} threshold.
\begin{center}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/tikz436}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/towerconstruction1}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/towerconstruction2}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/towerconstruction3}
\end{center}
With our initial parameters, the output looks like a flattened version of the earlier diagram, with its tower and plateau. To get the desired behaviour, we increase the parameter $h$ until it becomes large. That gives the if-then-else thresholding behaviour. Second, to get the threshold right, we'll choose $b\approx-3h/2$. Try it, and see how it works!

Here's what it looks like, when we use $h=10$:
\begin{center}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/towerconstruction4}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/towerconstruction5}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/towerconstruction6}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/towerconstruction7}
\end{center}
Even for this relatively modest value of $h$, we get a pretty good tower function. And, of course, we can make it as good as we want by increasing $h$ still further, and keeping the bias as $b=-3h/2$.

Let's try gluing two such networks together, in order to compute two different tower functions. To make the respective roles of the two sub-networks clear I've put them in separate boxes, below: each box computes a tower function, using the technique described above. The graph on the right shows the weighted output from the second hidden layer, that is, it's a weighted combination of tower functions.
\begin{center}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/tikz437}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/twotowers1}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/twotowers2}
	\includegraphics[width=0.24\linewidth]{./figures/ch4/3d/twotowers3}
\end{center}
%
%x=1y=1Weighted output
%
In particular, you can see that by modifying the weights in the final layer you can change the height of the output towers.

The same idea can be used to compute as many towers as we like. We can also make them as thin as we like, and whatever height we like. As a result, we can ensure that the weighted output from the second hidden layer approximates any desired function of two variables:
\begin{center}
\includegraphics[width=0.32\linewidth]{figures/ch4/3d/manytowers2}
\end{center}
%\begin{lstlisting}
%ANIMATION
%\end{lstlisting}
%%
%x=1y=1Many towers
In particular, by making the weighted output from the second hidden layer a good approximation to $\sigma^{-1}\circ f$, we ensure the output from our network will be a good approximation to any desired function, $f$.

What about functions of more than two variables?

Let's try three variables $x_1,x_2,x_3$. The following network can be used to compute a tower function in four dimensions:
\begin{center}
\includegraphics[width=0.35\linewidth]{figures/ch4/tikz441}
\end{center}
Here, the $x_1,x_2,x_3$ denote inputs to the network. The $s_1,t_1$ and so on are step points for neurons -- that is, all the weights in the first layer are large, and the biases are set to give the step points $s_1,t_1,s_2,\ldots$. The weights in the second layer alternate $+h$,$-h$, where $h$ is some very large number. And the output bias is $-5h/2$.

This network computes a function which is 1 provided three conditions are met: $x_1$ is between $s_1$ and $t_1$; $x_2$ is between $s_2$ and $t_2$; and $x_3$ is between $s_3$ and $t_3$. The network is 0 everywhere else. That is, it's a kind of tower which is 1 in a little region of input space, and 0 everywhere else.

By gluing together many such networks we can get as many towers as we want, and so approximate an arbitrary function of three variables. Exactly the same idea works in $m$ dimensions. The only change needed is to make the output bias $(-m+1/2)h$, in order to get the right kind of sandwiching behavior to level the plateau.

Okay, so we now know how to use neural networks to approximate a real-valued function of many variables. What about vector-valued functions $f(x_1,\ldots,x_m)\in R^n$? Of course, such a function can be regarded as just $n$ separate real-valued functions, $f^1(x_1,\ldots,x_m)$,$f^2(x_1,\ldots,x_m)$, and so on. So we create a network approximating $f^1$, another network for $f^2$, and so on. And then we simply glue all the networks together. So that's also easy to cope with.

\begin{exercize}{Problem}
\item We've seen how to use networks with two hidden layers to approximate an arbitrary function. Can you find a proof showing that it's possible with just a single hidden layer? As a hint, try working in the case of just two input variables, and showing that: (a)~it's possible to get step functions not just in the $x$ or $y$ directions, but in an arbitrary direction; (b)~by adding up many of the constructions from part (a) it's possible to approximate a tower function which is circular in shape, rather than rectangular; (c)~using these circular towers, it's possible to approximate an arbitrary function. To do part (c) it may help to use ideas from a bit later in this chapter.
\end{exercize}
\section{Extension beyond sigmoid neurons}
We've proved that networks made up of sigmoid neurons can compute any function. Recall that in a sigmoid neuron the inputs $x_1,x_2,\ldots$ result in the output $\sigma$($\sum_jw_jx_j+b$), where $w_j$ are the weights, $b$ is the bias, and $\sigma$ is the sigmoid function:
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch4/tikz451}
\end{center}
What if we consider a different type of neuron, one using some other activation function, $s(z)$:
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch4/tikz452}
\end{center}
That is, we'll assume that if our neurons have inputs $x_1,x_2,\ldots$, weights $w_1,w_2,\ldots$ and bias $b$, then the output is $s(\sum_jw_jx_j+b)$.

We can use this activation function to get a step function, just as we did with the sigmoid. Try ramping up the weight in the following, say to $w=100$:

\begin{tikzpicture}[declare function = {sigma(\x) = 1/(1+exp(-\x)); neuron(\x) = sigma(\x)+0.2*sin(10*deg(\x))*exp(-abs(\x)); f(\x,\w,\b) = neuron(\w*\x+\b);}]
	\pgfplotsset{every axis legend/.append style={at={(1.03,1)},anchor=north west}}

		\begin{axis}[width=0.75\linewidth,height=0.45\linewidth,title={Output from neuron},axis x line=bottom,axis y line=left,xtick={0,1},ytick={0,1},xmax=1.05,ymax=1.05]
		%\addplot[blue!90!red,domain=0:1,samples=101]{f(x,9,-3)};
		\addplot[black,domain=0:0.2,samples=200]{f(x,100,-3)};
		\addplot[blue!90!red,domain=0:1,samples=200]{f(x,15,-3)};
		\addplot[blue!75!red,domain=0:1,samples=100]{f(x,8,-3)};
		\addplot[blue!50!red,domain=0:1,samples=100]{f(x,6,-3)};
		\addplot[blue!25!red,domain=0:1,samples=100]{f(x,4,-3)};
		\addplot[black,domain=0.2:1,samples=50]{1};
		%\addplot[blue!10!red,domain=0:1,samples=101]{f(x,3,-3)};
		\addlegendentry{$w=100, b = -3$}
		\addlegendentry{$w=15, b = -3$}
		%\addlegendentry{$w=9, b = -3$}
		\addlegendentry{$w=8, b = -3$}
		\addlegendentry{$w=6, b = -3$}
		\addlegendentry{$w=4, b = -3$}
		%\addlegendentry{$w=3, b = -3$}

	\end{axis}
\end{tikzpicture}

\begin{tikzpicture}[declare function = {sigma(\x) = 1/(1+exp(-\x)); neuron(\x) = sigma(\x)+0.2*sin(10*deg(\x))*exp(-abs(\x)); f(\x,\w,\b) = neuron(\w*\x+\b);}]
\pgfplotsset{every axis legend/.append style={at={(1.03,1)},anchor=north west}}
\begin{axis}[width=0.75\linewidth,height=0.45\linewidth,axis x line=bottom,axis y line=left,xtick={0,1},ytick={0,1},xmax=1.05,ymax=1.05]
	\addplot[blue!90!red,domain=0:1,samples=101]{f(x,6,-5)};
	%\addplot[blue!75!red,domain=0:1,samples=101]{f(x,6,-4)};
	\addplot[blue!50!red,domain=0:1,samples=101]{f(x,6,-3)};
	%\addplot[blue!25!red,domain=0:1,samples=101]{f(x,6,-2)};
	\addplot[blue!10!red,domain=0:1,samples=101]{f(x,6,-1)};
	\addlegendentry{$w=6, b = -5$}
	%\addlegendentry{$w=6, b = -4$}
	\addlegendentry{$w=6, b = -3$}
	%\addlegendentry{$w=6, b = -2$}
	\addlegendentry{$w=6, b = -1$}
\end{axis}
\end{tikzpicture}

\noindent Just as with the sigmoid, this causes the activation function to contract, and ultimately it becomes a very good approximation to a step function. Try changing the bias, and you'll see that we can set the position of the step to be wherever we choose. And so we can use all the same tricks as before to compute any desired function.

What properties does $s(z)$ need to satisfy in order for this to work? We do need to assume that $s(z)$ is well-defined as $z\to\infty$ and $z\to\infty$. These two limits are the two values taken on by our step function. We also need to assume that these limits are different from one another. If they weren't, there'd be no step, simply a flat graph! But provided the activation function $s(z)$ satisfies these properties, neurons based on such an activation function are universal for computation.

\begin{exercize}{Problems}
\item Earlier in the book we met another type of neuron known as a rectified linear unit. Explain why such neurons don't satisfy the conditions just given for universality. Find a proof of universality showing that rectified linear units are universal for computation.
\item Suppose we consider linear neurons, i.e., neurons with the activation function $s(z)=z$. Explain why linear neurons don't satisfy the conditions just given for universality. Show that such neurons can't be used to do universal computation.
\end{exercize}
\section{Fixing up the step functions}
Up to now, we've been assuming that our neurons can produce step functions exactly. That's a pretty good approximation, but it is only an approximation. In fact, there will be a narrow window of failure, illustrated in the following graph, in which the function behaves very differently from a step function:
\begin{center}
	\includegraphics[width=0.4\linewidth]{figures/ch4/tikz453}
\end{center}
In these windows of failure the explanation I've given for universality will fail.

Now, it's not a terrible failure. By making the weights input to the neurons big enough we can make these windows of failure as small as we like. Certainly, we can make the window much narrower than I've shown above -- narrower, indeed, than our eye could see. So perhaps we might not worry too much about this problem.

Nonetheless, it'd be nice to have some way of addressing the problem.

In fact, the problem turns out to be easy to fix. Let's look at the fix for neural networks computing functions with just one input and one output. The same ideas work also to address the problem when there are more inputs and outputs.

In particular, suppose we want our network to compute some function, $f$. As before, we do this by trying to design our network so that the weighted output from our hidden layer of neurons is $\sigma^{-1}\circ f(x)$:
\begin{center}
	\includegraphics[width=0.6\linewidth]{figures/ch4/tikz454}
\end{center}
If we were to do this using the technique described earlier, we'd use the hidden neurons to produce a sequence of bump functions:
\begin{center}
	\includegraphics[width=0.4\linewidth]{figures/ch4/tikz455}
\end{center}
Again, I've exaggerated the size of the windows of failure, in order to make them easier to see. It should be pretty clear that if we add all these bump functions up we'll end up with a reasonable approximation to $\sigma^{-1}\circ f(x)$, except within the windows of failure.

Suppose that instead of using the approximation just described, we use a set of hidden neurons to compute an approximation to half our original goal function, i.e., to $\sigma^{-1}\circ f(x)/2$. Of course, this looks just like a scaled down version of the last graph:
\begin{center}
	\includegraphics[width=0.4\linewidth]{figures/ch4/tikz456}
\end{center}
And suppose we use another set of hidden neurons to compute an approximation to $\sigma^{-1}\circ f(x)/2$, but with the bases of the bumps \textit{shifted} by half the width of a bump:
\begin{center}
	\includegraphics[width=0.4\linewidth]{figures/ch4/tikz457}
\end{center}
Now we have two different approximations to $\sigma^{-1}\circ f(x)/2$. If we add up the two approximations we'll get an overall approximation to $\sigma^{-1}\circ f(x)$. That overall approximation will still have failures in small windows. But the problem will be much less than before. The reason is that points in a failure window for one approximation won't be in a failure window for the other. And so the approximation will be a factor roughly 2 better in those windows.

We could do even better by adding up a large number, $M$, of overlapping approximations to the function $\sigma^{-1}\circ f(x)/M$. Provided the windows of failure are narrow enough, a point will only ever be in one window of failure. And provided we're using a large enough number $M$ of overlapping approximations, the result will be an excellent overall approximation.

\section*{Conclusion}
The explanation for universality we've discussed is certainly not a practical prescription for how to compute using neural networks! In this, it's much like proofs of universality for NAND gates and the like. For this reason, I've focused mostly on trying to make the construction clear and easy to follow, and not on optimizing the details of the construction. However, you may find it a fun and instructive exercise to see if you can improve the construction.

Although the result isn't directly useful in constructing networks, it's important because it takes off the table the question of whether any particular function is computable using a neural network. The answer to that question is always ``yes''. So the right question to ask is not whether any particular function is computable, but rather what's a good way to compute the function.

The universality construction we've developed uses just two hidden layers to compute an arbitrary function. Furthermore, as we've discussed, it's possible to get the same result with just a single hidden layer. Given this, you might wonder why we would ever be interested in deep networks, i.e., networks with many hidden layers. Can't we simply replace those networks with shallow, single hidden layer networks?

\footnotetext{Chapter acknowledgments: Thanks to \href{http://jendodd.com/}{Jen Dodd} and \href{http://colah.github.io/about.html}{Chris Olah} for many discussions about universality in neural networks. My thanks, in particular, to Chris for suggesting the use of a lookup table to prove universality. The interactive visual form of the chapter is inspired by the work of people such as \href{http://bost.ocks.org/mike/algorithms/}{Mike Bostock}, \href{http://www-cs-students.stanford.edu/~amitp/}{Amit Patel}, \href{http://worrydream.com/}{Bret Victor}, and \href{http://acko.net/}{Steven Wittens}.}

While in principle that's possible, there are good practical reasons to use deep networks. As argued in \hyperref[sec:1.8]{Chapter 1}, deep networks have a hierarchical structure which makes them particularly well adapted to learn the hierarchies of knowledge that seem to be useful in solving real-world problems. Put more concretely, when attacking problems such as image recognition, it helps to use a system that understands not just individual pixels, but also increasingly more complex concepts: from edges to simple geometric shapes, all the way up through complex, multi-object scenes. In later chapters, we'll see evidence suggesting that deep networks do a better job than shallow networks at learning such hierarchies of knowledge. To sum up: universality tells us that neural networks can compute any function; and empirical evidence suggests that deep networks are the networks best adapted to learn the functions useful in solving many real-world problems.
\end{document}

### 2019-04-24 Bibliography/Machine Learning/neuralnetworksanddeeplearning.com.pdf-master/neuralnetworksanddeeplearning.com.pdf-master/ch5_whyhardtotrain.tex
Ameliorate bibliography

\documentclass[a4paper,twoside,10pt]{book}
\include{header}
\begin{document}
\chapter{Why are deep neural networks hard to train?}
Imagine you're an engineer who has been asked to design a computer from scratch. One day you're working away in your office, designing logical circuits, setting out \inline{AND} gates, \inline{OR} gates, and so on, when your boss walks in with bad news. The customer has just added a surprising design requirement: the circuit for the entire computer must be just two layers deep:
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch5/shallow_circuit}
\end{center}
You're dumbfounded, and tell your boss: ``The customer is crazy!''

Your boss replies: ``I think they're crazy, too. But what the customer wants, they get.''

In fact, there's a limited sense in which the customer isn't crazy. Suppose you're allowed to use a special logical gate which lets you \inline{AND} together as many inputs as you want. And you're also allowed a many-input \inline{NAND} gate, that is, a gate which can \inline{AND} multiple inputs and then negate the output. With these special gates it turns out to be possible to compute any function at all using a circuit that's just two layers deep.

But just because something is possible doesn't make it a good idea. In practice, when solving circuit design problems (or most any kind of algorithmic problem), we usually start by figuring out how to solve sub-problems, and then gradually integrate the solutions. In other words, we build up to a solution through multiple layers of abstraction.

For instance, suppose we're designing a logical circuit to multiply two numbers. Chances are we want to build it up out of sub-circuits doing operations like adding two numbers. The sub-circuits for adding two numbers will, in turn, be built up out of sub-sub-circuits for adding two bits. Very roughly speaking our circuit will look like:
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch5/circuit_multiplication}
\end{center}
That is, our final circuit contains at least three layers of circuit elements. In fact, it'll probably contain more than three layers, as we break the sub-tasks down into smaller units than I've described. But you get the general idea.

So deep circuits make the process of design easier. But they're not just helpful for design. There are, in fact, mathematical proofs showing that for some functions very shallow circuits require exponentially more circuit elements to compute than do deep circuits. For instance, a famous series of papers in the early 1980s\footnote{The history is somewhat complex, so I won't give detailed references. See Johan H\aa{}stad's 2012 paper \href{http://eccc.hpi-web.de/report/2012/137/}{On the correlation of parity and small-depth circuits} for an account of the early history and references.} showed that computing the parity of a set of bits requires exponentially many gates, if done with a shallow circuit. On the other hand, if you use deeper circuits it's easy to compute the parity using a small circuit: you just compute the parity of pairs of bits, then use those results to compute the parity of pairs of pairs of bits, and so on, building up quickly to the overall parity. Deep circuits thus can be intrinsically much more powerful than shallow circuits.

Up to now, this book has approached neural networks like the crazy customer. Almost all the networks we've worked with have just a single hidden layer of neurons (plus the input and output layers):
\begin{center}
	\includegraphics[width=0.6\linewidth]{figures/ch5/tikz35}
\end{center}
These simple networks have been remarkably useful: in earlier chapters we used networks like this to classify handwritten digits with better than 98 percent accuracy! Nonetheless, intuitively we'd expect networks with many more hidden layers to be more powerful:
\begin{center}
	\includegraphics[width=\linewidth]{figures/ch5/tikz36}
\end{center}
Such networks could use the intermediate layers to build up multiple layers of abstraction, just as we do in Boolean circuits. For instance, if we're doing visual pattern recognition, then the neurons in the first layer might learn to recognize edges, the neurons in the second layer could learn to recognize more complex shapes, say triangle or rectangles, built up from edges. The third layer would then recognize still more complex shapes. And so on. These multiple layers of abstraction seem likely to give deep networks a compelling advantage in learning to solve complex pattern recognition problems. Moreover, just as in the case of circuits, there are theoretical results suggesting that deep networks are intrinsically more powerful than shallow networks\footnote{For certain problems and network architectures this is proved in \href{http://arxiv.org/pdf/1312.6098.pdf}{On the number of response regions of deep feed forward networks with piece-wise linear activations}, by Razvan Pascanu, Guido Mont\'ufar, and Yoshua Bengio (2014). See also the more informal discussion in section 2 of \href{http://www.iro.umontreal.ca/~bengioy/papers/ftml_book.pdf}{Learning deep architectures for AI}, by Yoshua Bengio (2009).}.

How can we train such deep networks? In this chapter, we'll try training deep networks using our workhorse learning algorithm -- stochastic gradient descent by backpropagation. But we'll run into trouble, with our deep networks not performing much (if at all) better than shallow networks.

That failure seems surprising in the light of the discussion above. Rather than give up on deep networks, we'll dig down and try to understand what's making our deep networks hard to train. When we look closely, we'll discover that the different layers in our deep network are learning at vastly different speeds. In particular, when later layers in the network are learning well, early layers often get stuck during training, learning almost nothing at all. This stuckness isn't simply due to bad luck. Rather, we'll discover there are fundamental reasons the learning slowdown occurs, connected to our use of gradient-based learning techniques.

As we delve into the problem more deeply, we'll learn that the opposite phenomenon can also occur: the early layers may be learning well, but later layers can become stuck. In fact, we'll find that there's an intrinsic instability associated to learning by gradient descent in deep, many-layer neural networks. This instability tends to result in either the early or the later layers getting stuck during training.

This all sounds like bad news. But by delving into these difficulties, we can begin to gain insight into what's required to train deep networks effectively. And so these investigations are good preparation for the next chapter, where we'll use deep learning to attack image recognition problems.

\section{The vanishing gradient problem}
So, what goes wrong when we try to train a deep network?

To answer that question, let's first revisit the case of a network with just a single hidden layer. As per usual, we'll use the MNIST digit classification problem as our playground for learning and experimentation\footnote{I introduced the MNIST problem and data here~\ref{sec:1.5} and here~\ref{sec:1.7}.}.

If you wish, you can follow along by training networks on your computer. It is also, of course, fine to just read along. If you do wish to follow live, then you'll need Python 2.7, Numpy, and a copy of the code, which you can get by cloning the relevant repository from the command line:

\begin{lstlisting}
git clone https://github.com/mnielsen/neural-networks-and-deep-learning.git
\end{lstlisting}
If you don't use git then you can download the data and code \href{https://github.com/mnielsen/neural-networks-and-deep-learning/archive/master.zip}{here}. You'll need to change into the \inline{src} subdirectory.
Then, from a Python shell we load the MNIST data:

\begin{lstlisting}
>>> import mnist_loader
>>> training_data, validation_data, test_data = \
... mnist_loader.load_data_wrapper()
\end{lstlisting}
We set up our network:

\begin{lstlisting}
>>> import network2
>>> net = network2.Network([784, 30, 10])
\end{lstlisting}
This network has 784 neurons in the input layer, corresponding to the $28\times28=784$ pixels in the input image. We use 30 hidden neurons, as well as 10 output neurons, corresponding to the 10 possible classifications for the MNIST digits (`0', `1', `2',..., `9').

Let's try training our network for 30 complete epochs, using mini-batches of 10 training examples at a time, a learning rate $\eta=0.1$, and regularization parameter $\lambda=5.0$. As we train we'll monitor the classification accuracy on the \inline{validation_data}\footnote{Note that the networks is likely to take some minutes to train, depending on the speed of your machine. So if you're running the code you may wish to continue reading and return later, not wait for the code to finish executing.}:
\begin{lstlisting}
>>> net.SGD(training_data, 30, 10, 0.1, lmbda=5.0, 
... evaluation_data=validation_data, monitor_evaluation_accuracy=True)
\end{lstlisting}
We get a classification accuracy of 96.48 percent (or thereabouts -- it'll vary a bit from run to run), comparable to our earlier results with a similar configuration.

Now, let's add another hidden layer, also with 30 neurons in it, and try training with the same hyper-parameters:

\begin{lstlisting}
>>> net = network2.Network([784, 30, 30, 10])
>>> net.SGD(training_data, 30, 10, 0.1, lmbda=5.0, 
... evaluation_data=validation_data, monitor_evaluation_accuracy=True)
\end{lstlisting}
This gives an improved classification accuracy, 96.90 percent. That's encouraging: a little more depth is helping. Let's add another 30-neuron hidden layer:

\begin{lstlisting}
>>> net = network2.Network([784, 30, 30, 30, 10])
>>> net.SGD(training_data, 30, 10, 0.1, lmbda=5.0, 
... evaluation_data=validation_data, monitor_evaluation_accuracy=True)
\end{lstlisting}
That doesn't help at all. In fact, the result drops back down to 96.57 percent, close to our original shallow network. And suppose we insert one further hidden layer:

\begin{lstlisting}
>>> net = network2.Network([784, 30, 30, 30, 30, 10])
>>> net.SGD(training_data, 30, 10, 0.1, lmbda=5.0, 
... evaluation_data=validation_data, monitor_evaluation_accuracy=True)
\end{lstlisting}
The classification accuracy drops again, to 96.53 percent. That's probably not a statistically significant drop, but it's not encouraging, either.

This behaviour seems strange. Intuitively, extra hidden layers ought to make the network able to learn more complex classification functions, and thus do a better job classifying. Certainly, things shouldn't get worse, since the extra layers can, in the worst case, simply do nothing\footnote{See this later problem~\ref{sec:5.2.1} to understand how to build a hidden layer that does nothing.}. But that's not what's going on.

So what is going on? Let's assume that the extra hidden layers really could help in principle, and the problem is that our learning algorithm isn't finding the right weights and biases. We'd like to figure out what's going wrong in our learning algorithm, and how to do better.

To get some insight into what's going wrong, let's visualize how the network learns. Below, I've plotted part of a [784,30,30,10] network, i.e., a network with two hidden layers, each containing 30 hidden neurons. Each neuron in the diagram has a little bar on it, representing how quickly that neuron is changing as the network learns. A big bar means the neuron's weights and bias are changing rapidly, while a small bar means the weights and bias are changing slowly. More precisely, the bars denote the gradient $\partial C/\partial b$ for each neuron, i.e., the rate of change of the cost with respect to the neuron's bias. Back in Chapter 2 we saw that this gradient quantity controlled not just how rapidly the bias changes during learning, but also how rapidly the weights input to the neuron change, too. Don't worry if you don't recall the details: the thing to keep in mind is simply that these bars show how quickly each neuron's weights and bias are changing as the network learns.

To keep the diagram simple, I've shown just the top six neurons in the two hidden layers. I've omitted the input neurons, since they've got no weights or biases to learn. I've also omitted the output neurons, since we're doing layer-wise comparisons, and it makes most sense to compare layers with the same number of neurons. The results are plotted at the very beginning of training, i.e., immediately after the network is initialized. Here they are\footnote{The data plotted is generated using the program \inline{generate_gradient.py}. The same program is also used to generate the results quoted later in this section.}:
\begin{center}
	\includegraphics[width=0.45\linewidth]{figures/ch5/tikz500}
\end{center}
The network was initialized randomly, and so it's not surprising that there's a lot of variation in how rapidly the neurons learn. Still, one thing that jumps out is that the bars in the second hidden layer are mostly much larger than the bars in the first hidden layer. As a result, the neurons in the second hidden layer will learn quite a bit faster than the neurons in the first hidden layer. Is this merely a coincidence, or are the neurons in the second hidden layer likely to learn faster than neurons in the first hidden layer in general?

To determine whether this is the case, it helps to have a global way of comparing the speed of learning in the first and second hidden layers. To do this, let's denote the gradient as $\delta^l_j=\partial C/\partial b^l_j$, i.e., the gradient for the $j$-th neuron in the $l$-th layer\footnote{Back in Chapter 2 we referred to this as the error, but here we'll adopt the informal term ``gradient''. I say ``informal'' because of course this doesn't explicitly include the partial derivatives of the cost with respect to the weights, $\partial C/\partial w$.}  We can think of the gradient $\delta^1$ as a vector whose entries determine how quickly the first hidden layer learns, and $\delta^2$ as a vector whose entries determine how quickly the second hidden layer learns. We'll then use the lengths of these vectors as (rough!) global measures of the speed at which the layers are learning. So,for instance, the length $\|\delta^1\|$ measures the speed at which the first hidden layer is learning, while the length $\|\delta^2\|$ measures the speed at which the second hidden layer is learning.

With these definitions, and in the same configuration as was plotted above, we find $\|\delta^1\| = 0.07...$ and $\|\delta^2\|=0.31...$. So this confirms our earlier suspicion: the neurons in the second hidden layer really are learning much faster than the neurons in the first hidden layer.

What happens if we add more hidden layers? If we have three hidden layers, in a [784,30,30,30,10] network, then the respective speeds of learning turn out to be 0.012, 0.060, and 0.283. Again, earlier hidden layers are learning much slower than later hidden layers. Suppose we add yet another layer with 30 hidden neurons. In that case, the respective speeds of learning are 0.003, 0.017, 0.070, and 0.285. The pattern holds: early layers learn slower than later layers.

We've been looking at the speed of learning at the start of training, that is, just after the networks are initialized. How does the speed of learning change as we train our networks? Let's return to look at the network with just two hidden layers. The speed of learning changes as follows:
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch5/training_speed_2_layers}
\end{center}
To generate these results, I used batch gradient descent with just 1,000 training images, trained over 500 epochs. This is a bit different than the way we usually train -- I've used no mini-batches, and just 1,000 training images, rather than the full 50,000 image training set. I'm not trying to do anything sneaky, or pull the wool over your eyes, but it turns out that using mini-batch stochastic gradient descent gives much noisier (albeit very similar, when you average away the noise) results. Using the parameters I've chosen is an easy way of smoothing the results out, so we can see what's going on.

In any case, as you can see the two layers start out learning at very different speeds (as we already know). The speed in both layers then drops very quickly, before rebounding. But through it all, the first hidden layer learns much more slowly than the second hidden layer.

What about more complex networks? Here's the results of a similar experiment, but this time with three hidden layers (a [784,30,30,30,10] network):
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch5/training_speed_3_layers}
\end{center}
Again, early hidden layers learn much more slowly than later hidden layers. Finally, let's add a fourth hidden layer (a [784,30,30,30,30,10] network), and see what happens when we train:
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch5/training_speed_4_layers}
\end{center}
Again, early hidden layers learn much more slowly than later hidden layers. In this case, the first hidden layer is learning roughly 100 times slower than the final hidden layer. No wonder we were having trouble training these networks earlier!

We have here an important observation: in at least some deep neural networks, the gradient tends to get smaller as we move backward through the hidden layers. This means that neurons in the earlier layers learn much more slowly than neurons in later layers. And while we've seen this in just a single network, there are fundamental reasons why this happens in many neural networks. The phenomenon is known as the vanishing gradient problem\footnote{See \href{http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.24.7321}{Gradient flow in recurrent nets: the difficulty of learning long-term dependencies}, by Sepp Hochreiter, Yoshua Bengio, Paolo Frasconi, and J\"urgen Schmidhuber (2001). This paper studied recurrent neural nets, but the essential phenomenon is the same as in the feedforward networks we are studying. See also Sepp Hochreiter's earlier Diploma Thesis, \href{http://www.idsia.ch/~juergen/SeppHochreiter1991ThesisAdvisorSchmidhuber.pdf}{Untersuchungen zu dynamischen neuronalen Netzen} (1991, in German).}.

Why does the vanishing gradient problem occur? Are there ways we can avoid it? And how should we deal with it in training deep neural networks? In fact, we'll learn shortly that it's not inevitable, although the alternative is not very attractive, either: sometimes the gradient gets much larger in earlier layers! This is the exploding gradient problem, and it's not much better news than the vanishing gradient problem. More generally, it turns out that the gradient in deep neural networks is unstable, tending to either explode or vanish in earlier layers. This instability is a fundamental problem for gradient-based learning in deep neural networks. It's something we need to understand, and, if possible, take steps to address.

One response to vanishing (or unstable) gradients is to wonder if they're really such a problem. Momentarily stepping away from neural nets, imagine we were trying to numerically minimize a function $f(x)$ of a single variable. Wouldn't it be good news if the derivative $f'(x)$ was small? Wouldn't that mean we were already near an extremum? In a similar way, might the small gradient in early layers of a deep network mean that we don't need to do much adjustment of the weights and biases?

Of course, this isn't the case. Recall that we randomly initialized the weight and biases in the network. It is extremely unlikely our initial weights and biases will do a good job at whatever it is we want our network to do. To be concrete, consider the first layer of weights in a [784,30,30,30,10] network for the MNIST problem. The random initialization means the first layer throws away most information about the input image. Even if later layers have been extensively trained, they will still find it extremely difficult to identify the input image, simply because they don't have enough information. And so it can't possibly be the case that not much learning needs to be done in the first layer. If we're going to train deep networks, we need to figure out how to address the vanishing gradient problem.

\section{What's causing the vanishing gradient problem? Unstable gradients in deep neural nets}
To get insight into why the vanishing gradient problem occurs, let's consider the simplest deep neural network: one with just a single neuron in each layer. Here's a network with three hidden layers:
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch5/tikz37}
\end{center}
Here, $w_1,w_2,\ldots$ are the weights, $b_1,b_2,\ldots$ are the biases, and $C$ is some cost function. Just to remind you how this works, the output $a_j$ from the $j$-th neuron is $\sigma(z_j)$, where $\sigma$ is the usual sigmoid activation function, and $z_j=w_ja_{j-1}+b_j$ is the weighted input to the neuron. I've drawn the cost $C$ at the end to emphasize that the cost is a function of the network's output, $a_4$: if the actual output from the network is close to the desired output, then the cost will be low, while if it's far away, the cost will be high.

We're going to study the gradient $\partial C/\partial b_1$ associated to the first hidden neuron. We'll figure out an expression for $\partial C/\partial b_1$, and by studying that expression we'll understand why the vanishing gradient problem occurs.

I'll start by simply showing you the expression for $\partial C/\partial b_1$. It looks forbidding, but it's actually got a simple structure, which I'll describe in a moment. Here's the expression (ignore the network, for now, and note that $\sigma'$ is just the derivative of the $\sigma$ function):
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch5/tikz38}
\end{center}
The structure in the expression is as follows: there is a $\sigma'(z_j)$ term in the product for each neuron in the network; a weight $w_j$ term for each weight in the network; and a final $\partial C/\partial a_4$ term, corresponding to the cost function at the end. Notice that I've placed each term in the expression above the corresponding part of the network. So the network itself is a mnemonic for the expression.

You're welcome to take this expression for granted, and skip to the discussion of how it relates to the vanishing gradient problem. There's no harm in doing this, since the expression is a special case of our earlier discussion of backpropagation. But there's also a simple explanation of why the expression is true, and so it's fun (and perhaps enlightening) to take a look at that explanation.

Imagine we make a small change $\Delta b_1$ in the bias $b_1$. That will set off a cascading series of changes in the rest of the network. First, it causes a change $\Delta a_1$ in the output from the first hidden neuron. That, in turn, will cause a change $\Delta z_2$ in the weighted input to the second hidden neuron. Then a change $\Delta a_2$ in the output from the second hidden neuron. And so on, all the way through to a change $\Delta C$ in the cost at the output. We have
%$\partial C\partial b1≈\deltaC\delta b_1.(114)
\begin{align}
\frac{\partial C}{\partial b_1} \approx \frac{\Delta C}{\Delta b_1}.\tag{114}\label{eq:114}
\end{align}
This suggests that we can figure out an expression for the gradient $\partial C/\partial b_1$ by carefully tracking the effect of each step in this cascade.

To do this, let's think about how $\Delta b_1$ causes the output $a_1$ from the first hidden neuron to change. We have $a_1=\sigma(z_1)=\sigma(w_1a_0+b_1)$, so
%\delta a_1≈=\partial \sigma(w1a0+b1)\partial b1\delta b_1\sigma'(z1)\delta b_1.(115)(116)
\begin{align}
\Delta a_1 \approx \frac{\partial \sigma(w_1 a_0+b_1)}{\partial b_1} \Delta b_1 = \sigma'(z_1) \Delta b_1. \tag{116}\label{eq:116}
\end{align}
That $\sigma'(z_1)$ term should look familiar: it's the first term in our claimed expression for the gradient $\partial C/\partial b_1$. Intuitively, this term converts a change $\Delta b_1$ in the bias into a change $\Delta a_1$ in the output activation. That change $\Delta a_1$ in turn causes a change in the weighted input $z_2=w_2a_1+b_2$ to the second hidden neuron:
\begin{align}
\Delta z_2  \approx \frac{\partial z_2}{\partial a_1} \Delta a_1 =  w_2 \Delta a_1. \tag{118}\label{eq:118}
\end{align}
Combining our expressions for $\Delta z_2$ and $\Delta a_1$, we see how the change in the bias $b_1$ propagates along the network to affect $z_2$:
\begin{align}
\Delta z_2 \approx \sigma'(z_1) w_2 \Delta b_1.\tag{119}\label{eq:119}
\end{align}
%\delta z_2≈\sigma'(z1)w2\delta b_1.(119)
Again, that should look familiar: we've now got the first two terms in our claimed expression for the gradient $\partial C/\partial b_1$.

We can keep going in this fashion, tracking the way changes propagate through the rest of the network. At each neuron we pick up a $\sigma'(z_j)$ term, and through each weight we pick up a $w_j$ term. The end result is an expression relating the final change $\Delta C$ in cost to the initial change $\Delta b_1$ in the bias:
\begin{align}
\Delta C  \approx  \sigma'(z_1) w_2 \sigma'(z_2) \ldots \sigma'(z_4) \frac{\partial C}{\partial a_4} \Delta b_1. \tag{120}\label{eq:120}
\end{align}
%\deltaC≈\sigma'(z1)w2\sigma'(z2)…\sigma'(z4)\partial C\partial a4\delta b_1.(120)
Dividing by $\Delta b_1$ we do indeed get the desired expression for the gradient:
%\partial C\partial b1=\sigma'(z1)w2\sigma'(z2)…\sigma'(z4)\partial C\partial a4.(121)
\begin{align}
\frac{\partial C}{\partial b_1} = \sigma'(z_1) w_2 \sigma'(z_2) \ldots \sigma'(z_4) \frac{\partial C}{\partial a_4}. \tag{121}\label{eq:121}
\end{align}

\textbf{Why the vanishing gradient problem occurs:} To understand why the vanishing gradient problem occurs, let's explicitly write out the entire expression for the gradient:
\begin{align}
\frac{\partial C}{\partial b_1} = \sigma'(z_1) \, w_2 \sigma'(z_2) \,w_3 \sigma'(z_3) \, w_4 \sigma'(z_4) \, \frac{\partial C}{\partial a_4}.\tag{122}\label{eq:122}
\end{align}
Excepting the very last term, this expression is a product of terms of the form $w_j\sigma'(z_j)$. To understand how each of those terms behave, let's look at a plot of the function $\sigma'$:
\begin{center} 
\begin{tikzpicture}
\begin{axis}[title={Derivative of sigmoid function},axis x line=bottom,axis y line=left,width=0.75\linewidth,height=0.35\linewidth,ymin=0,ymax=0.27,scaled y ticks = false,y tick label style={/pgf/number format/fixed},xmin=-5.5,xmax=5.5]
\addplot[blue,domain=-5:5,samples=101]{exp(-x)/(1+exp(-x))/(1+exp(-x))};
\end{axis}
\end{tikzpicture}
\end{center}%
The derivative reaches a maximum at $\sigma'(0)=1/4$. Now, if we use our standard approach to initializing the weights in the network, then we'll choose the weights using a Gaussian with mean 0 and standard deviation 1. So the weights will usually satisfy $|w_j|<1$. Putting these observations together, we see that the terms $w_j\sigma'(z_j)$ will usually satisfy $|w_j\sigma'(z_j)|<1/4$. And when we take a product of many such terms, the product will tend to exponentially decrease: the more terms, the smaller the product will be. This is starting to smell like a possible explanation for the vanishing gradient problem.

To make this all a bit more explicit, let's compare the expression for $\partial C/\partial b_1$ to an expression for the gradient with respect to a later bias, say $\partial C/\partial b_3$. Of course, we haven't explicitly worked out an expression for $\partial C/\partial b_3$, but it follows the same pattern described above for $\partial C/\partial b_1$. Here's the comparison of the two expressions:
\begin{center}
	\includegraphics[width=0.7\linewidth]{figures/ch5/tikz39}
\end{center}
The two expressions share many terms. But the gradient $\partial C/\partial b_1$ includes two extra terms each of the form $w_j\sigma'(z_j)$. As we've seen, such terms are typically less than 1/4 in magnitude. And so the gradient $\partial C/\partial b_1$ will usually be a factor of 16 (or more) smaller than $\partial C/\partial b_3$. This is the essential origin of the vanishing gradient problem.

Of course, this is an informal argument, not a rigorous proof that the vanishing gradient problem will occur. There are several possible escape clauses. In particular, we might wonder whether the weights $w_j$ could grow during training. If they do, it's possible the terms $w_j\sigma'(z_j)$ in the product will no longer satisfy $|w_j\sigma'(z_j)|<1/4$. Indeed, if the terms get large enough -- greater than 1 -- then we will no longer have a vanishing gradient problem. Instead, the gradient will actually grow exponentially as we move backward through the layers. Instead of a vanishing gradient problem, we'll have an exploding gradient problem.

\textbf{The exploding gradient problem:} Let's look at an explicit example where exploding gradients occur. The example is somewhat contrived: I'm going to fix parameters in the network in just the right way to ensure we get an exploding gradient. But even though the example is contrived, it has the virtue of firmly establishing that exploding gradients aren't merely a hypothetical possibility, they really can happen.

There are two steps to getting an exploding gradient. First, we choose all the weights in the network to be large, say $w_1=w_2=w_3=w_4=100$. Second, we'll choose the biases so that the $\sigma'(z_j)$ terms are not too small. That's actually pretty easy to do: all we need do is choose the biases to ensure that the weighted input to each neuron is $z_j=0$ (and so $\sigma'(z_j)=1/4$). So, for instance, we want $z_1=w_1a_0+b_1=0$. We can achieve this by setting $b_1=-100\times a_0$. We can use the same idea to select the other biases. When we do this, we see that all the terms $w_j\sigma'(z_j)$ are equal to $100\times1/4=25$. With these choices we get an exploding gradient.

\textbf{The unstable gradient problem:} The fundamental problem here isn't so much the vanishing gradient problem or the exploding gradient problem. It's that the gradient in early layers is the product of terms from all the later layers. When there are many layers, that's an intrinsically unstable situation. The only way all layers can learn at close to the same speed is if all those products of terms come close to balancing out. Without some mechanism or underlying reason for that balancing to occur, it's highly unlikely to happen simply by chance. In short, the real problem here is that neural networks suffer from an \textit{unstable gradient problem}. As a result, if we use standard gradient-based learning techniques, different layers in the network will tend to learn at wildly different speeds.

\begin{exercize}{Exercise}
\item In our discussion of the vanishing gradient problem, we made use of the fact that $|\sigma'(z)|<1/4$. Suppose we used a different activation function, one whose derivative could be much larger. Would that help us avoid the unstable gradient problem?
\end{exercize}
\textbf{The prevalence of the vanishing gradient problem:} We've seen that the gradient can either vanish or explode in the early layers of a deep network. In fact, when using sigmoid neurons the gradient will usually vanish. To see why, consider again the expression $|w\sigma'(z)|$. To avoid the vanishing gradient problem we need $|w\sigma'(z)|\geq1$. You might think this could happen easily if $w$ is very large. However, it's more difficult than it looks. The reason is that the $\sigma'(z)$ term also depends on $w$: $\sigma'(z)=\sigma'(wa+b)$, where $a$ is the input activation. So when we make $w$ large, we need to be careful that we're not simultaneously making $\sigma'(wa+b)$ small. That turns out to be a considerable constraint. The reason is that when we make $w$ large we tend to make $wa+b$ very large. Looking at the graph of $\sigma'$ you can see that this puts us off in the ``wings'' of the $\sigma'$ function, where it takes very small values. The only way to avoid this is if the input activation falls within a fairly narrow range of values (this qualitative explanation is made quantitative in the first problem below). Sometimes that will chance to happen. More often, though, it does not happen. And so in the generic case we have vanishing gradients.

\begin{exercize}{Problems}
\item Consider the product $|w\sigma'(wa+b)|$. Suppose $|w\sigma'(wa+b)|\geq1$. (1) Argue that this can only ever occur if $|w|≥\geq4$. (2) Supposing that $|w|\geq4$, consider the set of input activations a for which $|w\sigma'(wa+b)|\geq1$. Show that the set of a satisfying that constraint can range over an interval no greater in width than
\begin{equation}
%	2|w|ln(|w|(1+1-4/|w|--------√)2-1).(123)
	\frac2{|w|}\ln\left(\frac{|w|(1+\sqrt{1-4/|w|})}{2}-1\right).\tag{123}\label{eq:123}
\end{equation}
(3) Show numerically that the above expression bounding the width of the range is greatest at $|w|\approx6.9$, where it takes a value $\approx0.45$. And so even given that everything lines up just perfectly, we still have a fairly narrow range of input activations which can avoid the vanishing gradient problem.
\item \textbf{Identity neuron:}\label{sec:5.2.1} Consider a neuron with a single input, $x$, a corresponding weight, $w_1$, a bias $b$, and a weight $w_2$ on the output. Show that by choosing the weights and bias appropriately, we can ensure $w_2\sigma(w_1x+b)\approx x$ for $x\in[0,1]$. Such a neuron can thus be used as a kind of identity neuron, that is, a neuron whose output is the same (up to rescaling by a weight factor) as its input. \textit{Hint: It helps to rewrite $x=1/2+\Delta$, to assume $w_1$ is small, and to use a Taylor series expansion in $w_1\Delta$.}
\end{exercize}
\section{Unstable gradients in more complex networks}
We've been studying toy networks, with just one neuron in each hidden layer. What about more complex deep networks, with many neurons in each hidden layer?
\begin{center}
	\includegraphics[width=\linewidth]{figures/ch5/tikz40}
\end{center}
In fact, much the same behaviour occurs in such networks. In the earlier chapter on backpropagation we saw that the gradient in the $l$-th layer of an L layer network is given by:

%\deltal=\sigma'(zl)(wl+1)T\sigma'(zl+1)(wl+2)T…\sigma'(zL)∇aC(124)
\begin{align}
\delta^l = \Sigma'(z^l) (w^{l+1})^T \Sigma'(z^{l+1}) (w^{l+2})^T \ldots\Sigma'(z^L) \nabla_a C
\tag{124}\label{eq:124}
\end{align}
Here, $\Sigma'(z^l)$ is a diagonal matrix whose entries are the $\sigma'(z)$ values for the weighted inputs to the $l$-th layer. The $w^l$ are the weight matrices for the different layers. And $\nabla_aC$ is the vector of partial derivatives of $C$ with respect to the output activations.

This is a much more complicated expression than in the single-neuron case. Still, if you look closely, the essential form is very similar, with lots of pairs of the form $(w^j)^T\Sigma'(z^j)$. What's more, the matrices $\Sigma'(z^j)$ have small entries on the diagonal, none larger than 1/4. Provided the weight matrices $w^j$ aren't too large, each additional term $(w^j)^T\Sigma'(z^l)$ tends to make the gradient vector smaller, leading to a vanishing gradient. More generally, the large number of terms in the product tends to lead to an unstable gradient, just as in our earlier example. In practice, empirically it is typically found in sigmoid networks that gradients vanish exponentially quickly in earlier layers. As a result, learning slows down in those layers. This slowdown isn't merely an accident or an inconvenience: it's a fundamental consequence of the approach we're taking to learning.

\section{Other obstacles to deep learning}
In this chapter we've focused on vanishing gradients -- and, more generally, unstable gradients -- as an obstacle to deep learning. In fact, unstable gradients are just one obstacle to deep learning, albeit an important fundamental obstacle. Much ongoing research aims to better understand the challenges that can occur when training deep networks. I won't comprehensively summarize that work here, but just want to briefly mention a couple of papers, to give you the flavor of some of the questions people are asking.

As a first example, in 2010 Glorot and Bengio\footnote{\href{http://jmlr.org/proceedings/papers/v9/glorot10a/glorot10a.pdf}{Understanding the difficulty of training deep feedforward neural networks}, by Xavier Glorot and Yoshua Bengio (2010). See also the earlier discussion of the use of sigmoids in \href{http://yann.lecun.com/exdb/publis/pdf/lecun-98b.pdf}{Efficient BackProp}, by Yann LeCun, L\'eon Bottou, Genevieve Orr and Klaus-Robert M\"uller (1998).} found evidence suggesting that the use of sigmoid activation functions can cause problems training deep networks. In particular, they found evidence that the use of sigmoids will cause the activations in the final hidden layer to saturate near 0 early in training, substantially slowing down learning. They suggested some alternative activation functions, which appear not to suffer as much from this saturation problem.

As a second example, in 2013 Sutskever, Martens, Dahl and Hinton\footnote{\href{http://www.cs.toronto.edu/~hinton/absps/momentum.pdf}{On the importance of initialization and momentum in deep learning}, by Ilya Sutskever, James Martens, George Dahl and Geoffrey Hinton (2013).} studied the impact on deep learning of both the random weight initialization and the momentum schedule in momentum-based stochastic gradient descent. In both cases, making good choices made a substantial difference in the ability to train deep networks.

These examples suggest that ``What makes deep networks hard to train?'' is a complex question. In this chapter, we've focused on the instabilities associated to gradient-based learning in deep networks. The results in the last two paragraphs suggest that there is also a role played by the choice of activation function, the way weights are initialized, and even details of how learning by gradient descent is implemented. And, of course, choice of network architecture and other hyper-parameters is also important. Thus, many factors can play a role in making deep networks hard to train, and understanding all those factors is still a subject of ongoing research. This all seems rather downbeat and pessimism-inducing. But the good news is that in the next chapter we'll turn that around, and develop several approaches to deep learning that to some extent manage to overcome or route around all these challenges.

\end{document}

### 2019-04-24 Bibliography/Machine Learning/neuralnetworksanddeeplearning.com.pdf-master/neuralnetworksanddeeplearning.com.pdf-master/scripts/animation_32.py

@@ -1,0 +1,59 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Aug  3 20:04:02 2018
+
+@author:
+"""
+
+# chapter 3, animation 1
+import math
+def cost(x):    return x**2/2
+def costd(x):   return x*x*(1-x)
+def s(x):       return 1/(1+math.exp(-x))
+def ov(w,b):    return s(w+b)
+x = 1
+w0 = 2
+b0 = 2
+eta = 0.15
+w = w0
+b = b0
+Y = []
+X = []
+A = []
+W = []
+B = []
+while x <= 300:
+    a = ov(w,b)
+    d = costd(a)
+    w = w - eta * d
+    b = b - eta * d
+    y = cost(a)
+    X.append(x)
+    Y.append(y)
+    W.append(w)
+    B.append(b)
+    A.append(a)
+    x = x + 1
+
+import matplotlib.pyplot as plt
+plt.figure(figsize=(10,5))
+plt.subplot(2,2,1)
+plt.plot(X,Y)
+plt.legend(['Cost'],loc=0)
+plt.subplot(2,2,3)
+plt.plot(X,A,'red')
+plt.xlabel('Epoch')
+plt.legend(['Output'],loc=0)
+plt.subplot(2,2,2)
+plt.plot(X,W,'green')
+plt.legend(['Weight'],loc=0)
+plt.subplot(2,2,4)
+plt.plot(X,B,'black')
+plt.xlabel('Epoch')
+plt.legend(['Bias'],loc=0)
+plt.tight_layout()
+plt.savefig('animation_32',dpi=300,pad_inches=0,bbox_inches=0)
+#plt.plot(X,B)
+#plt.plot(X,W)
+
+    

### 2019-04-24 Bibliography/Machine Learning/neuralnetworksanddeeplearning.com.pdf-master/neuralnetworksanddeeplearning.com.pdf-master/scripts/animation_33.py

@@ -1,0 +1,59 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Aug  3 20:04:02 2018
+
+@author:
+"""
+
+# chapter 3, animation 1
+import math
+def cost(x):    return x**2/2
+def costd(x):   return x*x*(1-x)
+def s(x):       return 1/(1+math.exp(-x))
+def ov(w,b):    return s(w+b)
+x = 1
+w0 = 0.6
+b0 = 0.9
+eta = 0.15
+w = w0
+b = b0
+Y = []
+X = []
+A = []
+W = []
+B = []
+while x <= 300:
+    a = ov(w,b)
+    d = costd(a)
+    w = w - eta * d
+    b = b - eta * d
+    y = cost(a)
+    X.append(x)
+    Y.append(y)
+    W.append(w)
+    B.append(b)
+    A.append(a)
+    x = x + 1
+
+import matplotlib.pyplot as plt
+plt.figure(figsize=(10,5))
+plt.subplot(2,2,1)
+plt.plot(X,Y)
+plt.legend(['Cost'],loc=0)
+plt.subplot(2,2,3)
+plt.plot(X,A,'red')
+plt.xlabel('Epoch')
+plt.legend(['Output'],loc=0)
+plt.subplot(2,2,2)
+plt.plot(X,W,'green')
+plt.legend(['Weight'],loc=0)
+plt.subplot(2,2,4)
+plt.plot(X,B,'black')
+plt.xlabel('Epoch')
+plt.legend(['Bias'],loc=0)
+plt.tight_layout()
+plt.savefig('animation_31',dpi=300,pad_inches=0,bbox_inches=0)
+#plt.plot(X,B)
+#plt.plot(X,W)
+
+    

### 2019-04-24 Bibliography/Machine Learning/neuralnetworksanddeeplearning.com.pdf-master/neuralnetworksanddeeplearning.com.pdf-master/scripts/animation_overfitting1.py

@@ -1,0 +1,56 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Aug  3 20:04:02 2018
+
+@author:
+"""
+
+# chapter 3, animation 1
+import numpy as np
+import matplotlib.pyplot as plt
+data = np.array([[ 0.7, 1.8], 
+	    [ 1.3, 2.2], 
+	    [ 1.9, 4.0], 
+	    [ 2.6, 5.0], 
+	    [ 2.9, 6.1], 
+	    [ 3.6, 7.0], 
+	    [ 3.8, 7.4], 
+	    [ 3.95, 8.0], 
+	    [ 4.4, 9.1], 
+	    [ 4.9, 10.0] ])
+X = data[:,0]
+Y = data[:,1]
+plt.figure(figsize=(6,4))
+plt.scatter(X,Y)
+plt.xlabel('$x$')
+plt.ylabel('$y$')
+plt.xlim(0,5.5)
+plt.ylim(0,11)
+plt.tight_layout()
+plt.savefig('animation_overfitting1',dpi=300,pad_inches=0,bbox_inches=0)
+
+v = np.polyfit(X,Y,9)
+x = np.linspace(0,5.5,55*3+1)
+y = np.polyval(v,x)
+plt.figure(figsize=(6,4))
+plt.scatter(X,Y)
+plt.plot(x,y,'black')
+plt.xlabel('$x$')
+plt.ylabel('$y$')
+plt.xlim(0,5.5)
+plt.ylim(0,11)
+plt.tight_layout()
+plt.savefig('animation_overfitting2',dpi=300,pad_inches=0,bbox_inches=0)
+
+v = np.polyfit(X,Y,1)
+x = np.linspace(0,5.5,55*3+1)
+y = np.polyval(v,x)
+plt.figure(figsize=(6,4))
+plt.scatter(X,Y)
+plt.plot(x,y,'black')
+plt.xlabel('$x$')
+plt.ylabel('$y$')
+plt.xlim(0,5.5)
+plt.ylim(0,11)
+plt.tight_layout()
+plt.savefig('animation_overfitting3',dpi=300,pad_inches=0,bbox_inches=0)

### 2019-04-24 Bibliography/Machine Learning/neuralnetworksanddeeplearning.com.pdf-master/neuralnetworksanddeeplearning.com.pdf-master/scripts/animations.py

@@ -1,0 +1,58 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Aug  3 20:04:02 2018
+
+@author:
+"""
+
+# chapter 3, animation 1
+import math
+def cost(x):    return x**2/2
+def costd(x):   return x*x*(1-x)
+def s(x):       return 1/(1+math.exp(-x))
+def ov(w,b):    return s(w+b)
+x = 1
+w0 = 0.6
+b0 = 0.9
+eta = 0.15
+w = w0
+b = b0
+Y = []
+X = []
+A = []
+W = []
+B = []
+while x <= 300:
+    a = ov(w,b)
+    d = costd(a)
+    w = w - eta * d
+    b = b - eta * d
+    y = cost(a)
+    X.append(x)
+    Y.append(y)
+    W.append(w)
+    B.append(b)
+    A.append(a)
+    x = x + 1
+
+import matplotlib.pyplot as plt
+plt.subplot(2,2,1)
+plt.plot(X,Y)
+plt.xlabel('Epoch')
+plt.legend(['Cost'])
+plt.subplot(2,2,3)
+plt.plot(X,A)
+plt.xlabel('Epoch')
+plt.legend(['Output'])
+plt.subplot(2,2,2)
+plt.plot(X,W)
+plt.xlabel('Epoch')
+plt.legend(['Weight'])
+plt.subplot(2,2,4)
+plt.plot(X,B)
+plt.xlabel('Epoch')
+plt.legend(['Bias'])
+#plt.plot(X,B)
+#plt.plot(X,W)
+
+    

### 2019-04-24 MyCode/BPSK_modulation.py

@@ -85,7 +85,6 @@ y = sign(x)
 # find erroneous symbols
 
 
-
 #plt.semilogy(Eb_No_dB, Pe,'r',Eb_No_dB, BER,'s')
 plt.semilogy(Eb_No_dB, A_BER,'r',linewidth=2)
 plt.semilogy(Eb_No_dB, A_BER,'-s')

### 2019-04-24 MyCode/ML-python-environment.py

@@ -1,0 +1,24 @@
+# scipy
+import scipy
+print('scipy: %s' % scipy.__version__)
+# numpy
+import numpy
+print('numpy: %s' % numpy.__version__)
+# matplotlib
+import matplotlib
+print('matplotlib: %s' % matplotlib.__version__)
+# pandas
+import pandas
+print('pandas: %s' % pandas.__version__)
+# statsmodels
+import statsmodels
+print('statsmodels: %s' % statsmodels.__version__)
+# scikit-learn
+import sklearn
+print('sklearn: %s' % sklearn.__version__)
+# tensorflow
+import tensorflow
+print('tensorflow: %s' % tensorflow.__version__)
+# keras
+import keras
+print('keras: %s' % keras.__version__)

### 2019-04-24 MyCode/my_functions.py

@@ -65,12 +65,18 @@ def PSK2 (b, V):
             
     return um
 
-def pulsesignal(A, d): 
+def pulseSignal(A, d): 
     # d is the discretization of the pulse
     pulse1 = A*ones(int(d/2))
     pulse2 = zeros(int(d/2))
     
     return concatenate((pulse1, pulse2), axis = None)
+    
+def constantSignal(A, d): 
+    # d is the discretization of the pulse
+    pulse1 = A*ones(int(d))
+    return pulse1
+    
     
 def pulses2waveform (N,d,p,uk):
     uwv = empty(N*d)
@@ -91,5 +97,14 @@ def fourriertransform(x, interval, fc, BB):
     
     return concatenate((freq, xhat), axis=0)
     
+    
+def Q(x):
+    return 0.5*erfc(x/sqrt(2))
+
+def MPSK_BER(M,Eb,No):
+    n = log2(8)
+    Es = n*Eb # energy per symbol: nEb where 2^n = M
+    gamma_s = Es/No
+    return 2*Q(sqrt(2*gamma_s)*sin(pi/M))/n
     
     

### 2019-05-05 Rapport/TG_ITA/Bibliographic_study/conference_041818.tex
Paper PIR 2 first pages

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{verbatim} 
\usetikzlibrary{positioning}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Machine Learning Autoencoder Applied to a Binary Symmetric Channel}

\author{\IEEEauthorblockN{Eduardo Dadalto Camara Gomes}
\IEEEauthorblockA{\textit{Department of Electronics, Optronics and Signal Processing (DEOS) } \\
\textit{ISAE-SUPAERO}\\
Toulouse, France \\
Email: edadaltocg@gmail.com}}

\maketitle

\begin{abstract}

write in the end.

\end{abstract}

\begin{IEEEkeywords}
communication system, machine learning, autoencoder, binary symmetric channel, maximum a posteriori (MAP) decoder 
\end{IEEEkeywords}

\section{Introduction}
\subsection{Motivation}
Communication systems are widely available in the everyday life and are built over different structures. These structures are less or more complex; nevertheless, they share a common structure. The simpler real communication architecture defines the system as an exchange of information between two terminals through a noisy channel. As a consequence, we know the ultimate reliable data rate that can be transmitted over a noisy channel, theorized by Shannon \cite{b1} in 1948.

Since then, the research community in digital communication developed a series of algorithms to minimize bit error rate (BER) over a channel. However, the challenge of finding a \textit{efficient} solution , i.e., at same time low latency and with low error probability, for low signal to noise ratio (SNR) channels remains. Based on this vision, Tim O'Shea and Jakob Hoydis \cite{b2} pertinently remarked that traditional algorithms in the field have foundations in probability theory (e.g. maximum a posteriori (MAP) estimation used as a channel decoder), hence they are built on top of mathematically convenient models. Often, these models does not account for all the system's real imperfections, what leads to intrinsic errors, even though they are theoretically optimal.

Accordingly, machine learning (ML) algorithms does not require rigidly designed models and can take effortlessly account for non-linearities. Moreover, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate in an end-to-end manner. As a result, an optimized autoencoder with a stochastic layer that models channel's imperfections can substitute the block based representation. Therefore, ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths and with low decoding latency, resulting ultimately in gain of bandwidth.

Low latency and high bandwidth wireless communication are key to critical systems, such as airplanes, satellites, cellular communication and 5G operations. The latter was studied by F. D. Calabrese et al. in \cite{b3} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance. Thereby, ML algorithms are cardinal for future state of the art communication applications.

\subsection{Background and related work}

State of the art work in machine learning in communication systems
\subsection{Problem statement}

This research intend to implement a ML autoencoder for binary symmetric channels that outperforms the MAP decoder for a range of SNR. Using state-of-the art DNN algorithms to find the best solution of the problem, contributing to set up a higher standard in terms of performance in bit-error correction for digital communication applications. In a near future, this disruptive methodology for error correction using ML could replace mathematically optimal decoders which are the current guideline.

%\subsection{Notation}
%Special notations used throughout the paper

\section{Theoretical Background}

\subsection{Maximum a posteriori decoder}

The concept of a MAP decoder or Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is choosing the message value which maximizes the MAP probability of the corrupted information received \cite{b4}. This algorithm is known for its optimality in error correction for white noise interference. MAP decoding algorithms and it's variants Log-MAP and Max-Log-MAP are widely used as decoders for turbo codes \cite{b7}.

The MAP decoder is practically nonviable to implement \cite{b6}. Thus, for the purpose of implementation, its variants cited in the paragraph above are better suited. The Log-MAP is an optimal decoder, having equivalent performance to the MAP decoder. The Max-Log-MAP is a suboptimal decoder and will not be treated in this paper. Refer to \cite{b6} for specific details in the derivation of the Log-MAP algorithm.

Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a binary symmetric channel (BSC) was studied in \cite{b5}. They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{b6}. It means that if the real SNR is smaller, the decoder will not perform well. Hence, for channel characteristics that change over time, the application of a MAP algorithm is debatable. This conclusion serve as motivation for ML autoencoders, which could adapt to different SNR conditions, outperforming the MAP decoder for realistic applications.

\subsection{Neural network basics}

Neural networks (NN) are able to approximate any kind of function through an architecture composed of several single processing units (neurons) connected together forming a network. They are defined by \cite{b9} as are parallel distributed, learning- and self-organizing information processing systems.

In the context of learning algorithms implemented by a NN, the key issue is to find a suitable architecture that delivers the best results. To define and build this infrastructure, the activation function, the loss function and the structure must be decided.  M. Nielson \cite{b8} in his book explores a vast amount of NN architectures and how they work.

Multi-layer feed forward neural network (MLNN) as shown in Fig. \ref{fig:NN} will be treated in this work. They are known for the universal approximation property \cite{b9} and for the versatility of increasing the number of layers, creating a DNN able to undertake complex classification problems with low classification error. In more general terms, a MLNN with $L$ layers and parameter $\theta$ is a mapping of an input $\textbf{r}_0 \in \mathbb{R}^{N_0}$ to an output $\textbf{r}_L \in \mathbb{R}^{N_L}$ through $L$ iterative steps. For a fully-connected network, each layer vector is calculated in a forward propagation iterative manner, where each neuron is composed of a linear combinator and an activation function  defined as

\begin{equation}\label{eq:eqFP}
	f_{l}\left( \textbf{r}_{l-1};\theta _{l}\right) = \sigma \left( \textbf{W}_{l}\textbf{r}_{l-1}+\textbf{b}_{l}\right)
\end{equation}
where $\textbf{W}_{l}\in \mathbb{R} ^{N_{l}\times N_{l-1}}$ is the weight matrix between layers $l-1$ and $l$,  $\textbf{b}_{l}\in \mathbb{R} ^{N_l}$ is the bias vector and $\sigma$ is the activation function \cite{b2}.

In order to correct the errors from a BSC, a \textit{perceptron} processing unit is chosen to solve a classification problem for each input message. A perceptron maps a binary vector input into a single binary output digit \cite{b8}. Mathematically, it means that the function defined in \eqref{eq:eqFP}, outputs $1$ or $0$, depending on a threshold value.

For a classification problem in communication, the \textit{categorical cross-entropy} loss function $l(\textbf{p},\textbf{q}):\mathbb{R} ^{N_L}\times \mathbb{R} ^{N_L}\mapsto \mathbb{R}$ defined in \eqref{eq:c-e} is most common. It is derived from the log-likelihood of a training set where $q_j$ is the estimated probability of the outcome $j$ and $p_{j}$ is the true probability. Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{b10} where $j=1,...,N_L$.
\begin{equation}\label{eq:c-e}
	l(\textbf{p},\textbf{q})=-\sum _{j}p_{j}\log \left( q_{j}\right)
\end{equation}

With all these elements set, training the neural network to calculate an accurate weight matrix $\textbf{W}$ for forward propagation requires a labeled training data set. This set is composed of pairs $ (\textbf{r}_{0,i}, \textbf{p}_i) $, where $i=1,...,S$ and $S$ is the number of training sets. It matches the input and its respective desired output. The objective is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{b2}. 
\begin{equation}\label{eq:o-c-e}
L\left( \theta \right) =\dfrac {1}{S}\sum ^{S}_{i=1}l\left( \textbf{p}_{L,i},\textbf{r}_{L,i}\right)	
\end{equation}

This problem is an optimization problem and can be solved through different algorithms. For NN, an efficient way of computing this minimization is implementing the back-propagation (BP) algorithm. 	BP is classified as a supervised learning algorithm that is able to optimize a function based on some parameter and is highly parallelizable in computational terms \cite{b8}\cite{b9}.

Libraries \cite{b11} \cite{b12}.

\tikzset{%
  every neuron/.style={
    circle,
    draw,
    minimum size=0.666cm
  },
  neuron missing/.style={
    draw=none, 
    scale=2,
    text height=0.333cm,
    execute at begin node=\color{black}$\vdots$
  },
}

\begin{figure}
\centering

\begin{tikzpicture}[x=1.1cm, y=1.1cm, >=stealth]

\foreach \m/\l [count=\y] in {1,2,3,missing,4}
  \node [every neuron/.try, neuron \m/.try] (input-\m) at (0,2.5-\y) {};

\foreach \m [count=\y] in {1,missing,2}
  \node [every neuron/.try, neuron \m/.try ] (hidden-\m) at (2,2-\y*1.25) {};

\foreach \m [count=\y] in {1,missing,2}
  \node [every neuron/.try, neuron \m/.try ] (output-\m) at (4,1.5-\y) {};

\foreach \l [count=\i] in {1,2,3,m}
  \draw [<-] (input-\i) -- ++(-1,0)
    node [above, midway] {$x_\l$};

\foreach \l [count=\i] in {1,j}
  \node [above] at (hidden-\i.north) {$r_\l$};

\foreach \l [count=\i] in {1,n}
  \draw [->] (output-\i) -- ++(1,0)
    node [above, midway] {$y_\l$};

\foreach \i in {1,...,4}
  \foreach \j in {1,...,2}
    \draw [->] (input-\i) -- (hidden-\j);

\foreach \i in {1,...,2}
  \foreach \j in {1,...,2}
    \draw [->] (hidden-\i) -- (output-\j);

\foreach \l [count=\x from 0] in {Input, Hidden, Ouput}
  \node [align=center, above] at (\x*2,2) {\l \\ layer};

\end{tikzpicture}

\caption{MLNN representative diagram. Where } \label{fig:NN}
\end{figure}

\subsection{Autoencoders}

%\section{Procedure and methods}
%Explain what I Should do to achieve my goal
\begin{comment}

\subsection{Figures and Tables}

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}
\end{comment}

\begin{thebibliography}{00}
\bibitem{b1} C. E. Shannon, "A mathematical theory of communication," Bell System Technical Journal, vol. 27, no. 3, pp. 379–423, Jul. 1948.

\bibitem{b2} O'Shea, Tim \& Hoydis, Jakob, "An Introduction to Machine Learning Communications Systems", 2017.
\bibitem{b3} F. D. Calabrese, L. Wang, E. Ghadimi, G. Peters, and P. Soldati,
"Learning radio resource management in 5G networks: Framework,
opportunities and challenges," arXiv preprint arXiv:1611.10253, 2016
\bibitem{b4} W. Alexander, H. Peter \& W. Norbert, "Turbo-Decoding Without SNR Estimation",IEEE Communications Letter, vol. 4, no. 6, pp. 193-195, 2000.
\bibitem{b5} M. Jordan and R. Nichols, “The effects of channel characteristics on turbo code performance,” in Proc. Milcom’96, McLean, VA, Oct. 1996,
pp. 17–21.
\bibitem{b6} Robertson, P. , Hoeher, P. and Villebrun, E. (1997), Optimal and sub‐optimal maximum a posteriori algorithms suitable for turbo decoding. Eur. Trans. Telecomm., 8: 119-125. doi:10.1002/ett.4460080202.
\bibitem{b7} J. Hagenauer,P. Robertson, L. Papke: iterative (“Turbo”) decoding of systematic convolutional codes with the MAP and SOVA
algorithms. In: ITG-Fachbericht 130. October 1994, p. 21-29.
\bibitem{b8} M. Nielsen, "Neural Networks and Deep Learning".
\bibitem{b9} I. Mohamed, "Applications of neural networks to digital communications - a survey", 1997.
\bibitem{b10} M. Kevin, "Machine Learning: A Probabilistic Perspective", MIT, ISBN 978-0262018029, 2012.
\bibitem{b11} M. Abadi, A. Agarwal, P. Barham, et al., 
"TensorFlow: Large-scale machine learning on heterogeneous systems", 2015. Software available from tensorflow.org.
\bibitem{b12} F. Chollet, “keras,” https://github.com/fchollet/keras, 2015.

\end{thebibliography}

\end{document}

### 2019-05-08 Rapport/TG_ITA/Bibliographic_study/conference_041818.tex
2 page paper

@@
 \usepackage{verbatim}
+
 \usetikzlibrary{positioning}
@@
 
-\title{Machine Learning Autoencoder Applied to a Binary Symmetric Channel}
+\title{Machine Learning Autoencoder Applied to Communication Channels}
 
@@
 \begin{IEEEkeywords}
-communication system, machine learning, autoencoder, binary symmetric channel, maximum a posteriori (MAP) decoder 
+communication system, machine learning, autoencoder,channel decoding, maximum a posteriori (MAP) decoder 
 \end{IEEEkeywords}
@@
 \subsection{Motivation}
-Communication systems are widely available in the everyday life and are built over different structures.
-These structures are less or more complex; nevertheless, they share a common structure.
+Communication systems are widely present in  everyday life and have different structures.
+These designs are less or more complex; nevertheless, they share a common core.
 The simpler real communication architecture defines the system as an exchange of information between two terminals through a noisy channel.
-As a consequence, we know the ultimate reliable data rate that can be transmitted over a noisy channel, theorized by Shannon \cite{b1} in 1948.
+As a result of imperfections, Shannon theorized \cite{b1} in 1948 the ultimate reliable data rate that can be transmitted through a communication systems of this kind.
 
 Since then, the research community in digital communication developed a series of algorithms to minimize bit error rate (BER) over a channel.
-However, the challenge of finding a \textit{efficient} solution , i.e., at same time low latency and with low error probability, for low signal to noise ratio (SNR) channels remains.
-Based on this vision, Tim O'Shea and Jakob Hoydis \cite{b2} pertinently remarked that traditional algorithms in the field have foundations in probability theory (e.g.
-maximum a posteriori (MAP) estimation used as a channel decoder), hence they are built on top of mathematically convenient models.
-Often, these models does not account for all the system's real imperfections, what leads to intrinsic errors, even though they are theoretically optimal.
+However, the challenge of finding an \textit{efficient} solution  (i.e., at same time low latency and with low error probability) for low signal to noise ratio (SNR) channels, remains.
+Based on this vision, Tim O'Shea and Jakob Hoydis \cite{b2} pertinently noted that traditional algorithms in the field have foundations in probability theory (e.g.
+maximum a posteriori (MAP) estimation used as a channel decoder), hence they are usually built on top of mathematically convenient models.
+Often, these models does not account for all the real system's imperfections, what leads to intrinsic errors in practice, even though they are theoretically optimal error corrector machines.
 
 Accordingly, machine learning (ML) algorithms does not require rigidly designed models and can take effortlessly account for non-linearities.
-Moreover, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate in an end-to-end manner.
+These characteristics suits them well for being used as bit error correctors.
+Moreover, with ML based channel encoder and channel decoder, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate in an end-to-end optimization of the system.
 As a result, an optimized autoencoder with a stochastic layer that models channel's imperfections can substitute the block based representation.
-Therefore, ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths and with low decoding latency, resulting ultimately in gain of bandwidth.
+Therefore, ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths and with low decoding latency, resulting ultimately in gain of bandwidth over standard methods.
 
@@
 Calabrese et al.
-in \cite{b3} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance.
+in \cite{b3} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance of a network.
 Thereby, ML algorithms are cardinal for future state of the art communication applications.
 
-\subsection{Background and related work}
+\subsection{Related work}
 
-State of the art work in machine learning in communication systems
+Recently, a significant amount of work in radio communication theory has emerged, introducing ML elements to the communication system.
+O'Shea et al.
+in \cite{osheaautoencoder} developed a channel autoencoder with optimized impairment and regularization layers to emulate channel impairments.
+They studied this architecture over a BSC and other channels, founding ``some promising initial capacity" for this schema.
+In their research, results in terms of BER over SNR of a DNN based autoencoder and a convolutional neural network (CNN) based autoencoder were treated.
+They used a range of SNR from $-10\text{dB}$ to $15\text{dB}$ with QPSK and QAM16 theoretical capacity as benchmarks.
+This analysis was conducted for a binary input message.
+
+T.
+Gruber et.
+al.
+in a 2017 paper \cite{2018} proved that a deep learning-based channel decoder could actually learn a decoding algorithm rather than just being a simple classifier.
+They introduced codewords that were not been used in the training set, and the trained NN was able to correctly decode it.
+They also observed that structured codes are easier to learn than unstructured ones.
+NN for structured codes are able to generalize to the full codebook even if they have not seen all the training examples.
+They trained the NN for very short blocklengths ($N \leq 64$) in order to compare with MAP decoding performance.
+
 \subsection{Problem statement}
 
-This research intend to implement a ML autoencoder for binary symmetric channels that outperforms the MAP decoder for a range of SNR.
-Using state-of-the art DNN algorithms to find the best solution of the problem, contributing to set up a higher standard in terms of performance in bit-error correction for digital communication applications.
+This research will implement a ML autoencoder for BSC that outperforms the MAP decoder for a range of SNR.
+In its most simple form, a channel autoencoder includes an encoder, a noisy channel and an decoder.
+Using state-of-the art DNN algorithms to find the best solution of the problem, this work hopes to contribe to set up a higher standard in terms of performance in bit-error correction and reduced delay for digital communication applications.
 In a near future, this disruptive methodology for error correction using ML could replace mathematically optimal decoders which are the current guideline.
@@
 
-The concept of a MAP decoder or Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is choosing the message value which maximizes the MAP probability of the corrupted information received \cite{b4}.
+The concept of a MAP decoder or Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is choosing a message which maximizes the MAP probability of the corrupted information received  by the decoder \cite{b4}.
 This algorithm is known for its optimality in error correction for white noise interference.
-MAP decoding algorithms and it's variants Log-MAP and Max-Log-MAP are widely used as decoders for turbo codes \cite{b7}.
+MAP decoding algorithms and it's variants: Log-MAP and Max-Log-MAP, are widely used as decoders for turbo codes \cite{b7}.
 
@@
 
-Neural networks (NN) are able to approximate any kind of function through an architecture composed of several single processing units (neurons) connected together forming a network.
-They are defined by \cite{b9} as are parallel distributed, learning- and self-organizing information processing systems.
-
-In the context of learning algorithms implemented by a NN, the key issue is to find a suitable architecture that delivers the best results.
-To define and build this infrastructure, the activation function, the loss function and the structure must be decided.
-M.
-Nielson \cite{b8} in his book explores a vast amount of NN architectures and how they work.
-
-Multi-layer feed forward neural network (MLNN) as shown in Fig.
-\ref{fig:NN} will be treated in this work.
-They are known for the universal approximation property \cite{b9} and for the versatility of increasing the number of layers, creating a DNN able to undertake complex classification problems with low classification error.
-In more general terms, a MLNN with $L$ layers and parameter $\theta$ is a mapping of an input $\textbf{r}_0 \in \mathbb{R}^{N_0}$ to an output $\textbf{r}_L \in \mathbb{R}^{N_L}$ through $L$ iterative steps.
-For a fully-connected network, each layer vector is calculated in a forward propagation iterative manner, where each neuron is composed of a linear combinator and an activation function  defined as
-
-\begin{equation}\label{eq:eqFP}
-	f_{l}\left( \textbf{r}_{l-1};\theta _{l}\right) = \sigma \left( \textbf{W}_{l}\textbf{r}_{l-1}+\textbf{b}_{l}\right)
-\end{equation}
-where $\textbf{W}_{l}\in \mathbb{R} ^{N_{l}\times N_{l-1}}$ is the weight matrix between layers $l-1$ and $l$,  $\textbf{b}_{l}\in \mathbb{R} ^{N_l}$ is the bias vector and $\sigma$ is the activation function \cite{b2}.
-
-In order to correct the errors from a BSC, a \textit{perceptron} processing unit is chosen to solve a classification problem for each input message.
-A perceptron maps a binary vector input into a single binary output digit \cite{b8}.
-Mathematically, it means that the function defined in \eqref{eq:eqFP}, outputs $1$ or $0$, depending on a threshold value.
-
-For a classification problem in communication, the \textit{categorical cross-entropy} loss function $l(\textbf{p},\textbf{q}):\mathbb{R} ^{N_L}\times \mathbb{R} ^{N_L}\mapsto \mathbb{R}$ defined in \eqref{eq:c-e} is most common.
-It is derived from the log-likelihood of a training set where $q_j$ is the estimated probability of the outcome $j$ and $p_{j}$ is the true probability.
-Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{b10} where $j=1,...,N_L$.
-\begin{equation}\label{eq:c-e}
-	l(\textbf{p},\textbf{q})=-\sum _{j}p_{j}\log \left( q_{j}\right)
-\end{equation}
-
-With all these elements set, training the neural network to calculate an accurate weight matrix $\textbf{W}$ for forward propagation requires a labeled training data set.
-This set is composed of pairs $ (\textbf{r}_{0,i}, \textbf{p}_i) $, where $i=1,...,S$ and $S$ is the number of training sets.
-It matches the input and its respective desired output.
-The objective is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{b2}.
-\begin{equation}\label{eq:o-c-e}
-L\left( \theta \right) =\dfrac {1}{S}\sum ^{S}_{i=1}l\left( \textbf{p}_{L,i},\textbf{r}_{L,i}\right)	
-\end{equation}
-
-This problem is an optimization problem and can be solved through different algorithms.
-For NN, an efficient way of computing this minimization is implementing the back-propagation (BP) algorithm.
-	BP is classified as a supervised learning algorithm that is able to optimize a function based on some parameter and is highly parallelizable in computational terms \cite{b8}\cite{b9}.
-
-Libraries \cite{b11} \cite{b12}.
-
 \tikzset{%
@@
 
-\begin{figure}
+\begin{figure}[!hbp]
 \centering
@@
 
-\caption{MLNN representative diagram. Where } \label{fig:NN}
+\caption{MLNN representative diagram. Where \textbf{x} is the input vector of the NN, $\textbf{r}_{l}$ is a hidden layer vector
+and \textbf{y} is the output vector.} \label{fig:NN}
 \end{figure}
 
+Neural networks (NN) are a set of connected \textit{neurons} able to approximate any kind of function through an architecture composed of several single processing units (neurons) connected together forming a network.
+They are defined by \cite{b9} as parallel distributed, learning- and self-organizing information processing systems.
+
+In the context of learning algorithms implemented by a NN, the key issue is to find a suitable architecture that delivers the best results.
+To define and build this infrastructure, the activation function, the loss function and the structure must be decided.
+M.
+Nielson \cite{b8} in his book explores a vast amount of NN architectures and how they work.
+
+Multi-layer feed forward neural network (MLNN) as shown in Fig.
+\ref{fig:NN} will be applied in this work.
+They are known for the universal approximation property \cite{b9} and for the versatility of increasing the number of layers, creating a DNN able to undertake complex classification problems with low classification error and low dimensionality.
+In more general terms, a MLNN with $L$ layers, also called \textit{depth}, and parameter $\theta$ is a mapping of an input $\textbf{r}_0 \in \mathbb{R}^{N_0}$ to an output $\textbf{r}_L \in \mathbb{R}^{N_L}$ through $L$ iterative steps.
+For a fully-connected network, each layer vector is calculated in an iterative forward propagation, where each neuron is composed of a linear combinator and an activation function defined as
+
+\begin{equation}\label{eq:eqFP}
+	f_{l}\left( \textbf{r}_{l-1};\theta _{l}\right) = \sigma \left( \textbf{W}_{l}\textbf{r}_{l-1}+\textbf{b}_{l}\right)
+\end{equation}
+where $\textbf{W}_{l}\in \mathbb{R} ^{N_{l}\times N_{l-1}}$ is the weight matrix between layers $l-1$ and $l$,  $\textbf{b}_{l}\in \mathbb{R} ^{N_l}$ is the bias vector and $\sigma$ is the activation function \cite{b2}.
+
+In order to accurately classify an input, a \textit{perceptron} processing unit is chosen to solve a classification problem for each input message.
+A perceptron maps a binary vector input into a single binary output digit \cite{b8}.
+Mathematically, it means that the function defined in \eqref{eq:eqFP}, outputs $1$ or $0$, depending on a threshold value.
+
+For a classification problem in communication, the \textit{categorical cross-entropy} loss function $l(\textbf{p},\textbf{q}):\mathbb{R} ^{N_L}\times \mathbb{R} ^{N_L}\mapsto \mathbb{R}$ defined in \eqref{eq:c-e} is most common.
+It is derived from the log-likelihood of a training set where $q_j$ is the estimated probability of the outcome $j$ and $p_{j}$ is the true probability.
+Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{b10} where $j=1,...,N_L$.
+\begin{equation}\label{eq:c-e}
+	l(\textbf{p},\textbf{q})=-\sum _{j}p_{j}\log \left( q_{j}\right)
+\end{equation}
+
+With all these elements set, training the neural network to calculate an accurate weight matrix $\textbf{W}$ for forward propagation requires a labeled \textit{training data set}.
+This set is composed of pairs $ (\textbf{r}_{0,i}, \textbf{p}_i) $, where $i=1,...,S$ and $S$ is the number of training sets.
+It matches the input and its respective desired output.
+The objective is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{b2}.
+\begin{equation}\label{eq:o-c-e}
+L\left( \theta \right) =\dfrac {1}{S}\sum ^{S}_{i=1}l\left( \textbf{p}_{L,i},\textbf{r}_{L,i}\right)	
+\end{equation}
+
+This problem is an optimization problem and can be solved through different algorithms.
+For NN, an efficient way of computing this minimization is implementing the back-propagation (BP) algorithm.
+	BP is classified as a supervised learning algorithm that is able to optimize a function based on some parameter and is highly parallelizable in computational terms \cite{b8}\cite{b9}.
+
+The great advantage of NN is its ability do generalize the training set, delivering the right results even for data not contained in the training set.
+With the weight matrix trained, the NN find the correct putputs for unseen inputs .
+In order to validate this principle, a \textit{validation data set} is used and the loss is mesured.
+Based on this value, we can evaluate the NN overall accuracy.
+
+For implementing all the algorithms, Keras, a high-level python API for ML, will be used together with TensorFlow 1.13.1, a ML python friendly open-source library \cite{b11} \cite{b12}.
+Both are available for free.
+
 \subsection{Autoencoders}
+TO PARAPHRASE: The parallel computing
+capabilities of NNs allow us to solve or, at least, approximate
+the MLD problem in polynomial time [7]. Moreover, the
+weights of the NN are precomputed during training and the
+decoding step itself is then relatively simple.
+
+NNs which can learn an appropriate
+mapping between noisy input patterns and codewords. No
+assumption has to be made about the statistics of the channel
+noise because the NN is able to learn the mapping or to
+extract the channel statistics during the learning process
+
+decoding problems have far more possibilities than conventional pattern recognition problems
+
+In 2006, a new training technique, called layer-by-layer
+unsupervised pre-training followed by gradient descent finetuning [14], led to the renaissance of NNs because it made
+training of NNs with more layers feasible.
+
+new hardware such as graphical processing units (GPUs) are available
+to speed up learning as well as inference. In this renaissance
+of NNs, new NND ideas emerge
+
+The curse of dimensionality
 
@@
 \begin{thebibliography}{00}
-\bibitem{b1} C. E. Shannon, "A mathematical theory of communication," Bell System Technical Journal, vol. 27, no. 3, pp. 379–423, Jul. 1948.
+\bibitem{b1} C. E. Shannon, ``A mathematical theory of communication," Bell System Technical Journal, vol. 27, no. 3, pp. 379–423, Jul. 1948.
 
-\bibitem{b2} O'Shea, Tim \& Hoydis, Jakob, "An Introduction to Machine Learning Communications Systems", 2017.
-\bibitem{b3} F. D. Calabrese, L. Wang, E. Ghadimi, G. Peters, and P. Soldati,
-"Learning radio resource management in 5G networks: Framework,
+\bibitem{b2} O'Shea, Tim \& Hoydis, Jakob, ``An Introduction to Machine Learning Communications Systems", 2017.
+\bibitem{b3} F. D. Calabrese, L. Wang, E. Ghadimi, G. Peters, and P. Soldati, ``Learning radio resource management in 5G networks: Framework,
 opportunities and challenges," arXiv preprint arXiv:1611.10253, 2016
-\bibitem{b4} W. Alexander, H. Peter \& W. Norbert, "Turbo-Decoding Without SNR Estimation",IEEE Communications Letter, vol. 4, no. 6, pp. 193-195, 2000.
-\bibitem{b5} M. Jordan and R. Nichols, “The effects of channel characteristics on turbo code performance,” in Proc. Milcom’96, McLean, VA, Oct. 1996,
-pp. 17–21.
+\bibitem{b4} W. Alexander, H. Peter \& W. Norbert, ``Turbo-Decoding Without SNR Estimation",IEEE Communications Letter, vol. 4, no. 6, pp. 193-195, 2000.
+\bibitem{b5} M. Jordan and R. Nichols, “The effects of channel characteristics on turbo code performance,” in Proc. Milcom’96, McLean, VA, Oct. 1996, pp. 17–21.
 \bibitem{b6} Robertson, P.
@@
 algorithms. In: ITG-Fachbericht 130. October 1994, p. 21-29.
-\bibitem{b8} M. Nielsen, "Neural Networks and Deep Learning".
-\bibitem{b9} I. Mohamed, "Applications of neural networks to digital communications - a survey", 1997.
-\bibitem{b10} M. Kevin, "Machine Learning: A Probabilistic Perspective", MIT, ISBN 978-0262018029, 2012.
+\bibitem{b8} M. Nielsen, \textit{Neural Networks and Deep Learning}.
+\bibitem{b9} I. Mohamed, ``Applications of neural networks to digital communications - a survey", 1997.
+\bibitem{b10} M. Kevin, ``Machine Learning: A Probabilistic Perspective", MIT, ISBN 978-0262018029, 2012.
 \bibitem{b11} M. Abadi, A. Agarwal, P. Barham, et al., 
-"TensorFlow: Large-scale machine learning on heterogeneous systems", 2015. Software available from tensorflow.org.
+``TensorFlow: Large-scale machine learning on heterogeneous systems", 2015. Software available from tensorflow.org.
 \bibitem{b12} F. Chollet, “keras,” https://github.com/fchollet/keras, 2015.
+\bibitem{osheaautoencoder} T. O'Shea, K. arra, T. C. Clancy, ``Learning to Communicate: Channel Auto-encoders, Domain Specific Regularizers, and Attention", 2016.
+\bibitem{2018} T. Gruber, S. Cammerer, J. Hoydis, and S. ten Brink, “On deep learning based channel decoding,” accepted for CISS 2017, arXiv preprint arXiv:1701.07738, 2017.
 

### 2019-05-08 MyCode/ML-python-environment.py

@@ -16,9 +16,10 @@ print('statsmodels: %s' % statsmodels.__version__)
 # scikit-learn
 import sklearn
 print('sklearn: %s' % sklearn.__version__)
-# tensorflow
-import tensorflow
-print('tensorflow: %s' % tensorflow.__version__)
 # keras
 import keras
 print('keras: %s' % keras.__version__)
+# tensorflow
+#import tensorflow
+#print('tensorflow: %s' % tensorflow.__version__)
+

### 2019-05-08 Rapport/TG_ITA/Bibliographic_study/conference_041818.tex
update bibliography review

@@
 
-write in the end.
+Communication channel error correction is key to enable digital critical digital communication systems to works efficiently.
+Although noisy channels has already a mathematically proven optimal solution for error error rate minimization: the maximum a posteriori (MAP) decoder; it`s implementation cripples  system`s applicability, since it introduces a large delay for large codewords.
+In this context, a deep neural network (DNN) is proposed to optimize the channel with an end-to-end autoencoder.
+The DNN autoencoder outperforms MAP decoder in terms of delay, because of its \textit{one-shot} capability when it is trained.
+Moreover, it has similar BER compared to MAP for a range of signal to noise ratio (SNR).
 
@@
 \begin{IEEEkeywords}
-communication system, machine learning, autoencoder,channel decoding, maximum a posteriori (MAP) decoder 
+communication system, machine learning,  autoencoder, channel decoding, maximum a posteriori (MAP) decoder 
 \end{IEEEkeywords}
@@
 \subsection{Motivation}
-Communication systems are widely present in  everyday life and have different structures.
+Communication systems are present in everyday life and have different forms.
 These designs are less or more complex; nevertheless, they share a common core.
-The simpler real communication architecture defines the system as an exchange of information between two terminals through a noisy channel.
-As a result of imperfections, Shannon theorized \cite{b1} in 1948 the ultimate reliable data rate that can be transmitted through a communication systems of this kind.
+The simpler architecture defines the system as an exchange of information between two terminals through a noisy channel.
+As a result of channel imperfections, Shannon theorized \cite{b1} in 1948 the ultimate reliable data rate that can be transmitted through a communication system.
 
@@
 Based on this vision, Tim O'Shea and Jakob Hoydis \cite{b2} pertinently noted that traditional algorithms in the field have foundations in probability theory (e.g.
-maximum a posteriori (MAP) estimation used as a channel decoder), hence they are usually built on top of mathematically convenient models.
-Often, these models does not account for all the real system's imperfections, what leads to intrinsic errors in practice, even though they are theoretically optimal error corrector machines.
+maximum a posteriori (MAP) estimation used as a channel decoder).
+Hence, they are usually built on top of mathematically convenient models.
+Even though they are theoretically optimal error corrector machines, these models oftenly does not account for all the real system's imperfections, what leads to errors when in practice.
 
-Accordingly, machine learning (ML) algorithms does not require rigidly designed models and can take effortlessly account for non-linearities.
-These characteristics suits them well for being used as bit error correctors.
-Moreover, with ML based channel encoder and channel decoder, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate in an end-to-end optimization of the system.
+Accordingly, machine learning (ML) algorithms does not require rigidly designed models and can take non-linearities effortlessly into account.
+These characteristics make these algorithms candidates for being used as bit error corrector.
+Moreover, with ML based channel encoder and channel decoder, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate end-to-end in the system.
 As a result, an optimized autoencoder with a stochastic layer that models channel's imperfections can substitute the block based representation.
@@
 Calabrese et al.
-in \cite{b3} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance of a network.
-Thereby, ML algorithms are cardinal for future state of the art communication applications.
+in \cite{b3} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance of the network.
+Thereby, ML algorithms are cardinal for state of the art communication applications.
 
@@
 in \cite{osheaautoencoder} developed a channel autoencoder with optimized impairment and regularization layers to emulate channel impairments.
-They studied this architecture over a BSC and other channels, founding ``some promising initial capacity" for this schema.
-In their research, results in terms of BER over SNR of a DNN based autoencoder and a convolutional neural network (CNN) based autoencoder were treated.
-They used a range of SNR from $-10\text{dB}$ to $15\text{dB}$ with QPSK and QAM16 theoretical capacity as benchmarks.
+They studied this architecture over a binary symmetric channel (BSC) and others, founding ``some promising initial capacity" for this schema.
+In their research, results in terms of BER over SNR for a DNN based autoencoder and for a convolutional neural network (CNN) based autoencoder were treated.
+They used a range of SNR - from $-10 \text{dB}$ to $15 \text{dB}$ - with QPSK and QAM16 modulation as benchmarks.
 This analysis was conducted for a binary input message.
@@
 
-This research will implement a ML autoencoder for BSC that outperforms the MAP decoder for a range of SNR.
+This research will implement a ML autoencoder for a  BSC that performs similar to the MAP decoder in real applications for a range of SNR from $-10 \text{dB}$ to $10 \text{dB}$.
 In its most simple form, a channel autoencoder includes an encoder, a noisy channel and an decoder.
@@
 The concept of a MAP decoder or Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is choosing a message which maximizes the MAP probability of the corrupted information received  by the decoder \cite{b4}.
-This algorithm is known for its optimality in error correction for white noise interference.
-MAP decoding algorithms and it's variants: Log-MAP and Max-Log-MAP, are widely used as decoders for turbo codes \cite{b7}.
+This algorithm is known for its optimality in error correction for white noise interference and it's variants: Log-MAP and Max-Log-MAP, are widely used as decoders for turbo codes \cite{b7}.
 
 The MAP decoder is practically nonviable to implement \cite{b6}.
-Thus, for the purpose of implementation, its variants cited in the paragraph above are better suited.
+Thus, its variants cited in the paragraph above are better suited for practical cases.
 The Log-MAP is an optimal decoder, having equivalent performance to the MAP decoder.
-The Max-Log-MAP is a suboptimal decoder and will not be treated in this paper.
+While the Max-Log-MAP is a suboptimal decoder and will not be treated in this paper.
 Refer to \cite{b6} for specific details in the derivation of the Log-MAP algorithm.
 
-Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a binary symmetric channel (BSC) was studied in \cite{b5}.
+Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a BSC was studied in \cite{b5}.
 They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{b6}.
@@
 In more general terms, a MLNN with $L$ layers, also called \textit{depth}, and parameter $\theta$ is a mapping of an input $\textbf{r}_0 \in \mathbb{R}^{N_0}$ to an output $\textbf{r}_L \in \mathbb{R}^{N_L}$ through $L$ iterative steps.
-For a fully-connected network, each layer vector is calculated in an iterative forward propagation, where each neuron is composed of a linear combinator and an activation function defined as
+For a fully-connected network, each layer vector is calculated iteratively in the forward propagation.
+Each neuron is composed of a linear combinator and an activation function defined as
 
@@
 
-The great advantage of NN is its ability do generalize the training set, delivering the right results even for data not contained in the training set.
-With the weight matrix trained, the NN find the correct putputs for unseen inputs .
+One of the greatest advantage of NN is its ability do generalize the training set, delivering the right results even for input data not contained in the training set.
+With the weight matrix trained, the NN find the correct output values for unseen inputs.
 In order to validate this principle, a \textit{validation data set} is used and the loss is mesured.
@@
 
+\begin{comment}
 \subsection{Autoencoders}
@@
 
+\end{comment}
 %\section{Procedure and methods}

### 2019-05-13 Rapport/TG_ITA/Bibliographic_study/conference_041818.tex
Bibliographic correction + Hamming decoding

@@
 
-\author{\IEEEauthorblockN{Eduardo Dadalto Camara Gomes}
-\IEEEauthorblockA{\textit{Department of Electronics, Optronics and Signal Processing (DEOS) } \\
-\textit{ISAE-SUPAERO}\\
-Toulouse, France \\
-Email: edadaltocg@gmail.com}}
+\author{\IEEEauthorblockN{Eduardo Dadalto Camara Gomes\IEEEauthorrefmark{1}}
+\IEEEauthorblockA{\IEEEauthorrefmark{1}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
+Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr}
+}
 
@@
 
-Communication channel error correction is key to enable digital critical digital communication systems to works efficiently.
-Although noisy channels has already a mathematically proven optimal solution for error error rate minimization: the maximum a posteriori (MAP) decoder; it`s implementation cripples  system`s applicability, since it introduces a large delay for large codewords.
+Communication channel error correction is key to enable digital critical communication systems to work efficiently.
+Although the optimal decoder for noisy channels, the maximum a posteriori (MAP) decoder, has already a mathematically proven optimal error probability minimization; its implementation cripples system applicability, since it introduces a large delay for large codewords.
 In this context, a deep neural network (DNN) is proposed to optimize the channel with an end-to-end autoencoder.
-The DNN autoencoder outperforms MAP decoder in terms of delay, because of its \textit{one-shot} capability when it is trained.
-Moreover, it has similar BER compared to MAP for a range of signal to noise ratio (SNR).
+The DNN autoencoder outperforms a MAP decoder form a given encoder in terms of delay, because of the first`s \textit{one-shot} capability once trained.
 
@@
 \subsection{Motivation}
-Communication systems are present in everyday life and have different forms.
-These designs are less or more complex; nevertheless, they share a common core.
-The simpler architecture defines the system as an exchange of information between two terminals through a noisy channel.
-As a result of channel imperfections, Shannon theorized \cite{b1} in 1948 the ultimate reliable data rate that can be transmitted through a communication system.
+A point to point communication channel is a system in which  two terminals exchange information through a noisy channel as Fig.
+\ref{fig:cs} illustrates.
+As a result of channel imperfections, Shannon theorized in \cite{b1} the ultimate reliable data rate that can be transmitted through a communication system with an arbitrarily small probability error.
 
-Since then, the research community in digital communication developed a series of algorithms to minimize bit error rate (BER) over a channel.
-However, the challenge of finding an \textit{efficient} solution  (i.e., at same time low latency and with low error probability) for low signal to noise ratio (SNR) channels, remains.
-Based on this vision, Tim O'Shea and Jakob Hoydis \cite{b2} pertinently noted that traditional algorithms in the field have foundations in probability theory (e.g.
-maximum a posteriori (MAP) estimation used as a channel decoder).
-Hence, they are usually built on top of mathematically convenient models.
-Even though they are theoretically optimal error corrector machines, these models oftenly does not account for all the real system's imperfections, what leads to errors when in practice.
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{simple_sys}
+    \caption{Diagram of a simplified communication system.}\label{fig:cs}
+\end{figure}
 
-Accordingly, machine learning (ML) algorithms does not require rigidly designed models and can take non-linearities effortlessly into account.
-These characteristics make these algorithms candidates for being used as bit error corrector.
-Moreover, with ML based channel encoder and channel decoder, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate end-to-end in the system.
-As a result, an optimized autoencoder with a stochastic layer that models channel's imperfections can substitute the block based representation.
-Therefore, ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths and with low decoding latency, resulting ultimately in gain of bandwidth over standard methods.
-
+Since then, the research community in digital communication developed a series of algorithms to minimize error probability over a channel.
+In practice, the bit error rate (BER), an approximation of the probability of error, is targeted.
+However, the challenge of finding an \textit{efficient} solution  i.e., with both low latency and low error probability, for low signal to noise ratio (SNR) channels, remains.
 Low latency and high bandwidth wireless communication are key to critical systems, such as airplanes, satellites, cellular communication and 5G operations.
@@
 in \cite{b3} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance of the network.
+
+Based on this vision, Tim O'Shea and Jakob Hoydis \cite{b2} pertinently noted that traditional algorithms in the field have foundations in probability theory e.g.
+maximum a posteriori (MAP) rule, maximum likelihood decoder (MLD) and turbo codes.
+Hence, they are usually built on top of mathematically convenient models.
+Even though they are theoretically optimal error corrector codes, these models often do not account for all the real system's imperfections, which leads to errors when implemented in practice.
+
+As opposite to structured algorithms, machine learning (ML) algorithms do not require rigidly designed models and can take non-linearities effortlessly into account.
+These characteristics make these algorithms candidates for being used as channel decoder.
+Moreover, with ML based channel encoder and channel decoder, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate end-to-end in the system.
+As a result, an optimized autoencoder with a stochastic layer that models channel's imperfections can substitute the block based representation.
+Therefore, ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths and with low decoding latency, resulting ultimately in gain of bandwidth over standard methods.
 Thereby, ML algorithms are cardinal for state of the art communication applications.
@@
 in \cite{osheaautoencoder} developed a channel autoencoder with optimized impairment and regularization layers to emulate channel impairments.
-They studied this architecture over a binary symmetric channel (BSC) and others, founding ``some promising initial capacity" for this schema.
+They studied this architecture over an additive white Gaussian noise (AWGN) channel, founding ``some promising initial capacity" for this scheme.
 In their research, results in terms of BER over SNR for a DNN based autoencoder and for a convolutional neural network (CNN) based autoencoder were treated.
@@
 al.
-in a 2017 paper \cite{2018} proved that a deep learning-based channel decoder could actually learn a decoding algorithm rather than just being a simple classifier.
+in \cite{2018} proved that a deep learning-based channel decoder could actually learn a decoding algorithm rather than just being a simple classifier.
 They introduced codewords that were not been used in the training set, and the trained NN was able to correctly decode it.
@@
 
-This research will implement a ML autoencoder for a  BSC that performs similar to the MAP decoder in real applications for a range of SNR from $-10 \text{dB}$ to $10 \text{dB}$.
-In its most simple form, a channel autoencoder includes an encoder, a noisy channel and an decoder.
-Using state-of-the art DNN algorithms to find the best solution of the problem, this work hopes to contribe to set up a higher standard in terms of performance in bit-error correction and reduced delay for digital communication applications.
+In this work, we will implement a ML autoencoder for a  BSC that performs similar to the MAP decoder in real applications for a range of SNR from $-10 \text{dB}$ to $10 \text{dB}$.
+In its most simple form, a channel autoencoder includes an encoder, a noisy channel and a decoder.
+Using state-of-the art DNN algorithms to find the best solution of the problem, this work aims to contribute to set up a higher standard in terms of performance in bit-error correction and reduced delay for digital communication applications.
 In a near future, this disruptive methodology for error correction using ML could replace mathematically optimal decoders which are the current guideline.
@@
 %\subsection{Notation}
-%Special notations used throughout the paper
+%Special notations used throughout the paper as shown in figures tac tac
 
@@
 
-The concept of a MAP decoder or Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is choosing a message which maximizes the MAP probability of the corrupted information received  by the decoder \cite{b4}.
-This algorithm is known for its optimality in error correction for white noise interference and it's variants: Log-MAP and Max-Log-MAP, are widely used as decoders for turbo codes \cite{b7}.
+The concept of a MAP decoder algorithm for sequences, is choosing a message which maximizes the a posterior probability of the corrupted codeword $y$ received by the decoder \cite{b4}.
+This algorithm is known for its optimal error correction capability for white noise interference.
+This algorithm is also referred as the Viterbi decoder \cite{viterbi}.
 
-The MAP decoder is practically nonviable to implement \cite{b6}.
-Thus, its variants cited in the paragraph above are better suited for practical cases.
+Mathematically, we want to maximize the average probability of making a correct decision when analysing the channel output $\textbf{y}$, which is written as
+\begin{equation}\label{eq:MAP1}
+Pc := Pr[\textbf{X}=f({Y})]
+\end{equation}
+where $x \sim X$ and $y \sim Y$ are multivariate random variables. We derive below the decoder that maximizes $P_c$.
+
+\begin{align} 
+\Pr \left[ X=f\left( Y\right) \right]& =\sum_{f\left( y\right) y\in \mathcal{X}\times \mathcal{Y}} P_{XY}(f(y)y) \label{eq:MAP2} \\
+&=\sum _{y\in \mathcal{Y}}P_{Y}\left( y\right)P_{X|Y}\left( f(y)|y\right) \label{eq:MAP3}
+\end{align}
+
+The optimal decoder is finally given by
+\begin{align}
+f(y) &= \underset{x\in \mathcal{X}}{\text{arg max}} P_{X|Y}(x|y)\label{eq:MAPF1}\\
+&= \underset{x\in \mathcal{X}}{\text{arg max}}
+P_X(x)P_{Y|X}(y|x).\label{eq:MAPF2}
+\end{align}
+
+However, the MAP decoder is practically nonviable to implement \cite{b6}.
+Thus, its variants cited in this paragraph are better suited for practical cases.
 The Log-MAP is an optimal decoder, having equivalent performance to the MAP decoder.
 While the Max-Log-MAP is a suboptimal decoder and will not be treated in this paper.
+these variants are widely used as decoders for turbo codes \cite{b7}.
 Refer to \cite{b6} for specific details in the derivation of the Log-MAP algorithm.
+Furthermore, the Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is a bit-wise MAP rule with lower complexity widely used in practice, even for long data packets.
 
@@
 
-\foreach \l [count=\i] in {1,2,3,m}
+\foreach \l [count=\i] in {1,2,3,n}
   \draw [<-] (input-\i) -- ++(-1,0)
-    node [above, midway] {$x_\l$};
+    node [above, midway] {$y_\l$};
 
 \foreach \l [count=\i] in {1,j}
-  \node [above] at (hidden-\i.north) {$r_\l$};
+  \node [above] at (hidden-\i.north) {$r_{l,\l}$};
 
-\foreach \l [count=\i] in {1,n}
+\foreach \l [count=\i] in {1,k}
   \draw [->] (output-\i) -- ++(1,0)
-    node [above, midway] {$y_\l$};
+    node [above, midway] {$\hat{u}_\l$};
 
@@
 
-\caption{MLNN representative diagram. Where \textbf{x} is the input vector of the NN, $\textbf{r}_{l}$ is a hidden layer vector
-and \textbf{y} is the output vector.} \label{fig:NN}
+\caption{MLNN representative diagram, where $\textbf{y}^n$ is the input vector, $\textbf{r}_{l}^{j}$ is a hidden layer vector
+and $\hat{\textbf{u}}^{k}$ is the output vector.} \label{fig:NN}
 \end{figure}
@@
 To define and build this infrastructure, the activation function, the loss function and the structure must be decided.
-M.
 Nielson \cite{b8} in his book explores a vast amount of NN architectures and how they work.
@@
 They are known for the universal approximation property \cite{b9} and for the versatility of increasing the number of layers, creating a DNN able to undertake complex classification problems with low classification error and low dimensionality.
-In more general terms, a MLNN with $L$ layers, also called \textit{depth}, and parameter $\theta$ is a mapping of an input $\textbf{r}_0 \in \mathbb{R}^{N_0}$ to an output $\textbf{r}_L \in \mathbb{R}^{N_L}$ through $L$ iterative steps.
+In more general terms, a MLNN with $L$ layers, also called \textit{depth}, and parameter $\theta$ is a mapping of an input $\textbf{y}^n \in \mathbb{R}^{n}$ to an output $\textbf{û}^k \in \mathbb{R}^{k}$ through $L$ iterative steps.
 For a fully-connected network, each layer vector is calculated iteratively in the forward propagation.
-Each neuron is composed of a linear combinator and an activation function defined as
+Each neuron is composed of a linear combiner and an activation function defined as
 
@@
 \end{equation}
-where $\textbf{W}_{l}\in \mathbb{R} ^{N_{l}\times N_{l-1}}$ is the weight matrix between layers $l-1$ and $l$,  $\textbf{b}_{l}\in \mathbb{R} ^{N_l}$ is the bias vector and $\sigma$ is the activation function \cite{b2}.
+where $\textbf{W}_{l}\in \mathbb{R} ^{N_{l}\times N_{l-1}}$ is the weight matrix between layers $l-1$ and $l$, $\textbf{r}_{l-1} \in \mathbb{R} ^{N_{l-1}}$ is a vector containing the hidden layer $l-1$ values, $\textbf{b}_{l}\in \mathbb{R} ^{N_l}$ is the bias vector and $\sigma$ is the activation function \cite{b2}.
 
-In order to accurately classify an input, a \textit{perceptron} processing unit is chosen to solve a classification problem for each input message.
-A perceptron maps a binary vector input into a single binary output digit \cite{b8}.
-Mathematically, it means that the function defined in \eqref{eq:eqFP}, outputs $1$ or $0$, depending on a threshold value.
+% TODO: WRITE WHICH ACTIVATION FUNCTION I WILL USE
+%
 
@@
 With all these elements set, training the neural network to calculate an accurate weight matrix $\textbf{W}$ for forward propagation requires a labeled \textit{training data set}.
-This set is composed of pairs $ (\textbf{r}_{0,i}, \textbf{p}_i) $, where $i=1,...,S$ and $S$ is the number of training sets.
+This set is composed of pairs $ (\textbf{y}_{i}^{n}, {\mathbf{u}}_i^k) $, where $i=1,...,S$ and $S$ is the number of training sets.
 It matches the input and its respective desired output.
@@
 \begin{equation}\label{eq:o-c-e}
-L\left( \theta \right) =\dfrac {1}{S}\sum ^{S}_{i=1}l\left( \textbf{p}_{L,i},\textbf{r}_{L,i}\right)	
+L\left( \theta \right) =\dfrac {1}{S}\sum ^{S}_{i=1}l\left({\mathbf{u}}_i^k  , \textbf{y}_{i}^{n}\right)	
 \end{equation}
@@
 For NN, an efficient way of computing this minimization is implementing the back-propagation (BP) algorithm.
-	BP is classified as a supervised learning algorithm that is able to optimize a function based on some parameter and is highly parallelizable in computational terms \cite{b8}\cite{b9}.
+	BP is classified as a supervised learning algorithm that is able to optimize a function based on some parameter and is highly parallelizable in computational terms \cite{b9}\cite{b8}.
 
-One of the greatest advantage of NN is its ability do generalize the training set, delivering the right results even for input data not contained in the training set.
-With the weight matrix trained, the NN find the correct output values for unseen inputs.
+One of the greatest advantage of NN is its ability to generalize the training set, delivering the right results even for input data not contained in the training set.
+With the weight matrix trained, the NN finds the correct output values for unseen inputs.
 In order to validate this principle, a \textit{validation data set} is used and the loss is mesured.
@@
 For implementing all the algorithms, Keras, a high-level python API for ML, will be used together with TensorFlow 1.13.1, a ML python friendly open-source library \cite{b11} \cite{b12}.
-Both are available for free.
+Both are available on a free license terms.
 
@@
 
+\bibitem{b3} F.
+D.
+Calabrese, L.
+Wang, E.
+Ghadimi, G.
+Peters, and P.
+Soldati, ``Learning radio resource management in 5G networks: Framework, opportunities and challenges," arXiv preprint arXiv:1611.10253, 2016.
+
 \bibitem{b2} O'Shea, Tim \& Hoydis, Jakob, ``An Introduction to Machine Learning Communications Systems", 2017.
-\bibitem{b3} F. D. Calabrese, L. Wang, E. Ghadimi, G. Peters, and P. Soldati, ``Learning radio resource management in 5G networks: Framework,
-opportunities and challenges," arXiv preprint arXiv:1611.10253, 2016
+
+\bibitem{osheaautoencoder} T. O'Shea, K. arra, T. C. Clancy, ``Learning to Communicate: Channel Auto-encoders, Domain Specific Regularizers, and Attention", 2016.
+
+\bibitem{2018} T. Gruber, S. Cammerer, J. Hoydis, and S. ten Brink, “On deep learning based channel decoding,” accepted for CISS 2017, arXiv preprint arXiv:1701.07738, 2017.
+
 \bibitem{b4} W. Alexander, H. Peter \& W. Norbert, ``Turbo-Decoding Without SNR Estimation",IEEE Communications Letter, vol. 4, no. 6, pp. 193-195, 2000.
-\bibitem{b5} M. Jordan and R. Nichols, “The effects of channel characteristics on turbo code performance,” in Proc. Milcom’96, McLean, VA, Oct. 1996, pp. 17–21.
+
+\bibitem{viterbi} A. J. Viterbi, ``Error Bounds for Convolutional Codes and an Asymptotically Optimum Decoding Algorithm", IEEE Transactions on Information Theory, 13 (2): 260–269, 1967.
+
 \bibitem{b6} Robertson, P.
@@
 doi:10.1002/ett.4460080202.
+
 \bibitem{b7} J. Hagenauer,P. Robertson, L. Papke: iterative (“Turbo”) decoding of systematic convolutional codes with the MAP and SOVA
 algorithms. In: ITG-Fachbericht 130. October 1994, p. 21-29.
+
+\bibitem{b5} M. Jordan and R. Nichols, “The effects of channel characteristics on turbo code performance,” in Proc. Milcom’96, McLean, VA, Oct. 1996, pp. 17–21.
+
+\bibitem{b9} I. Mohamed, ``Applications of neural networks to digital communications - a survey", 1997.
+
 \bibitem{b8} M. Nielsen, \textit{Neural Networks and Deep Learning}.
-\bibitem{b9} I. Mohamed, ``Applications of neural networks to digital communications - a survey", 1997.
+
 \bibitem{b10} M. Kevin, ``Machine Learning: A Probabilistic Perspective", MIT, ISBN 978-0262018029, 2012.
+
 \bibitem{b11} M. Abadi, A. Agarwal, P. Barham, et al., 
 ``TensorFlow: Large-scale machine learning on heterogeneous systems", 2015. Software available from tensorflow.org.
+
 \bibitem{b12} F. Chollet, “keras,” https://github.com/fchollet/keras, 2015.
-\bibitem{osheaautoencoder} T. O'Shea, K. arra, T. C. Clancy, ``Learning to Communicate: Channel Auto-encoders, Domain Specific Regularizers, and Attention", 2016.
-\bibitem{2018} T. Gruber, S. Cammerer, J. Hoydis, and S. ten Brink, “On deep learning based channel decoding,” accepted for CISS 2017, arXiv preprint arXiv:1701.07738, 2017.
 

### 2019-05-13 MyCode/BSC-MAP-Decoder.py

@@ -45,14 +45,14 @@ print("Parity check: \n",dot(G,H.T)%2)
 dmin = 3
 
 # input
-N = 10 # number of messages sent
+N = 100 # number of messages sent
 
 # Channel
 CTYPE = "BSC"
-p = 0.1 # cross probability of BSC
+p = 0.01 # cross probability of BSC
 e = stats.bernoulli.rvs(p, size=n) # BSC noise
 He = fn.Hb(p)# binary entropy function
-C = 1- H# channel capacity
+C = 1- He # channel capacity
 #%% 
 '''
        Generate channel Input
@@ -91,22 +91,19 @@ S = empty([N,n-k]) # identify if there are errors - syndrome
 # k stored syndromes
 for i in range(N):
     S[i] = dot(y[i],H.T)%2 # syndrome row vector
-paux = arange(m)[::-1]
-Hindex = size(H,1) - sum(power(S*2,paux),axis=1) # index of H where there is a bit error
-
 
 D = y # error corrected codewords
 e = zeros([N,n])# bit error location
 # single error correction
 for i in range(N):
     if(sum(S[i,:])!=0):
-        e[i,int(Hindex[i])] += 1
+        print(i)
+        index = sum(S[i,:] == H.T,1).argmax() # Find position where H matrix says there is an error
+        e[i,int(index)] += 1
 e = e%2
 D = (D+e)%2 # decoded codewords
 
 E2cw = sum(D != x) # Codeword Error without hamming decoding
 uhat = D[:,:k]
 E2b = sum(uhat != u)
-
-    
 

### 2019-06-01 Rapport/Article/Article.tex
Simulation 01 06/01/2019

--- 2019-05-13 Rapport/TG_ITA/Bibliographic_study/conference_041818.tex
@@
 
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/test}
+    \caption{Diagram of a simplified communication system.}\label{fig:cs}
+\end{figure}
+
 Since then, the research community in digital communication developed a series of algorithms to minimize error probability over a channel.
@@
 
-%\subsection{Notation}
-%Special notations used throughout the paper as shown in figures tac tac
+\subsection{Notation}
+Special notations used throughout the paper as shown in figures tac tac
 
@@
 %
+Two common activation functions are the rectified linear unit (ReLU) and the sigmoid function. Both will be applied in the NN treated in this paper. Their mathematical expressions are shown below.
+
+\begin{align}
+	\text{ReLU: } \sigma_1 (x)& = max\{0, x\} \label{eq:relu}\\ 
+	\text{Sigmoid: } \sigma_2 (x)& = \frac{1}{1+e^{- \lambda x}} \label{eq:sig}
+\end{align}
+where $x \in \mathbb{R}$.
 
@@
 
-\begin{comment}
 \subsection{Autoencoders}
@@
 
-\end{comment}
-%\section{Procedure and methods}
-%Explain what I Should do to achieve my goal
+\section{Procedure and Methodology}
+Explain what I Should do to achieve my goal
+Talk about decoder + autoencoder and the OBJECTIVE of each one, G matrix of choice, Batch size, The reader should be able to replicate what I ve done
+
+\subsection{Array Decoder}
+Architecture, Training parameters, cite choice of p train \cite{b13}, number of epochs, Talk about noise layer and its removal in prediction,
+
+\subsection{One-hot Decoder}
+Architecture, Training parameters, Explain what is a One hot decoding, talk about the matrix transformation from one hot to messages, p train equals 0, number of epochs
+
+\subsection{Autoencoder}
+Architecture, Training and simulation parameters, procedure to find a M epoch suitable, choices of p train, Talk about NN separation,
+
+\section{Results and Analysis}
+
+\subsection{Decoders}
+BER Curves from both decoders, analyse why the number of epochs is so different.
+
+\subsection{Autoencoder}
+BER Curves of simulation varying p, best choice of p, analyse autoencoder with best choice of p
+
+\subsection{Decoding time analysis}
+compare time in decoding between map, autoencoder and a decoder.
+
+\section{Conclusions}
+
 \begin{comment}
@@
 
+\bibitem{b13} M. Benammar and P. Piantanida, “Optimal training channel statistics for neural-based decoders,” in Asilomar, Oct 2018, pp. 2157–2161.
+
 \end{thebibliography}

### 2019-06-01 MyCode/BSC-MAP-Decoder.py

@@ -47,7 +47,7 @@ for i_global in range(globalReps):
         globalError[i_global][i_p] = fn.codeErrorFunction(y, x)
         globalErrorMAP[i_global][i_p] = fn.bitErrorFunction(MAP, u)
 
-#%% plot
+#% plot
         
 avgGlobalError = np.average(globalError, 0)
 avgGlobalErrorMAP = np.average(globalErrorMAP, 0)

### 2019-06-01 MyCode/autoencoderArray.py

@@ -13,19 +13,6 @@ Created on Mon May 27 16:04:54 2019
 """
 #%%
 '''
-    DNN One Hot Model Autoencoder
-'''
-'''
-    One hot training and validation data
-'''
-u_train_labels = fn.messages2onehot(messages.copy())
-x_train_data = messages.copy()
-
-u_train_labels = np.repeat(u_train_labels, 1, axis=0)
-x_train_data = np.repeat(x_train_data, 1, axis=0)
-trainSize = np.size(x_train_data, 0)
-#%%
-'''
     Array training and validation data
 '''
 u_train_labels = messages.copy()
@@ -35,15 +22,6 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 #%%
-'''
-    Constants
-'''
-numEpochs = 2**15  #2**16 approx 65000
-batchSize = trainSize 
-train_p = 0.07
-timestr = time.strftime("%Y%m%d-%H%M%S")
-title = 'AutoencoderArray'
-
 '''
     Architecture
 '''
@@ -57,7 +35,7 @@ Encoder = tf.keras.Sequential([
         # Coded Layer
         layers.Dense(2*k, activation='sigmoid', name='Codedfloat'),
         # Rounded codeword
-        layers.Lambda(roundCode, input_shape=(2*k,), output_shape=(2*k,), name='Codeword'),
+        layers.Lambda(fn.roundCode, input_shape=(2*k,), output_shape=(2*k,), name='Codeword'),
         ], name='Encoder')
 
 NoiseL = tf.keras.Sequential([
@@ -89,7 +67,7 @@ Autoencoder.compile(loss=lossFunc ,
 '''
     Summaries and checkpoints (to do)
 '''
-summary = Autoencoder.summary()
+#summary = Autoencoder.summary()
 checkpoint = tf.keras.callbacks.ModelCheckpoint(
         'Checkpoints/'+timestr+'_'+title+'_weights.{epoch:02d}-{loss:.6f}.hdf5', monitor='loss', 
         verbose=0, save_best_only=True, save_weights_only=False, mode='min', period=2**11)
@@ -117,11 +95,12 @@ trainingFig.savefig('training_history/'+title+'/'+timestr + '_'+title+'_train.pn
 '''
[…]

### 2019-06-01 MyCode/autoencoderInitializer.py

@@ -1,0 +1,41 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Jun  1 13:42:07 2019
+
+@author: user
+"""
+#%%
+'''
+    DNN One Hot Model Autoencoder
+'''
+'''
+    One hot training and validation data
+'''
+u_train_labels = fn.messages2onehot(messages.copy())
+x_train_data = messages.copy()
+
+u_train_labels = np.repeat(u_train_labels, 1, axis=0)
+x_train_data = np.repeat(x_train_data, 1, axis=0)
+trainSize = np.size(x_train_data, 0)
+#%%
+'''
+    Array training and validation data
+'''
+u_train_labels = messages.copy()
+x_train_data = u_train_labels
+
+u_train_labels = np.repeat(u_train_labels, 1, axis=0)
+x_train_data = np.repeat(x_train_data, 1, axis=0)
+trainSize = np.size(x_train_data, 0)
+
+#%%
+'''
+    Constants
+'''
+numEpochs = 2**13  #2**16 approx 65000
+batchSize = 256
+train_p = 0.07
+timestr = time.strftime("%Y%m%d-%H%M%S")
+title = 'AutoencoderArray'
+path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
+

### 2019-06-01 MyCode/backup_code.py

@@ -211,7 +211,7 @@ fig.savefig('images/'+timestr+'MAP_MLNN_Mep_'+str(numEpochs)+'.png', bbox_inches
 import os
 
 # define the name of the directory to be created
-path = "tmp\oi"
+directory = "tmp"
 
 try:  
     os.mkdir(path)
@@ -219,4 +219,22 @@ except OSError:
     print ("Creation of the directory %s failed" % path)
 else:  
     print ("Successfully created the directory %s " % path)
+    
+#%%
+print('main')
+exec(open("autoencoder1H.py").read())
+
+#%%
+# load weights into new model
+loadedModel = tf.keras.models.load_model("Trained_AutoencoderArray\AutoencoderArray_Mep_32768_bs_256.h5")
+print("Loaded model from disk")
+plot_model(loadedModel, to_file='Debug\Loaded.pdf')
+loadedModel.summary()
+
+#%%
+t = TicToc('Training')
+t.tic()
+some code...
+t.toc()
+print(t.elapsed)
 

### 2019-06-01 MyCode/libs-and-global-consts.py

@@ -17,15 +17,18 @@ sys.path.append('C:\\Users\\user\\Desktop\\GitHub\\PIR\\MyCode')
 import my_functions as fn
 
 import time
+import os
+from ttictoc import TicToc
 
 #Plot setup
 #plt.rcParams['mathtext.fontset'] = 'dejavuserif'
 plt.rcParams["font.family"] = "Times New Roman"
 plt.rcParams.update({'font.weight':'normal'})
-plt.rc('xtick', labelsize=8)
-plt.rc('ytick', labelsize=8)
-plt.rc('axes', labelsize=8)
-plt.rc('legend', fontsize=8)
+letterSize = 6
+plt.rc('xtick', labelsize=letterSize)
+plt.rc('ytick', labelsize=letterSize)
+plt.rc('axes', labelsize=letterSize)
+plt.rc('legend', fontsize=letterSize)
 
 # width as measured in inkscape
 width = 3.487
@@ -93,7 +96,7 @@ def plotBERp(globalErrorMLNN, legendEntry):
     plt.plot(pOptions,avgGlobalError, color='b', linewidth=1, linestyle='--')
     plt.plot(pOptions,avgGlobalErrorMAP, color='r', linewidth=1)
     plt.grid(True, which='both')
-    #plt.title('Batch size = '+str(batchSize)+', train_p = ' + str(train_p))
+    #plt.title('Training $p = $ '+ str(train_p))
     plt.xlabel('$p$')
     plt.ylabel('BER')
     plt.yscale('log')
@@ -101,5 +104,5 @@ def plotBERp(globalErrorMLNN, legendEntry):
     plt.show()
     
     fig.set_size_inches(width, height)
-    fig.savefig('images/'+title+'/'+timestr+'_MAP_'+title+'_Mep_'+str(numEpochs)+'.png', bbox_inches='tight', dpi=300)
+    fig.savefig('images/'+title+'/'+timestr+'_MAP_'+title+'_Mep_'+str(numEpochs)+'_ptrain_'+str(train_p)+'.png', bbox_inches='tight', dpi=300)
 

### 2019-06-01 MyCode/multi-train.py

@@ -1,0 +1,104 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Jun  1 13:41:33 2019
+
+@author: user
+"""
+#%%
+'''
+    Multi Training
+'''
+trainTime = TicToc('Training')
+trainTimeT = TicToc('TrainingT')
+# Constants 
+numEpochs = 2**14  #2**16 approx 65000
+batchSize = trainSize 
+timestr = time.strftime("%Y%m%d-%H%M%S")
+title = 'AutoencoderArray'
+p_trainOptions = np.array([0, 0.01, 0.03, 0.05, 0.07, 0.09])
+
+myDir = 'Simulation_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'_pmin_'+str(min(p_trainOptions))+'_pmax_'+str(max(p_trainOptions))
+directory = 'Autoencoder_Simulations\\'+ myDir
+fn.createDir(directory)
+fn.createDir(directory+'\\models')
+
+paths = ["" for x in range(len(p_trainOptions))]
+trainTime.tic()
+for i in range(len(p_trainOptions)):
+    trainTime.tic()
+    train_p = p_trainOptions[i]
+        
+    paths[i] = directory+'\\models\\'+'i_'+str(i)+'_'+title+'_Mep_'+str(numEpochs)+'_p_'+str(train_p)+'.h5'
+    path = paths[i]
+    exec(open("autoencoderArray.py").read())
+    trainTime.toc()
+    print('Train time for model '+str(i)+': ',trainTime.elapsed)
+trainTimeT.toc()
+print('Total train time: ',trainTimeT.elapsed)
+    
+    
+#%%
+'''
+    Multi Prediction
+'''
+predictTime = TicToc('Predict')
+predictTimeT = TicToc('PredictT')
+# Initialize variables
+multiPredictions = np.empty([len(p_trainOptions),len(pOptions)])
+globalReps = 1000
+globalErrorAutoencoder = np.empty([globalReps, len(pOptions)])
+predictTime.tic()
+for i_train in range(len(p_trainOptions)):
+    predictTime.tic()
+    # load weights into new model
+    fileName = paths[i_train]
+    loadedModel = tf.keras.models.load_model(fileName)
+    print("Loaded model from disk")
+    plot_model(loadedModel, to_file='Debug\Loaded.pdf')
+    Encoder = loadedModel.layers[0]
+    Decoder = loadedModel.layers[2]
+    
+    for i_global in range(globalReps):
[…]

### 2019-06-01 MyCode/my_functions.py

@@ -14,6 +14,7 @@ from scipy import stats
 import tensorflow as tf
 from tensorflow.keras import layers
 from keras import backend as K
+import os
 
 def plotit(t, u, title="Title", X="time (s)", Y="Amplitude"):
     plt.figure()
@@ -323,6 +324,14 @@ def TensorOnehot2singleMessage(h):
 
 def roundCode(x):
     return tf.stop_gradient(K.round(x)-x)+x
+
+def createDir(path):
+    try:  
+        os.mkdir(path)
+    except OSError:  
+        print ("Creation of the directory %s failed" % path)
+    else:  
+        print ("Successfully created the directory %s " % path)
        
     
     

### 2019-06-03 Rapport/Article/Article.tex
Seance 06/03/2019

@@
 \usepackage{amsmath,amssymb,amsfonts}
+
 \usepackage{algorithmic}
@@
     \includegraphics[width=0.5\textwidth]{simple_sys}
-    \caption{Diagram of a simplified communication system.}\label{fig:cs}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/test}
     \caption{Diagram of a simplified communication system.}\label{fig:cs}
@@
 \subsection{Notation}
-Special notations used throughout the paper as shown in figures tac tac
+Throughout the work, vectors will be written in bold font weight and may  have a subscript which indicates their layer in a NN and may have a superscript which indicates their size length.
+In order to homogenize notation, for the encoding phase of the communication system, $\textbf{u}^k$ represents an input \textit{source message} and $\textbf{x}^n$ its correspondent codeword.
+For the decoding portion, $\textbf{y}^n$ represents the output of the noisy channel and $\hat{\textbf{u}}^k$ the estimated source message.
+These definitions are illustrated in Fig.
+\ref{fig:cs}.
+
+In addition, every indispensable concept is introduced either in italics or is defined by a commonly used abbreviation present in others scientific works.
 
@@
 \end{align}
-where $x \in \mathbb{R}$.
+where $x \in \mathbb{R}$ and $\lambda>0$.
 
@@
 \subsection{Autoencoders}
-TO PARAPHRASE: The parallel computing
-capabilities of NNs allow us to solve or, at least, approximate
-the MLD problem in polynomial time [7]. Moreover, the
-weights of the NN are precomputed during training and the
-decoding step itself is then relatively simple.
 
-NNs which can learn an appropriate
-mapping between noisy input patterns and codewords. No
-assumption has to be made about the statistics of the channel
-noise because the NN is able to learn the mapping or to
-extract the channel statistics during the learning process
+Autoencoders based on DNN can learn an appropriate correspondence between input source messages and estimated source message output through different noisy channels.
+The statistics of the channel's noise is shown in \cite{2018} to be irrelevant, since the NN extract it during training phase.
+Their applicability is tantamount compelling because the weights computed during training can be separated in two groups, one correspondent to the nodes before the channel and one posterior to the channel.
+As a consequence, we are able to design an non linear encoder and its correspondent non linear decoder with substantial ease.
+This new representation has ultimately far more possibilities than conventional encoding patterns.
 
-decoding problems have far more possibilities than conventional pattern recognition problems
-
-In 2006, a new training technique, called layer-by-layer
-unsupervised pre-training followed by gradient descent finetuning [14], led to the renaissance of NNs because it made
-training of NNs with more layers feasible.
-
-new hardware such as graphical processing units (GPUs) are available
-to speed up learning as well as inference. In this renaissance
-of NNs, new NND ideas emerge
-
-The curse of dimensionality
+Whilst autoencoders are more complex dimensional wise when compared to NN decoders, the readily access to graphical processing units (GPUs) contributed to the feasibility of this technique, tackling the \textit{curse of dimensionality} challenge in ML.
+This type of hardware made possible the approach of the layer-by-layer unsupervised pre-training followed by stochastic gradient descent applied in this work.
+\cite{GPU}
 
@@
 
-\subsection{Array Decoder}
+\subsection{Decoder}
+Hamming coding
+
+The linear encoding matrix $\textbf{G}_{8\times16}$ is shown below.
+
+\begin{flalign*}
+    &\textbf{G}=&
+\end{flalign*}
+
+\[
+\left [
+  \begin{array}{cccccccccccccccc}
+
+   1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+	1 & 1 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+	1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+	1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
+	1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 & 0 & 0 & 0 \\
+	1 & 1 & 0 & 0 & 1 & 1 & 0 & 0 & 1 & 1 & 0 & 0 & 1 & 1 & 0 & 0 \\
+	1 & 0 & 1 & 0 & 1 & 0 & 1 & 0 & 1 & 0 & 1 & 0 & 1 & 0 & 1 & 0 \\
+	1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1
+  \end{array}
+\right ]
+\]
+
+Introduce encoding equations
+
+\subsubsection{Array Decoder}
 Architecture, Training parameters, cite choice of p train \cite{b13}, number of epochs, Talk about noise layer and its removal in prediction, 
-
-\subsection{One-hot Decoder}
+\subsubsection{One-hot Decoder}
 Architecture, Training parameters, Explain what is a One hot decoding, talk about the matrix transformation from one hot to messages, p train equals 0, number of epochs
@@
 
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/MAP_MLNN1H_p_0}
+    \caption{Test caption to test image.}\label{fig:1HD}
+\end{figure}
+
 \subsection{Autoencoder}
 BER Curves of simulation varying p, best choice of p, analyse autoencoder with best choice of p
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/MAP_MAP_Mep_65536}
+    \caption{Test caption to test image.}\label{fig:ae1}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/MAP_AutoencoderArray_Mep_65536}
+    \caption{Test caption to test image.}\label{fig:ae1}
+\end{figure}
 
@@
 
+\bibitem{GPU} G. E. Hinton, S. Osindero, and Y.-W. Teh, “A fast learning algorithm for deep belief nets,” Neural Computation, vol. 18, no. 7, pp. 1527–1554, July 2006.
+
 \bibitem{b13} M. Benammar and P. Piantanida, “Optimal training channel statistics for neural-based decoders,” in Asilomar, Oct 2018, pp. 2157–2161.

### 2019-06-03 MyCode/DecoderInitializer.py

@@ -1,0 +1,14 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jun  3 14:55:00 2019
+
+@author: user
+"""
+#%%
+numEpochs = 2**10  #2**16 approx 65000
+batchSize = 256 # Mini batch size
+train_p = 0.07
+timestr = time.strftime("%Y%m%d-%H%M%S")
+title = 'MLNN'
+path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
+

### 2019-06-03 MyCode/MLNN-train.py

@@ -1,0 +1,84 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jun  3 15:15:35 2019
+
+@author: user
+"""
+#%%
+n=16
+k=8
+
+u_train_labels = messages.copy()
+x_train_data = possibleCodewords.copy()
+
+u_train_labels = np.repeat(u_train_labels, 1, axis=0)
+x_train_data = np.repeat(x_train_data, 1, axis=0)
+trainSize = np.size(x_train_data, 0)
+
+test_Size = 100
+u_val_labels = fn.generateU(test_Size,k)
+x_val = fn.generteCodeWord(test_Size, n, u_val_labels, G)
+
+'''
+    Sequential Model: most simple tf MLNN model
+'''
+NoiseL = tf.keras.Sequential([
+        # Noise Layer
+        layers.Lambda(tensorBSC,input_shape=(n,), output_shape=(n,), name='Noise'),
+        ], name='Noise')
+MLNNDecoder = tf.keras.Sequential([ # Array to define layers
+              # Adds a densely-connected layer with n units to the model: L1
+              layers.Dense(128, activation='relu', input_shape=(n,), name='HL1'),
+              # Add another: L2
+              layers.Dense(64, activation='relu', name='HL2'),
+              # Add another: L3
+              layers.Dense(32, activation='relu', name='HL3'),
+              # Add layer with k output units:
+              layers.Dense(k, activation='sigmoid', name='Output')
+              ], name='Array_Decoder')
+MLNN = tf.keras.Sequential([NoiseL, MLNNDecoder])
+plot_model(MLNN,to_file='GraphNN/'+title+'/'+timestr+'_'+title+'.pdf',show_shapes=True)
+    
+'''
+    Overall Settings/ Compilation
+'''
+lossFunc = 'binary_crossentropy'
+MLNN.compile(loss=lossFunc ,
+              optimizer='adam')
+'''
+    Summaries and checkpoints 
+'''
+summary = MLNN.summary()
+callbacks_list = []
+''' 
+    Training
+'''
+history = MLNN.fit(x_train_data, u_train_labels, epochs=numEpochs, 
+                   batch_size=batchSize, shuffle=True, verbose=0, callbacks=callbacks_list)
+#history = MLNN.fit(x_train, u_train_labels, epochs=numEpochs, batch_size=batchSize,
[…]

### 2019-06-03 MyCode/MLNN1H-train.py

@@ -1,0 +1,92 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jun  3 15:15:26 2019
+
+@author: user
+"""
+'''
+    DNN One Hot Model Decoder
+'''
+'''
+    Training and validation data
+'''
+u_train_labels = fn.messages2onehot(messages.copy())
+x_train_data = possibleCodewords.copy()
+
+u_train_labels = np.repeat(u_train_labels, 1, axis=0)
+x_train_data = np.repeat(x_train_data, 1, axis=0)
+trainSize = np.size(x_train_data, 0)
+
+test_Size = 100
+u_val_labels = fn.generateU(test_Size,k)
+x_val = fn.generteCodeWord(test_Size, n, u_val_labels, G)
+u_val_labels = fn.messages2onehot(u_val_labels)
+
+'''
+    Sequential Model: most simple tf MLNN model
+'''
+
+MLNN1H = tf.keras.Sequential([ # Array to define layers
+              # Adds a densely-connected layer with n units to the model: L1
+              #layers.Dense(32, activation='relu', input_shape=(n,), name='HL1'),
+              # Add another: L2
+              layers.Dense(64, activation='relu', input_shape=(n,), name='HL1'),
+              # Add another: L3
+              #layers.Dense(128, activation='relu',input_shape=(n,), name='HL1'),
+              # Add layer with k output units:
+              layers.Dense(256, activation='softmax', name='Output')
+])
+
+plot_model(MLNN1H,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',show_shapes=True)
+    
+'''
+    Overall Settings/ Compilation
+'''
+lossFunc = 'binary_crossentropy'
+MLNN1H.compile(loss=lossFunc ,
+              optimizer='adam')
+'''
+    Summaries and checkpoints (to do)
+'''
+summary = MLNN1H.summary()
+checkpoint = tf.keras.callbacks.ModelCheckpoint(
+        'Checkpoints/'+timestr+'_'+title+'_weights.{epoch:02d}-{loss:.6f}.hdf5', monitor='loss', 
+        verbose=0, save_best_only=True, save_weights_only=False, mode='min', period=2**11)
+callbacks_list = [checkpoint]
+''' 
+    Training
+'''
+history = MLNN1H.fit(x_train_data, u_train_labels, epochs=numEpochs, 
[…]

### 2019-06-03 MyCode/autoencoder1H.py

@@ -25,7 +25,7 @@ trainSize = np.size(x_train_data, 0)
 '''
 numEpochs = 2**16  #2**16 approx 65000
 batchSize = trainSize 
-train_p = 0.07
+train_p = 0.00
 timestr = time.strftime("%Y%m%d-%H%M%S")
 title = 'Autoencoder1H'
 
@@ -52,7 +52,7 @@ NoiseL = tf.keras.Sequential([
 
 Decoder1H = tf.keras.Sequential([ # Array to define layers
         layers.Dense(64, activation='relu', input_shape=(2*k,), name='DHL1'),
-        layers.Dense(128, activation='relu', input_shape=(2*k,), name='DHL2'),
+        #layers.Dense(128, activation='relu', input_shape=(2*k,), name='DHL2'),
         layers.Dense(256, activation='softmax', name='1H_Output')
         ], name = 'Decoder')
 

### 2019-06-03 MyCode/autoencoderArray.py

@@ -15,43 +15,50 @@ Created on Mon May 27 16:04:54 2019
 '''
     Array training and validation data
 '''
+timestr = time.strftime("%Y%m%d-%H%M%S")
+
 u_train_labels = messages.copy()
 x_train_data = u_train_labels
 
 u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
-#%%
+
+encoderNodes = np.array([256, 256, 256 ,2*k])
+DecoderNodes = [128, 64, 32 ,k]
+#%
 '''
     Architecture
 '''
 Encoder = tf.keras.Sequential([
         # Input Layer
-        layers.Dense(128, activation='relu', input_shape=(k,), name='Input'),
+        layers.Dense(encoderNodes[0], activation='relu', input_shape=(k,), name='Input'),
         # Hidden Layer
-        layers.Dense(64, activation='relu', name='EHL1'),
+        layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
         # Hidden Layer
-        layers.Dense(32, activation='relu', name='EHL2'),
+        layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
         # Coded Layer
-        layers.Dense(2*k, activation='sigmoid', name='Codedfloat'),
+        layers.Dense(encoderNodes[3], activation='sigmoid', name='Codedfloat'),
         # Rounded codeword
-        layers.Lambda(fn.roundCode, input_shape=(2*k,), output_shape=(2*k,), name='Codeword'),
+        layers.Lambda(fn.roundCode, input_shape=(encoderNodes[3],), 
+                      output_shape=(encoderNodes[3],), name='Codeword'),
         ], name='Encoder')
 
 NoiseL = tf.keras.Sequential([
         # Noise Layer
-        layers.Lambda(tensorBSC,input_shape=(2*k,), output_shape=(2*k,), name='Noise'),
+        layers.Lambda(tensorBSC,input_shape=(encoderNodes[3],), 
+                      output_shape=(encoderNodes[3],), name='Noise'),
         ], name='Noise')
 
 Decoder = tf.keras.Sequential([ # Array to define layers
         # Adds a densely-connected layer with n units to the model: L1
-        layers.Dense(32, activation='relu', input_shape=(2*k,), name='DHL1'),
[…]

### 2019-06-03 MyCode/autoencoderInitializer.py

@@ -32,7 +32,7 @@ trainSize = np.size(x_train_data, 0)
 '''
     Constants
 '''
-numEpochs = 2**13  #2**16 approx 65000
+numEpochs = 2**14  #2**16 approx 65000
 batchSize = 256
 train_p = 0.07
 timestr = time.strftime("%Y%m%d-%H%M%S")

### 2019-06-03 MyCode/libs-and-global-consts.py

@@ -24,7 +24,9 @@ from ttictoc import TicToc
 #plt.rcParams['mathtext.fontset'] = 'dejavuserif'
 plt.rcParams["font.family"] = "Times New Roman"
 plt.rcParams.update({'font.weight':'normal'})
-letterSize = 6
+letterSize = 7
+markerSize = 3
+lineWidth = 0.5
 plt.rc('xtick', labelsize=letterSize)
 plt.rc('ytick', labelsize=letterSize)
 plt.rc('axes', labelsize=letterSize)
@@ -92,9 +94,9 @@ def tensorBSC(x):
 def plotBERp(globalErrorMLNN, legendEntry):
     avgGlobalErrorMLNN = np.average(globalErrorMLNN,0)
     fig = plt.figure(figsize=(8, 6), dpi=80)
-    plt.scatter(pOptions,avgGlobalErrorMLNN, color='g', marker='^', zorder=3, s=8)
-    plt.plot(pOptions,avgGlobalError, color='b', linewidth=1, linestyle='--')
-    plt.plot(pOptions,avgGlobalErrorMAP, color='r', linewidth=1)
+    plt.scatter(pOptions,avgGlobalErrorMLNN, color='g', marker='^', zorder=3, s=markerSize)
+    plt.plot(pOptions,avgGlobalError, color='b', linewidth=lineWidth, linestyle='--')
+    plt.plot(pOptions,avgGlobalErrorMAP, color='r', linewidth=lineWidth)
     plt.grid(True, which='both')
     #plt.title('Training $p = $ '+ str(train_p))
     plt.xlabel('$p$')

### 2019-06-03 MyCode/multi-train.py

@@ -11,11 +11,11 @@ Created on Sat Jun  1 13:41:33 2019
 trainTime = TicToc('Training')
 trainTimeT = TicToc('TrainingT')
 # Constants 
-numEpochs = 2**14  #2**16 approx 65000
+numEpochs = 2**16  #2**16 approx 65000
 batchSize = trainSize 
 timestr = time.strftime("%Y%m%d-%H%M%S")
 title = 'AutoencoderArray'
-p_trainOptions = np.array([0, 0.01, 0.03, 0.05, 0.07, 0.09])
+p_trainOptions = np.array([2**16, 2**17, 2**18])
 
 myDir = 'Simulation_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'_pmin_'+str(min(p_trainOptions))+'_pmax_'+str(max(p_trainOptions))
 directory = 'Autoencoder_Simulations\\'+ myDir
@@ -23,18 +23,19 @@ fn.createDir(directory)
 fn.createDir(directory+'\\models')
 
 paths = ["" for x in range(len(p_trainOptions))]
-trainTime.tic()
+trainTimeT.tic()
 for i in range(len(p_trainOptions)):
     trainTime.tic()
-    train_p = p_trainOptions[i]
+    numEpochs = p_trainOptions[i]
         
     paths[i] = directory+'\\models\\'+'i_'+str(i)+'_'+title+'_Mep_'+str(numEpochs)+'_p_'+str(train_p)+'.h5'
     path = paths[i]
     exec(open("autoencoderArray.py").read())
+    
     trainTime.toc()
     print('Train time for model '+str(i)+': ',trainTime.elapsed)
 trainTimeT.toc()
-print('Total train time: ',trainTimeT.elapsed)
+
     
     
 #%%
@@ -47,7 +48,7 @@ predictTimeT = TicToc('PredictT')
 multiPredictions = np.empty([len(p_trainOptions),len(pOptions)])
 globalReps = 1000
 globalErrorAutoencoder = np.empty([globalReps, len(pOptions)])
-predictTime.tic()
+predictTimeT.tic()
 for i_train in range(len(p_trainOptions)):
     predictTime.tic()
     # load weights into new model
@@ -78,17 +79,17 @@ for i_train in range(len(p_trainOptions)):
     print('Predict time for model '+str(i_train)+ ': ',predictTime.elapsed)
 predictTimeT.toc()
[…]

### 2019-06-03 MyCode/multiDecoderTrain.py

@@ -1,0 +1,179 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Jun  1 13:41:33 2019
+
+@author: user
+"""
+#%%
+'''
+    Multi Training
+'''
+trainTime = TicToc('Training')
+trainTimeT = TicToc('TrainingT')
+# Constants 
+batchSize = trainSize 
+timestr = time.strftime("%Y%m%d-%H%M%S")
+Mep_trainOptions = np.array([2**17, 2**18])# 2**13, 2**14, 2**15, 2**16])
+train_p = 0.07
+
+myDir = 'Simulation'+'_bs_'+str(batchSize)+'_Mepmin_'+str(min(Mep_trainOptions))+'_Mepmax_'+str(max(Mep_trainOptions))
+fn.createDir('Decoder_Simulations')
+directory = 'Decoder_Simulations\\'+ myDir
+fn.createDir(directory)
+fn.createDir(directory+'\\ArrayModels')
+fn.createDir(directory+'\\1HModels')
+
+paths1 = ["" for x in range(len(Mep_trainOptions))]
+paths2 = ["" for x in range(len(Mep_trainOptions))]
+trainTimeT.tic()
+for i in range(len(Mep_trainOptions)):
+    trainTime.tic()
+    numEpochs = Mep_trainOptions[i]
+        
+    # Array decoder
+    title = 'MLNN'
+    paths1[i] = directory+'\\ArrayModels\\'+'i_'+str(i)+'_'+title+'_Mep_'+str(numEpochs)+'.h5'
+    path = paths1[i]
+    exec(open("MLNN-train.py").read())
+    
+    # One hot decoder
+    title = 'MLNN1H'
+    paths2[i] = directory+'\\1HModels\\'+'i_'+str(i)+'_'+title+'_Mep_'+str(numEpochs)+'.h5'
+    path = paths2[i]
+    exec(open("MLNN1H-train.py").read())
+    
+    trainTime.toc()
+    print('Train time for models of wave '+str(i)+': ',trainTime.elapsed)
+trainTimeT.toc()
+print('Total Train time for models: ',trainTimeT.elapsed)
+    
+    
+#%%
+'''
+    Multi Prediction Array Decoder
+'''
+predictTime = TicToc('Predict')
+predictTimeT = TicToc('PredictT')
+# Initialize variables
+multiPredictionsA = np.empty([len(Mep_trainOptions),len(pOptions)])
[…]

### 2019-06-05 Rapport/Article/Article.tex
Seance 06/05/2019

@@
 Communication channel error correction is key to enable digital critical communication systems to work efficiently.
-Although the optimal decoder for noisy channels, the maximum a posteriori (MAP) decoder, has already a mathematically proven optimal error probability minimization; its implementation cripples system applicability, since it introduces a large delay for large codewords.
+Although the optimal decoder for noisy channels, the maximum a posteriori (MAP) decoder, has already a mathematically proven optimal error probability minimization; its implementation cripples system applicability, since it introduces a large delay for large code words.
 In this context, a deep neural network (DNN) is proposed to optimize the channel with an end-to-end autoencoder.
@@
 in \cite{2018} proved that a deep learning-based channel decoder could actually learn a decoding algorithm rather than just being a simple classifier.
-They introduced codewords that were not been used in the training set, and the trained NN was able to correctly decode it.
+They introduced code words that were not been used in the training set, and the trained NN was able to correctly decode it.
 They also observed that structured codes are easier to learn than unstructured ones.
@@
 Throughout the work, vectors will be written in bold font weight and may  have a subscript which indicates their layer in a NN and may have a superscript which indicates their size length.
-In order to homogenize notation, for the encoding phase of the communication system, $\textbf{u}^k$ represents an input \textit{source message} and $\textbf{x}^n$ its correspondent codeword.
+In order to homogenize notation, for the encoding phase of the communication system, $\textbf{u}^k$ represents an input \textit{source message} and $\textbf{x}^n$ its correspondent code word.
 For the decoding portion, $\textbf{y}^n$ represents the output of the noisy channel and $\hat{\textbf{u}}^k$ the estimated source message.
@@
 
+\subsection{Channel coding}
+Consider the communication system illustrated in Fig.
+\ref{fig:cs}.
+The left portion of the chain, the transmitter, wants to communicate a message $\textbf{u}$ through a noisy channel.
+The right portion of the chain, the receiver, may interpret this possible corrupted stream of bits into the original message.
+Usually this prediction $\hat{\textbf{u}}$ carry errors which we aim to minimize its probability of occurring.
+In this scenario, the theory of channel coding discuss possible solutions.
+
+In particular, linear block codes represented by a pair $(n, k)$ called \textit{code name} will be implemented.
+Where $k$ is the length of a source message and $n$ is the length of a code word.
+Briefly, a linear encoder will index a message $\textbf{u}^k \in \mathcal{U} \subseteq \{0,1\}^k$ to a code word $\textbf{x}^n \in \mathcal{X} \subseteq \{0,1\}^n$.
+This mapping is done by the \textit{generator matrix} $ \textbf{G} $.
+\begin{equation}\label{eq:encoding}
+\mathbb{E}:\textbf{u} \mapsto \textbf{uG}
+\end{equation}
+
+Then, the objective of the receiver is to estimate the original message with an empirical error probability, or a BER, defined as
+
+\begin{equation}
+P_{eb} = \frac{1}{M} \underset{u}{\sum}Pr(\hat{u}\neq u)
+\end{equation}
+where $M$ is the total amount of bits transmitted, $u$ represents one bit transmitted and $\hat{u}$ its estimation.
+
 \subsection{Maximum a posteriori decoder}
 
-The concept of a MAP decoder algorithm for sequences, is choosing a message which maximizes the a posterior probability of the corrupted codeword $y$ received by the decoder \cite{b4}.
+The concept of a MAP decoder algorithm for sequences, is choosing a message which maximizes the a posterior probability of the corrupted code word $y$ received by the decoder \cite{b4}.
 This algorithm is known for its optimal error correction capability for white noise interference.
@@
 The statistics of the channel's noise is shown in \cite{2018} to be irrelevant, since the NN extract it during training phase.
-Their applicability is tantamount compelling because the weights computed during training can be separated in two groups, one correspondent to the nodes before the channel and one posterior to the channel.
-As a consequence, we are able to design an non linear encoder and its correspondent non linear decoder with substantial ease.
+The autoencoder applicability is compelling because the weights computed during training can be separated in two groups, one correspondent to the nodes before the channel and one posterior to the channel.
+As a consequence, we are able to design a non linear encoder and its correspondent non linear decoder with ease.
 This new representation has ultimately far more possibilities than conventional encoding patterns.
+
+Mathematically, the non linear autoencoder`s encoding function $\phi: \{0,1\}^k \mapsto \{0,1\}^n $ maps the message input to its code word.
+
+\begin{equation}
+x^n = \phi (u^k)
+\end{equation}
 
@@
 
-\section{Procedure and Methodology}
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/DNN_autoencoder}
+    \caption{Representation of a DNN autoencoder composed of Dense NNs.}\label{fig:DDNNAutoencoder}
+\end{figure}
+
+\section{Implementation and Methodology}
 Explain what I Should do to achieve my goal
@@
 \subsection{Decoder}
-Hamming coding
-
 The linear encoding matrix $\textbf{G}_{8\times16}$ is shown below.

### 2019-06-05 MyCode/autoencoder-retrain.py

@@ -1,0 +1,55 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Jun  5 13:47:36 2019
+
+@author: user
+"""
+#%%
+numEpochs = 2**16
+directory = 'Retrain\AutoencoderArray\\'
+fileName = 'i_1_AutoencoderArray_Mep_65536_p_0.03.h5'
+Autoencoder = tf.keras.models.load_model(directory+'OriginalModel\\'+fileName)
+print("Loaded models from disk")
+path = directory+'TrainedModel\\'+fileName
+
+'''
+    Overall Settings/ Compilationl
+'''
+lossFunc = 'mse'
+Autoencoder.compile(loss=lossFunc ,
+              optimizer='adam',
+              )
+'''
+    Summaries and checkpoints (to do)
+'''
+summary = Autoencoder.summary()
+checkpoint = tf.keras.callbacks.ModelCheckpoint(
+        checkpointPath, monitor='loss', 
+        verbose=0, save_best_only=True, save_weights_only=False, mode='min', period=checkpointPeriod)
+callbacks_list = [checkpoint]
+''' 
+    Training
+'''
+history = Autoencoder.fit(x_train_data, u_train_labels, epochs=numEpochs, 
+                   batch_size=batchSize, shuffle=True, verbose=0, callbacks=callbacks_list)
+
+plotTraining(history)
+
+'''
+    Saving model
+'''
+Autoencoder.save(path)  # creates a HDF5 file
+
+Encoder = Autoencoder.layers[0]
+Decoder = Autoencoder.layers[2]
+#% Prediction
+globalErrorAutoencoder=fn.arrayAutoencoderPrediction(Encoder, Decoder, pOptions, globalReps, N, n, k)
+
+#% Plotting
+figPath = directory+'\\images\\'+fileName
+fn.createDir(figPath)
+numEpochs = 2**17
+plotBERp(globalErrorAutoencoder, 'Array Autoencoder')
+
+
+

### 2019-06-05 MyCode/autoencoderInitializer.py

@@ -34,7 +34,7 @@ trainSize = np.size(x_train_data, 0)
 '''
 numEpochs = 2**16  #2**16 approx 65000
 batchSize = 256
-train_p = 0.07
+train_p = 0.03
 timestr = time.strftime("%Y%m%d-%H%M%S")
 title = 'AutoencoderArray'
 path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'

### 2019-06-05 MyCode/libs-and-global-consts.py

@@ -85,6 +85,7 @@ globalReps = 1000
 checkpointPath = 'Checkpoints/'+title+'/'+timestr+'_'+title+'_Mep_{epoch:02d}-{loss:.8f}.h5'
 checkpointPeriod = 2**12
 trainingPath = 'training_history/'+title+'/'+timestr + '_'+title+'_train.png'
+figPath = 'images/'+title+'/'+timestr+'_MAP_'+title+'_Mep_'+str(numEpochs)+'_ptrain_'+str(train_p)+'.png'
 
 #% Global Functions
 def tensorBSC(x):
@@ -111,7 +112,7 @@ def plotBERp(globalErrorMLNN, legendEntry):
     plt.show()
     
     fig.set_size_inches(width, height)
-    fig.savefig('images/'+title+'/'+timestr+'_MAP_'+title+'_Mep_'+str(numEpochs)+'_ptrain_'+str(train_p)+'.png', bbox_inches='tight', dpi=300)
+    fig.savefig(figPath, bbox_inches='tight', dpi=300)
 
 def plotTraining(history):
     # summarize history for loss

### 2019-06-05 MyCode/multi-train.py

@@ -90,7 +90,7 @@ plt.plot(pOptions,avgGlobalErrorMAP, color='r', linewidth=lineWidth, label='MAP'
 plt.grid(True, which='both')
 
 for i in range(len(p_trainOptions)):
-    plt.scatter(pOptions,multiPredictions[i], marker=markers[i], zorder=3+i, s=markerSize, label='Autoen., Mep = %s' % p_trainOptions[i])
+    plt.scatter(pOptions,multiPredictions[i], marker=markers[i], zorder=3+i, s=markerSize, label='Autoen., p = %s' % p_trainOptions[i])
     
 plt.xlabel('$p$')
 plt.ylabel('BER')

### 2019-06-07 Rapport/Article/Article.tex
Seance 06/07/2019

@@
 
-Whilst autoencoders are more complex dimensional wise when compared to NN decoders, the readily access to graphical processing units (GPUs) contributed to the feasibility of this technique, tackling the \textit{curse of dimensionality} challenge in ML.
-This type of hardware made possible the approach of the layer-by-layer unsupervised pre-training followed by stochastic gradient descent applied in this work.
-\cite{GPU}
+Whilst autoencoders are more complex dimensional wise compared to NN decoders, the readily access to graphical processing units (GPUs) contributed to the feasibility of this technique, tackling the \textit{curse of dimensionality} challenge in ML.
+This type of hardware made possible the approach of the layer-by-layer supervised training by stochastic gradient descent optimization applied in this work.
+\cite{GPU}.
+
+A common architecture of DNN autoencoders is shown in Fig.
+\ref{fig:DDNNAutoencoder} where the encoder and the decoder are composed of dense layers of different widths.
+In addition, a noise layer and a rounding layer acts on the forward propagation but not on the BP, since their activation functions are not differentiable.
+More details on the autoencoder's architecture is discussed is section III.
 
@@
   \centering
-    \includegraphics[width=0.5\textwidth]{images/DNN_autoencoder}
-    \caption{Representation of a DNN autoencoder composed of Dense NNs.}\label{fig:DDNNAutoencoder}
+    \includegraphics[width=0.4\textwidth]{images/DNN_autoencoder}
+    \caption{Representation of a DNN autoencoder composed of dense layers.}\label{fig:DDNNAutoencoder}
 \end{figure}
@@
 \section{Implementation and Methodology}
-Explain what I Should do to achieve my goal
-Talk about decoder + autoencoder and the OBJECTIVE of each one, G matrix of choice, Batch size, The reader should be able to replicate what I ve done
+This section explains the main implementation procedures in order to replicate the results and understand in depth the methodology behind the work.
+For organizational purposes, the section is divided into two subsections, where the first concerns the experiments relating to the ML based decoders and the second treats the implementation of the ML based autoencoders.
 
@@
 BER Curves of simulation varying p, best choice of p, analyse autoencoder with best choice of p
-
-\begin{figure}[!ht]
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/MAP_MAP_Mep_65536}
-    \caption{Test caption to test image.}\label{fig:ae1}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/MAP_AutoencoderArray_Mep_65536}
-    \caption{Test caption to test image.}\label{fig:ae1}
-\end{figure}
 

### 2019-06-07 MyCode/autoencoder-retrain.py

@@ -46,10 +46,9 @@ Decoder = Autoencoder.layers[2]
 globalErrorAutoencoder=fn.arrayAutoencoderPrediction(Encoder, Decoder, pOptions, globalReps, N, n, k)
 
 #% Plotting
-figPath = directory+'\\images\\'+fileName
-fn.createDir(figPath)
+figPath = directory+'AutoencoderArray_64-128-256_Mep_131072_p_0.03.png'
 numEpochs = 2**17
-plotBERp(globalErrorAutoencoder, 'Array Autoencoder')
+plotBERp(globalErrorAutoencoder, 'Autoencoder')
 
 
 

### 2019-06-07 MyCode/autoencoderArray.py

@@ -25,7 +25,7 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([64, 128, 256, 16])
+encoderNodes = np.array([64, 64, 64, 16])
 DecoderNodes = [128, 64, 32, k]
 #%
 '''
@@ -97,7 +97,6 @@ Autoencoder.save(path)  # creates a HDF5 file
 '''
     Prediction Array
 '''
-'''
 globalReps = 1000
 globalErrorAutoencoder = np.empty([globalReps, len(pOptions)])
 for i_global in range(globalReps):
@@ -114,7 +113,7 @@ for i_global in range(globalReps):
 
 #% Plotting
 plotBERp(globalErrorAutoencoder, 'Array Autoencoder')
-'''
+
 
 
 

### 2019-06-07 MyCode/libs-and-global-consts.py

@@ -2,6 +2,7 @@
 '''
        Load libraries
 '''
+#%%%
 import numpy as np
 from scipy import stats
 import matplotlib.pyplot as plt
@@ -81,6 +82,7 @@ C = 1-He # channel capacity (R<C)
 # Simulation
 globalReps = 1000
 
+#%%
 # NN parameters
 checkpointPath = 'Checkpoints/'+title+'/'+timestr+'_'+title+'_Mep_{epoch:02d}-{loss:.8f}.h5'
 checkpointPeriod = 2**12
@@ -108,7 +110,7 @@ def plotBERp(globalErrorMLNN, legendEntry):
     plt.xlabel('$p$')
     plt.ylabel('BER')
     plt.yscale('log')
-    plt.legend(['No Decoding', 'MAP', legendEntry+ ', $M_{ep}=$'+str(numEpochs)])
+    plt.legend(['No Decoding', 'MAP', legendEntry+ ', $p_t=$'+str(train_p)])
     plt.show()
     
     fig.set_size_inches(width, height)

### 2019-06-07 Rapport/Article/Article.tex
Seance 06/07/2019

@@
 
-\usepackage{algorithmic}
+%\usepackage{algorithmic}
 \usepackage{graphicx}
@@
 \usepackage{verbatim}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+
+\makeatletter
+\def\BState{\State\hskip-\ALG@thistlm}
+\makeatother
 
@@
 
-Then, the objective of the receiver is to estimate the original message with an empirical error probability, or a BER, defined as
+The objective of the receiver is to estimate the original message. The empirical error probability, or a BER, in this process is defined as
 
@@
 where $M$ is the total amount of bits transmitted, $u$ represents one bit transmitted and $\hat{u}$ its estimation.
+
+The Hamming distance $d_H:\mathcal{X}\mapsto \mathbb{Z}$ is a metric in the word subset which in the context of linear coding is equal to equation \ref{eq:dh}.
+Noteworthy for the MAP rule, a non zero Hamming distance indicates that there is a block error, and ultimately at least a binary error in the received code word.
+\begin{equation}\label{eq:dh}
+d_H(\textbf{x},\textbf{y}) = \sum_{i=1}^{n} y_i \text{ XOR } x_i 
+\end{equation}
 
@@
 This type of hardware made possible the approach of the layer-by-layer supervised training by stochastic gradient descent optimization applied in this work.
-\cite{GPU}.
+\cite{GPU}
 
@@
 This section explains the main implementation procedures in order to replicate the results and understand in depth the methodology behind the work.
-For organizational purposes, the section is divided into two subsections, where the first concerns the experiments relating to the ML based decoders and the second treats the implementation of the ML based autoencoders.
+For organizational purposes, the section is divided into three subsections.
+First, the MAP algorithm for a linear code is shown, then the experiments relating to the ML based decoders are assessed and finally the implementation of the ML based autoencoder is explained.
+
+\subsection{MAP Rule}
+
+The pseudo code of the MAP decoder algorithm is available below.
+
+\begin{algorithm}
+\caption{MAP rule for BSC and linear block code.}\label{euclid}
+\hspace*{\algorithmicindent} \textbf{Input:} received block $\textbf{y}^n \in \{0,1\}^n$, the code word set $\mathcal{X}$ and generator matrix $G_{k \times n}$. \\
+\hspace*{\algorithmicindent} \textbf{Output:} message estimation $\hat{\textbf{u}}^k \in \{0,1\}^k$.
+\begin{algorithmic}[H]
+\Procedure{MAP Decoder}{$y,\mathcal{X}, G$}
+\State $\textit{p} \gets \text{channel crossover probability}$
+\For{i in \textit{range}($2^k$)}
+\State $\text{distances}[i] \gets d_H(\textbf{y},word[i] \in \mathcal{X})$
+\EndFor
+\State $\hat{\textbf{x}} \gets argmin(\text{distances})$
+\State $\hat{\textbf{u}} $
+\Return $\hat{\textbf{u}}$
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/20190607-160812_MAP}
+    \caption{MAP algorithm performance in terms of BER versus channel crossover probability.}\label{fig:MAP}
+\end{figure}
 
@@
     \includegraphics[width=0.5\textwidth]{images/MAP_MLNN1H_p_0}
-    \caption{Test caption to test image.}\label{fig:1HD}
+    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/20190529-142119_MAP_MLNN_Mep_65536}
+    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
 \end{figure}

### 2019-06-07 MyCode/BSC-MAP-Decoder.py

@@ -57,14 +57,14 @@ avgGlobalErrorMAP = np.average(globalErrorMAP, 0)
 
 fig = plt.figure(figsize=(8, 6), dpi=80)
 
-plt.plot(pOptions,avgGlobalError, color='b', linewidth=lineWidth)
+plt.plot(pOptions,avgGlobalError, color='b', linewidth=lineWidth, linestyle='--')
 plt.plot(pOptions,avgGlobalErrorMAP, color='r', linewidth=lineWidth)
 
 plt.grid(True, which="both")
 plt.xlabel('$p$')
 plt.ylabel('BER')
 plt.yscale('log')
-plt.legend(['No Decoding', 'MAP'])
+plt.legend(['No Decoding', 'MAP Algorithm'])
 plt.show()
 
 timestr = time.strftime("%Y%m%d-%H%M%S")

### 2019-06-07 MyCode/autoencoder-retrain.py

@@ -11,7 +11,7 @@ fileName = 'i_1_AutoencoderArray_Mep_65536_p_0.03.h5'
 Autoencoder = tf.keras.models.load_model(directory+'OriginalModel\\'+fileName)
 print("Loaded models from disk")
 path = directory+'TrainedModel\\'+fileName
-
+#%%
 '''
     Overall Settings/ Compilationl
 '''
@@ -39,7 +39,7 @@ plotTraining(history)
     Saving model
 '''
 Autoencoder.save(path)  # creates a HDF5 file
-
+#%%
 Encoder = Autoencoder.layers[0]
 Decoder = Autoencoder.layers[2]
 #% Prediction

### 2019-06-07 MyCode/autoencoderInitializer.py

@@ -32,7 +32,7 @@ trainSize = np.size(x_train_data, 0)
 '''
     Constants
 '''
-numEpochs = 2**16  #2**16 approx 65000
+numEpochs = 2**14  #2**16 approx 65000
 batchSize = 256
 train_p = 0.03
 timestr = time.strftime("%Y%m%d-%H%M%S")

### 2019-06-07 MyCode/libs-and-global-consts.py

@@ -25,7 +25,7 @@ from ttictoc import TicToc
 #plt.rcParams['mathtext.fontset'] = 'dejavuserif'
 plt.rcParams["font.family"] = "Times New Roman"
 plt.rcParams.update({'font.weight':'normal'})
-letterSize = 7
+letterSize = 6
 markerSize = 3
 lineWidth = 0.5
 plt.rc('xtick', labelsize=letterSize)
@@ -110,7 +110,7 @@ def plotBERp(globalErrorMLNN, legendEntry):
     plt.xlabel('$p$')
     plt.ylabel('BER')
     plt.yscale('log')
-    plt.legend(['No Decoding', 'MAP', legendEntry+ ', $p_t=$'+str(train_p)])
+    plt.legend(['No Decoding', 'MAP Algorithm', legendEntry+ ', $p_t=$'+str(train_p)])
     plt.show()
     
     fig.set_size_inches(width, height)

### 2019-06-12 Rapport/Article/Article.tex
Seance 06/12/2019

@@
 Communication channel error correction is key to enable digital critical communication systems to work efficiently.
-Although the optimal decoder for noisy channels, the maximum a posteriori (MAP) decoder, has already a mathematically proven optimal error probability minimization; its implementation cripples system applicability, since it introduces a large delay for large code words.
+The maximum a posteriori (MAP) decoder is proven to be the mathematically optimal solution for the problem.
+However, its implementation cripples the system applicability.
+This is because the MAP rule introduces a large delay for long code words.
 In this context, a deep neural network (DNN) is proposed to optimize the channel with an end-to-end autoencoder.
-The DNN autoencoder outperforms a MAP decoder form a given encoder in terms of delay, because of the first`s \textit{one-shot} capability once trained.
+The trained DNN autoencoder has a \textit{one-shot} capability which outperforms a MAP decoder for a given encoder in terms of delay.
+ONE PHRASE FOR RESULTS
 
@@
 \State $\hat{\textbf{x}} \gets argmin(\text{distances})$
-\State $\hat{\textbf{u}} $
+\State $\hat{\textbf{u}} \gets G^{-1}\hat{\textbf{x}}$
 \Return $\hat{\textbf{u}}$

### 2019-06-12 MyCode/autoencoder1H.py

@@ -19,7 +19,7 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([256, 256, 256])
+encoderNodes = np.array([256, 256])
 DecoderNodes = [256, 256]
 #%
 '''
@@ -32,19 +32,15 @@ path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(
 Encoder = tf.keras.Sequential([
         # Input Layer
         layers.Dense(encoderNodes[0], activation='relu', input_shape=(256,), name='Input'),
-        layers.BatchNormalization(axis=-1, momentum=0.99, epsilon=0.001, center=True, 
-            scale=True, beta_initializer='zeros', gamma_initializer='ones', 
-            moving_mean_initializer='zeros', moving_variance_initializer='ones'),
         # Hidden Layer
         layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
-        layers.BatchNormalization(axis=-1, momentum=0.99, epsilon=0.001, center=True, 
-            scale=True, beta_initializer='zeros', gamma_initializer='ones', 
-            moving_mean_initializer='zeros', moving_variance_initializer='ones'),
         # Hidden Layer
         #layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
         # Coded Layer
         layers.Dense(n, activation='sigmoid', name='Codedfloat'),
-        #layers.BatchNormalization(axis=-1, momentum=0.99, epsilon=0.001, center=True, scale=True, beta_initializer='zeros', gamma_initializer='ones', moving_mean_initializer='zeros', moving_variance_initializer='ones', beta_regularizer=None, gamma_regularizer=None, beta_constraint=None, gamma_constraint=None)
+        layers.BatchNormalization(axis=-1, momentum=0.99, epsilon=0.001, center=True, 
+            scale=True, beta_initializer='zeros', gamma_initializer='ones', 
+            moving_mean_initializer='zeros', moving_variance_initializer='ones'),
         ], name='Encoder')
 
 NoiseL = tf.keras.Sequential([

### 2019-06-12 MyCode/encoder.py

@@ -13,18 +13,18 @@ path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(
 u_train_labels = possibleCodewords.copy()
 x_train_data = fn.messages2onehot(messages)
 
-u_train_labels = np.repeat(u_train_labels, 2, axis=0)
-x_train_data = np.repeat(x_train_data, 2, axis=0)
+u_train_labels = np.repeat(u_train_labels, 1, axis=0)
+x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([256, 256])
+encoderNodes = np.array([32, 32])
 
 '''
     Architecture
 '''
 Encoder = tf.keras.Sequential([
         # Input Layer
-        layers.Dense(encoderNodes[0], activation='relu', input_shape=(k,), name='Input'),
+        layers.Dense(encoderNodes[0], activation='relu', input_shape=(256,), name='Input'),
         #layers.Conv1D(12, kernel_size= 4, activation='relu', input_shape=(k,), name='Input'),
         # Dropout Layer
         #layers.Dropout(rate=0.01),
@@ -40,7 +40,7 @@ Encoder = tf.keras.Sequential([
         #layers.BatchNormalization(axis=-1, momentum=0.99, epsilon=0.001, center=True, scale=True, beta_initializer='zeros', gamma_initializer='ones', moving_mean_initializer='zeros', moving_variance_initializer='ones')
         ], name='Encoder')
 
-trainingPath = 'training_history/'+title+'/'+timestr + '_'+title+'_train' + str(encoderNodes) +'conv1d.png'
+trainingPath = 'training_history/'+title+'/'+timestr + '_'+title+'_train_' + str(encoderNodes) +'_1H.png'
 plot_model(Encoder,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',show_shapes=True)
 
 '''
@@ -115,3 +115,23 @@ for i_global in range(globalReps):
 
 #% Plotting
 plotBERp(globalErrorMAP2, 'MAP with NN Encoder')
+
+#%%
+'''
+    Predicition one-hot encoder
+'''
+globalReps = 100
+globalErrorEncoder = np.empty([globalReps, len(pOptions)])
+for i_global in range(globalReps):
+    for i_p in range(np.size(pOptions)):
[…]

### 2019-06-14 Rapport/Article/Article.tex
Seance 06/14/2019

@@
 
-%\usepackage{algorithmic}
 \usepackage{graphicx}
@@
 \usepackage[noend]{algpseudocode}
+\usepackage{multirow}
+\usepackage{url}
+
+\usepackage[utf8]{inputenc}
+\usepackage{ifthen}
+\usepackage{filecontents}
+\usetikzlibrary{shapes,arrows,shadings,patterns}
+\usepackage{pgfplots}
+\pgfplotsset{compat=newest}
+\pgfplotsset{plot coordinates/math parser=false}
 
@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\makeatletter
+\def\endthebibliography{%
+  \def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}%
+  \endlist
+}
+\makeatother
+
+\makeatletter
+\def\thickhline{%
+  \noalign{\ifnum0=`}\fi\hrule \@height \thickarrayrulewidth \futurelet
+   \reserved@a\@xthickhline}
+\def\@xthickhline{\ifx\reserved@a\thickhline
+               \vskip\doublerulesep
+               \vskip-\thickarrayrulewidth
+             \fi
+      \ifnum0=`{\fi}}
+\makeatother
+
+\newlength{\thickarrayrulewidth}
+\setlength{\thickarrayrulewidth}{3\arrayrulewidth}
+
 \begin{document}
@@
 
-\author{\IEEEauthorblockN{Eduardo Dadalto Camara Gomes\IEEEauthorrefmark{1}}
+\author{
+\IEEEauthorblockN{Eduardo Dadalto Camara Gomes\IEEEauthorrefmark{1},
+Meryem Benammar\IEEEauthorrefmark{2}}
+
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
-Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr}
+Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr
 }
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
+Email: meryem.benammar@isae-supaero.fr
+}
+}
+
+\IEEEspecialpapernotice{(Final report)}
 
@@
 This algorithm is known for its optimal error correction capability for white noise interference.
-This algorithm is also referred as the Viterbi decoder \cite{viterbi}.
+This algorithm is also referred as the Viterbi decoder \cite{Viterbi}.
 
@@
 \begin{algorithm}
-\caption{MAP rule for BSC and linear block code.}\label{euclid}
-\hspace*{\algorithmicindent} \textbf{Input:} received block $\textbf{y}^n \in \{0,1\}^n$, the code word set $\mathcal{X}$ and generator matrix $G_{k \times n}$. \\
+\caption{MAP rule for BSC and linear block code.}\label{alg:MAP}
+\hspace*{\algorithmicindent} \textbf{Input:} received block $\textbf{y}^n \in \{0,1\}^n$, code word set $\mathcal{X}$ and generator matrix $G_{k \times n}$. \\
 \hspace*{\algorithmicindent} \textbf{Output:} message estimation $\hat{\textbf{u}}^k \in \{0,1\}^k$.
@@
 Architecture, Training parameters, cite choice of p train \cite{b13}, number of epochs, Talk about noise layer and its removal in prediction,
+
+\begin{table}
+\caption{DNN array decoder architecture and parameters.} 
+\begin{center}
+
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{3}{*}{Decoder} & Dense: X, activation: ReLU, input size: $n$ \\
+ & Dense: X, activation: ReLU  \\
+ & Dense: $k$, activation: Sigmoid  \\ \hline
+    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:arraydecoder}
+\end{center}
+\end{table}
+
 \subsubsection{One-hot Decoder}
@@
 
+\begin{table}
+\caption{DNN one-hot decoder architecture and parameters.} 
+\begin{center}
+
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{3}{*}{Decoder} & Dense: X, activation: ReLU, input size: $n$ \\
+ & Dense: X, activation: ReLU  \\
+ & Dense: $2^k$, activation: Sigmoid  \\ \hline
+    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:onehotdecoder}
+\end{center}
+\end{table}
+
 \subsection{Autoencoder}
 Architecture, Training and simulation parameters, procedure to find a M epoch suitable, choices of p train, Talk about NN separation,
+
+\begin{table}
+\caption{DNN array autoencoder architecture and parameters.} 
+\begin{center}
+
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: $k$ \\
+ & Dense: X, activation: ReLU  \\
+ & Dense: $n$, activation: Sigmoid  \\ \hline
+
+    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $n$ \\
+ & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
+
+ \multirow{2}{*}{Decoder} & Dense: X, input size: $n$ \\
+ & Dense: X, activation: ReLU  \\
+ & Dense: $k$, activation: Sigmoid  \\ \hline
+
+    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:arrayautoencoder}
+\end{center}
+\end{table}
 
@@
 
-\begin{comment}
+\section*{Acknowledgment}
+The authors would like to thank...
 
-\subsection{Figures and Tables}
+\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18},
 
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
+\bibliographystyle{plain}
+\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18}
 
-\begin{figure}[htbp]
-\centerline{\includegraphics{fig1.png}}
-\caption{Example of a figure caption.}
-\label{fig}
-\end{figure}
-\end{comment}
+%\bibitem{b1} C. E. Shannon, ``A mathematical theory of communication," Bell System Technical Journal, vol. 27, no. 3, pp. 379–423, Jul. 1948.
 
-\begin{thebibliography}{00}
-\bibitem{b1} C. E. Shannon, ``A mathematical theory of communication," Bell System Technical Journal, vol. 27, no. 3, pp. 379–423, Jul. 1948.
-
-\bibitem{b3} F.
+%\bibitem{b3} F.
 D.
@@
 
-\bibitem{b2} O'Shea, Tim \& Hoydis, Jakob, ``An Introduction to Machine Learning Communications Systems", 2017.
+%\bibitem{b2} O'Shea, Tim \& Hoydis, Jakob, ``An Introduction to Machine Learning Communications Systems", 2017.
 
-\bibitem{osheaautoencoder} T. O'Shea, K. arra, T. C. Clancy, ``Learning to Communicate: Channel Auto-encoders, Domain Specific Regularizers, and Attention", 2016.
+%\bibitem{osheaautoencoder} T. O'Shea, K. arra, T. C. Clancy, ``Learning to Communicate: Channel Auto-encoders, Domain Specific Regularizers, and Attention", 2016.
 
-\bibitem{2018} T. Gruber, S. Cammerer, J. Hoydis, and S. ten Brink, “On deep learning based channel decoding,” accepted for CISS 2017, arXiv preprint arXiv:1701.07738, 2017.
+%\bibitem{2018} T. Gruber, S. Cammerer, J. Hoydis, and S. ten Brink, “On deep learning based channel decoding,” accepted for CISS 2017, arXiv preprint arXiv:1701.07738, 2017.
 
-\bibitem{b4} W. Alexander, H. Peter \& W. Norbert, ``Turbo-Decoding Without SNR Estimation",IEEE Communications Letter, vol. 4, no. 6, pp. 193-195, 2000.
+%\bibitem{b4} W. Alexander, H. Peter \& W. Norbert, ``Turbo-Decoding Without SNR Estimation",IEEE Communications Letter, vol. 4, no. 6, pp. 193-195, 2000.
 
-\bibitem{viterbi} A. J. Viterbi, ``Error Bounds for Convolutional Codes and an Asymptotically Optimum Decoding Algorithm", IEEE Transactions on Information Theory, 13 (2): 260–269, 1967.
+%\bibitem{viterbi} A. J. Viterbi, ``Error Bounds for Convolutional Codes and an Asymptotically Optimum Decoding Algorithm", IEEE Transactions on Information Theory, 13 (2): 260–269, 1967.
 
-\bibitem{b6} Robertson, P.
+%\bibitem{b6} Robertson, P.
 , Hoeher, P.
@@
 
-\bibitem{b7} J. Hagenauer,P. Robertson, L. Papke: iterative (“Turbo”) decoding of systematic convolutional codes with the MAP and SOVA
-algorithms. In: ITG-Fachbericht 130. October 1994, p. 21-29.
+%\bibitem{b7} J. Hagenauer,P. Robertson, L. Papke: iterative (“Turbo”) decoding of systematic convolutional codes with the MAP and SOVA algorithms. In: ITG-Fachbericht 130. October 1994, p. 21-29.
 
-\bibitem{b5} M. Jordan and R. Nichols, “The effects of channel characteristics on turbo code performance,” in Proc. Milcom’96, McLean, VA, Oct. 1996, pp. 17–21.
+%\bibitem{b5} M. Jordan and R. Nichols, “The effects of channel characteristics on turbo code performance,” in Proc. Milcom’96, McLean, VA, Oct. 1996, pp. 17–21.
 
-\bibitem{b9} I. Mohamed, ``Applications of neural networks to digital communications - a survey", 1997.
+%\bibitem{b9} I. Mohamed, ``Applications of neural networks to digital communications - a survey", 1997.
 
-\bibitem{b8} M. Nielsen, \textit{Neural Networks and Deep Learning}.
+%\bibitem{b8} M. Nielsen, \textit{Neural Networks and Deep Learning}.
 
-\bibitem{b10} M. Kevin, ``Machine Learning: A Probabilistic Perspective", MIT, ISBN 978-0262018029, 2012.
+%%\bibitem{b10} M. Kevin, ``Machine Learning: A Probabilistic Perspective", MIT, ISBN 978-0262018029, 2012.
 
-\bibitem{b11} M. Abadi, A. Agarwal, P. Barham, et al., 
-``TensorFlow: Large-scale machine learning on heterogeneous systems", 2015. Software available from tensorflow.org.
+%\bibitem{b11} M. Abadi, A. Agarwal, P. Barham, et al.,``TensorFlow: Large-scale machine learning on heterogeneous systems", 2015. Software available from tensorflow.org.
 
-\bibitem{b12} F. Chollet, “keras,” https://github.com/fchollet/keras, 2015.
+%\bibitem{b12} F. Chollet, “keras,” https://github.com/fchollet/keras, 2015.
 
-\bibitem{GPU} G. E. Hinton, S. Osindero, and Y.-W. Teh, “A fast learning algorithm for deep belief nets,” Neural Computation, vol. 18, no. 7, pp. 1527–1554, July 2006.
+%\bibitem{GPU} G. E. Hinton, S. Osindero, and Y.-W. Teh, “A fast learning algorithm for deep belief nets,” Neural Computation, vol. 18, no. 7, pp. 1527–1554, July 2006.
 
-\bibitem{b13} M. Benammar and P. Piantanida, “Optimal training channel statistics for neural-based decoders,” in Asilomar, Oct 2018, pp. 2157–2161.
-
-\end{thebibliography}
+%\bibitem{b13} M. Benammar and P. Piantanida, “Optimal training channel statistics for neural-based decoders,” in Asilomar, Oct 2018, pp. 2157–2161.
 

### 2019-06-14 MyCode/autoencoder1H.py

@@ -19,13 +19,9 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([256, 256])
-DecoderNodes = [256, 256]
+encoderNodes = np.array([256, 1024])
+DecoderNodes = [128, 64]
 #%
-'''
-    Constants
-'''
-path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
 '''
     Architecture
 '''
@@ -38,9 +34,6 @@ Encoder = tf.keras.Sequential([
         #layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
         # Coded Layer
         layers.Dense(n, activation='sigmoid', name='Codedfloat'),
-        layers.BatchNormalization(axis=-1, momentum=0.99, epsilon=0.001, center=True, 
-            scale=True, beta_initializer='zeros', gamma_initializer='ones', 
-            moving_mean_initializer='zeros', moving_variance_initializer='ones'),
         ], name='Encoder')
 
 NoiseL = tf.keras.Sequential([
@@ -88,12 +81,12 @@ plotTraining(history)
 '''
     Saving model
 '''
-Autoencoder1H.save('Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5')  # creates a HDF5 file
+Autoencoder1H.save(path)  # creates a HDF5 file
 
-#%
+#%%%
 '''
     Prediction 1H
-'''
+
 t = TicToc('name')
 t.tic()
 
@@ -105,7 +98,7 @@ for i_global in range(globalReps):
         p = pOptions[i_p]
         u = fn.generateU(N,k)
         u1h = fn.messages2onehot(u)
-        x = Encoder.predict(u1h)
+        x = np.round(Encoder.predict(u1h))
         xflat = np.reshape(x, [-1])
         yflat = fn.BSC(xflat,p)
         y = yflat.reshape(N,2*k) # noisy codewords
@@ -119,3 +112,4 @@ plotBERp(globalErrorAutoencoder1H, 'One-hot Autoencoder')
 
 t.toc()
 print(t.elapsed)
+'''

### 2019-06-14 MyCode/libs-and-global-consts.py

@@ -86,10 +86,11 @@ globalReps = 1000
 # NN parameters
 title = 'Autoencoder1H'
 timestr = time.strftime("%Y%m%d-%H%M%S")
-numEpochs = 2**16
+numEpochs = 2**17
 batchSize = 256
 train_p = 0.0
 
+path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
 checkpointPath = 'Checkpoints/'+title+'/'+timestr+'_'+title+'_Mep_{epoch:02d}-{loss:.8f}.h5'
 checkpointPeriod = 2**11
 trainingPath = 'training_history/'+title+'/'+timestr + '_'+title+'_train.png'

### 2019-06-14 MyCode/multi-train.py

@@ -11,14 +11,14 @@ Created on Sat Jun  1 13:41:33 2019
 trainTime = TicToc('Training')
 trainTimeT = TicToc('TrainingT')
 # Constants 
-numEpochs = 2**16  #2**16 approx 65000
+numEpochs = 2**17  #2**16 approx 65000
 batchSize = trainSize 
 timestr = time.strftime("%Y%m%d-%H%M%S")
-title = 'AutoencoderArray'
-p_trainOptions = np.array([0.020, 0.025, 0.03, 0.035, 0.04, 0.07])
+title = 'Autoencoder1H'
+p_trainOptions = np.array([0.0, 0.01, 0.02, 0.03, 0.05, 0.07])
 
-myDir = 'Simulation_Mep_64_128_256_'+str(numEpochs)+'_bs_'+str(batchSize)+'_pmin_'+str(min(p_trainOptions))+'_pmax_'+str(max(p_trainOptions))
-directory = 'Autoencoder_Simulations\\'+ myDir
+myDir = 'Simulation_ESTRUTURA1_-_ESTRUTURA2_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'_pmin_'+str(min(p_trainOptions))+'_pmax_'+str(max(p_trainOptions))
+directory = 'Autoencoder1H_Simulations\\'+ myDir
 fn.createDir(directory)
 fn.createDir(directory+'\\models')
 
@@ -31,7 +31,7 @@ for i in range(len(p_trainOptions)):
     paths[i] = directory+'\\models\\'+'i_'+str(i)+'_'+title+'_Mep_'+str(numEpochs)+'_p_'+str(train_p)+'.h5'
     path = paths[i]
     checkpointPath = directory+'\\models\\'+'i_'+str(i)+'_'+title+'_p_'+str(train_p)+'_Mep_{epoch:02d}-{loss:.6f}.h5'
-    exec(open("autoencoderArray.py").read())
+    exec(open("autoencoder1H.py").read())
     
     trainTime.toc()
     print('Train time for model '+str(i)+': ',trainTime.elapsed)
@@ -64,18 +64,19 @@ for i_train in range(len(p_trainOptions)):
         for i_p in range(np.size(pOptions)):
             p = pOptions[i_p]
             u = fn.generateU(N,k)
-            x = Encoder.predict(u)
+            u1h = fn.messages2onehot(u)
+            x = np.round(Encoder.predict(u1h))
[…]

### 2019-06-14 Rapport/Article/Article.tex
Seance 06/14/2019

@@
 \usepackage{url}
-
+%\usepackage[sorting=none]{biblatex}
 \usepackage[utf8]{inputenc}
@@
 \ref{fig:cs} illustrates.
-As a result of channel imperfections, Shannon theorized in \cite{b1} the ultimate reliable data rate that can be transmitted through a communication system with an arbitrarily small probability error.
+As a result of channel imperfections, Shannon theorized in \cite{Shannon:2001:MTC:584091.584093} the ultimate reliable data rate that can be transmitted through a communication system with an arbitrarily small probability error.
 
@@
 Calabrese et al.
-in \cite{b3} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance of the network.
+in \cite{DBLP:journals/corr/CalabreseWGPS16} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance of the network.
 
-Based on this vision, Tim O'Shea and Jakob Hoydis \cite{b2} pertinently noted that traditional algorithms in the field have foundations in probability theory e.g.
+Based on this vision, Tim O'Shea and Jakob Hoydis \cite{DBLP:journals/corr/OSheaH17} pertinently noted that traditional algorithms in the field have foundations in probability theory e.g.
 maximum a posteriori (MAP) rule, maximum likelihood decoder (MLD) and turbo codes.
@@
 O'Shea et al.
-in \cite{osheaautoencoder} developed a channel autoencoder with optimized impairment and regularization layers to emulate channel impairments.
+in \cite{2016arXiv160806409O} developed a channel autoencoder with optimized impairment and regularization layers to emulate channel impairments.
 They studied this architecture over an additive white Gaussian noise (AWGN) channel, founding ``some promising initial capacity" for this scheme.
@@
 al.
-in \cite{2018} proved that a deep learning-based channel decoder could actually learn a decoding algorithm rather than just being a simple classifier.
+in \cite{2017arXiv171008379G} proved that a deep learning-based channel decoder could actually learn a decoding algorithm rather than just being a simple classifier.
 They introduced code words that were not been used in the training set, and the trained NN was able to correctly decode it.
@@
 
-The concept of a MAP decoder algorithm for sequences, is choosing a message which maximizes the a posterior probability of the corrupted code word $y$ received by the decoder \cite{b4}.
+The concept of a MAP decoder algorithm for sequences, is choosing a message which maximizes the a posterior probability of the corrupted code word $y$ received by the decoder \cite{Worm00turbo-decodingwithout}.
 This algorithm is known for its optimal error correction capability for white noise interference.
@@
 
-However, the MAP decoder is practically nonviable to implement \cite{b6}.
+However, the MAP decoder is practically nonviable to implement \cite{journals/ett/RobertsonHV97}.
 Thus, its variants cited in this paragraph are better suited for practical cases.
@@
 these variants are widely used as decoders for turbo codes \cite{b7}.
-Refer to \cite{b6} for specific details in the derivation of the Log-MAP algorithm.
+Refer to \cite{journals/ett/RobertsonHV97} for specific details in the derivation of the Log-MAP algorithm.
 Furthermore, the Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is a bit-wise MAP rule with lower complexity widely used in practice, even for long data packets.
@@
 Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a BSC was studied in \cite{b5}.
-They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{b6}.
+They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{journals/ett/RobertsonHV97}.
 It means that if the real SNR is smaller, the decoder will not perform well.
@@
 Neural networks (NN) are a set of connected \textit{neurons} able to approximate any kind of function through an architecture composed of several single processing units (neurons) connected together forming a network.
-They are defined by \cite{b9} as parallel distributed, learning- and self-organizing information processing systems.
+They are defined by \cite{Ibnkahla} as parallel distributed, learning- and self-organizing information processing systems.
 
@@
 To define and build this infrastructure, the activation function, the loss function and the structure must be decided.
-Nielson \cite{b8} in his book explores a vast amount of NN architectures and how they work.
+Nielson \cite{nielsenneural} in his book explores a vast amount of NN architectures and how they work.
 
@@
 \ref{fig:NN} will be applied in this work.
-They are known for the universal approximation property \cite{b9} and for the versatility of increasing the number of layers, creating a DNN able to undertake complex classification problems with low classification error and low dimensionality.
+They are known for the universal approximation property \cite{Ibnkahla} and for the versatility of increasing the number of layers, creating a DNN able to undertake complex classification problems with low classification error and low dimensionality.
 In more general terms, a MLNN with $L$ layers, also called \textit{depth}, and parameter $\theta$ is a mapping of an input $\textbf{y}^n \in \mathbb{R}^{n}$ to an output $\textbf{û}^k \in \mathbb{R}^{k}$ through $L$ iterative steps.
@@
 \end{equation}
-where $\textbf{W}_{l}\in \mathbb{R} ^{N_{l}\times N_{l-1}}$ is the weight matrix between layers $l-1$ and $l$, $\textbf{r}_{l-1} \in \mathbb{R} ^{N_{l-1}}$ is a vector containing the hidden layer $l-1$ values, $\textbf{b}_{l}\in \mathbb{R} ^{N_l}$ is the bias vector and $\sigma$ is the activation function \cite{b2}.
+where $\textbf{W}_{l}\in \mathbb{R} ^{N_{l}\times N_{l-1}}$ is the weight matrix between layers $l-1$ and $l$, $\textbf{r}_{l-1} \in \mathbb{R} ^{N_{l-1}}$ is a vector containing the hidden layer $l-1$ values, $\textbf{b}_{l}\in \mathbb{R} ^{N_l}$ is the bias vector and $\sigma$ is the activation function \cite{DBLP:journals/corr/OSheaH17}.
 
@@
 It is derived from the log-likelihood of a training set where $q_j$ is the estimated probability of the outcome $j$ and $p_{j}$ is the true probability.
-Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{b10} where $j=1,...,N_L$.
+Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{murphy2013machine} where $j=1,...,N_L$.
 \begin{equation}\label{eq:c-e}
@@
 It matches the input and its respective desired output.
-The objective is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{b2}.
+The objective is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{DBLP:journals/corr/OSheaH17}.
 \begin{equation}\label{eq:o-c-e}
@@
 For NN, an efficient way of computing this minimization is implementing the back-propagation (BP) algorithm.
-	BP is classified as a supervised learning algorithm that is able to optimize a function based on some parameter and is highly parallelizable in computational terms \cite{b9}\cite{b8}.
+	BP is classified as a supervised learning algorithm that is able to optimize a function based on some parameter and is highly parallelizable in computational terms \cite{Ibnkahla}\cite{nielsenneural}.
 
@@
 
-For implementing all the algorithms, Keras, a high-level python API for ML, will be used together with TensorFlow 1.13.1, a ML python friendly open-source library \cite{b11} \cite{b12}.
+For implementing all the algorithms, Keras, a high-level python API for ML, will be used together with TensorFlow 1.13.1, a ML python friendly open-source library \cite{DBLP:journals/corr/AbadiABBCCCDDDG16} \cite{chollet2015keras}.
 Both are available on a free license terms.
@@
 Autoencoders based on DNN can learn an appropriate correspondence between input source messages and estimated source message output through different noisy channels.
-The statistics of the channel's noise is shown in \cite{2018} to be irrelevant, since the NN extract it during training phase.
+The statistics of the channel's noise is shown in \cite{2017arXiv171008379G} to be irrelevant, since the NN extract it during training phase.
 The autoencoder applicability is compelling because the weights computed during training can be separated in two groups, one correspondent to the nodes before the channel and one posterior to the channel.
@@
 This type of hardware made possible the approach of the layer-by-layer supervised training by stochastic gradient descent optimization applied in this work.
-\cite{GPU}
+\cite{doi:10.1162/neco.2006.18.7.1527}
 
@@
 \State $\hat{\textbf{x}} \gets argmin(\text{distances})$
-\State $\hat{\textbf{u}} \gets G^{-1}\hat{\textbf{x}}$
+\State $\hat{\textbf{u}} \gets \hat{\textbf{x}} G^{-1}$
 \Return $\hat{\textbf{u}}$
@@
 \subsubsection{Array Decoder}
-Architecture, Training parameters, cite choice of p train \cite{b13}, number of epochs, Talk about noise layer and its removal in prediction,
+Architecture, Training parameters, cite choice of p train \cite{DBLP:conf/acssc/BenammarP18}, number of epochs, Talk about noise layer and its removal in prediction,
 
@@
 
-\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18},
+\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}
 
-\bibliographystyle{plain}
-\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18}
+\bibliographystyle{ieeetr}
+\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18}
 

### 2019-06-14 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/14/2019

\documentclass{beamer}

% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}

\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{multirow}
\usepackage{url}
%\usepackage[sorting=none]{biblatex}
\usepackage[utf8]{inputenc}
\usepackage{ifthen}
\usepackage{filecontents}
\usetikzlibrary{shapes,arrows,shadings,patterns}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\pgfplotsset{plot coordinates/math parser=false}

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\usetikzlibrary{positioning}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\makeatletter
\def\endthebibliography{%
  \def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}%
  \endlist
}
\makeatother

\makeatletter
\def\thickhline{%
  \noalign{\ifnum0=`}\fi\hrule \@height \thickarrayrulewidth \futurelet
   \reserved@a\@xthickhline}
\def\@xthickhline{\ifx\reserved@a\thickhline
               \vskip\doublerulesep
               \vskip-\thickarrayrulewidth
             \fi
      \ifnum0=`{\fi}}
\makeatother

\newlength{\thickarrayrulewidth}
\setlength{\thickarrayrulewidth}{3\arrayrulewidth}

\title[Machine Learning and Communication]{Machine Learning Autoencoder Applied to Communication Channels}

\author{E.~Dadalto Camara Gomes\inst{1} \and M.~Benammar\inst{2}}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[] % (optional, but mostly needed)
{
  \inst{1}%
   ISAE-SUPAERO\\
   Université de Toulouse\\
    31055, Toulouse, France\\
Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr
  \and
  \inst{2}%
  Department of Electronics, Optronics, and Signal processing\\
  ISAE-SUPAERO\\
  31055, Toulouse, France\\
  Email: meryem.benammar@isae-supaero.fr}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date{ISAE-SUPAERO, 2019}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

\subject{Machine Learning and Communication}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

\pgfdeclareimage[height=0.5cm]{university-logo}{images/ISAE}
\logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
%\AtBeginSubsection[]
%{
%  \begin{frame}<beamer>{Outline}
%    \tableofcontents[currentsection,currentsubsection]
%  \end{frame}
%}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Introduction}

\begin{frame}{Context}{Communication system context in general - what field will I be treating}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Context}{Machine Learning applications - what could we do in communication system}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Relevance \& Challenges}{Explain why the work is relevant and explain what are the challenges}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Problem Statement}{What exactly I will solve in this work}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

% You can reveal the parts of a slide one at a time
% with the \pause command:
\begin{frame}{Second Slide Title}
  \begin{itemize}
  \item {
    First item.
    \pause % The slide will pause after showing the first item
  }
  \item {   
    Second item.
  }
  % You can also specify when the content should appear
  % by using <n->:
  \item<3-> {
    Third item.
  }
  \item<4-> {
    Fourth item.
  }
  % or you can use the \uncover command to reveal general
  % content (not just \items):
  \item<5-> {
    Fifth item. \uncover<6->{Extra text in the fifth item.}
  }
  \end{itemize}
\end{frame}

\section{Methodology}

\subsection{Reference Model}
\begin{frame}{Maximum a Posterior (MAP) Rule}{Implementation of a MAP decoder for a linear block code through a BSC.}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\subsection{Design \& Architecture}
\begin{frame}{Neural Network's Architecture}{Show the architecture used for each case and remarks some important parameters}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\subsection{Training}
\begin{frame}{Neural Network's Training}{Show the best training parameters for each structure}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\subsection{Predictions}
\begin{frame}{Monte Carlo Simulations}{Explain how we could use NN to predict the results with certain confidence.}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Blocks}
\begin{block}{Block Title}
You can also highlight sections of your presentation in a block, with it's own title
\end{block}
\begin{theorem}
There are separate environments for theorems, examples, definitions and proofs.
\end{theorem}
\begin{example}
Here is an example of an example block.
\end{example}
\end{frame}

% Placing a * after \section means it will not show in the
% outline or table of contents.

\section{Results \& Discussions}
\subsection{DNN Decoders}
\begin{frame}{DNN Array Decoder}{Show the results for the array decoder in terms of train p, Mep, Parameters, etc}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\begin{frame}{DNN One-hot Decoder}{Show the results for the one-hot decoder in terms of train p, Mep, Parameters, etc}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\subsection{DNN Autoencoders}
\begin{frame}{DNN Autoencoder}{Show the results for the autoencoder in terms of train p, Mep, Parameters, etc}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\section{Conclusions}
\begin{frame}{Conclusions}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\section{Future Work}
\begin{frame}{Future Work}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\section*{Acknowledgment}

\begin{frame}{Acknowledgment}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

% All of the following is optional and typically not needed. 
\appendix
\section<presentation>*{\appendixname}
\subsection<presentation>*{Bibliography}

\begin{frame}[allowframebreaks]
  \frametitle<presentation>{Bibliography}

  \cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}

  \bibliographystyle{ieeetr}
  \bibliography{../../Article/bib/ShannonCE1948,../../Article/bib/CalabreseWGPS16,../../Article/bib/OSheaH17,../../Article/bib/Oshea2,../../Article/bib/Gruber,../../Article/bib/Worm,../../Article/bib/Viterbi,../../Article/bib/RobertsonHV97,../../Article/bib/Ibnkahla,../../Article/bib/Nielsen,../../Article/bib/Murphy,../../Article/bib/AbadiABBCCCDDDG16,../../Article/bib/Keras,../../Article/bib/mit_neco18_1527,../../Article/bib/BenammarP18}

\end{frame}

\end{document}

### 2019-06-14 MyCode/autoencoder-retrain.py

@@ -6,8 +6,8 @@ Created on Wed Jun  5 13:47:36 2019
 """
 #%%
 numEpochs = 2**16
-directory = 'Retrain\AutoencoderArray\\'
-fileName = 'i_1_AutoencoderArray_Mep_65536_p_0.03.h5'
+directory = 'Retrain\Autoencoder1H\\'
+fileName = 'Autoencoder1H_Mep_65536_bs_256_32_32-32_32.h5'
 Autoencoder = tf.keras.models.load_model(directory+'OriginalModel\\'+fileName)
 print("Loaded models from disk")
 path = directory+'TrainedModel\\'+fileName
@@ -20,7 +20,7 @@ Autoencoder.compile(loss=lossFunc ,
               optimizer='adam',
               )
 '''
-    Summaries and checkpoints (to do)
+    Summaries and checkpoints
 '''
 summary = Autoencoder.summary()
 checkpoint = tf.keras.callbacks.ModelCheckpoint(
@@ -43,12 +43,12 @@ Autoencoder.save(path)  # creates a HDF5 file
 Encoder = Autoencoder.layers[0]
 Decoder = Autoencoder.layers[2]
 #% Prediction
-globalErrorAutoencoder=fn.arrayAutoencoderPrediction(Encoder, Decoder, pOptions, globalReps, N, n, k)
-
+#globalErrorAutoencoder=fn.arrayAutoencoderPrediction(Encoder, Decoder, pOptions, globalReps, N, n, k)
+globalErrorAutoencoder=fn.onehotAutoencoderPrediction(Encoder, Decoder, messages, pOptions, globalReps, N, n, k)
 #% Plotting
-figPath = directory+'AutoencoderArray_64-128-256_Mep_131072_p_0.03.png'
 numEpochs = 2**17
-plotBERp(globalErrorAutoencoder, 'Autoencoder')
+figPath = directory+'Autoencoder1H_Mep_'+str(numEpochs)+'_bs_256_32_32-32_32_p_train_0.png'
+plotBERp(globalErrorAutoencoder, 'One-hot Autoencoder')
 
 
 

### 2019-06-14 MyCode/autoencoder1H.py

@@ -19,8 +19,8 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([256, 1024])
-DecoderNodes = [128, 64]
+encoderNodes = np.array([32, 32])
+DecoderNodes = [32, 32]
 #%
 '''
     Architecture
@@ -86,12 +86,15 @@ Autoencoder1H.save(path)  # creates a HDF5 file
 #%%%
 '''
     Prediction 1H
-
+'''
 t = TicToc('name')
 t.tic()
 
 
 globalReps = 100
+globalErrorAutoencoder1H = fn.onehotAutoencoderPrediction(Encoder, Decoder, 
+                               messages, pOptions, globalReps, N, n, k)
+'''
 globalErrorAutoencoder1H = np.empty([globalReps, len(pOptions)])
 for i_global in range(globalReps):
     for i_p in range(np.size(pOptions)):
@@ -106,10 +109,10 @@ for i_global in range(globalReps):
         predictedMessages = fn.multipleOneshot2messages(prediction, messages)
 
         globalErrorAutoencoder1H[i_global][i_p] = fn.bitErrorFunction(predictedMessages, u)
-
+'''
 #% Plotting
 plotBERp(globalErrorAutoencoder1H, 'One-hot Autoencoder')
 
 t.toc()
 print(t.elapsed)
-'''
+

### 2019-06-16 Rapport/Article/Article.tex
Seance 06/16/2019

@@
 While the Max-Log-MAP is a suboptimal decoder and will not be treated in this paper.
-these variants are widely used as decoders for turbo codes \cite{b7}.
+these variants are widely used as decoders for turbo codes \cite{HagenauerJ}.
 Refer to \cite{journals/ett/RobertsonHV97} for specific details in the derivation of the Log-MAP algorithm.
@@
 
-Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a BSC was studied in \cite{b5}.
+Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a BSC was studied in \cite{JordanMA}.
 They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{journals/ett/RobertsonHV97}.
@@
 %
-Two common activation functions are the rectified linear unit (ReLU) and the sigmoid function. Both will be applied in the NN treated in this paper. Their mathematical expressions are shown below.
+Two common activation functions are the rectified linear unit (ReLU) and the sigmoid function.
+Both will be applied to the DNNs treated in this paper.
+In addition, the Softmax activation function is commonly used in multi classification problems and will be used to the One-hot decoding DNN architectures.
+Their mathematical expressions are shown below.
 
@@
 	\text{ReLU: } \sigma_1 (x)& = max\{0, x\} \label{eq:relu}\\ 
-	\text{Sigmoid: } \sigma_2 (x)& = \frac{1}{1+e^{- \lambda x}} \label{eq:sig}
+	\text{Sigmoid: } \sigma_2 (x)& = \frac{1}{1+e^{- \lambda x}} \label{eq:sig}\\
+	\text{Softmax: } \sigma_3 (\textbf{z})& = \frac{e^{z_i}}{\sum_{j=1}^{K}e^{z_j}}
 \end{align}
-where $x \in \mathbb{R}$ and $\lambda>0$.
+where $x \in \mathbb{R}$, $\lambda>0$, $i = 1,...,K$ and $\textbf{z} = (z_1,...,z_K) \in \mathbb{R}^K$.
 
@@
 With the weight matrix trained, the NN finds the correct output values for unseen inputs.
-In order to validate this principle, a \textit{validation data set} is used and the loss is mesured.
+In order to validate this principle, a \textit{validation data set} is used and the loss is measured.
 Based on this value, we can evaluate the NN overall accuracy.
@@
   \centering
-    \includegraphics[width=0.5\textwidth]{images/20190607-160812_MAP}
+    \includegraphics[width=0.5\textwidth]{images/MAP-alone-curve}
     \caption{MAP algorithm performance in terms of BER versus channel crossover probability.}\label{fig:MAP}
@@
     \thickhline
-    \multirow{3}{*}{Decoder} & Dense: X, activation: ReLU, input size: $n$ \\
- & Dense: X, activation: ReLU  \\
+    \multirow{3}{*}{Decoder} & Dense: 128, activation: ReLU, input size: $n$ \\
+ & Dense: 64, activation: ReLU  \\
+ & Dense: 32, activation: ReLU  \\
  & Dense: $k$, activation: Sigmoid  \\ \hline
-    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \multicolumn{2}{l}{\textbf{Total parameters: $12776$}}\\
     \thickhline
@@
     \thickhline
-    \multirow{3}{*}{Decoder} & Dense: X, activation: ReLU, input size: $n$ \\
- & Dense: X, activation: ReLU  \\
- & Dense: $2^k$, activation: Sigmoid  \\ \hline
-    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \multirow{1}{*}{Decoder} & Dense: 256, activation: Softmax, input size: $n$ \\ \hline
+    \multicolumn{2}{l}{\textbf{Total parameters: }$4352$}\\
     \thickhline
@@
 
+\subsection{Monte Carlo Simulation}
+Explain the prediction simulation (global reps, values of p, etc)
 \section{Results and Analysis}
@@
   \centering
-    \includegraphics[width=0.5\textwidth]{images/MAP_MLNN1H_p_0}
-    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
+    \includegraphics[width=0.5\textwidth]{images/MLNN_Mep_65536_ptrain_007}
+    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
 \end{figure}
@@
   \centering
-    \includegraphics[width=0.5\textwidth]{images/20190529-142119_MAP_MLNN_Mep_65536}
-    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
+    \includegraphics[width=0.5\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
+    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
 \end{figure}
@@
 
+\begin{table}
+\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the mean MAP algorithm decoding time. } 
+\begin{center}
+
+  \begin{tabular}{ l | l | l | l | l }
+    \thickhline
+    MAP & A. Dec. & One-hot Dec. & A. Auto. & One-hot Auto. \\ \hline
+    $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ & $ \pm $ & $ \pm $ \\ 
+    \thickhline
+  \end{tabular}
+    \label{tab:timeanalysis}
+\end{center}
+\end{table}
+
 \section{Conclusions}
@@
 
-\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}
+\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}
 
 \bibliographystyle{ieeetr}
-\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18}
+\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18}
 

### 2019-06-16 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/16/2019

@@
   \item {
-    My first point.
+
   }
   \item {
-    My second point.
+
+  }
+  \item {
+
   }
@@
 \begin{frame}{DNN Array Decoder}{Show the results for the array decoder in terms of train p, Mep, Parameters, etc}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+
+    \begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{../../Article/images/MLNN_Mep_65536_ptrain_007}
+    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
+\end{figure}
+
 \end{frame}
@@
 \begin{frame}{DNN One-hot Decoder}{Show the results for the one-hot decoder in terms of train p, Mep, Parameters, etc}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+
+    \begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{../../Article/images/MLNN1H_Mep_16384_ptrain_0}
+    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
+\end{figure}
+
 \end{frame}
@@
 % All of the following is optional and typically not needed. 
-\appendix
-\section<presentation>*{\appendixname}
-\subsection<presentation>*{Bibliography}
+%\appendix
+%\section<presentation>*{Bibliography}
 
-\begin{frame}[allowframebreaks]
-  \frametitle<presentation>{Bibliography}
+%\begin{frame}[allowframebreaks]
+%  \frametitle<presentation>{Bibliography}
 
-  \cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}
+ % \cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}
 
-  \bibliographystyle{ieeetr}
-  \bibliography{../../Article/bib/ShannonCE1948,../../Article/bib/CalabreseWGPS16,../../Article/bib/OSheaH17,../../Article/bib/Oshea2,../../Article/bib/Gruber,../../Article/bib/Worm,../../Article/bib/Viterbi,../../Article/bib/RobertsonHV97,../../Article/bib/Ibnkahla,../../Article/bib/Nielsen,../../Article/bib/Murphy,../../Article/bib/AbadiABBCCCDDDG16,../../Article/bib/Keras,../../Article/bib/mit_neco18_1527,../../Article/bib/BenammarP18}
+ % \bibliographystyle{ieeetr}
+ % \bibliography{../../Article/bib/ShannonCE1948,../../Article/bib/CalabreseWGPS16,../../Article/bib/OSheaH17,../../Article/bib/Oshea2,../../Article/bib/Gruber,../../Article/bib/Worm,../../Article/bib/Viterbi,../../Article/bib/RobertsonHV97,../../Article/bib/Ibnkahla,../../Article/bib/Nielsen,../../Article/bib/Murphy,../../Article/bib/AbadiABBCCCDDDG16,../../Article/bib/Keras,../../Article/bib/mit_neco18_1527,../../Article/bib/BenammarP18}
 
-\end{frame}
+%\end{frame}
 

### 2019-06-16 MyCode/BSC-MAP-Decoder.py

@@ -5,6 +5,7 @@ Created on Sun May 12 12:02:48 2019
 @author: user
 """
 #%% Initialization
+globalReps = 1000
 MAPpredictTime = TicToc('MAPPredict')
 globalError = np.empty([globalReps, len(pOptions)])
 globalErrorHamming = np.empty([globalReps, len(pOptions)])
@@ -57,8 +58,8 @@ avgGlobalErrorMAP = np.average(globalErrorMAP, 0)
 
 fig = plt.figure(figsize=(8, 6), dpi=80)
 
-plt.plot(pOptions,avgGlobalError, color='b', linewidth=lineWidth, linestyle='--')
-plt.plot(pOptions,avgGlobalErrorMAP, color='r', linewidth=lineWidth)
+plt.plot(pOptions,avgGlobalError, color='k', linewidth=lineWidth, linestyle='--')
+plt.plot(pOptions,avgGlobalErrorMAP, color='k', linewidth=lineWidth)
 
 plt.grid(True, which="both")
 plt.xlabel('$p$')

### 2019-06-16 MyCode/MLNN-train.py

@@ -5,9 +5,6 @@ Created on Mon Jun  3 15:15:35 2019
 @author: user
 """
 #%%
-n=16
-k=8
-
 u_train_labels = messages.copy()
 x_train_data = possibleCodewords.copy()
 
@@ -72,11 +69,6 @@ plt.show()
 
 timestr = time.strftime("%Y%m%d-%H%M%S")
 trainingFig.savefig('training_history/'+title+'/'+timestr + '_'+title+'_train.png', bbox_inches='tight', dpi=300)
-'''
-    evaluate the inference-model
-''' 
-
-#evaluation = MLNN.evaluate(x_val, u_val_labels)
 
 '''
     Saving model

### 2019-06-16 MyCode/MLNN1H-train.py

@@ -4,6 +4,7 @@ Created on Mon Jun  3 15:15:26 2019
 
 @author: user
 """
+#%%
 '''
     DNN One Hot Model Decoder
 '''
@@ -27,14 +28,9 @@ u_val_labels = fn.messages2onehot(u_val_labels)
 '''
 
 MLNN1H = tf.keras.Sequential([ # Array to define layers
-              # Adds a densely-connected layer with n units to the model: L1
               #layers.Dense(32, activation='relu', input_shape=(n,), name='HL1'),
-              # Add another: L2
-              layers.Dense(64, activation='relu', input_shape=(n,), name='HL1'),
-              # Add another: L3
-              #layers.Dense(128, activation='relu',input_shape=(n,), name='HL1'),
-              # Add layer with k output units:
-              layers.Dense(256, activation='softmax', name='Output')
+              # Add layer with 2**k output units:
+              layers.Dense(256, activation='softmax', input_shape=(n,),name='Output')
 ])
 
 plot_model(MLNN1H,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',show_shapes=True)
@@ -62,28 +58,7 @@ history = MLNN1H.fit(x_train_data, u_train_labels, epochs=numEpochs,
 #          validation_data=(x_val, u_val_labels))
 
 # summarize history for loss
-trainingFig = plt.figure(figsize=(8, 6), dpi=80)
-plt.title('Batch size = '+str(batchSize))
-plt.plot(history.history['loss']) # all outputs: ['acc', 'loss', 'val_acc', 'val_loss']
-#plt.plot(history.history['metricBER1H'])
-plt.grid(True, which='both')
-#plt.plot(history.history['val_loss'])
-plt.xlabel('$M_{ep}$')
-plt.xscale('log')
-plt.legend([lossFunc + ' loss'])
-plt.show()
-trainingFig.set_size_inches(width, height)
-trainingFig.savefig('training_history/'+title+'/'+timestr + '_'+title+'_train.png', bbox_inches='tight', dpi=300)
-'''
-    evaluate the inference-model
-''' 
-
-evaluation = MLNN1H.evaluate(x_val, u_val_labels)
-
-u = fn.generateU(1,k)
-y = fn.generteCodeWord(1, n, u, G)
-prediction = MLNN1H.predict(y)
-predictedMessage = fn.onehot2singleMessage(prediction, messages)
+plotTraining(history)
 
 '''
     Saving model

### 2019-06-16 MyCode/autoencoder1H.py

@@ -19,8 +19,8 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([32, 32])
-DecoderNodes = [32, 32]
+encoderNodes = np.array([256, 256, 64])
+DecoderNodes = [256, 256]
 #%
 '''
     Architecture
@@ -34,7 +34,7 @@ Encoder = tf.keras.Sequential([
         #layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
         # Coded Layer
         layers.Dense(n, activation='sigmoid', name='Codedfloat'),
-        ], name='Encoder')
+        ], name='Encoder')  
 
 NoiseL = tf.keras.Sequential([
         # Rounded codeword
@@ -45,13 +45,13 @@ NoiseL = tf.keras.Sequential([
                       output_shape=(n,), name='Noise'),
         ], name='Noise')
 
-Decoder1H = tf.keras.Sequential([ # Array to define layers
+Decoder = tf.keras.Sequential([ # Array to define layers
         layers.Dense(DecoderNodes[0], activation='relu', input_shape=(n,), name='DHL1'),
         layers.Dense(DecoderNodes[1], activation='relu', name='DHL2'),
         layers.Dense(256, activation='softmax', name='1H_Output')
         ], name = 'Decoder')
 
-Autoencoder1H = tf.keras.Sequential([Encoder,NoiseL, Decoder1H])
+Autoencoder1H = tf.keras.Sequential([Encoder,NoiseL, Decoder])
 plot_model(Autoencoder1H,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',show_shapes=True)
 
 '''
@@ -94,22 +94,7 @@ t.tic()
 globalReps = 100
 globalErrorAutoencoder1H = fn.onehotAutoencoderPrediction(Encoder, Decoder, 
                                messages, pOptions, globalReps, N, n, k)
-'''
-globalErrorAutoencoder1H = np.empty([globalReps, len(pOptions)])
-for i_global in range(globalReps):
-    for i_p in range(np.size(pOptions)):
-        p = pOptions[i_p]
-        u = fn.generateU(N,k)
-        u1h = fn.messages2onehot(u)
-        x = np.round(Encoder.predict(u1h))
[…]

### 2019-06-16 MyCode/autoencoderArray.py

@@ -26,9 +26,8 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([256, 128, 64, 16])
-DecoderNodes = [128, 64, 32, k]
-n = 8
+encoderNodes = np.array([256, 256, 256])
+DecoderNodes = [256, 256, 32]
 #%
 '''
     Architecture
@@ -61,9 +60,9 @@ Decoder = tf.keras.Sequential([ # Array to define layers
         # Add another: L2
         layers.Dense(DecoderNodes[1], activation='relu', name='DHL2'),
         # Add another: L3
-        layers.Dense(DecoderNodes[2], activation='relu', name='DHL3'),
+        #layers.Dense(DecoderNodes[2], activation='relu', name='DHL3'),
         # Add layer with k output units: output
-        layers.Dense(DecoderNodes[3], activation='sigmoid', name='Output')
+        layers.Dense(k, activation='sigmoid', name='Output')
         ], name = 'Decoder')
 
 Autoencoder = tf.keras.Sequential([Encoder,NoiseL, Decoder])

### 2019-06-16 MyCode/libs-and-global-consts.py

@@ -27,7 +27,7 @@ plt.rcParams["font.family"] = "Times New Roman"
 plt.rcParams.update({'font.weight':'normal'})
 letterSize = 6
 markerSize = 3
-lineWidth = 0.5
+lineWidth = 0.75
 plt.rc('xtick', labelsize=letterSize)
 plt.rc('ytick', labelsize=letterSize)
 plt.rc('axes', labelsize=letterSize)
@@ -84,11 +84,11 @@ globalReps = 1000
 
 #%%
 # NN parameters
-title = 'Autoencoder1H'
+title = 'AutoencoderArray'
 timestr = time.strftime("%Y%m%d-%H%M%S")
 numEpochs = 2**16
 batchSize = 256
-train_p = 0.0
+train_p = 0.03
 
 path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
 checkpointPath = 'Checkpoints/'+title+'/'+timestr+'_'+title+'_Mep_{epoch:02d}-{loss:.8f}.h5'
@@ -110,16 +110,16 @@ def tensorBSC(x):
 def plotBERp(globalErrorMLNN, legendEntry):
     avgGlobalErrorMLNN = np.average(globalErrorMLNN,0)
     fig = plt.figure(figsize=(8, 6), dpi=80)
-    plt.scatter(pOptions,avgGlobalErrorMLNN, color='g', marker='^', zorder=3, s=markerSize)
-    plt.plot(pOptions,avgGlobalError, color='b', linewidth=lineWidth, linestyle='--')
-    plt.plot(pOptions,avgGlobalErrorMAP, color='r', linewidth=lineWidth)
+    plt.plot(pOptions,avgGlobalErrorMLNN, '--g', marker='^', zorder=3, markersize=markerSize, linewidth=lineWidth)
+    plt.plot(pOptions,avgGlobalError, color='k', linewidth=lineWidth, linestyle='--')
+    plt.plot(pOptions,avgGlobalErrorMAP, color='k', linewidth=lineWidth)
     plt.grid(True, which='both')
     #plt.title('Training $p = $ '+ str(train_p))
     plt.xlabel('$p$')
     plt.ylabel('BER')
     plt.yscale('log')
     #plt.legend(['No Decoding', 'MAP Algorithm', legendEntry+ ', $p_t=$'+str(train_p)])
[…]

### 2019-06-16 MyCode/my_functions.py

@@ -336,6 +336,38 @@ def createDir(path):
 def messages2customEncoding(messages, Encoder):
     return Encoder.predict(messages)
 
+def arrayDecoderPrediction(G, Decoder, pOptions, globalReps, N, n, k):
+    globalErrorDecoder = np.empty([globalReps, len(pOptions)])
+    for i_global in range(globalReps):
+        for i_p in range(np.size(pOptions)):
+            p = pOptions[i_p]
+            u = generateU(N,k)
+            x = generteCodeWord(N, n, u, G)
+            xflat = np.reshape(x, [-1])
+            yflat = BSC(xflat,p)
+            y = yflat.reshape(N,n) # noisy codewords
+            prediction = Decoder.predict(y)
+            predictedMessages = np.round(prediction)
+            globalErrorDecoder[i_global][i_p] = bitErrorFunction(predictedMessages, u)
+            
+    return globalErrorDecoder
+
+def onehotDecoderPrediction(G, Decoder, pOptions, globalReps, N, n, k, messages):
+    globalErrorDecoder = np.empty([globalReps, len(pOptions)])
+    for i_global in range(globalReps):
+        for i_p in range(np.size(pOptions)):
+            p = pOptions[i_p]
+            u = generateU(N,k)
+            x = generteCodeWord(N, n, u, G)
+            xflat = np.reshape(x, [-1])
+            yflat = BSC(xflat,p)
+            y = yflat.reshape(N,n) # noisy codewords
+            prediction = Decoder.predict(y)
+            predictedMessages = multipleOneshot2messages(prediction, messages)
+            globalErrorDecoder[i_global][i_p] = bitErrorFunction(predictedMessages, u)
+            
+    return globalErrorDecoder
+
 def arrayAutoencoderPrediction(Encoder, Decoder, pOptions, globalReps, N, n, k):
     globalErrorAutoencoder = np.empty([globalReps, len(pOptions)])
     for i_global in range(globalReps):
@@ -345,7 +377,7 @@ def arrayAutoencoderPrediction(Encoder, Decoder, pOptions, globalReps, N, n, k):
             x = Encoder.predict(u)
             xflat = np.reshape(x, [-1])
             yflat = BSC(xflat,p)
-            y = yflat.reshape(N,2*k) # noisy codewords
+            y = yflat.reshape(N,n) # noisy codewords
             prediction = Decoder.predict(y)
             predictedMessages = np.round(prediction)
[…]

### 2019-06-16 MyCode/time_analysis.py

@@ -1,0 +1,38 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sun Jun 16 11:09:37 2019
+
+@author: user
+"""
+#%%
+'''
+    MAP Decoder
+'''
+for i in range(10):
+    exec(open("BSC-MAP-Decoder.py").read())
+
+#%%
+'''
+    Array Decoder
+'''
+# NN characteristics:
+predictTime = TicToc('Predict')
+for i in range(10):
+    predictTime.tic()
+    globalErrorMLNN = fn.arrayDecoderPrediction(G, MLNNDecoder, pOptions, globalReps, N, n, k)
+    predictTime.toc()
+    print('Total predict time: ', predictTime.elapsed)
+    plotBERp(globalErrorMLNN, 'Array Decoder')
+    
+    #%%
+'''
+    One-hot Decoder
+'''
+# NN characteristics:
+predictTime = TicToc('Predict')
+for i in range(10):
+    predictTime.tic()
+    globalErrorMLNN1H = fn.onehotDecoderPrediction(G, MLNN1H, pOptions, globalReps, N, n, k, messages)
+    predictTime.toc()
+    print('Total predict time: ', predictTime.elapsed)
+    plotBERp(globalErrorMLNN1H, 'One-hot Decoder')

### 2019-06-17 Rapport/Article/Article.tex
Seance 06/17/2019

@@
 
+\subsection{Binary Symmetric Channel}
+
+Briefly, a BSC is a noisy channel with a binary input $X\in{0,1}$ and a binary output $Y\in{0,1}$.
+They are linked through a stochastic model characterized by a crossover probability of value $p$ as illustrated in Fig.
+\ref{fig:BSC}.
+Typically $p$ is smaller than $0.10$.
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/BSC}
+    \caption{Illustration of a binary symmetric channel with crossover probability $p$.}\label{fig:BSC}
+\end{figure}
+
+A BSC is translated mathematically as follows.
+
+\begin{align*}
+Pr[ Y = 0 | X = 0 ] &= 1 - p \\
+Pr[ Y = 0 | X = 1 ] &= p \\
+Pr[ Y = 1 | X = 0 ] &= p \\
+Pr[ Y = 1 | X = 1 ] &= 1 - p
+\end{align*}
+
 \subsection{Maximum a posteriori decoder}
@@
 	\text{Sigmoid: } \sigma_2 (x)& = \frac{1}{1+e^{- \lambda x}} \label{eq:sig}\\
-	\text{Softmax: } \sigma_3 (\textbf{z})& = \frac{e^{z_i}}{\sum_{j=1}^{K}e^{z_j}}
+	\text{Softmax: } \sigma_3 (\textbf{z})& = \frac{e^{z_i}}{\sum_{j=1}^{K}e^{z_j}}\label{eq:soft}
 \end{align}
@@
 This section explains the main implementation procedures in order to replicate the results and understand in depth the methodology behind the work.
-For organizational purposes, the section is divided into three subsections.
-First, the MAP algorithm for a linear code is shown, then the experiments relating to the ML based decoders are assessed and finally the implementation of the ML based autoencoder is explained.
+For organizational purposes, the section is divided into four subsections.
+First, the predictions and error corrections statistical relevance are justified through the Monte Carlo simulations conducted.
+Second, the MAP algorithm for a linear code is shown.
+Then the experiments relating to the ML based decoders are assessed.
+Finally the implementation of the ML based autoencoder is explained.
+
+\subsection{Monte Carlo Simulation}
+
+For calculating the BER versus the crossover probability $p$ with confidence over the results, a Monte Carlo simulation was elaborated.
+This stochastic computational algorithm uses pseudo-random sampling to solve different problems.
+Hence, we can work with a wide variety of transmission scenarios.
+
+Applied to the context of this research, the messages are repeatedly randomly generated for a specific BSC of crossover probability $p$.
+They are coded by a linear block code or by a DNN encoder and then transmitted through the channel.
+Their information arrives corrupted by noise in the decoder and is subsequently partially corrected by a decoder.
+Finally the average BER is calculated and plotted.
+The interval of the horizontal axis is in between $0$ and $0.10$.
+
+Every channel is tested with $100000$ messages. As a consequence of the \textit{law of large numbers}, the empirical sampling of the BER represents sufficiently well the real distribution.
 
@@
 
-The pseudo code of the MAP decoder algorithm is available below.
+The MAP algorithm as cited previously, has a high time complexity $(\mathcal{O}(2^k))$, since it computes the Hamming distance between the received code word and every code word from the code book.
+For more details, the pseudo code of the algorithm is available below.
 
@@
 
-\begin{figure}[!ht]
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/MAP-alone-curve}
-    \caption{MAP algorithm performance in terms of BER versus channel crossover probability.}\label{fig:MAP}
-\end{figure}
-
-\subsection{Decoder}
-The linear encoding matrix $\textbf{G}_{8\times16}$ is shown below.
+The linear encoding matrix $\textbf{G}_{8\times16}$ was used for the encoding of the messages.
+Its dimension is equal to the linear block code name represented by the double $(n,k)=(16,8)$.
+Thus, the \textit{rate} of the code is equal to $0.5$.
+The linear block name and = rate is constant throughout the work.
 
@@
 
-Introduce encoding equations
+To establish a reference model to use as comparison base to the DNN models and MAP decoder, the BER results for a scenario without codification was calculated.
+The ensemble of the no decoding curve and the MAP algorithm curve of Fig.
+\ref{fig:MAP} are replicated for comparison purposes with the methods of interest.
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/MAP-alone-curve}
+    \caption{No decoding and MAP algorithm performance in terms of BER versus channel crossover probability.}\label{fig:MAP}
+\end{figure}
+
+\subsection{Decoder}
+
+This section is divided in two for better identify the differences between the implementation of the array DNN decoder and the one-hot DNN decoder.
 
 \subsubsection{Array Decoder}
-Architecture, Training parameters, cite choice of p train \cite{DBLP:conf/acssc/BenammarP18}, number of epochs, Talk about noise layer and its removal in prediction,
+
+The architecture of the DNN necessary to achieve a great performance in terms of BER is available in table \ref{tab:arraydecoder}.
+What should be remarked is the first layer of the DNN.
+A \textit{Lambda} layer is a layer in which the activation operations are user defined and the parameters are non-trainable.
+Thus, we were able to define a layer that act as a BSC to force our DNN to learn its effects in the training data.
+The fact that the parameters are non-trainable means that the noise layer is applied in forward propagation only.
+
+The choice of the crossover probability of the training channel, $p_{t}$, was based on \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the best in terms of decoding precision after the DNN is trained.
+In addition, a deeper NN showed best results in \cite{ DBLP:journals/corr/OSheaH17} and were adopted to solve the problem.
+The training parameters are displayed in table \ref{tab:arraydecoderTrain}.
+
+The choices of training parameters is a mix between what other works in literature uses and a hyper parametric analytic approach, where several trials lead to the decision.
+First, the binary cross-entropy as loss function is widely used in literature and  showed to converge to the desired result consistently during tests.
+For the optimizer, the \textit{Adam} built-in Keras optimizer was chosen.
+Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}.
+Next, the BATCH SIZE.
+Finally, the necessary number of epochs to training is sensitive to the problem, since it impacts directly the training computational cost.
+As a result, we tried several training epoch sizes and found $2^{16}$ to have precise results and be computationally viable.
 
 \begin{table}
-\caption{DNN array decoder architecture and parameters.} 
+\caption{DNN array decoder architecture.} 
 \begin{center}
@@
     \thickhline
-    \multirow{3}{*}{Decoder} & Dense: 128, activation: ReLU, input size: $n$ \\
+    Channel &  Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
+    \multirow{3}{*}{Decoder} & Dense: 128, activation: ReLU, input size: 16 \\
  & Dense: 64, activation: ReLU  \\
  & Dense: 32, activation: ReLU  \\
- & Dense: $k$, activation: Sigmoid  \\ \hline
+ & Dense: 8, activation: Sigmoid  \\ \hline
     \multicolumn{2}{l}{\textbf{Total parameters: $12776$}}\\
@@
 
+\begin{table}
+\caption{DNN array decoder training parameters.} 
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Binary cross-entropy & Adam & $2^{16}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:arraydecoderTrain}
+\end{center}
+\end{table}
+
 \subsubsection{One-hot Decoder}
-Architecture, Training parameters, Explain what is a One hot decoding, talk about the matrix transformation from one hot to messages, p train equals 0, number of epochs
+
+The architecture of the one-hot decoder is much simpler when compared to the array decoder.
+It contains only one layer of $256$ units which maps the input of size 16 to the $2^8=256$ possible messages.
+Hence, this characterizes a multi-class classification problem, meaning that the chosen activation function is the Softmax (Eq.
+\ref{eq:soft}).
+This training technique is known in the literature as \textit{one-hot}.
+
+The denomination one-hot comes from the output vector that is usually a long vector of zeros with one element equals to one.
+This \textit{hot} vector element indicates which class the NN has classified the input.
+In this context, which message the code word refers to.
+
+Tranformation from one-hot to messagges
+
+Abscence of noise layer
+
+Lower complexity of the NN
+
+Loss function choice
+
+Optimizer choice
+
+Num epochs choice
+
+Batchsize choice
 
 \begin{table}
-\caption{DNN one-hot decoder architecture and parameters.} 
+\caption{DNN one-hot decoder architecture.} 
 \begin{center}
@@
     \thickhline
-    \multirow{1}{*}{Decoder} & Dense: 256, activation: Softmax, input size: $n$ \\ \hline
+    \multirow{1}{*}{Decoder} & Dense: 256, activation: Softmax, input size: 16 \\ \hline
     \multicolumn{2}{l}{\textbf{Total parameters: }$4352$}\\
@@
 
+\begin{table}
+\caption{DNN one-hot decoder training parameters.} 
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Binary cross-entropy & Adam & $2^{14}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:onehotdecoderTrain}
+\end{center}
+\end{table}
+
 \subsection{Autoencoder}
@@
 \begin{table}
-\caption{DNN array autoencoder architecture and parameters.} 
+\caption{DNN array autoencoder architecture.} 
 \begin{center}
@@
     \thickhline
-    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: $k$ \\
+    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: 8 \\
  & Dense: X, activation: ReLU  \\
- & Dense: $n$, activation: Sigmoid  \\ \hline
+ & Dense: 16, activation: Sigmoid  \\ \hline
 
@@
 
- \multirow{2}{*}{Decoder} & Dense: X, input size: $n$ \\
+ \multirow{2}{*}{Decoder} & Dense: X, input size: 16 \\
  & Dense: X, activation: ReLU  \\
- & Dense: $k$, activation: Sigmoid  \\ \hline
+ & Dense: 8, activation: Sigmoid  \\ \hline
 
@@
 
-\subsection{Monte Carlo Simulation}
-Explain the prediction simulation (global reps, values of p, etc)
+\begin{table}
+\caption{DNN array autodecoder training parameters.} 
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Binary cross-entropy & Adam & $2^{14}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:arrayautoencoderTrain}
+\end{center}
+\end{table}
+
 \section{Results and Analysis}

### 2019-06-17 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/17/2019

--- 2019-06-14 Rapport/Presentation/Presentation/Presentation.tex
@@
 \begin{frame}{Maximum a Posterior (MAP) Rule}{Implementation of a MAP decoder for a linear block code through a BSC.}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+
+ \begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.8\textwidth]{images/algorithm}
+\end{figure}
+
 \end{frame}
@@
 \subsection{Design \& Architecture}
-\begin{frame}{Neural Network's Architecture}{Show the architecture used for each case and remarks some important parameters}
+\begin{frame}[allowframebreaks]{Neural Network's Architecture}{Show the architecture used for each case and remarks some important parameters}
   \begin{itemize}
   \item {
-    My first point.
+    \begin{table}
+\caption{DNN array decoder architecture and parameters.} 
+\begin{center}
+
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{3}{*}{Decoder} & Dense: 128, activation: ReLU, input size: $n$ \\
+ & Dense: 64, activation: ReLU  \\
+ & Dense: 32, activation: ReLU  \\
+ & Dense: $k$, activation: Sigmoid  \\ \hline
+    \multicolumn{2}{l}{\textbf{Total parameters: $12776$}}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:arraydecoder}
+\end{center}
+\end{table}
   }
   \item {
-    My second point.
+    \begin{table}
+\caption{DNN one-hot decoder architecture and parameters.} 
+\begin{center}
+
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{1}{*}{Decoder} & Dense: 256, activation: Softmax, input size: $n$ \\ \hline
+    \multicolumn{2}{l}{\textbf{Total parameters: }$4352$}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:onehotdecoder}
+\end{center}
+\end{table}
+
+  }
+  \item {
+    \begin{table}
+\caption{DNN array autoencoder architecture and parameters.} 
+\begin{center}
+
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: $k$ \\
+ & Dense: X, activation: ReLU  \\
+ & Dense: $n$, activation: Sigmoid  \\ \hline
+
+    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $n$ \\
+ & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
+
+ \multirow{2}{*}{Decoder} & Dense: X, input size: $n$ \\
+ & Dense: X, activation: ReLU  \\
+ & Dense: $k$, activation: Sigmoid  \\ \hline
+
+    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:arrayautoencoder}
+\end{center}
+\end{table}
+
   }
@@
 \begin{frame}{Neural Network's Training}{Show the best training parameters for each structure}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+
+    \begin{table}
+\caption{DNN array decoder training parameters.} 
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Binary cross-entropy & Adam & $2^{16}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:arraydecoderTrain}
+\end{center}
+\end{table}
+
+    \begin{table}
+\caption{DNN one-hot decoder training parameters.} 
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Binary cross-entropy & Adam & $2^{14}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:onehotdecoderTrain}
+\end{center}
+\end{table}
+
+\begin{table}
+\caption{DNN array autodecoder training parameters.} 
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Binary cross-entropy & Adam & $2^{14}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:arrayautoencoderTrain}
+\end{center}
+\end{table}
+
 \end{frame}
 
-\subsection{Predictions}
-\begin{frame}{Monte Carlo Simulations}{Explain how we could use NN to predict the results with certain confidence.}
+\subsection{Error Correction \& Predictions}
+\begin{frame}{Error Correction and Monte Carlo Simulations}{Explain how we could use NN to predict the results with certain confidence.}
   \begin{itemize}
@@
 \begin{frame}{DNN Array Decoder}{Show the results for the array decoder in terms of train p, Mep, Parameters, etc}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+
+ \begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.8\textwidth]{images/MLNN_Mep_65536_ptrain_007}
+    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
+\end{figure}
+
 \end{frame}
@@
 \begin{frame}{DNN One-hot Decoder}{Show the results for the one-hot decoder in terms of train p, Mep, Parameters, etc}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+
+   \begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.8\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
+    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
+\end{figure}
+
 \end{frame}
@@
   \end{itemize}
+\end{frame}
+
+\subsection{Time Analysis}
+\begin{frame}{Delay Time Analysis}{Comparison of decoding time for each method.}
+  \begin{table}
+\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the mean MAP algorithm decoding time. } 
+\begin{center}
+
+  \begin{tabular}{ l | l | l | l | l }
+    \thickhline
+    MAP & A. Dec. & One-hot Dec. & A. Auto. & One-hot Auto. \\ \hline
+    $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ & $ \pm $ & $ \pm $ \\ 
+    \thickhline
+  \end{tabular}
+    \label{tab:timeanalysis}
+\end{center}
+\end{table}
 \end{frame}
@@
 \appendix
-\section<presentation>*{\appendixname}
-\subsection<presentation>*{Bibliography}
+\section<presentation>*{Bibliography}
 

### 2019-06-17 MyCode/autoencoderArray.py

@@ -26,8 +26,9 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([256, 256, 256])
-DecoderNodes = [256, 256, 32]
+#encoderNodes = np.array([512, 256]) - best so far
+encoderNodes = np.array([8192, 256, 256])
+DecoderNodes = [128, 64, 32]
 #%
 '''
     Architecture
@@ -37,10 +38,10 @@ Encoder = tf.keras.Sequential([
         layers.Dense(encoderNodes[0], activation='relu', input_shape=(k,), name='Input'),
         #layers.Dropout(rate=0.1),
         # Hidden Layer
-        layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
+        #layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
         #layers.Dropout(rate=0.1), 
         # Hidden Layer
-        layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
+        #layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
         # Coded Layer
         layers.Dense(n, activation='sigmoid', name='Codedfloat')
         ], name='Encoder')
@@ -71,7 +72,7 @@ plot_model(Autoencoder,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',sho
 '''
     Overall Settings/ Compilationl
 '''
-lossFunc = 'mse'
+lossFunc = 'logcosh'
 Autoencoder.compile(loss=lossFunc ,
               optimizer='adam',
               )
@@ -101,7 +102,7 @@ Autoencoder.save(path)  # creates a HDF5 file
 '''
     Prediction Array
 '''
-globalReps = 100
+globalReps = 1000
 globalErrorAutoencoder = np.empty([globalReps, len(pOptions)])
 for i_global in range(globalReps):
     for i_p in range(np.size(pOptions)):

### 2019-06-17 MyCode/libs-and-global-consts.py

@@ -86,7 +86,7 @@ globalReps = 1000
 # NN parameters
 title = 'AutoencoderArray'
 timestr = time.strftime("%Y%m%d-%H%M%S")
-numEpochs = 2**16
+numEpochs = 2**15
 batchSize = 256
 train_p = 0.03
 

### 2019-06-17 Rapport/Article/Article.tex
Seance 06/17/2019

@@
 
-The denomination one-hot comes from the output vector that is usually a long vector of zeros with one element equals to one.
+The denomination one-hot comes from the output vector which is usually a long vector of zeros with one element equals to one.
 This \textit{hot} vector element indicates which class the NN has classified the input.
-In this context, which message the code word refers to.
+In this context, which message encoded in 256 different classes the code word refers to.
+Basically, a message is mapped into a class and a class only.
 
-Tranformation from one-hot to messagges
+A straight forward consequence of this shallower NN is the number of training parameters almost three times less than the array decoder.
+Because of that, we could expect a lower training time.
+Also, we observed that the number of training epochs is substantially smaller for the one-hot decoder.
+It was necessary only $2^{14}$ epochs to achieve the results for the same batch size of $256$ and Adam optimizer.
 
-Abscence of noise layer
+Despite the NN complexity benefits of the one-hot decoder, it is required to transform the predicted vector into a message.
+Consequently, an additional operation is needed.
+To transform from a one-hot vector to a message, a ordered collection of the messages was stored in memory and the correspondent message is determined by a $argmax$ operation in the one-hot vector, which identifies which message is the most likely to have been transmitted through the channel.
 
-Lower complexity of the NN
-
-Loss function choice
-
-Optimizer choice
-
-Num epochs choice
-
-Batchsize choice
+The one-hot decoder was trained both with and without a noise layer.
+The model without a integrated channel converged faster to the solution.
+Tables \ref{tab:onehotdecoder} and \ref{tab:onehotdecoderTrain} resume all the training variables.
 

### 2019-06-17 MyCode/autoencoderArray.py

@@ -27,7 +27,7 @@ x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
 #encoderNodes = np.array([512, 256]) - best so far
-encoderNodes = np.array([8192, 256, 256])
+encoderNodes = np.array([512, 256, 256])
 DecoderNodes = [128, 64, 32]
 #%
 '''
@@ -38,10 +38,13 @@ Encoder = tf.keras.Sequential([
         layers.Dense(encoderNodes[0], activation='relu', input_shape=(k,), name='Input'),
         #layers.Dropout(rate=0.1),
         # Hidden Layer
-        #layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
+        layers.BatchNormalization(),
+        layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
+        layers.BatchNormalization(),
         #layers.Dropout(rate=0.1), 
         # Hidden Layer
-        #layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
+        layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
+        layers.BatchNormalization(),
         # Coded Layer
         layers.Dense(n, activation='sigmoid', name='Codedfloat')
         ], name='Encoder')

### 2019-06-18 Rapport/Article/Article.tex
Seance 06/18/2019

@@
 
-Briefly, a BSC is a noisy channel with a binary input $X\in{0,1}$ and a binary output $Y\in{0,1}$.
+Briefly, a BSC is a noisy channel with a binary input $X\in \{0,1\}$ on the left and a binary output $Y\in \{0,1\}$ on the right.
 They are linked through a stochastic model characterized by a crossover probability of value $p$ as illustrated in Fig.
 \ref{fig:BSC}.
-Typically $p$ is smaller than $0.10$.
+Typically, $p$ is smaller than $0.10$.
 
@@
   \centering
-    \includegraphics[width=0.5\textwidth]{images/BSC}
+    \includegraphics[width=0.4\textwidth]{images/BSC}
     \caption{Illustration of a binary symmetric channel with crossover probability $p$.}\label{fig:BSC}
@@
 
-A BSC is translated mathematically as follows.
+A BSC could be represented mathematically through conditional probabilities as follows.
 
@@
 
-The choice of the crossover probability of the training channel, $p_{t}$, was based on \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the best in terms of decoding precision after the DNN is trained.
-In addition, a deeper NN showed best results in \cite{ DBLP:journals/corr/OSheaH17} and were adopted to solve the problem.
+The choice of the crossover probability of the training channel, $p_{t}$, was based in the work of \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the best in terms of decoding precision after the DNN is trained.
+In addition, a deeper NN showed best results in \cite{DBLP:journals/corr/OSheaH17} and were adopted to solve the problem.
 The training parameters are displayed in table \ref{tab:arraydecoderTrain}.
@@
 Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}.
-Next, the BATCH SIZE.
+Next, a batch size matching the size of the training set which is sufficiently small to still operate in the regime of small batch was chosen.
+In \cite{DBLP:journals/corr/KeskarMNST16}, the authors demonstrated better performance for a NN training with stochastic gradient descent and its variants operating in a small-batch regime, usually between 32-512 data points.
 Finally, the necessary number of epochs to training is sensitive to the problem, since it impacts directly the training computational cost.
@@
 The model without a integrated channel converged faster to the solution.
-Tables \ref{tab:onehotdecoder} and \ref{tab:onehotdecoderTrain} resume all the training variables.
+Tables \ref{tab:onehotdecoder} and \ref{tab:onehotdecoderTrain} resume the architecture and training variables, respectively.
 
@@
 
+Explain any different training procedure
+
 \subsection{Decoders}
-BER Curves from both decoders, analyse why the number of epochs is so different.
+BER Curves from both decoders, analyze why the number of epochs is so different, .
+
+We can infer from figures \ref{fig:ArrayD} and \ref{fig:1HD} that both the array decoder and the one-hot decoder are able to learn a decodification function.
+Moreover,  its performances replicate the MAP algorithm's decoding capability.
+Hence, DNN decoders could substitute the mathematically optimal MAP decoder for a stochastic channel, without compromising the bit error probability.
 
@@
     \includegraphics[width=0.5\textwidth]{images/MLNN_Mep_65536_ptrain_007}
-    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
+    \caption{Array decoding BER performance. DNN array decoder trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
 \end{figure}
@@
     \includegraphics[width=0.5\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
-    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
+    \caption{One hot decoding BER performance. DNN one-hot decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
 \end{figure}
@@
 
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/AutoencoderArray_Mep_65536_ptrain_003_logcosh}
+    \caption{Array autoencoder BER performance. DNN array autoencoder trained with a channel crossover probability error of $p_t=0.03$.}\label{fig:arrayautoencoder}
+\end{figure}
+
 \subsection{Decoding time analysis}
+
+In order to analyze the delay in decoding for each technique studied, a monte carlo simulation was conducted.
+There were selected 20 different BSC channels.
+For each channel, $100000$ messages were transmitted and decodified for each decoder.
+This process was timed and repeated several times.
+In the end, an average time and standard deviation of the results were recorded.
+A normalization for the average MAP decoding time was done.
+The results are available in the table \ref{tab:timeanalysis}.
+
 compare time in decoding between map, autoencoder and a decoder.
@@
 
-  \begin{tabular}{ l | l | l | l | l }
+  \begin{tabular}{ c | c | c}
     \thickhline
-    MAP & A. Dec. & One-hot Dec. & A. Auto. & One-hot Auto. \\ \hline
-    $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ & $ \pm $ & $ \pm $ \\ 
+    MAP & Array Decoder & One-hot Decoder\\
+    $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ \\ \hline
+    \multicolumn{2}{c|}{Array. Autoencoder}  & One-hot Autoencoder\\
+
+     \multicolumn{2}{c|}{$0.00 \pm 0.00 $} & $0.00 \pm 0.00$ \\ 
     \thickhline
@@
 
+Talk about feasibility of decoders, improved performance of the channel delay with DNN decoders, challenges in implementing this architecture in real life, challenges in training the NNs, Future work, what else could be done with NN in error correction,
+
 \section*{Acknowledgment}
@@
 
-\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}
+\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16}
 
 \bibliographystyle{ieeetr}
-\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18}
+\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18,bib/KeskarMNST16}
 

### 2019-06-18 MyCode/autoencoder-retrain.py

@@ -5,9 +5,10 @@ Created on Wed Jun  5 13:47:36 2019
 @author: user
 """
 #%%
+
 numEpochs = 2**16
-directory = 'Retrain\Autoencoder1H\\'
-fileName = 'Autoencoder1H_Mep_65536_bs_256_32_32-32_32.h5'
+directory = 'Retrain/Autoencoder1H/'
+fileName = 'Autoencoder1H_Mep_65536_bs_256_1024-256.h5'
 Autoencoder = tf.keras.models.load_model(directory+'OriginalModel\\'+fileName)
 print("Loaded models from disk")
 path = directory+'TrainedModel\\'+fileName
@@ -15,7 +16,7 @@ path = directory+'TrainedModel\\'+fileName
 '''
     Overall Settings/ Compilationl
 '''
-lossFunc = 'mse'
+lossFunc = 'logcosh'
 Autoencoder.compile(loss=lossFunc ,
               optimizer='adam',
               )
@@ -47,7 +48,7 @@ Decoder = Autoencoder.layers[2]
 globalErrorAutoencoder=fn.onehotAutoencoderPrediction(Encoder, Decoder, messages, pOptions, globalReps, N, n, k)
 #% Plotting
 numEpochs = 2**17
-figPath = directory+'Autoencoder1H_Mep_'+str(numEpochs)+'_bs_256_32_32-32_32_p_train_0.png'
+figPath = directory+'Autoencoder1H_Mep_'+str(numEpochs)+'_bs_1024-256_p_train_0.png'
 plotBERp(globalErrorAutoencoder, 'One-hot Autoencoder')
 
 

### 2019-06-18 MyCode/autoencoder1H.py

@@ -19,8 +19,8 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([256, 256, 64])
-DecoderNodes = [256, 256]
+encoderNodes = np.array([2048, 128, 64])
+DecoderNodes = [128, 256]
 #%
 '''
     Architecture
@@ -29,7 +29,7 @@ Encoder = tf.keras.Sequential([
         # Input Layer
         layers.Dense(encoderNodes[0], activation='relu', input_shape=(256,), name='Input'),
         # Hidden Layer
-        layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
+        #layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
         # Hidden Layer
         #layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
         # Coded Layer
@@ -38,7 +38,7 @@ Encoder = tf.keras.Sequential([
 
 NoiseL = tf.keras.Sequential([
         # Rounded codeword
-        layers.Lambda(fn.roundCode, input_shape=(n,), 
+        layers.Lambda(roundCode, input_shape=(n,), 
                       output_shape=(n,), name='Codeword'),
         # Noise Layer
         layers.Lambda(tensorBSC,input_shape=(n,), 
@@ -47,7 +47,7 @@ NoiseL = tf.keras.Sequential([
 
 Decoder = tf.keras.Sequential([ # Array to define layers
         layers.Dense(DecoderNodes[0], activation='relu', input_shape=(n,), name='DHL1'),
-        layers.Dense(DecoderNodes[1], activation='relu', name='DHL2'),
+        #layers.Dense(DecoderNodes[1], activation='relu', name='DHL2'),
         layers.Dense(256, activation='softmax', name='1H_Output')
         ], name = 'Decoder')
 
@@ -57,7 +57,7 @@ plot_model(Autoencoder1H,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',s
 '''
     Overall Settings/ Compilation
 '''
-lossFunc = 'mse'
+lossFunc = 'logcosh'
 Autoencoder1H.compile(loss=lossFunc ,
               optimizer='adam',
               )
@@ -91,7 +91,7 @@ t = TicToc('name')
 t.tic()
 
 
[…]

### 2019-06-18 MyCode/autoencoderArray.py

@@ -51,7 +51,7 @@ Encoder = tf.keras.Sequential([
 
 NoiseL = tf.keras.Sequential([
         # Rounded codeword
-        layers.Lambda(fn.roundCode, input_shape=(n,), 
+        layers.Lambda(roundCode, input_shape=(n,), 
                       output_shape=(n,), name='Codeword'),
         # Noise Layer
         layers.Lambda(tensorBSC,input_shape=(n,), 

### 2019-06-18 MyCode/libs-and-global-consts.py

@@ -84,11 +84,11 @@ globalReps = 1000
 
 #%%
 # NN parameters
-title = 'AutoencoderArray'
+title = 'Autoencoder1H'
 timestr = time.strftime("%Y%m%d-%H%M%S")
 numEpochs = 2**16
 batchSize = 256
-train_p = 0.03
+train_p = 0.0
 
 path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
 checkpointPath = 'Checkpoints/'+title+'/'+timestr+'_'+title+'_Mep_{epoch:02d}-{loss:.8f}.h5'
@@ -106,6 +106,9 @@ def tensorBSC(x):
     noiseFloat = K.cast(noise, dtype=tf.float32)
     result = tf.math.add(noiseFloat, x)%2
     return result
+
+def roundCode(x):
+    return tf.stop_gradient(K.round(x)-x)+x
 
 def plotBERp(globalErrorMLNN, legendEntry):
     avgGlobalErrorMLNN = np.average(globalErrorMLNN,0)

### 2019-06-18 Rapport/Article/Article.tex
Seance 06/18/2019

@@
 
+Trainin set for array decoding.
+
 The choices of training parameters is a mix between what other works in literature uses and a hyper parametric analytic approach, where several trials lead to the decision.
@@
 
+Trainin set for one-hot decoding, say how was pseudo-randomly generated, sa that the stastistic was checked to have 1/2 of 1s and 1/2 of 0s.
+
 The one-hot decoder was trained both with and without a noise layer.
@@
 
+However, we still need a decoding matrix.... Introduces autoencoder that could train without G
+
 \subsection{Autoencoder}
-Architecture, Training and simulation parameters, procedure to find a M epoch suitable, choices of p train, Talk about NN separation,
+Architecture, Training and simulation parameters, procedure to find a M epoch suitable, choices of p train, Talk about NN separation, no necessity for G, a slight different trainin approach (training data and labels are identical), same channel rate (16 units) for a message of 8 bits,
 

### 2019-06-18 MyCode/autoencoder-retrain.py

@@ -8,7 +8,7 @@ Created on Wed Jun  5 13:47:36 2019
 
 numEpochs = 2**16
 directory = 'Retrain/Autoencoder1H/'
-fileName = 'Autoencoder1H_Mep_65536_bs_256_1024-256.h5'
+fileName = 'Autoencoder1H_Mep_65536_bs_256_-2048-128.h5'
 Autoencoder = tf.keras.models.load_model(directory+'OriginalModel\\'+fileName)
 print("Loaded models from disk")
 path = directory+'TrainedModel\\'+fileName
@@ -48,7 +48,7 @@ Decoder = Autoencoder.layers[2]
 globalErrorAutoencoder=fn.onehotAutoencoderPrediction(Encoder, Decoder, messages, pOptions, globalReps, N, n, k)
 #% Plotting
 numEpochs = 2**17
-figPath = directory+'Autoencoder1H_Mep_'+str(numEpochs)+'_bs_1024-256_p_train_0.png'
+figPath = directory+'images/Autoencoder1H_Mep_'+str(numEpochs)+'_bs_2048-128_p_train_0.png'
 plotBERp(globalErrorAutoencoder, 'One-hot Autoencoder')
 
 

### 2019-06-18 MyCode/autoencoder1H.py

@@ -19,7 +19,7 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([2048, 128, 64])
+encoderNodes = np.array([256, 128, 96, 64, 32]) # add a 192 layer at the beginning
 DecoderNodes = [128, 256]
 #%
 '''
@@ -28,10 +28,19 @@ DecoderNodes = [128, 256]
 Encoder = tf.keras.Sequential([
         # Input Layer
         layers.Dense(encoderNodes[0], activation='relu', input_shape=(256,), name='Input'),
+        layers.BatchNormalization(),
         # Hidden Layer
-        #layers.Dense(encoderNodes[1], activation='relu', name='EHL1'),
+        layers.Dense(encoderNodes[1], activation='relu', name='EHL'),
+        layers.BatchNormalization(),
         # Hidden Layer
-        #layers.Dense(encoderNodes[2], activation='relu', name='EHL2'),
+        layers.Dense(encoderNodes[2], activation='relu', name='EHL1'),
+        layers.BatchNormalization(),
+        # Hidden Layer
+        layers.Dense(encoderNodes[3], activation='relu', name='EHL2'),
+        layers.BatchNormalization(),
+        # Hidden Layer
+        layers.Dense(encoderNodes[4], activation='relu', name='EHL3'),
+        layers.BatchNormalization(),
         # Coded Layer
         layers.Dense(n, activation='sigmoid', name='Codedfloat'),
         ], name='Encoder')  
@@ -60,6 +69,7 @@ plot_model(Autoencoder1H,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',s
 lossFunc = 'logcosh'
 Autoencoder1H.compile(loss=lossFunc ,
               optimizer='adam',
+              metrics=[fn.metricBER1H]
               )
 '''
     Summaries and checkpoints (to do)

### 2019-06-18 MyCode/autoencoderArray.py

@@ -78,16 +78,17 @@ plot_model(Autoencoder,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',sho
 lossFunc = 'logcosh'
 Autoencoder.compile(loss=lossFunc ,
               optimizer='adam',
+              metric=[fn.metricBER]
               )
 '''
     Summaries and checkpoints (to do)
 '''
 summary = Autoencoder.summary()
-'''checkpoint = tf.keras.callbacks.ModelCheckpoint(
+checkpoint = tf.keras.callbacks.ModelCheckpoint(
         checkpointPath, monitor='loss', 
         verbose=0, save_best_only=True, save_weights_only=False, mode='min', period=checkpointPeriod)
-'''
-callbacks_list = []
+
+callbacks_list = [checkpoint]
 ''' 
     Training
 '''

### 2019-06-18 MyCode/my_functions.py

@@ -259,7 +259,7 @@ def metricBER(y_true, y_pred):
 
 def metricBER1H(y_true, y_pred):
     #return K.mean(K.not_equal(K.argmax(y_true),K.argmax(y_pred)))
-    return K.mean(K.not_equal(y_true,K.round(y_pred)))
+    return K.mean(K.not_equal(y_true,K.round(y_pred))) #????
 
 def tensorPossibleMessages(name):
     n = name[0]

### 2019-06-19 Rapport/Article/Article.tex
Seance 06/19/2019

@@
 
-For a classification problem in communication, the \textit{categorical cross-entropy} loss function $l(\textbf{p},\textbf{q}):\mathbb{R} ^{N_L}\times \mathbb{R} ^{N_L}\mapsto \mathbb{R}$ defined in \eqref{eq:c-e} is most common.
+For a classification problem in communication, the \textit{categorical cross-entropy} loss function $l(\textbf{p},\textbf{q}):\mathbb{R} ^{N_L}\times \mathbb{R} ^{N_L}\mapsto \mathbb{R}$ defined in equation \eqref{eq:c-e} is most common.
 It is derived from the log-likelihood of a training set where $q_j$ is the estimated probability of the outcome $j$ and $p_{j}$ is the true probability.
 Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{murphy2013machine} where $j=1,...,N_L$.
-\begin{equation}\label{eq:c-e}
-	l(\textbf{p},\textbf{q})=-\sum _{j}p_{j}\log \left( q_{j}\right)
-\end{equation}
+Another commonly used loss function, indispensable for this research, is the \textit{mean squared error} (MSE).
+The MSE is the  sum of squared distances between our target variable and predicted values and it is defined in equation \ref{eq:mse}.
+
+\begin{align}
+\
+	\text{Cross-entropy: }l_1(\textbf{p},\textbf{q}) &=-\sum _{j}p_{j}\log \left( q_{j}\right)\label{eq:c-e} \\
+	\text{MSE: } l_2(\textbf{p},\textbf{q}) &=\frac{1}{N_L}\sum _{j}(p_{j}-q_{j})^{2} \label{eq:mse}
+\end{align}
 
@@
 
-The architecture of the DNN necessary to achieve a great performance in terms of BER is available in table \ref{tab:arraydecoder}.
-What should be remarked is the first layer of the DNN.
+The architecture of the DNN necessary to achieve a great performance in terms of BER is not evident.
+What should be remarked though is the first layer of the DNN.
 A \textit{Lambda} layer is a layer in which the activation operations are user defined and the parameters are non-trainable.
@@
 The choice of the crossover probability of the training channel, $p_{t}$, was based in the work of \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the best in terms of decoding precision after the DNN is trained.
-In addition, a deeper NN showed best results in \cite{DBLP:journals/corr/OSheaH17} and were adopted to solve the problem.
-The training parameters are displayed in table \ref{tab:arraydecoderTrain}.
-
-Trainin set for array decoding.
-
-The choices of training parameters is a mix between what other works in literature uses and a hyper parametric analytic approach, where several trials lead to the decision.
-First, the binary cross-entropy as loss function is widely used in literature and  showed to converge to the desired result consistently during tests.
-For the optimizer, the \textit{Adam} built-in Keras optimizer was chosen.
-Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}.
-Next, a batch size matching the size of the training set which is sufficiently small to still operate in the regime of small batch was chosen.
-In \cite{DBLP:journals/corr/KeskarMNST16}, the authors demonstrated better performance for a NN training with stochastic gradient descent and its variants operating in a small-batch regime, usually between 32-512 data points.
-Finally, the necessary number of epochs to training is sensitive to the problem, since it impacts directly the training computational cost.
-As a result, we tried several training epoch sizes and found $2^{16}$ to have precise results and be computationally viable.
+In addition, a deeper DNN showed best results in \cite{DBLP:journals/corr/OSheaH17} where 3 hidden layers were used.
+Consequently, the same architecture were adopted to solve the problem as shown in table \ref{tab:arraydecoder}.
 
@@
 \end{table}
+
+The training set for the array decoder contained all the possible messages as training labels and its correspondents code words as training data.
+Because of the chosen encoding, the training data is a $256\times16$ matrix and the training label is a $256\times8$ matrix.
+To generate a possible message, a stochastic algorithm was developed to fabricate random 8-length vectors of integers 1s and 0s which the average value considering all elements was around $0.5$.
+In this way we could guarantee that our training environment were representative of a real life application.
+Where different bits are transmitted with the same probability.
+
+Even though it is not recommended to train a NN with all the available data - because it might lead to over-fitting -, for this case does not seems to be a problem.
+WHY?????
+(noisy channel?
+onehot decodeR?)
+
+The choices of training parameters is a mix between what other works in literature uses and a hyper parametric analytic approach, where several trials lead to the decision.
+First, the binary cross-entropy as loss function is widely used in literature and  showed to converge to the desired result consistently during tests.
+For the optimizer, the \textit{Adam} built-in Keras optimizer was chosen.
+Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}.
+Next, a batch size matching the size of the training set which is sufficiently small to still operate in the regime of small batch was chosen.
+In \cite{DBLP:journals/corr/KeskarMNST16}, the authors demonstrated better performance for a NN training with stochastic gradient descent and its variants operating in a small-batch regime, usually between 32-512 data points.
+Finally, the necessary number of epochs to training is sensitive to the problem, since it impacts directly the training computational cost.
+As a result, we tried several training epoch sizes and found $2^{16}$ to have precise results and be computationally viable.
+The training parameters are displayed in table \ref{tab:arraydecoderTrain}.
 
@@
 
-Trainin set for one-hot decoding, say how was pseudo-randomly generated, sa that the stastistic was checked to have 1/2 of 1s and 1/2 of 0s.
+The training set in this case differs from the array decoder.
+The training labels are one-hot vectors which maps into real messages through a bijective function.
+Thus, it is a matrix of dimension $256\times256$ composed of 256 different on-hot vectors.
+The training data is still the linear encoding of the real messages by the matrix $\textbf{G}$.
 
@@
 
-However, we still need a decoding matrix.... Introduces autoencoder that could train without G
+Despite of this, an encoding matrix is still needed.
+A commonly used method in ML for image recognition without training labels is the autoencoder.
+In this context, the NN is able to learn not only a decoding algorithm, but also a decoding one as introduced in the following section.
 
@@
 Architecture, Training and simulation parameters, procedure to find a M epoch suitable, choices of p train, Talk about NN separation, no necessity for G, a slight different trainin approach (training data and labels are identical), same channel rate (16 units) for a message of 8 bits,
+
+\subsubsection{Array Autoencoder}
 
@@
  Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
- Binary cross-entropy & Adam & $2^{14}$ & $256$ \\
+ MSE & Adam & $2^{17}$ & $256$ \\
     \thickhline
@@
 
+\subsubsection{One-hot Autoencoder}
+
+Briefly state that it is basically the same, apart from the input and output layers AND it demands a tremendously different architecture.
+
+\begin{table}
+\caption{DNN one-hot autoencoder architecture.} 
+\begin{center}
+
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: 256 \\
+ & Dense: X, activation: ReLU  \\
+ & Dense: 16, activation: Sigmoid  \\ \hline
+
+    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $n$ \\
+ & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
+
+ \multirow{2}{*}{Decoder} & Dense: 128, input size: 16 \\
+ & Dense: 256, activation: Sigmoid  \\ \hline
+
+    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:onehotautoencoder}
+\end{center}
+\end{table}
+
+\begin{table}
+\caption{DNN one-hot autodecoder training parameters.} 
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ MSE & Adam & $2^{17}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:onehotautoencoderTrain}
+\end{center}
+\end{table}
+
 \section{Results and Analysis}
 
-Explain any different training procedure
+This section is divided in three parts.
+The first shows the results obtained to the two implemented decoders.
+Observations during the training of the DNN are made.
+Secondly, the results for the autoencoders is brought with some observations during the training period.
+The discussion and analysis for these two subsections is around the BER performance for the decoders and autoencoders.
+Finally, a decoding time performance analysis is explained and the results are shown and discussed.
 
@@
     \includegraphics[width=0.5\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
-    \caption{One hot decoding BER performance. DNN one-hot decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
+    \caption{One hot decoding BER performance. DNN one-hot decoder trained without a noisy channel.}\label{fig:1HD}
 \end{figure}
 
-\subsection{Autoencoder}
+\subsection{Autoencoders}
 BER Curves of simulation varying p, best choice of p, analyse autoencoder with best choice of p
@@
 
-compare time in decoding between map, autoencoder and a decoder.
+say that only the decoding part was measured  for autoencoders
 

### 2019-06-19 MyCode/autoencoder-array-fine-tuning.py

@@ -165,4 +165,42 @@ fn.createDir(figPath)
 fig.set_size_inches(width, height)
 fig.savefig(figPath+'\MAP_'+title+'_Mep_'+str(numEpochs)+'.png', bbox_inches='tight', dpi=300)
 
+#%%%
+'''
+    Check checkedpoints autoencoders
+'''
+'''
+    Constants
+'''
+numEpochs = 2**13  #2**16 approx 65000
+batchSize = 256
+train_p = 0.0
+timestr = time.strftime("%Y%m%d-%H%M%S")
+title = 'Autoencoder1H'
+path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
+directory = 'Fine-tuning/full-models/'
+fileNames = ['20190619-143023_Autoencoder1H_Mep_12288-0.00000000.h5', 
+             '20190619-143023_Autoencoder1H_Mep_16384-0.00000000.h5',
+             '20190619-143023_Autoencoder1H_Mep_20480-0.00000000.h5',
+             '20190619-143023_Autoencoder1H_Mep_40960-0.00000000.h5',
+             '20190619-143023_Autoencoder1H_Mep_45056-0.00000000.h5',
+             '20190619-143023_Autoencoder1H_Mep_49152-0.00000000.h5']
+
+for i in range(len(fileNames)):  
+    '''
+        Load model
+    '''
+    fileName = fileNames[i]
+    loadedModel = tf.keras.models.load_model(directory+fileName)
+    print("Loaded model from disk")
+    
+    Encoder = loadedModel.layers[0]
+    Decoder = loadedModel.layers[2]
+    
+    globalReps = 100
+    globalErrorAutoencoder1H = fn.onehotAutoencoderPrediction(Encoder, Decoder, 
+                               messages, pOptions, globalReps, N, n, k)
+    
+    plotBERp(globalErrorAutoencoder1H, 'One-hot Autoencoder')
+
 

### 2019-06-19 MyCode/autoencoder-retrain.py

@@ -6,9 +6,9 @@ Created on Wed Jun  5 13:47:36 2019
 """
 #%%
 
-numEpochs = 2**16
+numEpochs = 2**10
 directory = 'Retrain/Autoencoder1H/'
-fileName = 'Autoencoder1H_Mep_65536_bs_256_-2048-128.h5'
+fileName = 'Autoencoder1H_Mep_131072_bs_256_-196_128_96_64_32-128.h5'
 Autoencoder = tf.keras.models.load_model(directory+'OriginalModel\\'+fileName)
 print("Loaded models from disk")
 path = directory+'TrainedModel\\'+fileName
@@ -19,6 +19,7 @@ path = directory+'TrainedModel\\'+fileName
 lossFunc = 'logcosh'
 Autoencoder.compile(loss=lossFunc ,
               optimizer='adam',
+              metrics=[fn.metricBER1H]
               )
 '''
     Summaries and checkpoints
@@ -47,7 +48,7 @@ Decoder = Autoencoder.layers[2]
 #globalErrorAutoencoder=fn.arrayAutoencoderPrediction(Encoder, Decoder, pOptions, globalReps, N, n, k)
 globalErrorAutoencoder=fn.onehotAutoencoderPrediction(Encoder, Decoder, messages, pOptions, globalReps, N, n, k)
 #% Plotting
-numEpochs = 2**17
+numEpochs = 2**17 + 2**10
 figPath = directory+'images/Autoencoder1H_Mep_'+str(numEpochs)+'_bs_2048-128_p_train_0.png'
 plotBERp(globalErrorAutoencoder, 'One-hot Autoencoder')
 

### 2019-06-19 MyCode/autoencoder1H.py

@@ -19,8 +19,8 @@ u_train_labels = np.repeat(u_train_labels, 1, axis=0)
 x_train_data = np.repeat(x_train_data, 1, axis=0)
 trainSize = np.size(x_train_data, 0)
 
-encoderNodes = np.array([256, 128, 96, 64, 32]) # add a 192 layer at the beginning
-DecoderNodes = [128, 256]
+encoderNodes = np.array([196, 128, 96, 64, 32]) # add a 192 layer at the beginning
+DecoderNodes = [128, 192]
 #%
 '''
     Architecture
@@ -56,7 +56,7 @@ NoiseL = tf.keras.Sequential([
 
 Decoder = tf.keras.Sequential([ # Array to define layers
         layers.Dense(DecoderNodes[0], activation='relu', input_shape=(n,), name='DHL1'),
-        #layers.Dense(DecoderNodes[1], activation='relu', name='DHL2'),
+        layers.BatchNormalization(),
         layers.Dense(256, activation='softmax', name='1H_Output')
         ], name = 'Decoder')
 
@@ -68,9 +68,9 @@ plot_model(Autoencoder1H,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',s
 '''
 lossFunc = 'logcosh'
 Autoencoder1H.compile(loss=lossFunc ,
-              optimizer='adam',
-              metrics=[fn.metricBER1H]
+              optimizer='adam'
               )
+#metrics=[fn.metricBER1H]
 '''
     Summaries and checkpoints (to do)
 '''
@@ -78,6 +78,9 @@ summary = Autoencoder1H.summary()
 checkpoint = tf.keras.callbacks.ModelCheckpoint(
         checkpointPath, monitor='loss', 
         verbose=0, save_best_only=True, save_weights_only=False, mode='min', period=checkpointPeriod)
+            
+es =  tf.keras.callbacks.EarlyStopping(monitor='loss', min_delta=0.0000000001, patience=256, verbose=1, mode='auto')          
+#es = tf.keras.callbacks.EarlyStopping(monitor='loss', mode='min', verbose=1, patience=1024)
 callbacks_list = [checkpoint]
 ''' 
     Training
@@ -101,7 +104,7 @@ t = TicToc('name')
 t.tic()
 
 
-globalReps = 1000
+globalReps = 100
 globalErrorAutoencoder1H = fn.onehotAutoencoderPrediction(Encoder, Decoder, 
                                messages, pOptions, globalReps, N, n, k)
 

### 2019-06-20 Rapport/Article/Article.tex
Seance 06/20/2019

@@
 The architecture of the DNN necessary to achieve a great performance in terms of BER is not evident.
-What should be remarked though is the first layer of the DNN.
-A \textit{Lambda} layer is a layer in which the activation operations are user defined and the parameters are non-trainable.
+To begin we must define which layers were used and why.
+A \textit{Lambda} layer is a layer in which the activation function is user defined and the parameters are non-trainable.
 Thus, we were able to define a layer that act as a BSC to force our DNN to learn its effects in the training data.
 The fact that the parameters are non-trainable means that the noise layer is applied in forward propagation only.
+A Dense layer is a densely connected layer, which means that each neuron receives input from all the neurons in the previous layer.
+Thus it is often used to change the dimensions of a vector.
+In other words, it applies a rotation, scaling and translation transform to a vector as shown in equation \ref{eq:eqFP}.
+The values in the weight matrix are trainable parameters which are updated during BP \cite{chollet2015keras}.
 
@@
 
+\begin{table}
+\caption{DNN array decoder training parameters.} 
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Binary cross-entropy & Adam & $2^{16}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:arraydecoderTrain}
+\end{center}
+\end{table}
+
 The training set for the array decoder contained all the possible messages as training labels and its correspondents code words as training data.
@@
 The training parameters are displayed in table \ref{tab:arraydecoderTrain}.
-
-\begin{table}
-\caption{DNN array decoder training parameters.} 
-\begin{center}
-\begin{tabular}{ l  |  l | l | l  }
-    \thickhline
- Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
- Binary cross-entropy & Adam & $2^{16}$ & $256$ \\
-    \thickhline
-  \end{tabular}
-    \label{tab:arraydecoderTrain}
-\end{center}
-\end{table}
 
@@
 
-Despite of this, an encoding matrix is still needed.
-A commonly used method in ML for image recognition without training labels is the autoencoder.
-In this context, the NN is able to learn not only a decoding algorithm, but also a decoding one as introduced in the following section.
+Even though DNN encoders are useful, an encoding matrix is still needed.
+With a view to eliminate completely the need of an explicit encoding, a technique commonly used method in ML for image recognition is explored.
+The autoencoder is able to learn not only a decoding algorithm, but also a decoding one as introduced in the following section.
 
@@
 
-Briefly state that it is basically the same, apart from the input and output layers AND it demands a tremendously different architecture.
+Briefly state that it is basically the same, apart from the input and output layers AND it demands a tremendously different architecture, batch normalization.
 
@@
  \multirow{2}{*}{Decoder} & Dense: 128, input size: 16 \\
- & Dense: 256, activation: Sigmoid  \\ \hline
+ & Dense: 256, activation: Softmax  \\ \hline
 
@@
 \begin{table}
-\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the mean MAP algorithm decoding time. } 
+\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. } 
 \begin{center}
@@
 
-Talk about feasibility of decoders, improved performance of the channel delay with DNN decoders, challenges in implementing this architecture in real life, challenges in training the NNs, Future work, what else could be done with NN in error correction,
+Talk about feasibility of decoders, improved performance of the channel delay with DNN decoders, challenges in implementing this architecture in real life, challenges in training the NNs, Future work, what else could be done with NN in error correction, Advantages and disadvantages in comparison with linear decoding.
 

### 2019-06-20 MyCode/autoencoder-array-fine-tuning.py

@@ -172,19 +172,19 @@ fig.savefig(figPath+'\MAP_'+title+'_Mep_'+str(numEpochs)+'.png', bbox_inches='ti
 '''
     Constants
 '''
-numEpochs = 2**13  #2**16 approx 65000
+numEpochs = 2**14  #2**16 approx 65000
 batchSize = 256
-train_p = 0.0
+train_p = 0.03
 timestr = time.strftime("%Y%m%d-%H%M%S")
 title = 'Autoencoder1H'
 path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
 directory = 'Fine-tuning/full-models/'
-fileNames = ['20190619-143023_Autoencoder1H_Mep_12288-0.00000000.h5', 
-             '20190619-143023_Autoencoder1H_Mep_16384-0.00000000.h5',
-             '20190619-143023_Autoencoder1H_Mep_20480-0.00000000.h5',
-             '20190619-143023_Autoencoder1H_Mep_40960-0.00000000.h5',
-             '20190619-143023_Autoencoder1H_Mep_45056-0.00000000.h5',
-             '20190619-143023_Autoencoder1H_Mep_49152-0.00000000.h5']
+fileNames = ['20190619-161454_Autoencoder1H_Mep_32768-0.00000000.h5', 
+             '20190619-161454_Autoencoder1H_Mep_49152-0.00000000.h5',
+             '20190619-161454_Autoencoder1H_Mep_57344-0.00000000.h5',
+             '20190619-161454_Autoencoder1H_Mep_69632-0.00000000.h5',
+             '20190619-161454_Autoencoder1H_Mep_118784-0.00000000.h5',
+             '20190619-161454_Autoencoder1H_Mep_126976-0.00000000.h5']
 
 for i in range(len(fileNames)):  
     '''
@@ -197,7 +197,7 @@ for i in range(len(fileNames)):
     Encoder = loadedModel.layers[0]
     Decoder = loadedModel.layers[2]
     
-    globalReps = 100
+    globalReps = 1000
     globalErrorAutoencoder1H = fn.onehotAutoencoderPrediction(Encoder, Decoder, 
                                messages, pOptions, globalReps, N, n, k)
     

### 2019-06-20 MyCode/autoencoder1H.py

@@ -56,7 +56,6 @@ NoiseL = tf.keras.Sequential([
 
 Decoder = tf.keras.Sequential([ # Array to define layers
         layers.Dense(DecoderNodes[0], activation='relu', input_shape=(n,), name='DHL1'),
-        layers.BatchNormalization(),
         layers.Dense(256, activation='softmax', name='1H_Output')
         ], name = 'Decoder')
 
@@ -66,7 +65,7 @@ plot_model(Autoencoder1H,to_file='graphNN/'+title+'/'+timestr+'_'+title+'.pdf',s
 '''
     Overall Settings/ Compilation
 '''
-lossFunc = 'logcosh'
+lossFunc = 'mse'
 Autoencoder1H.compile(loss=lossFunc ,
               optimizer='adam'
               )

### 2019-06-20 MyCode/libs-and-global-consts.py

@@ -88,7 +88,7 @@ title = 'Autoencoder1H'
 timestr = time.strftime("%Y%m%d-%H%M%S")
 numEpochs = 2**17
 batchSize = 256
-train_p = 0.0
+train_p = 0.03
 
 path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
 checkpointPath = 'Checkpoints/'+title+'/'+timestr+'_'+title+'_Mep_{epoch:02d}-{loss:.8f}.h5'

### 2019-06-21 Rapport/Article/Article.tex
Seance 06/21/2019

@@
 \subsection{Autoencoder}
-Architecture, Training and simulation parameters, procedure to find a M epoch suitable, choices of p train, Talk about NN separation, no necessity for G, a slight different trainin approach (training data and labels are identical), same channel rate (16 units) for a message of 8 bits,
+For the DNN autoencoder, two different concepts were studied.
+The array autoencoder takes as input the messages and it tries to replicate them, encoding it in a 16 bit code word, transmitting through a noisy channel and decoding it in the second portion of the NN.
+A one-hot autoencoder is inputted by a one-hot vector representing one unique k-bit message.
+It encodes the vector into a n-bit long code word, transmits it through a noisy channel and decode it in the second part of the NN.
+
+This end-to-end approach for both architectures is promising for digital communication, since there is no need for a predefined linear encoding.
+Theoretically, the DNN could learn the best encoding and associated decoding to a specific channel.
+Once trained, they could perform optimally in terms of encoding and decoding for a channel.
+To be clear during the discussion of this and the following sessions, the autoencoder will be divided in: the encoder, the channel and the decoder portions as shown in Fig.
+\ref{fig:DDNNAutoencoder}.
+In total, there were tested 71 array and 80 one-hot autoencoders different architecture or training configuration.
 
 \subsubsection{Array Autoencoder}
+
+Even though in theory the autoencoder could be able to achieve at least the same MAP performance, finding the parameters which unveils its capabilities requires multiple testing and a rigorous hyper-parametric analysis.
+For the array autoencoder, we first experimented with the depth of the encoder.
+The number of hidden layers in a DNN has influence in the BP algorithm.
+Extremely deep NN, with more than 4 hidden layers, may encounter problems in the gradient calculations due to vanishing weights.
+When the weights becomes too small, they may be source of underflow, depending on the machine precision.
+
+The encoder portion was first analyzed in a standalone session.
+We trained the encoder to replicate the linear encoding proposed in the preceding sections as we studied the effects of each parameter in the training performance.
+A more wide architecture showed faster convergence and the use of three hidden layers had no positive effects.
+In the end, for the integrated model we opted for using two hidden layers, with 512 and 256 densely connected units as detailed in table \ref{tab:arrayautoencoder}.
+The activation functions remained the same as in the DNN decoders discussed in the previous section.
+
+Integrating the channel in the autoencoder model posed a challenge.
+The prediction of the encoder outputs a real vector with elements in the codomain $(0,1)$ of the sigmoid function.
+Henceforth, a rounding function is mandatory for the channel transmission.
+The issue is that a rounding function is not differentiable in zero.
+This would stop the BP as no gradient would be computed.
+The solution was to only round values in the forward propagation and ignore this function during BP.
+The command \textit{tf.keras.backend.stop\textunderscore gradient} in a python environment with \textit{tf.keras} as backend does exactly what we wanted and was implemented in the rounding lambda layer.
+
+The decoder architecture was also based on the standalone array decoder.
+In spite of this, there were no difference outstanding in decoding with or without the last hidden layer of the decoder.
+In the other hand, the tests with a condensate NN with one hidden layer demanded a greater training time to arrive in the same results.
+The final test was done with the entire autoencoder and the final tweaks were done.
+
+While training the autoencoder we noticed that the binary cross-entropy loss function would need more training epochs than the MSE and \textit{logcosh} loss functions, even though they converged in similar results.
+The batch size was maintained as the size of the first dimension of the training set and there was no reason for changing the optimizer.
+Notwithstanding, the ensemble of changes and optimizations were not enough to find a suitable architecture for channels with $p<0.2$ as the MAP algorithm for the 1/2 rate encoder would do better in this scenario.
+Table \ref{tab:arrayautoencoderTrain} summarizes the array autoencoder training parameters.
+
+Furthermore, adding batch normalization layers inside the DNN was fundamental.
+With greater training parameters, achieving convergence in a reasonable time requires more computing power without any specific technique.
+The constant changes of weight values in the intermediate layers in every training step is the principal contributor to training cost.
+This problem is known as \textit{internal covariance shift}.
+What batch normalization does is normalize the input of a layer by its batch statistics and then scale and shift it.
+For further details, refer to \cite{DBLP:journals/corr/IoffeS15}.
+
+The final parameter to determine is the training channel cross over probability.
+Once the architecture and training parameters were defined, a simulation for the same DNN with different $p_t$ was done and the trained models were analyzed in terms of BER.
+A $p_t=0.03$ yielded best performance.
 
@@
     \thickhline
-    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: 8 \\
- & Dense: X, activation: ReLU  \\
+    \multirow{3}{*}{Encoder} & Dense: 512, activation: ReLU, input size: 8 \\
+   &  Batch normalization \\
+ & Dense: 256, activation: ReLU \\
+ &  Batch normalization \\
  & Dense: 16, activation: Sigmoid  \\ \hline
 
-    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $n$ \\
+    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $16$ \\
  & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
 
- \multirow{2}{*}{Decoder} & Dense: X, input size: 16 \\
- & Dense: X, activation: ReLU  \\
+ \multirow{2}{*}{Decoder} & Dense: 128, input size: 16 \\
+ &  Batch normalization \\
+ & Dense: 64, activation: ReLU  \\
+ &  Batch normalization \\
  & Dense: 8, activation: Sigmoid  \\ \hline
 
-    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \multicolumn{2}{l}{\textbf{Total parameters: }154072}\\
     \thickhline
@@
 
-Briefly state that it is basically the same, apart from the input and output layers AND it demands a tremendously different architecture, batch normalization.
+The procedure to implement the one-hot autoencoder was similar to the array's autoencoder one.
+Briefly, it was shown experimentally that for this model, a deeper encoder with layers of gradually decreasing number of units were more suited.
+We tried both a wider and shallower architecture and a deeper and narrow architecture.
+The second had better BER performance.
+The final architecture is available in table \ref{tab:onehotautoencoder}.
+
+In terms of noise, their presence in the channel degraded vastly the results, as the DNN was not able to learn.
+The challenge of the one-hot encoder implementation is the fact that we are trying to represent a 256 size length vector into a 16 length vector to transmit through the specified channel.
+The hypothesis is that it requires several steps to converge into a representative state.
+Thus, no noise channel was implemented, only a rounding layer.
+
+The decoder utilized has fewer layers than the array autoencoder, however more parameters.
+In terms of activation functions, there is no difference from the DNN one-hot decoder, which utilized a multi-class classification framework as well.
+In addition, batch normalization proved to be valuable and accelerated the overall training of the DNN.
+
+Apart from the number of training epochs, the other training parameters are identical to the array autoencoder.
+As table \ref{tab:onehotautoencoderTrain} shows, the one-hot autoencoder utilized two times less training epochs to converge.
 
@@
     \thickhline
-    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: 256 \\
- & Dense: X, activation: ReLU  \\
+    \multirow{3}{*}{Encoder} & Dense: 196, activation: ReLU, input size: 256\\
+   & Batch normalization \\
+ & Dense: 128, activation: ReLU  \\
+ & Batch normalization \\
+ & Dense: 96, activation: ReLU  \\
+ & Batch normalization \\
+ & Dense: 64, activation: ReLU  \\
+ & Batch normalization \\
+ & Dense: 32, activation: ReLU  \\
+ & Batch normalization \\
  & Dense: 16, activation: Sigmoid  \\ \hline
 
-    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $n$ \\
- & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
+   Channel & Lambda: $Round(\textbf{x})$, input size: $16$ \\ \hline
 
- \multirow{2}{*}{Decoder} & Dense: 128, input size: 16 \\
+ Decoder & Dense: 128, activation: ReLU, input size: 16 \\
+ & Batch normalization \\
  & Dense: 256, activation: Softmax  \\ \hline
 
-    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
+    \multicolumn{2}{l}{\textbf{Total parameters: }134052}\\
     \thickhline
@@
  Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
- MSE & Adam & $2^{17}$ & $256$ \\
+ MSE & Adam & $2^{16}$ & $256$ \\
     \thickhline
@@
   \centering
+    \includegraphics[width=0.5\textwidth]{images/MAP_AutoencoderArray_Mep_65536_64_128_256_p_analysis}
+    \caption{CAPTION}\label{fig:parrayanalysis}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
     \includegraphics[width=0.5\textwidth]{images/AutoencoderArray_Mep_65536_ptrain_003_logcosh}
     \caption{Array autoencoder BER performance. DNN array autoencoder trained with a channel crossover probability error of $p_t=0.03$.}\label{fig:arrayautoencoder}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
+    \caption{CAPTION.}\label{fig:1hautoencoder}
 \end{figure}
@@
 
-\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16}
+\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15}
 
 \bibliographystyle{ieeetr}
-\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18,bib/KeskarMNST16}
+\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18,bib/KeskarMNST16,bib/IoffeS15}
 

### 2019-06-21 MyCode/libs-and-global-consts.py

@@ -84,9 +84,9 @@ globalReps = 1000
 
 #%%
 # NN parameters
-title = 'Autoencoder1H'
+title = 'AutoencoderArray'
 timestr = time.strftime("%Y%m%d-%H%M%S")
-numEpochs = 2**17
+numEpochs = 2**16
 batchSize = 256
 train_p = 0.03
 

### 2019-06-21 Rapport/Article/Article.tex
Seance 06/21/2019

@@
 The batch size was maintained as the size of the first dimension of the training set and there was no reason for changing the optimizer.
-Notwithstanding, the ensemble of changes and optimizations were not enough to find a suitable architecture for channels with $p<0.2$ as the MAP algorithm for the 1/2 rate encoder would do better in this scenario.
+Notwithstanding, the ensemble of changes and optimizations were not enough to find a suitable architecture for channels with $p<0.02$ as the MAP algorithm for the 1/2 rate encoder would do better in this scenario.
 Table \ref{tab:arrayautoencoderTrain} summarizes the array autoencoder training parameters.
@@
 This section is divided in three parts.
-The first shows the results obtained to the two implemented decoders.
-Observations during the training of the DNN are made.
+The first shows the results obtained to the two decoders.
 Secondly, the results for the autoencoders is brought with some observations during the training period.
-The discussion and analysis for these two subsections is around the BER performance for the decoders and autoencoders.
-Finally, a decoding time performance analysis is explained and the results are shown and discussed.
+The discussion and analysis for these two subsections is around the BER performance.
+Finally, a decoding time performance analysis is done and the results are shown and discussed.
 
 \subsection{Decoders}
-BER Curves from both decoders, analyze why the number of epochs is so different, .
 
 We can infer from figures \ref{fig:ArrayD} and \ref{fig:1HD} that both the array decoder and the one-hot decoder are able to learn a decodification function.
-Moreover,  its performances replicate the MAP algorithm's decoding capability.
+Moreover,  its performances replicate the MAP algorithm's decoding capability for this specific encoding.
 Hence, DNN decoders could substitute the mathematically optimal MAP decoder for a stochastic channel, without compromising the bit error probability.
@@
 
+The MAP decoder and the DNN decoders were tested with the same number of messages and channels.
+As a result, since they express the same results, we can assume that the DNN decoder was able to learn the MAP algorithm during training phase.
+More interestingly, is the convergence period for each decoder.
+The one-hot decoder reached its peak performance much faster in comparison with the array decoder in terms of training epochs.
+I would attribute this increased performance to the activation function of the last layer.
+The sigmoid function used in the array decoder may lead to sharp damp gradients, gradient saturation or slow convergence during BP.
+While the one-hot decoder uses the softmax function which its partial derivatives have a better behavior during BP.
+Refer to \cite{2018arXiv181103378N} for deeper details in each activation function characteristics.
+
 \subsection{Autoencoders}
 BER Curves of simulation varying p, best choice of p, analyse autoencoder with best choice of p
+
+In the beginning we expected that the autoencoders would beat the MAP algorithm performance for a fixed linear block encoding.
+They were supposed to learn a non linear encoding and its particular MAP rule which would perform better than the studied MAP for a low crossover probability range.
+It turned out, however, since there are an enormous amount of possible non-linear codes, it is difficult to the learning algorithm to find the best one.
+Often during training, the DNN would find a local minimum which its cost function would be very close to zero, even though its performance during prediction phase was far from optimal.
+
+We observed during training, while experimenting with different architecture, that the array autoencoders would have a BER performance which looked like the no decoding curve, but translated in the y axis.
+This phenomenon can be observed in figure \ref{fig:parrayanalysis} where we tested several $p_t$ options to find the best one for implementing a more refined training.
+It is remarkable how the curves tend to replicate the same type of behavior.
+I believe it shows how the local minimums of the solution space of a NN are indeed close to each other, and the weight values may not be indeed that different.
 
@@
     \includegraphics[width=0.5\textwidth]{images/MAP_AutoencoderArray_Mep_65536_64_128_256_p_analysis}
-    \caption{CAPTION}\label{fig:parrayanalysis}
+    \caption{Training crossover probability simulation for the array autoencoder.
+$P_t=0.03$ demonstrated to have best performance to this particular architecture.
+The DNNs were trained with $2^{16}$ training epochs.}\label{fig:parrayanalysis}
 \end{figure}
@@
     \includegraphics[width=0.5\textwidth]{images/AutoencoderArray_Mep_65536_ptrain_003_logcosh}
-    \caption{Array autoencoder BER performance. DNN array autoencoder trained with a channel crossover probability error of $p_t=0.03$.}\label{fig:arrayautoencoder}
+    \caption{Array autoencoder BER performance.
+DNN array autoencoder trained with a channel crossover probability error of $p_t=0.03$.
+The DNN was trained with $2^{17}$ training epochs.}\label{fig:arrayautoencoder}
 \end{figure}
+
+However, it is noticeable the difference in error decoding in terms of a DNN trained with a $p_t=0$ and $p_t=0.3$ in Fig.
+\ref{fig:parrayanalysis}.
+Just similar to the array decoder, introducing a channel noise of some kind is beneficial to having a better error correction capability.
+Using noise also helps in the regularization of the DNN and approximates the training to the real application.
+
+We can observe from Fig.
+\ref{fig:arrayautoencoder} that for $p<0.02$, the autoencoder performance is actually worse than the tested MAP algorithm.
+Several hypothesis were considered for this comportment.
+In order to test them, we calculated the MAP rule for the encoder learned by the array autoencoder.
+The error correction yielded the same results as the DNN decoder learned by the autoencoder.
+Thus, the decoder replicates well the MAP for this specific encoding.
+Then, we experimented with training parameters and architecture, but could not find a configuration which would perform better than the traced MAP for low values of $p$.
+
+\ref{fig:1Hautoencoder}
 
@@
     \includegraphics[width=0.5\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
-    \caption{CAPTION.}\label{fig:1hautoencoder}
+    \caption{One-hot autoencdoer BER performance. Trained without a noise channel.}\label{fig:1Hautoencoder}
 \end{figure}
@@
 
-In order to analyze the delay in decoding for each technique studied, a monte carlo simulation was conducted.
+In order to analyze the delay in the communication for each technique studied, a monte carlo simulation was conducted.
 There were selected 20 different BSC channels.
-For each channel, $100000$ messages were transmitted and decodified for each decoder.
+For each channel, $100000$ messages were codified, transmitted and decodified for each decoder, autoencoder and MAP.
 This process was timed and repeated several times.
@@
 
-say that only the decoding part was measured  for autoencoders
+Both decoders using the linear encoding performed around 25\% faster than the MAP rule.
+This is due their fairly low dimensional complexity and one shot feature.
+While the MAP rule must calculate the distance between the received codeword to every possible codeword, the decoders once trained are able to decode the received code word into the message of highest probability with simple matrix multiplications.
+The latter efficiency is potentialized when using decoding computers equipped with GPUs.
+
+In spite of the one-hot decoder has considerably fewer parameters than the array decoder, it must calculate a $argmax$ function to deliver the actual message that was transmitted.
+This extra costly computation is also the reason for the increased time for the one-hot autoencoder.
+It must first encode the message, then transmit it and finally decode it, performing two times per message a $argmax$ operation.
 
@@
     $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ \\ \hline
-    \multicolumn{2}{c|}{Array. Autoencoder}  & One-hot Autoencoder\\
+    \multicolumn{2}{c|}{Array Autoencoder}  & One-hot Autoencoder\\
 
-     \multicolumn{2}{c|}{$0.00 \pm 0.00 $} & $0.00 \pm 0.00$ \\ 
+     \multicolumn{2}{c|}{$1.33 \pm 0.05 $} & $3.02 \pm 0.06$ \\ 
     \thickhline
@@
 
-\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15}
+\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15,2018arXiv181103378N}
 
 \bibliographystyle{ieeetr}
-\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18,bib/KeskarMNST16,bib/IoffeS15}
+\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18,bib/KeskarMNST16,bib/IoffeS15,bib/ActivationF}
 

### 2019-06-21 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/21/2019

@@
 \subsection{Design \& Architecture}
-\begin{frame}[allowframebreaks]{Neural Network's Architecture}{Show the architecture used for each case and remarks some important parameters}
-  \begin{itemize}
-  \item {
+\begin{frame}[allowframebreaks]{Neural Network's Design and Architecture}{Show the architecture used for each case and remarks some important parameters}
+
+	\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.4\textwidth]{images/MLNN}
+    \caption{MLNN representative diagram, where $\textbf{y}^n$ is the input vector, $\textbf{r}_{l}^{j}$ is a hidden layer vector
+and $\hat{\textbf{u}}^{k}$ is the output vector.} \label{fig:NN}
+\end{figure}
+
+	\framebreak
+
     \begin{table}
@@
 \end{table}
-  }
-  \item {
+
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Binary cross-entropy & Adam & $2^{16}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:arraydecoderTrain}
+\end{center}
+
+  \framebreak
+
     \begin{table}
@@
 
-  }
-  \item {
-    \begin{table}
-\caption{DNN array autoencoder architecture and parameters.} 
-\begin{center}
-
-  \begin{tabular}{ l  |  l }
-    \thickhline
-    \multirow{3}{*}{Encoder} & Dense: X, activation: ReLU, input size: $k$ \\
- & Dense: X, activation: ReLU  \\
- & Dense: $n$, activation: Sigmoid  \\ \hline
-
-    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $n$ \\
- & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
-
- \multirow{2}{*}{Decoder} & Dense: X, input size: $n$ \\
- & Dense: X, activation: ReLU  \\
- & Dense: $k$, activation: Sigmoid  \\ \hline
-
-    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
-    \thickhline
-  \end{tabular}
-    \label{tab:arrayautoencoder}
-\end{center}
-\end{table}
-
-  }
-  \end{itemize}
-\end{frame}
-
-\subsection{Training}
-\begin{frame}{Neural Network's Training}{Show the best training parameters for each structure}
-
-    \begin{table}
-\caption{DNN array decoder training parameters.} 
-\begin{center}
-\begin{tabular}{ l  |  l | l | l  }
-    \thickhline
- Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
- Binary cross-entropy & Adam & $2^{16}$ & $256$ \\
-    \thickhline
-  \end{tabular}
-    \label{tab:arraydecoderTrain}
-\end{center}
-\end{table}
-
-    \begin{table}
-\caption{DNN one-hot decoder training parameters.} 
 \begin{center}
@@
 \end{center}
+
+	\framebreak
+
+	\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.8\textwidth]{images/DNN_autoencoder}
+    \caption{Representation of a DNN autoencoder composed of dense layers.}\label{fig:DDNNAutoencoder}
+\end{figure}
+
+	\framebreak
+
+    \begin{table}
+\caption{DNN array autoencoder architecture.} 
+\begin{center}
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{3}{*}{Encoder} & Dense: 512, activation: ReLU, BN\footnotemark, input size: 8 \\
+ & Dense: 256, activation: ReLU, BN \\
+ & Dense: 16, activation: Sigmoid  \\ \hline
+    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $16$ \\
+ & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
+ \multirow{2}{*}{Decoder} & Dense: 128, BN, input size: 16 \\
+ & Dense: 64, activation: ReLU, BN  \\
+ & Dense: 8, activation: Sigmoid  \\ \hline
+    \multicolumn{2}{l}{\textbf{Total parameters: }154072}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:arrayautoencoder}
+\end{center}
 \end{table}
 
-\begin{table}
-\caption{DNN array autodecoder training parameters.} 
 \begin{center}
@@
 \end{center}
+
+\footnotetext{Batch Normalization (BN)}
+
+	\framebreak
+
+    \begin{table}
+\caption{DNN one-hot autoencoder architecture.} 
+\begin{center}
+
+  \begin{tabular}{ l  |  l }
+    \thickhline
+    \multirow{3}{*}{Encoder} & Dense: 196, activation: ReLU, BN, input size: 256\\
+ & Dense: 128, activation: ReLU, BN  \\
+ & Dense: 96, activation: ReLU, BN \\
+ & Dense: 64, activation: ReLU, BN  \\
+ & Dense: 32, activation: ReLU, BN  \\
+ & Dense: 16, activation: Sigmoid  \\ \hline
+   Channel & Lambda: $Round(\textbf{x})$, input size: $16$ \\ \hline
+ Decoder & Dense: 128, activation: ReLU, BN, input size: 16 \\
+ & Dense: 256, activation: Softmax  \\ \hline
+    \multicolumn{2}{l}{\textbf{Total parameters: }134052}\\
+    \thickhline
+  \end{tabular}
+    \label{tab:onehotautoencoder}
+\end{center}
 \end{table}
+
+\begin{center}
+\begin{tabular}{ l  |  l | l | l  }
+    \thickhline
+ Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ MSE & Adam & $2^{16}$ & $256$ \\
+    \thickhline
+  \end{tabular}
+    \label{tab:onehotautoencoderTrain}
+\end{center}
 
@@
 \subsection{DNN Autoencoders}
-\begin{frame}{DNN Autoencoder}{Show the results for the autoencoder in terms of train p, Mep, Parameters, etc}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+\begin{frame}[allowframebreaks]{DNN Array Autoencoder}{Show the results for the autoencoder in terms of train p, Mep, Parameters, etc}
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.8\textwidth]{images/MAP_AutoencoderArray_Mep_65536_64_128_256_p_analysis}
+    \caption{Training crossover probability simulation for the array autoencoder. $P_t=0.3$ demonstrated to have best performance to this particular architecture.}\label{fig:parrayanalysis}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.8\textwidth]{images/AutoencoderArray_Mep_65536_ptrain_003_logcosh}
+    \caption{Array autoencoder BER performance. DNN array autoencoder trained with a channel crossover probability error of $p_t=0.03$.}\label{fig:arrayautoencoder}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}{DNN One-hot Autoencoder}{Show the results for the autoencoder in terms of train p, Mep, Parameters, etc}
+
+\begin{figure}[!ht]
+  \centering
+    \includegraphics[width=0.8\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
+    \caption{One-hot autoencdoer BER performance. Trained without a noise channel.}\label{fig:1hautoencoder}
+\end{figure}
+
 \end{frame}
@@
   \begin{table}
-\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the mean MAP algorithm decoding time. } 
+\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. } 
 \begin{center}
 
-  \begin{tabular}{ l | l | l | l | l }
+  \begin{tabular}{ c | c | c}
     \thickhline
-    MAP & A. Dec. & One-hot Dec. & A. Auto. & One-hot Auto. \\ \hline
-    $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ & $ \pm $ & $ \pm $ \\ 
+    MAP & Array Decoder & One-hot Decoder\\
+    $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ \\ \hline
+    \multicolumn{2}{c|}{Array Autoencoder}  & One-hot Autoencoder\\
+
+     \multicolumn{2}{c|}{$1.33 \pm 0.05 $} & $3.02 \pm 0.06$ \\ 
     \thickhline
@@
 \end{table}
+
 \end{frame}
@@
 
- \cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}
+\color{white}
+\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15}
 
 \bibliographystyle{ieeetr}
-  \bibliography{../../Article/bib/ShannonCE1948,../../Article/bib/CalabreseWGPS16,../../Article/bib/OSheaH17,../../Article/bib/Oshea2,../../Article/bib/Gruber,../../Article/bib/Worm,../../Article/bib/Viterbi,../../Article/bib/RobertsonHV97,../../Article/bib/Ibnkahla,../../Article/bib/Nielsen,../../Article/bib/Murphy,../../Article/bib/AbadiABBCCCDDDG16,../../Article/bib/Keras,../../Article/bib/mit_neco18_1527,../../Article/bib/BenammarP18}
+\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18,bib/KeskarMNST16,bib/IoffeS15}
 

### 2019-06-21 MyCode/libs-and-global-consts.py

@@ -84,11 +84,11 @@ globalReps = 1000
 
 #%%
 # NN parameters
-title = 'AutoencoderArray'
+title = 'Autoencoder1H'
 timestr = time.strftime("%Y%m%d-%H%M%S")
 numEpochs = 2**16
 batchSize = 256
-train_p = 0.03
+train_p = 0.0
 
 path = 'Trained_'+title+'/'+timestr+'_'+title+'_Mep_'+str(numEpochs)+'_bs_'+str(batchSize)+'.h5'
 checkpointPath = 'Checkpoints/'+title+'/'+timestr+'_'+title+'_Mep_{epoch:02d}-{loss:.8f}.h5'

### 2019-06-21 Rapport/Article/Article.tex
Seance 06/22/2019

@@
 \author{
-\IEEEauthorblockN{Eduardo Dadalto Camara Gomes\IEEEauthorrefmark{1},
-Meryem Benammar\IEEEauthorrefmark{2}}
+\IEEEauthorblockN{Eduardo Dadalto Camara Gomes\IEEEauthorrefmark{1}}
+
+%%Meryem Benammar\IEEEauthorrefmark{2}}
 
@@
 }
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
-Email: meryem.benammar@isae-supaero.fr
-}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
+%Email: meryem.benammar@isae-supaero.fr
+%}
 }
 
-\IEEEspecialpapernotice{(Final report)}
+%\IEEEspecialpapernotice{(Final report)}
 
@@
 The trained DNN autoencoder has a \textit{one-shot} capability which outperforms a MAP decoder for a given encoder in terms of delay.
-ONE PHRASE FOR RESULTS
+The decoders proved to be around 25\% faster than the MAP and they achieved the same BER performance.
+There are still opportunities for enhancement for the autoencoders.
 
@@
 
-We observed during training, while experimenting with different architecture, that the array autoencoders would have a BER performance which looked like the no decoding curve, but translated in the y axis.
+We observed during training, while experimenting with different architecture, that the array autoencoders would have a BER performance which looked like the no decoding curve, but translated in the $y$ axis.
 This phenomenon can be observed in figure \ref{fig:parrayanalysis} where we tested several $p_t$ options to find the best one for implementing a more refined training.
@@
 
-\ref{fig:1Hautoencoder}
+The one-hot autoencoder could not outperform the MAP traced for any value of $p$.
+As observed in Fig.
+\ref{fig:1Hautoencoder}, even though it learned a similar encoding, the resulting BER estimation is not optimal.
+Equally to the array autoencoder, we tested a MAP rule for the learned encoder and found that the DNN decoder portion performs just like the MAP.
+Thus, once again the problem is in the learned encoding.
+We tried 80 different configurations, though we could not find one which could outperform the reference model.
+We can also highlight the trace of the BER curve, approximating the reference model one.
+Other one-hot autoencoders showed similar curves but translated up in the $y$ axis.
 
@@
     \includegraphics[width=0.5\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
-    \caption{One-hot autoencdoer BER performance. Trained without a noise channel.}\label{fig:1Hautoencoder}
+    \caption{One-hot autoencdoer BER performance. Trained without a noisy channel. The DNN was trained with $2^{16}$ training epochs.}\label{fig:1Hautoencoder}
 \end{figure}
@@
 
-In order to analyze the delay in the communication for each technique studied, a monte carlo simulation was conducted.
+In order to analyze the delay in the communication for each technique studied, a Monte Carlo simulation was conducted.
 There were selected 20 different BSC channels.
@@
 \begin{table}
-\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. } 
+\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders.
+The data is normalized to the average MAP algorithm decoding time.
+The simulation was based on CPU calculations.
+}
 \begin{center}
@@
 
+paragraph that talks about CPU, not GPU.
+
 \section{Conclusions}
 
-Talk about feasibility of decoders, improved performance of the channel delay with DNN decoders, challenges in implementing this architecture in real life, challenges in training the NNs, Future work, what else could be done with NN in error correction, Advantages and disadvantages in comparison with linear decoding.
+The feasibility of a machine learning based channel decoder and the end-to-end application of a deep neural network autoencoder were demonstrated for a communication system with a binary symmetric channel.
+Their performances were assessed and analyzed in terms of bit error rate performance and communication delay.
+A few hyper-parametric analysis were conducted to find the best solution for each technique.
+
+The designed decoders improved around 25\% the channel delay when compared to the maximum a posteriori rule for a $(n,k)$ code with $\textbf{G}$ aforementioned as linear encoding matrix.
+Their small number of parameters aligned with a GPU equipped decoding computer is promissory to be a substitute of the MAP rule and achieve same bit error correction rate.
+In addition, their training and architecture are straightforward and can be trained with a commercial computer.
+Once trained, they work as a non-linear decoder to a specific channel encoding.
+
+We also proved that autoencoders can learn a non linear encoding function and its decoding function for a binary symmetric channel.
+Thus there is no need for a generator matrix for error correction.
+There are still some challenges in training to solve, such as ameliorate the architecture so it can outperform the reference model in every channel condition.
+In addition, because of the number of parameters, their calculation in CPU is slow and they showed a greater delay when compared to the MAP rule.
+Though, they must take the most advantage of GPU based calculations and the delay time could be improved using this technology.
+
+A more rigorous hyper-parametric analysis could find a combination of parameters which yields better BER performance for the autoencoders and is strongly suggested for future work.
+In addition, in the future, a real implementation of the DNN decoders could be tested to confirm their capabilities as it is not the main focus of this work.
+We suggest using more advanced NN topology such as a recurrent neural network (RNN) or a generative adversarial network (GAN).
+Their implementation and training are more complex, however they have recently demonstrated outstanding results in other fields.
+
+%Talk about feasibility of decoders, improved performance of the channel delay with DNN decoders, challenges in implementing this architecture in real life, challenges in training the NNs, Future work, what else could be done with NN in error correction, Advantages and disadvantages in comparison with linear decoding.
 
 \section*{Acknowledgment}
-The authors would like to thank...
 
-\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15,2018arXiv181103378N}
+I would like to thank the guidance and advice from my supervisor Meryem Benammar, Ph.D, my colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research and Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project.
+
+%\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15,2018arXiv181103378N}
 

### 2019-06-21 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/22/2019

@@
 
-\pgfdeclareimage[height=0.5cm]{university-logo}{images/ISAE}
-\logo{\pgfuseimage{university-logo}}
+%\pgfdeclareimage[height=0.5cm]{university-logo}{images/ISAE}
+%\logo{\pgfuseimage{university-logo}}
 
@@
   \centering
-    \includegraphics[width=0.8\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
-    \caption{One-hot autoencdoer BER performance. Trained without a noise channel.}\label{fig:1hautoencoder}
+    \includegraphics[width=0.75\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
+    \caption{One-hot autoencdoer BER performance. Trained without a noisy channel.}\label{fig:1hautoencoder}
 \end{figure}
@@
 \subsection{Time Analysis}
-\begin{frame}{Delay Time Analysis}{Comparison of decoding time for each method.}
+\begin{frame}{Delay Time Analysis}{Comparison of encoding, transmission and decoding time for each method.}
   \begin{table}
@@
 \begin{frame}{Acknowledgment}
+  I would like to thank ...
   \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
+   \item{The guidance and advice from my supervisor Meryem Benammar, Ph.D;}
+  \item{ My colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research;}
+   \item{Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project; and }
+   \item{The supervisors, the LACS member and the audience present today.}
   \end{itemize}
+
 \end{frame}
 
-% All of the following is optional and typically not needed. 
 \appendix

### 2019-06-22 Rapport/Article/Article.tex
Seance 06/22/2019

@@
 
-Communication channel error correction is key to enable digital critical communication systems to work efficiently.
+Communication channel bit error correction is key to enable digital  communication systems to work efficiently.
 The maximum a posteriori (MAP) decoder is proven to be the mathematically optimal solution for the problem.
@@
 The decoders proved to be around 25\% faster than the MAP and they achieved the same BER performance.
-There are still opportunities for enhancement for the autoencoders.
+There are still opportunities for enhancement of the autoencoders.
 
@@
 \ref{fig:cs} illustrates.
-As a result of channel imperfections, Shannon theorized in \cite{Shannon:2001:MTC:584091.584093} the ultimate reliable data rate that can be transmitted through a communication system with an arbitrarily small probability error.
+As a result of channel imperfections, Shannon theorized in \cite{Shannon:2001:MTC:584091.584093} the ultimate reliable data rate which can be transmitted through a communication system with an arbitrarily small probability error.
 
@@
 Since then, the research community in digital communication developed a series of algorithms to minimize error probability over a channel.
-In practice, the bit error rate (BER), an approximation of the probability of error, is targeted.
+In practice, the bit error rate (BER), an empirical approximation of the probability of error, is targeted.
 However, the challenge of finding an \textit{efficient} solution  i.e., with both low latency and low error probability, for low signal to noise ratio (SNR) channels, remains.
-Low latency and high bandwidth wireless communication are key to critical systems, such as airplanes, satellites, cellular communication and 5G operations.
+Low latency and high bandwidth wireless communication are key to critical systems functioning, such as airplanes, satellites, cellular communication and 5G operations.
 The latter was studied by F.
@@
 Moreover, with ML based channel encoder and channel decoder, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate end-to-end in the system.
-As a result, an optimized autoencoder with a stochastic layer that models channel's imperfections can substitute the block based representation.
+As a result, an optimized autoencoder with a stochastic layer which models channel's imperfections can substitute the block based representation.
 Therefore, ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths and with low decoding latency, resulting ultimately in gain of bandwidth over standard methods.
@@
 
-In this work, we will implement a ML autoencoder for a  BSC that performs similar to the MAP decoder in real applications for a range of SNR from $-10 \text{dB}$ to $10 \text{dB}$.
+In this work, we will implement a ML autoencoder for a BSC which performs similar to the MAP decoder in real applications for a range of cross over probability $p$ from $0$ to $0.1$.
 In its most simple form, a channel autoencoder includes an encoder, a noisy channel and a decoder.
@@
 
+The main theoretical concepts will be briefly introduced. They are organized by theme and is aimed to a reader less familiar with the particular subject.
+
 \subsection{Channel coding}
@@
 \ref{fig:cs}.
-The left portion of the chain, the transmitter, wants to communicate a message $\textbf{u}$ through a noisy channel.
-The right portion of the chain, the receiver, may interpret this possible corrupted stream of bits into the original message.
-Usually this prediction $\hat{\textbf{u}}$ carry errors which we aim to minimize its probability of occurring.
+The left portion of the chain, the transmitter, communicates a message $\textbf{u}$ through a noisy channel.
+The right portion of the chain, the receiver, aims to interpret this possibly corrupted stream of bits into the original message.
+Usually this prediction $\hat{\textbf{u}}$ carry errors which we want to minimize their probability of occurring.
 In this scenario, the theory of channel coding discuss possible solutions.
@@
 Briefly, a linear encoder will index a message $\textbf{u}^k \in \mathcal{U} \subseteq \{0,1\}^k$ to a code word $\textbf{x}^n \in \mathcal{X} \subseteq \{0,1\}^n$.
-This mapping is done by the \textit{generator matrix} $ \textbf{G} $.
+This mapping is done by the \textit{generator matrix} $ \textbf{G} $ through equation \ref{eq:encoding}.
+
 \begin{equation}\label{eq:encoding}
@@
 
-The objective of the receiver is to estimate the original message. The empirical error probability, or a BER, in this process is defined as
+The objective of the receiver is to estimate the original message. The empirical error probability, or BER, in this process is defined as
 
@@
 
-The Hamming distance $d_H:\mathcal{X}\mapsto \mathbb{Z}$ is a metric in the word subset which in the context of linear coding is equal to equation \ref{eq:dh}.
-Noteworthy for the MAP rule, a non zero Hamming distance indicates that there is a block error, and ultimately at least a binary error in the received code word.
+The Hamming distance $d_H:\mathcal{X}\mapsto \mathbb{Z}$ is a metric in the word set which in the context of linear coding is equal to equation \ref{eq:dh}.
+Noteworthy for the MAP rule, a non zero Hamming distance indicates that there is a block error, and ultimately at least one binary error in the received code word.
+
 \begin{equation}\label{eq:dh}
@@
 
-Briefly, a BSC is a noisy channel with a binary input $X\in \{0,1\}$ on the left and a binary output $Y\in \{0,1\}$ on the right.
+In short, a BSC is a noisy channel with a binary input $X\in \{0,1\}$ on the left and a binary output $Y\in \{0,1\}$ on the right.
 They are linked through a stochastic model characterized by a crossover probability of value $p$ as illustrated in Fig.
@@
 
-The concept of a MAP decoder algorithm for sequences, is choosing a message which maximizes the a posterior probability of the corrupted code word $y$ received by the decoder \cite{Worm00turbo-decodingwithout}.
+The concept of a MAP decoder algorithm for sequences is choosing a message which maximizes the a posterior probability of the corrupted code word $y$ received by the decoder \cite{Worm00turbo-decodingwithout}.
 This algorithm is known for its optimal error correction capability for white noise interference.
@@
 While the Max-Log-MAP is a suboptimal decoder and will not be treated in this paper.
-these variants are widely used as decoders for turbo codes \cite{HagenauerJ}.
+These variants are widely used as decoders for turbo codes \cite{HagenauerJ}.
 Refer to \cite{journals/ett/RobertsonHV97} for specific details in the derivation of the Log-MAP algorithm.
@@
 It means that if the real SNR is smaller, the decoder will not perform well.
-Hence, for channel characteristics that change over time, the application of a MAP algorithm is debatable.
-This conclusion serve as motivation for ML autoencoders, which could adapt to different SNR conditions, outperforming the MAP decoder for realistic applications.
 
@@
 Neural networks (NN) are a set of connected \textit{neurons} able to approximate any kind of function through an architecture composed of several single processing units (neurons) connected together forming a network.
-They are defined by \cite{Ibnkahla} as parallel distributed, learning- and self-organizing information processing systems.
+They are defined in \cite{Ibnkahla} as parallel distributed, learning- and self-organizing information processing systems.
 
-In the context of learning algorithms implemented by a NN, the key issue is to find a suitable architecture that delivers the best results.
+In the context of learning algorithms implemented by a NN, the key issue is to find a suitable architecture which delivers the best results.
 To define and build this infrastructure, the activation function, the loss function and the structure must be decided.
@@
 
-% TODO: WRITE WHICH ACTIVATION FUNCTION I WILL USE
-%
 Two common activation functions are the rectified linear unit (ReLU) and the sigmoid function.
@@
 The objective is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{DBLP:journals/corr/OSheaH17}.
+
 \begin{equation}\label{eq:o-c-e}
@@
 
-One of the greatest advantage of NN is its ability to generalize the training set, delivering the right results even for input data not contained in the training set.
+One of the greatest advantage of NNs is their ability to generalize the training set, delivering the right results even for input data not contained in the training set.
 With the weight matrix trained, the NN finds the correct output values for unseen inputs.
@@
 The statistics of the channel's noise is shown in \cite{2017arXiv171008379G} to be irrelevant, since the NN extract it during training phase.
-The autoencoder applicability is compelling because the weights computed during training can be separated in two groups, one correspondent to the nodes before the channel and one posterior to the channel.
+The autoencoder applicability is compelling because the weights computed during training can be separated in two groups.
+One corresponding to the nodes before the channel and one to the nodes posterior to the channel.
 As a consequence, we are able to design a non linear encoder and its correspondent non linear decoder with ease.
@@
 
-Whilst autoencoders are more complex dimensional wise compared to NN decoders, the readily access to graphical processing units (GPUs) contributed to the feasibility of this technique, tackling the \textit{curse of dimensionality} challenge in ML.
+Whilst autoencoders are more complex dimensional wise compared to NN decoders, the readily access to graphical processing units (GPUs) contributes to the feasibility of this technique, tackling the \textit{curse of dimensionality} challenge in ML.
 This type of hardware made possible the approach of the layer-by-layer supervised training by stochastic gradient descent optimization applied in this work.
@@
 For organizational purposes, the section is divided into four subsections.
-First, the predictions and error corrections statistical relevance are justified through the Monte Carlo simulations conducted.
+First, prediction and error correction statistical relevance are justified through the Monte Carlo simulations conducted.
 Second, the MAP algorithm for a linear code is shown.
-Then the experiments relating to the ML based decoders are assessed.
-Finally the implementation of the ML based autoencoder is explained.
+Then, the experiments relating to the ML based decoders are assessed.
+Finally, the implementation of the ML based autoencoder is explained.
 
@@
 This stochastic computational algorithm uses pseudo-random sampling to solve different problems.
-Hence, we can work with a wide variety of transmission scenarios.
+Hence, we can work with a wide variety of transmission scenarios of the communication system.
 
-Applied to the context of this research, the messages are repeatedly randomly generated for a specific BSC of crossover probability $p$.
-They are coded by a linear block code or by a DNN encoder and then transmitted through the channel.
+Applied to the context of this research, the messages are repeatedly randomly generated.
+They are coded by a generator matrix or by a DNN encoder and then transmitted through the BSC channel of crossover probability $p$.
 Their information arrives corrupted by noise in the decoder and is subsequently partially corrected by a decoder.
-Finally the average BER is calculated and plotted.
-The interval of the horizontal axis is in between $0$ and $0.10$.
+Finally, the average BER is calculated and plotted.
+The interval of values of $p$ is in between $0$ and $0.10$.
 
@@
 
-The MAP algorithm as cited previously, has a high time complexity $(\mathcal{O}(2^k))$, since it computes the Hamming distance between the received code word and every code word from the code book.
+Although the MAP algorithm is the mathematically optimal decoder as cited previously, it has a high time complexity $(\mathcal{O}(2^k))$, since it computes the Hamming distance between the received code word and every code word from the code book.
 For more details, the pseudo code of the algorithm is available below.
@@
 The linear encoding matrix $\textbf{G}_{8\times16}$ was used for the encoding of the messages.
-Its dimension is equal to the linear block code name represented by the double $(n,k)=(16,8)$.
+It's dimension is equal to the linear block code name represented by the ordered double $(n,k)=(16,8)$.
 Thus, the \textit{rate} of the code is equal to $0.5$.
-The linear block name and = rate is constant throughout the work.
+The linear block name and rate is constant throughout the work.
 
@@
 
-This section is divided in two for better identify the differences between the implementation of the array DNN decoder and the one-hot DNN decoder.
+This section is divided in two for better identification of the differences between the implementation of the array DNN decoder and the one-hot DNN decoder.
 
@@
 
-The architecture of the DNN necessary to achieve a great performance in terms of BER is not evident.
-To begin we must define which layers were used and why.
+The architecture of the DNN decoder necessary to achieve a great performance in terms of BER is not evident.
+To begin, we must define which layers were used and why.
 A \textit{Lambda} layer is a layer in which the activation function is user defined and the parameters are non-trainable.
@@
 The fact that the parameters are non-trainable means that the noise layer is applied in forward propagation only.
+
 A Dense layer is a densely connected layer, which means that each neuron receives input from all the neurons in the previous layer.
@@
 
-The choice of the crossover probability of the training channel, $p_{t}$, was based in the work of \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the best in terms of decoding precision after the DNN is trained.
+The choice of the crossover probability of the training channel, $p_{t}$, was based in the work of \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the optimal in terms of decoding precision after the DNN is trained.
 In addition, a deeper DNN showed best results in \cite{DBLP:journals/corr/OSheaH17} where 3 hidden layers were used.
@@
 
-The training set for the array decoder contained all the possible messages as training labels and its correspondents code words as training data.
+The training set contained all the possible messages as training labels and its code words as training data.
 Because of the chosen encoding, the training data is a $256\times16$ matrix and the training label is a $256\times8$ matrix.
-To generate a possible message, a stochastic algorithm was developed to fabricate random 8-length vectors of integers 1s and 0s which the average value considering all elements was around $0.5$.
-In this way we could guarantee that our training environment were representative of a real life application.
+To generate a message, a stochastic algorithm was developed to fabricate random 8-length vectors of integers 1s and 0s which the average value considering all elements was around $0.5$.
+In this way, we could guarantee that our training environment were representative of a real life application.
 Where different bits are transmitted with the same probability.
@@
 Even though it is not recommended to train a NN with all the available data - because it might lead to over-fitting -, for this case does not seems to be a problem.
-WHY?????
-(noisy channel?
-onehot decodeR?)
+Apparently, the presence of a BSC is enough to regularize the training procedure and prevent over-fitting.
 
-The choices of training parameters is a mix between what other works in literature uses and a hyper parametric analytic approach, where several trials lead to the decision.
+The choice of training parameters is a mix between what is standardized in literature and a hyper parametric analytic approach, where several trials lead to a decision.
 First, the binary cross-entropy as loss function is widely used in literature and showed to converge to the desired result consistently during tests.
 For the optimizer, the \textit{Adam} built-in Keras optimizer was chosen.
-Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}.
+The Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}.
 Next, a batch size matching the size of the training set which is sufficiently small to still operate in the regime of small batch was chosen.
@@
 The architecture of the one-hot decoder is much simpler when compared to the array decoder.
-It contains only one layer of $256$ units which maps the input of size 16 to the $2^8=256$ possible messages.
+It contains only one hidden layer of $256$ units which maps the input of size 16 to the $2^8=256$ possible messages.
 Hence, this characterizes a multi-class classification problem, meaning that the chosen activation function is the Softmax (Eq.
 \ref{eq:soft}).
-This training technique is known in the literature as \textit{one-hot}.
+This training technique is known in the literature as \textit{one-hot} training.
 
-The denomination one-hot comes from the output vector which is usually a long vector of zeros with one element equals to one.
-This \textit{hot} vector element indicates which class the NN has classified the input.
+The denomination one-hot comes from the output vector which is usually a long vector of elements close to zeros with one element close to one.
+This \textit{hot} vector element indicates which class the NN has classified the input with the most probability.
 In this context, which message encoded in 256 different classes the code word refers to.
-Basically, a message is mapped into a class and a class only.
+Basically, a message is mapped into a class and a class only and the decoder determines this class based on a code word input.
 
-A straight forward consequence of this shallower NN is the number of training parameters almost three times less than the array decoder.
+A straight forward consequence of this shallower NN is the number of training parameters.
+Almost three times less than the array decoder.
 Because of that, we could expect a lower training time.
@@
 Consequently, an additional operation is needed.
-To transform from a one-hot vector to a message, a ordered collection of the messages was stored in memory and the correspondent message is determined by a $argmax$ operation in the one-hot vector, which identifies which message is the most likely to have been transmitted through the channel.
+To transform from a one-hot vector to a message, a ordered collection of the messages was stored in memory.
+The target message is determined by a $argmax$ operation in the one-hot vector, which identifies which message is the most likely to have been transmitted through the channel.
 
-The training set in this case differs from the array decoder.
-The training labels are one-hot vectors which maps into real messages through a bijective function.
+The training labels in this case differs from the array decoder.
+They are one-hot vectors which maps into real messages through a bijective function.
 Thus, it is a matrix of dimension $256\times256$ composed of 256 different on-hot vectors.
@@
 
-Even though DNN encoders are useful, an encoding matrix is still needed.
-With a view to eliminate completely the need of an explicit encoding, a technique commonly used method in ML for image recognition is explored.
-The autoencoder is able to learn not only a decoding algorithm, but also a decoding one as introduced in the following section.
+Even though DNN decoders are useful, an encoding matrix is still needed.
+Aiming to eliminate completely the need of an explicit encoding, a technique commonly used in ML for image recognition is explored.
+The autoencoder is able to learn not only a decoding algorithm, but also a encoding one as introduced in the following section.
 
@@
 For the DNN autoencoder, two different concepts were studied.
-The array autoencoder takes as input the messages and it tries to replicate them, encoding it in a 16 bit code word, transmitting through a noisy channel and decoding it in the second portion of the NN.
-A one-hot autoencoder is inputted by a one-hot vector representing one unique k-bit message.
-It encodes the vector into a n-bit long code word, transmits it through a noisy channel and decode it in the second part of the NN.
+The array autoencoder takes as input the messages and it tries to replicate them.
+It encodea the messages into a 16 bit code word, transmits the code word through a noisy channel and decodes them in the second portion of the DNN.
+A one-hot autoencoder is inputted by a one-hot vector representing one unique $k$-bit message.
+It encodes the vector into a $n$-bit long code word, transmits them through a noisy channel and decode them in the second part of the DNN.
 
@@
 Once trained, they could perform optimally in terms of encoding and decoding for a channel.
-To be clear during the discussion of this and the following sessions, the autoencoder will be divided in: the encoder, the channel and the decoder portions as shown in Fig.
-\ref{fig:DDNNAutoencoder}.
-In total, there were tested 71 array and 80 one-hot autoencoders different architecture or training configuration.
+
+To be clear in the discussion of this and the following sessions, the autoencoder will be divided in three: the encoder, the channel and the decoder portions.
+Fig.
+\ref{fig:DDNNAutoencoder} illustrates this division.
+Even though the autoencoder could in theory be able to achieve at least the same performance as the MAP, finding the parameters which unveils its capabilities requires multiple testing and a rigorous hyper-parametric analysis.
+In total, there were tested 71 array and 80 one-hot autoencoders different architectures or training configurations.
 
@@
 
-Even though in theory the autoencoder could be able to achieve at least the same MAP performance, finding the parameters which unveils its capabilities requires multiple testing and a rigorous hyper-parametric analysis.
 For the array autoencoder, we first experimented with the depth of the encoder.
-The number of hidden layers in a DNN has influence in the BP algorithm.
+The number of hidden layers in a DNN influences the BP algorithm performance.
 Extremely deep NN, with more than 4 hidden layers, may encounter problems in the gradient calculations due to vanishing weights.
@@
 We trained the encoder to replicate the linear encoding proposed in the preceding sections as we studied the effects of each parameter in the training performance.
-A more wide architecture showed faster convergence and the use of three hidden layers had no positive effects.
-In the end, for the integrated model we opted for using two hidden layers, with 512 and 256 densely connected units as detailed in table \ref{tab:arrayautoencoder}.
+A more wide architecture showed faster convergence and the use of three or more hidden layers had no positive effects.
+In the end, for the integrated model, we opted for using two hidden layers, with 512 and 256 densely connected units as detailed in table \ref{tab:arrayautoencoder}.
 The activation functions remained the same as in the DNN decoders discussed in the previous section.
@@
 The decoder architecture was also based on the standalone array decoder.
-In spite of this, there were no difference outstanding in decoding with or without the last hidden layer of the decoder.
+In spite of this, there were no outstanding difference in decoding with or without the last hidden layer of the decoder.
 In the other hand, the tests with a condensate NN with one hidden layer demanded a greater training time to arrive in the same results.
+Hence, we dropped one hidden laeyer from the cedoder and kept two
+
 The final test was done with the entire autoencoder and the final tweaks were done.
-
-While training the autoencoder we noticed that the binary cross-entropy loss function would need more training epochs than the MSE and \textit{logcosh} loss functions, even though they converged in similar results.
+While training the autoencoder, we noticed that the binary cross-entropy loss function would need more training epochs than the MSE and \textit{logcosh} loss functions, even though they converged in similar results.
 The batch size was maintained as the size of the first dimension of the training set and there was no reason for changing the optimizer.
@@
 Furthermore, adding batch normalization layers inside the DNN was fundamental.
-With greater training parameters, achieving convergence in a reasonable time requires more computing power without any specific technique.
-The constant changes of weight values in the intermediate layers in every training step is the principal contributor to training cost.
+With greater training parameters, achieving convergence in a reasonable time requires more computing power of no specific technique is implemented.
+The frequent changes of weight values in the intermediate layers during each training step   is the principal contributor to the training cost.
 This problem is known as \textit{internal covariance shift}.
 What batch normalization does is normalize the input of a layer by its batch statistics and then scale and shift it.
-For further details, refer to \cite{DBLP:journals/corr/IoffeS15}.
+For further details, refer to \cite{DBLP:journals/corr/IoffeS15} which explain their usability in demanding training scenarios.
 
-The final parameter to determine is the training channel cross over probability.
-Once the architecture and training parameters were defined, a simulation for the same DNN with different $p_t$ was done and the trained models were analyzed in terms of BER.
-A $p_t=0.03$ yielded best performance.
+The final parameter to set is the training channel cross over probability $p_t$.
+Once the architecture and training parameters were defined, a simulation for the same DNN with different $p_t$ was done.
+The trained models were analyzed in terms of BER and these with $p_t=0.03$ yielded best performance.
 
@@
 
-The procedure to implement the one-hot autoencoder was similar to the array's autoencoder one.
-Briefly, it was shown experimentally that for this model, a deeper encoder with layers of gradually decreasing number of units were more suited.
-We tried both a wider and shallower architecture and a deeper and narrow architecture.
+The procedure to implement the one-hot autoencoder was similar to the array autoencoder's.
+Briefly, it was shown experimentally that for the one-hot model, a deeper encoder with layers of gradually decreasing number of units were more suited.
+We tried both a wider and shallower architecture and a deeper and narrower architecture.
 The second had better BER performance.
@@
 
-In terms of noise, their presence in the channel degraded vastly the results, as the DNN was not able to learn.
+In terms of noise, their presence in the channel during training degraded vastly the results, as the DNN was not able to learn.
 The challenge of the one-hot encoder implementation is the fact that we are trying to represent a 256 size length vector into a 16 length vector to transmit through the specified channel.
-The hypothesis is that it requires several steps to converge into a representative state.
-Thus, no noise channel was implemented, only a rounding layer.
+Thus, only a rounding layer was implemented.
 
-The decoder utilized has fewer layers than the array autoencoder, however more parameters.
+The decoder utilized has only one hidden layer, but more parameters than the array autoencoder's decoder.
 In terms of activation functions, there is no difference from the DNN one-hot decoder, which utilized a multi-class classification framework as well.
@@
 This section is divided in three parts.
-The first shows the results obtained to the two decoders.
+The first shows the results obtained for the two decoders.
 Secondly, the results for the autoencoders is brought with some observations during the training period.
-The discussion and analysis for these two subsections is around the BER performance.
+The discussion and analysis of these two subsections is around the BER performance.
 Finally, a decoding time performance analysis is done and the results are shown and discussed.
@@
 We can infer from figures \ref{fig:ArrayD} and \ref{fig:1HD} that both the array decoder and the one-hot decoder are able to learn a decodification function.
-Moreover,  its performances replicate the MAP algorithm's decoding capability for this specific encoding.
-Hence, DNN decoders could substitute the mathematically optimal MAP decoder for a stochastic channel, without compromising the bit error probability.
+Moreover, their performances replicate the MAP algorithm's decoding capability for this specific encoding.
+Hence, DNN decoders could substitute the mathematically optimal MAP decoder for a stochastic channel, without compromising the bit error rate.
 
@@
 As a result, since they express the same results, we can assume that the DNN decoder was able to learn the MAP algorithm during training phase.
-More interestingly, is the convergence period for each decoder.
+More interestingly, is the difference between the convergence period of the decoders.
 The one-hot decoder reached its peak performance much faster in comparison with the array decoder in terms of training epochs.
-I would attribute this increased performance to the activation function of the last layer.
+We would attribute this increased performance to the activation function of the last layer and the implementation of the BSC in the first decoder.
 The sigmoid function used in the array decoder may lead to sharp damp gradients, gradient saturation or slow convergence during BP.
@@
 Refer to \cite{2018arXiv181103378N} for deeper details in each activation function characteristics.
+And the noise drawback in training a NN is enlarging its training time.
 
 \subsection{Autoencoders}
-BER Curves of simulation varying p, best choice of p, analyse autoencoder with best choice of p
 
-In the beginning we expected that the autoencoders would beat the MAP algorithm performance for a fixed linear block encoding.
+In the beginning, we expected that the autoencoders would outperform the MAP algorithm for a fixed linear block encoding.
 They were supposed to learn a non linear encoding and its particular MAP rule which would perform better than the studied MAP for a low crossover probability range.
-It turned out, however, since there are an enormous amount of possible non-linear codes, it is difficult to the learning algorithm to find the best one.
+It turned out however that, since there are an enormous amount of possible non-linear codes, it is difficult to the learning algorithm to converge to the best one.
 Often during training, the DNN would find a local minimum which its cost function would be very close to zero, even though its performance during prediction phase was far from optimal.
@@
 It is remarkable how the curves tend to replicate the same type of behavior.
-I believe it shows how the local minimums of the solution space of a NN are indeed close to each other, and the weight values may not be indeed that different.
+We believe it shows how the local minima of the solution space of a NN are indeed close to each other, and the weight values may not be indeed that different.
 
@@
 
-However, it is noticeable the difference in error decoding in terms of a DNN trained with a $p_t=0$ and $p_t=0.3$ in Fig.
+It is noticeable the error correction difference between a DNN trained with a $p_t=0$ and $p_t=0.3$ in Fig.
 \ref{fig:parrayanalysis}.
-Just similar to the array decoder, introducing a channel noise of some kind is beneficial to having a better error correction capability.
-Using noise also helps in the regularization of the DNN and approximates the training to the real application.
+Similarly to the array decoder, introducing a training noise is beneficial to having a better error correction capability.
 
 We can observe from Fig.
-\ref{fig:arrayautoencoder} that for $p<0.02$, the autoencoder performance is actually worse than the tested MAP algorithm.
+\ref{fig:arrayautoencoder} that, for $p<0.02$, the autoencoder performance is actually worse than the tested MAP algorithm.
 Several hypothesis were considered for this comportment.
@@
 Thus, the decoder replicates well the MAP for this specific encoding.
-Then, we experimented with training parameters and architecture, but could not find a configuration which would perform better than the traced MAP for low values of $p$.
+Next, we experimented with training parameters and architecture, but could not find a configuration which would perform better than the traced MAP for low values of $p$.
 
@@
 Equally to the array autoencoder, we tested a MAP rule for the learned encoder and found that the DNN decoder portion performs just like the MAP.
-Thus, once again the problem is in the learned encoding.
-We tried 80 different configurations, though we could not find one which could outperform the reference model.
-We can also highlight the trace of the BER curve, approximating the reference model one.
-Other one-hot autoencoders showed similar curves but translated up in the $y$ axis.
+Thus, once again, the problem may be in the learned encoding.
+We tried 80 different configurations, but we could not find one which could outperform the reference model.
+Other one-hot autoencoders generated similar curves but translated up in the $y$ axis.
 
@@
 This is due their fairly low dimensional complexity and one shot feature.
-While the MAP rule must calculate the distance between the received codeword to every possible codeword, the decoders once trained are able to decode the received code word into the message of highest probability with simple matrix multiplications.
+While the MAP rule must calculate the distance between the received codeword to every possible codeword, the trained decoders are able to decode the received code word into the message of highest probability with simple matrix multiplications.
 The latter efficiency is potentialized when using decoding computers equipped with GPUs.
 
-In spite of the one-hot decoder has considerably fewer parameters than the array decoder, it must calculate a $argmax$ function to deliver the actual message that was transmitted.
+Although the one-hot decoder has considerably fewer parameters than the array decoder, it must calculate a $argmax$ function to deliver the actual message that was transmitted.
 This extra costly computation is also the reason for the increased time for the one-hot autoencoder.
@@
 The data is normalized to the average MAP algorithm decoding time.
-The simulation was based on CPU calculations.
+The simulation was done entirely by CPU.
 }
@@
 \end{table}
-
-paragraph that talks about CPU, not GPU.
 

### 2019-06-25 Rapport/Article/Article.tex
Seance 06/25/2019

@@
 \ref{fig:cs} illustrates.
-As a result of channel imperfections, Shannon theorized in \cite{Shannon:2001:MTC:584091.584093} the ultimate reliable data rate which can be transmitted through a communication system with an arbitrarily small probability error.
+As a result of channel imperfections, Shannon theorized in \cite{Shannon:2001:MTC:584091.584093} the ultimate reliable data rate which can be transmitted through a communication system with an arbitrarily small probability of error.
 
@@
 In their research, results in terms of BER over SNR for a DNN based autoencoder and for a convolutional neural network (CNN) based autoencoder were treated.
-They used a range of SNR - from $-10 \text{dB}$ to $15 \text{dB}$ - with QPSK and QAM16 modulation as benchmarks.
+They used a range of SNR from $-10 \text{dB}$ to $15 \text{dB}$ with QPSK and QAM16 modulation as benchmarks.
 This analysis was conducted for a binary input message.
@@
 al.
-in \cite{2017arXiv171008379G} proved that a deep learning-based channel decoder could actually learn a decoding algorithm rather than just being a simple classifier.
+in \cite{2017arXiv171008379G} proved that a deep-learning based channel decoder could actually learn a decoding algorithm rather than just being a simple classifier.
 They introduced code words that were not been used in the training set, and the trained NN was able to correctly decode it.
@@
 In this work, we will implement a ML autoencoder for a BSC which performs similar to the MAP decoder in real applications for a range of cross over probability $p$ from $0$ to $0.1$.
-In its most simple form, a channel autoencoder includes an encoder, a noisy channel and a decoder.
-Using state-of-the art DNN algorithms to find the best solution of the problem, this work aims to contribute to set up a higher standard in terms of performance in bit-error correction and reduced delay for digital communication applications.
+In its most simple form, a channel autoencoder includes an encoder, a noisy channel and a decoder using state-of-the art DNN algorithms to find the best solution of the problem.
+This work aims to contribute to set up a higher standard in terms of performance in bit-error correction and reduce delay for digital communication applications.
 In a near future, this disruptive methodology for error correction using ML could replace mathematically optimal decoders which are the current guideline.
@@
 
-The main theoretical concepts will be briefly introduced. They are organized by theme and is aimed to a reader less familiar with the particular subject.
+The main theoretical concepts will be briefly introduced. They are organized by theme and are aimed to a reader less familiar with the particular subject.
 
@@
 Briefly, a linear encoder will index a message $\textbf{u}^k \in \mathcal{U} \subseteq \{0,1\}^k$ to a code word $\textbf{x}^n \in \mathcal{X} \subseteq \{0,1\}^n$.
-This mapping is done by the \textit{generator matrix} $ \textbf{G} $ through equation \ref{eq:encoding}.
+This mapping is done by the \textit{generator matrix} $ \textbf{G} $ through equation \eqref{eq:encoding}.
 
@@
 
-The Hamming distance $d_H:\mathcal{X}\mapsto \mathbb{Z}$ is a metric in the word set which in the context of linear coding is equal to equation \ref{eq:dh}.
+The Hamming distance $d_H:\mathcal{X}\mapsto \mathbb{Z}$ is a metric in the word set which in the context of linear coding in digital communication is equal to equation \eqref{eq:dh}.
 Noteworthy for the MAP rule, a non zero Hamming distance indicates that there is a block error, and ultimately at least one binary error in the received code word.
@@
 They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{journals/ett/RobertsonHV97}.
-It means that if the real SNR is smaller, the decoder will not perform well.
+It means that, the smaller the real SNR is, the worst is the decoder performance.
 
@@
 
-\caption{MLNN representative diagram, where $\textbf{y}^n$ is the input vector, $\textbf{r}_{l}^{j}$ is a hidden layer vector
+\caption{NN representative diagram, where $\textbf{y}^n$ is the input vector, $\textbf{r}_{l}^{j}$ is a hidden layer vector
 and $\hat{\textbf{u}}^{k}$ is the output vector.} \label{fig:NN}
@@
 Another commonly used loss function, indispensable for this research, is the \textit{mean squared error} (MSE).
-The MSE is the  sum of squared distances between our target variable and predicted values and it is defined in equation \ref{eq:mse}.
+The MSE defined in equation \eqref{eq:mse} is the sum of the squared distances between the target variable and the predicted values.
 
@@
 It matches the input and its respective desired output.
-The objective is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{DBLP:journals/corr/OSheaH17}.
+The objective of training is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{DBLP:journals/corr/OSheaH17}.
 
@@
 
-For calculating the BER versus the crossover probability $p$ with confidence over the results, a Monte Carlo simulation was elaborated.
+For calculating the BER versus crossover probability $p$ with confidence over the results, a Monte Carlo simulation was elaborated.
 This stochastic computational algorithm uses pseudo-random sampling to solve different problems.
@@
 They are coded by a generator matrix or by a DNN encoder and then transmitted through the BSC channel of crossover probability $p$.
-Their information arrives corrupted by noise in the decoder and is subsequently partially corrected by a decoder.
+Their information arrives corrupted by noise in the decoder and is subsequently completely or partially corrected by a decoder.
 Finally, the average BER is calculated and plotted.
@@
 
-Although the MAP algorithm is the mathematically optimal decoder as cited previously, it has a high time complexity $(\mathcal{O}(2^k))$, since it computes the Hamming distance between the received code word and every code word from the code book.
+Although the MAP algorithm is the mathematically optimal decoder as aforementioned, it has a high time complexity $(\mathcal{O}(2^k))$, since it computes the Hamming distance between the received code word and every code word from the code book.
 For more details, the pseudo code of the algorithm is available below.
@@
 The architecture of the DNN decoder necessary to achieve a great performance in terms of BER is not evident.
-To begin, we must define which layers were used and why.
+To begin, we must define which layers have been used and why.
 A \textit{Lambda} layer is a layer in which the activation function is user defined and the parameters are non-trainable.
@@
 Thus it is often used to change the dimensions of a vector.
-In other words, it applies a rotation, scaling and translation transform to a vector as shown in equation \ref{eq:eqFP}.
+In other words, it applies a rotation, scaling and translation transform to a vector as shown in equation \eqref{eq:eqFP}.
 The values in the weight matrix are trainable parameters which are updated during BP \cite{chollet2015keras}.
 
-The choice of the crossover probability of the training channel, $p_{t}$, was based in the work of \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the optimal in terms of decoding precision after the DNN is trained.
+The choice of the crossover probability of the training channel, $p_{t}$, was based in the work \cite{DBLP:conf/acssc/BenammarP18} which shows $p_{t}=0.07$ to be the optimal in terms of decoding precision after the DNN is trained.
 In addition, a deeper DNN showed best results in \cite{DBLP:journals/corr/OSheaH17} where 3 hidden layers were used.
@@
 In this way, we could guarantee that our training environment were representative of a real life application.
-Where different bits are transmitted with the same probability.
+Where different bits are transmitted with equal probability.
 
-Even though it is not recommended to train a NN with all the available data - because it might lead to over-fitting -, for this case does not seems to be a problem.
+Even though it is not recommended to train a NN with all the available data, because it might lead to over-fitting, for this case it does not seems to be a problem.
 Apparently, the presence of a BSC is enough to regularize the training procedure and prevent over-fitting.
@@
 The choice of training parameters is a mix between what is standardized in literature and a hyper parametric analytic approach, where several trials lead to a decision.
-First, the binary cross-entropy as loss function is widely used in literature and showed to converge to the desired result consistently during tests.
+First, the binary cross-entropy was used as loss function since it is widely used in literature and showed to converge to the desired result consistently during tests.
 For the optimizer, the \textit{Adam} built-in Keras optimizer was chosen.
-The Adam optimizer algorithm is similar to a classic stochastic gradient descent algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}.
+The Adam optimizer algorithm is similar to a classic stochastic gradient descent (SGD) algorithm, but computes individual adaptive learning rates for different parameters from estimates of first and second moments of the gradients \cite{chollet2015keras}.
+
 Next, a batch size matching the size of the training set which is sufficiently small to still operate in the regime of small batch was chosen.
@@
 Finally, the necessary number of epochs to training is sensitive to the problem, since it impacts directly the training computational cost.
-As a result, we tried several training epoch sizes and found $2^{16}$ to have precise results and be computationally viable.
+As a result, we tried several training epoch sizes and found $2^{16}$ to have precise results and still be computationally viable.
 The training parameters are displayed in table \ref{tab:arraydecoderTrain}.
@@
 It contains only one hidden layer of $256$ units which maps the input of size 16 to the $2^8=256$ possible messages.
-Hence, this characterizes a multi-class classification problem, meaning that the chosen activation function is the Softmax (Eq.
-\ref{eq:soft}).
+Hence, this characterizes a multi-class classification problem, meaning that the chosen activation function is the softmax defined in equation \eqref{eq:soft}.
 This training technique is known in the literature as \textit{one-hot} training.
 
-The denomination one-hot comes from the output vector which is usually a long vector of elements close to zeros with one element close to one.
+The denomination \textit{one-hot} comes from the output vector which is usually a long vector of elements close to zeros with one element close to one.
 This \textit{hot} vector element indicates which class the NN has classified the input with the most probability.
@@
 
-Despite the NN complexity benefits of the one-hot decoder, it is required to transform the predicted vector into a message.
+Despite the NN complexity benefits of the one-hot decoder, it must transform the predicted vector into a message.
 Consequently, an additional operation is needed.
@@
 They are one-hot vectors which maps into real messages through a bijective function.
-Thus, it is a matrix of dimension $256\times256$ composed of 256 different on-hot vectors.
+Thus, it is a matrix of dimension $256\times256$ composed of 256 different one-hot vectors.
 The training data is still the linear encoding of the real messages by the matrix $\textbf{G}$.
@@
 The one-hot decoder was trained both with and without a noise layer.
-The model without a integrated channel converged faster to the solution.
+The model without an integrated channel converged faster to the solution.
 Tables \ref{tab:onehotdecoder} and \ref{tab:onehotdecoderTrain} resume the architecture and training variables, respectively.
@@
 Aiming to eliminate completely the need of an explicit encoding, a technique commonly used in ML for image recognition is explored.
-The autoencoder is able to learn not only a decoding algorithm, but also a encoding one as introduced in the following section.
+The autoencoder is able to learn not only a decoding algorithm, but also a encoding one as demonstrated in the following section.
 
@@
 For the DNN autoencoder, two different concepts were studied.
-The array autoencoder takes as input the messages and it tries to replicate them.
-It encodea the messages into a 16 bit code word, transmits the code word through a noisy channel and decodes them in the second portion of the DNN.
-A one-hot autoencoder is inputted by a one-hot vector representing one unique $k$-bit message.
-It encodes the vector into a $n$-bit long code word, transmits them through a noisy channel and decode them in the second part of the DNN.
-
-This end-to-end approach for both architectures is promising for digital communication, since there is no need for a predefined linear encoding.
+The end-to-end approach for both architectures is promising for digital communication, since there is no need for a predefined linear encoding.
 Theoretically, the DNN could learn the best encoding and associated decoding to a specific channel.
@@
 
-To be clear in the discussion of this and the following sessions, the autoencoder will be divided in three: the encoder, the channel and the decoder portions.
+The autoencoder will be divided in three: the encoder, the channel and the decoder portions.
 Fig.
 \ref{fig:DDNNAutoencoder} illustrates this division.
-Even though the autoencoder could in theory be able to achieve at least the same performance as the MAP, finding the parameters which unveils its capabilities requires multiple testing and a rigorous hyper-parametric analysis.
-In total, there were tested 71 array and 80 one-hot autoencoders different architectures or training configurations.
+Even though the autoencoder could in theory be able to achieve at least the same performance as the MAP, finding the parameters which unveils its best capabilities requires multiple testing and a rigorous hyper-parametric analysis.
+In total, there were tested 71 array autoencoders and 80 one-hot autoencoders different architectures and training configurations.
 
@@
 
-For the array autoencoder, we first experimented with the depth of the encoder.
+The array autoencoder takes as input the messages and tries to replicate them as output.
+It encodes the messages into a 16 bit code word, transmits the code word through a noisy channel and decodes them in the second portion of the DNN.
+
+We first experimented with the depth of the encoder.
 The number of hidden layers in a DNN influences the BP algorithm performance.
@@
 The encoder portion was first analyzed in a standalone session.
-We trained the encoder to replicate the linear encoding proposed in the preceding sections as we studied the effects of each parameter in the training performance.
+We trained the encoder to replicate the linear encoding proposed in the reference model.
+We studied the effects of each parameter in the training performance.
 A more wide architecture showed faster convergence and the use of three or more hidden layers had no positive effects.
@@
 In spite of this, there were no outstanding difference in decoding with or without the last hidden layer of the decoder.
-In the other hand, the tests with a condensate NN with one hidden layer demanded a greater training time to arrive in the same results.
+In the other hand, the tests with a shallower NN with one hidden layer demanded a greater training time to arrive in the same results.
 Hence, we dropped one hidden laeyer from the cedoder and kept two
@@
 While training the autoencoder, we noticed that the binary cross-entropy loss function would need more training epochs than the MSE and \textit{logcosh} loss functions, even though they converged in similar results.
-The batch size was maintained as the size of the first dimension of the training set and there was no reason for changing the optimizer.
-Notwithstanding, the ensemble of changes and optimizations were not enough to find a suitable architecture for channels with $p<0.02$ as the MAP algorithm for the 1/2 rate encoder would do better in this scenario.
+The batch size and the optimizer were maintained as there was no reason for changing them.
+Notwithstanding, the changes and optimization were not enough to find a suitable architecture for channels with $p<0.02$ as the MAP algorithm for the 1/2 rate encoder would do better in this scenario.
 Table \ref{tab:arrayautoencoderTrain} summarizes the array autoencoder training parameters.
@@
 Furthermore, adding batch normalization layers inside the DNN was fundamental.
-With greater training parameters, achieving convergence in a reasonable time requires more computing power of no specific technique is implemented.
+With greater training parameters, achieving convergence in a reasonable time requires more computing power if no specific technique is implemented.
 The frequent changes of weight values in the intermediate layers during each training step   is the principal contributor to the training cost.
@@
 \subsubsection{One-hot Autoencoder}
+
+A one-hot autoencoder is inputted by one-hot vectors representing  unique $k$-bit message.
+It encodes the vectors into $n$-bit long code words, transmits them through a noisy channel and decode them in the second part of the DNN.
 
@@
 The first shows the results obtained for the two decoders.
-Secondly, the results for the autoencoders is brought with some observations during the training period.
+The second, the results for the autoencoders is brought with some observations during the training period.
 The discussion and analysis of these two subsections is around the BER performance.
@@
 The sigmoid function used in the array decoder may lead to sharp damp gradients, gradient saturation or slow convergence during BP.
-While the one-hot decoder uses the softmax function which its partial derivatives have a better behavior during BP.
-Refer to \cite{2018arXiv181103378N} for deeper details in each activation function characteristics.
-And the noise drawback in training a NN is enlarging its training time.
+While the one-hot decoder uses the softmax function, which its partial derivatives have a better behavior during BP.
+Refer to \cite{2018arXiv181103378N} for deeper details in activation function characteristics.
+And the drawback of using noise layers in training a NN is enlarging the training time.
 
@@
 There were selected 20 different BSC channels.
-For each channel, $100000$ messages were codified, transmitted and decodified for each decoder, autoencoder and MAP.
+For each channel, $100000$ messages were codified, transmitted and decodified for each model.
 This process was timed and repeated several times.
@@
 
-Although the one-hot decoder has considerably fewer parameters than the array decoder, it must calculate a $argmax$ function to deliver the actual message that was transmitted.
+Although the one-hot decoder has considerably fewer parameters than the array decoder, it must calculate an $argmax$ function to deliver the actual message that was transmitted.
 This extra costly computation is also the reason for the increased time for the one-hot autoencoder.
-It must first encode the message, then transmit it and finally decode it, performing two times per message a $argmax$ operation.
+It must first encode the message, then transmit it and finally decode it, performing two times per message an $argmax$ operation.
 
@@
 In addition, because of the number of parameters, their calculation in CPU is slow and they showed a greater delay when compared to the MAP rule.
-Though, they must take the most advantage of GPU based calculations and the delay time could be improved using this technology.
+Though, they must take the most advantage of GPU based calculations and the delay time could be improved by this technology.
 
-A more rigorous hyper-parametric analysis could find a combination of parameters which yields better BER performance for the autoencoders and is strongly suggested for future work.
-In addition, in the future, a real implementation of the DNN decoders could be tested to confirm their capabilities as it is not the main focus of this work.
-We suggest using more advanced NN topology such as a recurrent neural network (RNN) or a generative adversarial network (GAN).
+A more rigorous hyper-parametric analysis could find a combination of parameters which yields better BER performance for the autoencoders.
+In addition, a real implementation of the DNN decoders could be tested to confirm their capabilities in a critical scenario.
+We also suggest using more advanced NN topology, such as a recurrent neural network (RNN) or a generative adversarial network (GAN).
 Their implementation and training are more complex, however they have recently demonstrated outstanding results in other fields.
-
-%Talk about feasibility of decoders, improved performance of the channel delay with DNN decoders, challenges in implementing this architecture in real life, challenges in training the NNs, Future work, what else could be done with NN in error correction, Advantages and disadvantages in comparison with linear decoding.
 
@@
 
-I would like to thank the guidance and advice from my supervisor Meryem Benammar, Ph.D, my colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research and Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project.
+I would like to thank the guidance and advice from my supervisor Meryem Benammar, Ph.D, my colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research and Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the realization of the project.
 

### 2019-06-25 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/25/2019

@@
 \usepackage{amsmath,amssymb,amsfonts}
+\usepackage{outlines}
 
@@
 
-\author{E.~Dadalto Camara Gomes\inst{1} \and M.~Benammar\inst{2}}
+\author{E.~Dadalto Camara Gomes\inst{1}}
 % - Give the names in the same order as the appear in the paper.
@@
 Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr
-  \and
-  \inst{2}%
-  Department of Electronics, Optronics, and Signal processing\\
-  ISAE-SUPAERO\\
-  31055, Toulouse, France\\
-  Email: meryem.benammar@isae-supaero.fr}
+%  \and
+ % \inst{2}%
+ % Department of Electronics, Optronics, and Signal processing\\
+ % ISAE-SUPAERO\\
+ % 31055, Toulouse, France\\
+ % Email: meryem.benammar@isae-supaero.fr
+ }
 % - Use the \inst command only if there are several affiliations.
@@
 \section{Introduction}
+\subsection{Context}
+\begin{frame}{Context I}{Digital communication system definition}
 
-\begin{frame}{Context}{Communication system context in general - what field will I be treating}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+\begin{figure}
+  \centering
+   \includegraphics[width=0.6\textwidth]{images/Communication_system_schema_from_book}
+  \caption{Block diagram representation of a communication system.}%%\label{fig:cs}
+\end{figure}
+
+\begin{figure}
+  \centering
+   \includegraphics[width=0.65\textwidth]{images/simple_sys}
+  \caption{Simplified communication system.
+Where $\textbf{u}^k$ is a source message, $\textbf{x}^n$ a code word, $\textbf{y}^n$ the received code word and $\hat{\textbf{u}}^k$ the decoded message.}\label{fig:cs}
+\end{figure}
+
 \end{frame}
 
-\begin{frame}{Context}{Machine Learning applications - what could we do in communication system}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+\begin{frame}{Context II}{Machine learning applied to communication system}
+
+  \begin{figure}
+  \centering
+   \includegraphics[width=0.35\textwidth]{images/BSC}
+  \caption{Binary symmetric channel (BSC) representation.}%%\label{fig:cs}
+\end{figure}
+
+  \begin{figure}
+  \centering
+   \includegraphics[width=0.85\textwidth]{images/autoencoder}
+  \caption{Deep neural network (DNN) based autoencoder.}%%\label{fig:cs}
+\end{figure}
+
 \end{frame}
 
-\begin{frame}{Relevance \& Challenges}{Explain why the work is relevant and explain what are the challenges}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+\begin{frame}{Motivation \& Challenges}
+
+\begin{outline}
+ \1 Low latency and high bandwidth wireless communication are key to critical systems.
+   \2 i.g. airplanes, satellites, cellular communication and 5G operations.
+\pause
+ \1 As opposite to structured algorithms, machine learning (ML) algorithms do not require rigidly designed models and can take non-linearities effortlessly into account.
+\pause
+ \1  ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths.
+   \2 It translates to lower decoding latency, resulting ultimately in gain of bandwidth over standard methods.
+
+\end{outline}
+
 \end{frame}
 
-\begin{frame}{Problem Statement}{What exactly I will solve in this work}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
+\subsection{Problem Statement}
+\begin{frame}{Problem Statement}
+
+\begin{outline}
+\1 Could ML based decoders and autoencoders for a BSC perform similarly to the MAP decoder in real applications and have lower communication delay?
+\pause
+\1 This work aims to contribute to set up a higher standard in terms of performance in bit-error correction and demonstrate that ML based models can reduce delay for digital communication applications.
+\end{outline}
+
 \end{frame}
@@
 % with the \pause command:
-\begin{frame}{Second Slide Title}
-  \begin{itemize}
-  \item {
-    First item.
-    \pause % The slide will pause after showing the first item
-  }
-  \item {   
-    Second item.
-  }
+%\begin{frame}{Second Slide Title}
+%%  \begin{itemize}
+ % \item {
+%    First item.
+%%    \pause % The slide will pause after showing the first item
+ % }
+ %% \item {   
+  %  Second item.
+%  }
   % You can also specify when the content should appear
   % by using <n->:
-  \item<3-> {
-    Third item.
-  }
-  \item<4-> {
-    Fourth item.
-  }
+%  \item<3-> {
+%    Third item.
+%  }
+%  \item<4-> {
+%    Fourth item.
+%  }
   % or you can use the \uncover command to reveal general
   % content (not just \items):
-  \item<5-> {
-    Fifth item. \uncover<6->{Extra text in the fifth item.}
-  }
-  \end{itemize}
-\end{frame}
+%  \item<5-> {
+ %   Fifth item. \uncover<6->{Extra text in the fifth item.}
+%  }
+%  \end{itemize}
+%\end{frame}
 
@@
 \begin{frame}{Maximum a Posterior (MAP) Rule}{Implementation of a MAP decoder for a linear block code through a BSC.}
+
+\begin{itemize}
+\item{The concept of a MAP decoder algorithm for sequences is choosing a message which maximizes the a posterior probability.}
+\end{itemize}
+
+\begin{center}
+$f(y) = \underset{x\in \mathcal{X}}{\text{arg max}} P_{X|Y}(x|y)$
+\end{center}
 
@@
 \subsection{Design \& Architecture}
-\begin{frame}[allowframebreaks]{Neural Network's Design and Architecture}{Show the architecture used for each case and remarks some important parameters}
-
+\begin{frame}{Neural Network's Design and Architecture Basics}{}
+%Show the architecture used for each case and remarks some important parameters
 	\begin{figure}[!ht]
@@
     \includegraphics[width=0.4\textwidth]{images/MLNN}
-    \caption{MLNN representative diagram, where $\textbf{y}^n$ is the input vector, $\textbf{r}_{l}^{j}$ is a hidden layer vector
+    \caption{NN representative diagram, where $\textbf{y}^n$ is the input vector, $\textbf{r}_{l}^{j}$ is a hidden layer vector
 and $\hat{\textbf{u}}^{k}$ is the output vector.} \label{fig:NN}
 \end{figure}
-
+\end{frame}
 	\framebreak
-
+\begin{frame}{Array Decoder DNN Architecture and Implementation}{}
     \begin{table}
-\caption{DNN array decoder architecture and parameters.} 
+\caption{DNN array decoder architecture and training parameters. The input is a code word of size $n$ and the output is the decoded message of size $k$.} 
 \begin{center}
@@
     \thickhline
- Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Loss function & Optimizer & N. Epochs & Batch Size \\ \hline
  Binary cross-entropy & Adam & $2^{16}$ & $256$ \\
@@
 \end{center}
-
+\end{frame}
   \framebreak
-
+\begin{frame}{One-hot Decoder DNN Architecture and Implementation}
     \begin{table}
-\caption{DNN one-hot decoder architecture and parameters.} 
+\caption{DNN one-hot decoder architecture and parameters. The input is a code word of size $n$ and the output is a unique one-hot vector of size $2^k$ which is mapped into its correspondent message.} 
 \begin{center}
@@
     \thickhline
- Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Loss function & Optimizer & N. Epochs & Batch Size \\ \hline
  Binary cross-entropy & Adam & $2^{14}$ & $256$ \\
@@
 \end{center}
+\end{frame}
 
-	\framebreak
-
+\begin{frame}{DNN Autoencoder Design}{}
 	\begin{figure}[!ht]
   \centering
-    \includegraphics[width=0.8\textwidth]{images/DNN_autoencoder}
-    \caption{Representation of a DNN autoencoder composed of dense layers.}\label{fig:DDNNAutoencoder}
+    \includegraphics[width=0.6\textwidth]{images/DNN_autoencoder}
+    \caption{Representation of a DNN autoencoder composed of dense layers.
+Each block represents a DNN which together learn the channel encoding and decoding in an end-to-end manner.
+They also operate together in bit error correction.}\label{fig:DDNNAutoencoder}
 \end{figure}
+\end{frame}
 
-	\framebreak
-
+\begin{frame}{Array Autoencoder DNN Architecture and Implementation}{This design yielded best error correction results out of 80 models tried}
     \begin{table}
-\caption{DNN array autoencoder architecture.} 
+\caption{DNN array autoencoder architecture and training parameters.} 
 \begin{center}
@@
     \thickhline
- Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Loss function & Optimizer & N. Epochs & Batch Size \\ \hline
  MSE & Adam & $2^{17}$ & $256$ \\
@@
 \footnotetext{Batch Normalization (BN)}
+\end{frame}
 
-	\framebreak
-
+\begin{frame}{One-hot Autoencoder DNN Architecture}{This design yielded best error correction results out of 71 models tried}
     \begin{table}
-\caption{DNN one-hot autoencoder architecture.} 
+\caption{DNN one-hot autoencoder architecture and training parameters.} 
 \begin{center}
@@
     \thickhline
- Loss func. & Optimizer & N. Epochs & Batch Size \\ \hline
+ Loss function & Optimizer & N. Epochs & Batch Size \\ \hline
  MSE & Adam & $2^{16}$ & $256$ \\
@@
 
-\subsection{Error Correction \& Predictions}
-\begin{frame}{Error Correction and Monte Carlo Simulations}{Explain how we could use NN to predict the results with certain confidence.}
-  \begin{itemize}
-  \item {
-    My first point.
-  }
-  \item {
-    My second point.
-  }
-  \end{itemize}
-\end{frame}
+%\subsection{Error Correction \& Predictions}
+%\begin{frame}{Error Correction and Monte Carlo Simulations}{Explain how we could use NN to predict the results with certain confidence.}
 
-\begin{frame}{Blocks}
-\begin{block}{Block Title}
-You can also highlight sections of your presentation in a block, with it's own title
-\end{block}
-\begin{theorem}
-There are separate environments for theorems, examples, definitions and proofs.
-\end{theorem}
-\begin{example}
-Here is an example of an example block.
-\end{example}
-\end{frame}
+%\end{frame}
+
+%\begin{frame}{Blocks}
+%\begin{block}{Block Title}
+%You can also highlight sections of your presentation in a block, %%with it's own title
+%\end{block}
+%\begin{theorem}
+%%There are separate environments for theorems, examples, definitions and proofs.
+%\end{theorem}
+%%\begin{example}
+%Here is an example of an example block.
+%\end{example}
+%\end{frame}
 
@@
 \section{Results \& Discussions}
-\subsection{DNN Decoders}
-\begin{frame}{DNN Array Decoder}{Show the results for the array decoder in terms of train p, Mep, Parameters, etc}
+\subsection{Deep Neural Network Based Decoders}
+\begin{frame}{DNN Array Decoder Error Correction Performance}{For each of 20 different channels, 100000 messages were sent.
+The average error was calculated and traced.
+The same procedure was applied throughout the work.}
 
@@
   \centering
-    \includegraphics[width=0.8\textwidth]{images/MLNN_Mep_65536_ptrain_007}
-    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
+    \includegraphics[width=0.7\textwidth]{images/MLNN_Mep_65536_ptrain_007}
+    \caption{Array decoding BER performance. DNN trained with a channel crossover probability error of $p_t=0.07$. The decoder successfully learned the reference MAP algorithm.}\label{fig:ArrayD}
 \end{figure}
@@
 
-\begin{frame}{DNN One-hot Decoder}{Show the results for the one-hot decoder in terms of train p, Mep, Parameters, etc}
+\begin{frame}{DNN One-hot Decoder Error Correction Performance}
 
@@
   \centering
-    \includegraphics[width=0.8\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
-    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
+    \includegraphics[width=0.7\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
+    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$. The decoder successfully learned the reference MAP algorithm.}\label{fig:1HD}
 \end{figure}
@@
 
-\subsection{DNN Autoencoders}
-\begin{frame}[allowframebreaks]{DNN Array Autoencoder}{Show the results for the autoencoder in terms of train p, Mep, Parameters, etc}
+\subsection{Deep Neural Network Based Autoencoders}
+\begin{frame}[allowframebreaks]{DNN Array Autoencoder Error Correction Performance}
 
@@
   \centering
-    \includegraphics[width=0.8\textwidth]{images/MAP_AutoencoderArray_Mep_65536_64_128_256_p_analysis}
-    \caption{Training crossover probability simulation for the array autoencoder. $P_t=0.3$ demonstrated to have best performance to this particular architecture.}\label{fig:parrayanalysis}
+    \includegraphics[width=0.7\textwidth]{images/MAP_AutoencoderArray_Mep_65536_64_128_256_p_analysis}
+    \caption{Training crossover probability simulation for the array autoencoder. $P_t=0.03$ demonstrated to have best performance to this particular architecture.}\label{fig:parrayanalysis}
 \end{figure}
@@
   \centering
-    \includegraphics[width=0.8\textwidth]{images/AutoencoderArray_Mep_65536_ptrain_003_logcosh}
-    \caption{Array autoencoder BER performance. DNN array autoencoder trained with a channel crossover probability error of $p_t=0.03$.}\label{fig:arrayautoencoder}
+    \includegraphics[width=0.7\textwidth]{images/AutoencoderArray_Mep_65536_ptrain_003_logcosh}
+    \caption{Array autoencoder BER performance.
+DNN array autoencoder trained with a channel crossover probability error of $p_t=0.03$.
+The array autoencoder could not outperform the reference MAP algorithm for channels with $p<0.02$.}\label{fig:arrayautoencoder}
 \end{figure}
@@
 
-\begin{frame}{DNN One-hot Autoencoder}{Show the results for the autoencoder in terms of train p, Mep, Parameters, etc}
+\begin{frame}{DNN One-hot Autoencoder Error Correction Performance}
 
@@
   \centering
-    \includegraphics[width=0.75\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
-    \caption{One-hot autoencdoer BER performance. Trained without a noisy channel.}\label{fig:1hautoencoder}
+    \includegraphics[width=0.7\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
+    \caption{One-hot autoencdoer BER performance. Trained without a noisy channel. No architecture could outperform the reference MAP algorithm in terms of BER.}\label{fig:1hautoencoder}
 \end{figure}
@@
 \subsection{Time Analysis}
-\begin{frame}{Delay Time Analysis}{Comparison of encoding, transmission and decoding time for each method.}
+\begin{frame}{Delay Time Analysis}{Calculations done by CPU}
   \begin{table}
-\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. } 
+\caption{Decoding time comparison between the reference MAP algorithm and the DNN decoders and autoencoders.
+The data is normalized to the average MAP algorithm decoding time.
+The decoders had around a 25\% improvement in delay time in comparison to the MAP.}
 \begin{center}
@@
 \section{Conclusions}
+
 \begin{frame}{Conclusions}
@@
   \item {
-    My first point.
+    The feasibility of a machine learning based channel decoder were demonstrated for a communication system with a binary symmetric channel.
   }
   \item {
-    My second point.
+    The designed decoders improved around 25\% the channel delay when compared to the maximum a posteriori rule for a $(16,8)$ code.
+  }
+  \item{
+Their small number of parameters aligned with a GPU equipped decoding computer is promissory to be a substitute of the MAP rule and achieve same bit error correction rate in a real application. \\
+  }
+  \item{
+We also proved that autoencoders can learn a non linear encoding function and its decoding function for a BSC.  
+  }
+  \item{
+   There are still some challenges in training to solve.
   }
@@
   \item {
-    My first point.
+    A more rigorous hyper-parametric analysis could find a combination of parameters which yields better BER performance for the autoencoders.
+  }
+    \item{
+  Using more advanced NN topology such as a recurrent neural network (RNN) or a generative adversarial network (GAN).
   }
   \item {
-    My second point.
+     Real experimental implementation of the DNN decoders could be tested to confirm their capabilities.
   }
@@
 \section*{Acknowledgment}
-
 \begin{frame}{Acknowledgment}
@@
   \begin{itemize}
-   \item{The guidance and advice from my supervisor Meryem Benammar, Ph.D;}
+   \item{The guidance and advice from my supervisor Mme. Meryem Benammar, Ph.D;}
   \item{ My colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research;}
-   \item{Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project; and }
-   \item{The supervisors, the LACS member and the audience present today.}
+   \item{Mme. Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project; and }
+   \item{The evaluators, the LACS member and the audience.}
   \end{itemize}
-
 \end{frame}

### 2019-08-26 Rapport/Presentation PIR/Presentation/Presentation.tex
Inicio TG

--- 2019-06-25 Rapport/Presentation/Presentation/Presentation.tex
@@
 
-\author{E.~Dadalto Camara Gomes\inst{1}}
+\author{E.~Dadalto Camara Gomes\inst{1} \and M.~Benammar\inst{2}}
 % - Give the names in the same order as the appear in the paper.
@@
 Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr
-%  \and
- % \inst{2}%
- % Department of Electronics, Optronics, and Signal processing\\
- % ISAE-SUPAERO\\
- % 31055, Toulouse, France\\
- % Email: meryem.benammar@isae-supaero.fr
+  \and
+  \inst{2}%
+  Department of Electronics, Optronics, and Signal processing\\
+  ISAE-SUPAERO\\
+  31055, Toulouse, France\\
+  Email: meryem.benammar@isae-supaero.fr
  }
@@
  \1  ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths.
-   \2 It translates to lower decoding latency, resulting ultimately in gain of bandwidth over standard methods.
 
@@
 
-\begin{outline}
-\1 Could ML based decoders and autoencoders for a BSC perform similarly to the MAP decoder in real applications and have lower communication delay?
+\begin{block}{Hypothesis}
+Could ML based decoders and autoencoders for a BSC perform similarly to the MAP decoder in real applications and have lower communication delay?
+\end{block}
 \pause
-\1 This work aims to contribute to set up a higher standard in terms of performance in bit-error correction and demonstrate that ML based models can reduce delay for digital communication applications.
-\end{outline}
-
+\begin{block}{Objective}
+This work aims to contribute to set up a higher standard in terms of performance in bit-error correction and demonstrate that ML based models can reduce delay for digital communication applications. 
+\end{block}
 \end{frame}
@@
 
-\begin{frame}{DNN Autoencoder Design}{}
-	\begin{figure}[!ht]
-  \centering
-    \includegraphics[width=0.6\textwidth]{images/DNN_autoencoder}
-    \caption{Representation of a DNN autoencoder composed of dense layers.
+%%\begin{frame}{DNN Autoencoder Design}{}
+%	\begin{figure}[!ht]
+%  \centering
+%    \includegraphics[width=0.6\textwidth]{images/DNN_autoencoder}
+%%    \caption{Representation of a DNN autoencoder composed of dense layers.
 Each block represents a DNN which together learn the channel encoding and decoding in an end-to-end manner.
 They also operate together in bit error correction.}\label{fig:DDNNAutoencoder}
-\end{figure}
-\end{frame}
+%\end{figure}
+%\end{frame}
 
@@
  & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
- \multirow{2}{*}{Decoder} & Dense: 128, BN, input size: 16 \\
+ \multirow{2}{*}{Decoder} & Dense: 128, activation: ReLU, BN, input size: 16 \\
  & Dense: 64, activation: ReLU, BN  \\
@@
 
-\section*{Acknowledgment}
-\begin{frame}{Acknowledgment}
-  I would like to thank ...
-  \begin{itemize}
-   \item{The guidance and advice from my supervisor Mme. Meryem Benammar, Ph.D;}
-  \item{ My colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research;}
-   \item{Mme. Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project; and }
-   \item{The evaluators, the LACS member and the audience.}
-  \end{itemize}
-\end{frame}
+%%\section*{Acknowledgment}
+%\begin{frame}{Acknowledgment}
+%  I would like to thank ...
+%  \begin{itemize}
+%   \item{The guidance and advice from my supervisor Mme. Meryem Benammar, Ph.D;}
+%  \item{ My colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research;}
+%   \item{Mme. Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project; and }
+ %  \item{The evaluators, the LACS member and the audience.}
+ % \end{itemize}
+%\end{frame}
 
